---
title: "stylometryEnsemble: Introduction"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# This notebook assumes it is run from inside the notebooks/ folder.
# Folder layout:
#   stylometryEnsemble/
#     R/
#     data/
#     notebooks/

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

## 1. Confirm the raw KJV file exists

Place your raw file here:

```text
data/raw/kjv.csv
```

It should have this header:

```text
corpus_id,translation,book,chapter,verse,text
```

```{r check-file}
raw_path <- "./../data/raw/kjv.csv"

if(!file.exists(raw_path)){
  stop(
    paste0(
      "Could not find raw KJV file at: ", raw_path,
      "\nPlace kjv.csv in data/raw/kjv.csv"
    )
  )
}

file.exists(raw_path)
```

## 2. Import raw verses

```{r import}
kjv <- readr::read_csv(
  raw_path,
  show_col_types = FALSE
)

dplyr::glimpse(kjv)
```

## 3. Target books

```{r target-books}
target_books <- c(
  "Isaiah",
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings",
  "Psalms", "Job", "Proverbs"
)
```

## 4. Compatibility helper for chapter preparation

The architecture uses two prepared datasets:

- `chapter_docs`: source documents, usually one row per book chapter
- `section_docs`: target/reference documents, usually one row per section/book

If your local `R/prepare_data.R` does not yet include `prepare_kjv_chapters()`, this notebook defines it locally so the example can run.

```{r compatibility-prepare-chapters}
if(!exists("prepare_kjv_chapters")){

  prepare_kjv_chapters <- function(kjv, target_books = NULL){

    if(!is.null(target_books)){
      kjv <- kjv %>%
        dplyr::filter(book %in% target_books)
    }

    kjv <- kjv %>%
      dplyr::mutate(
        section = dplyr::case_when(
          book == "Isaiah" & chapter <= 35 ~ "Isaiah_1_35",
          book == "Isaiah" & chapter <= 39 ~ "Isaiah_36_39",
          book == "Isaiah" & chapter <= 55 ~ "Isaiah_40_55",
          book == "Isaiah" & chapter <= 66 ~ "Isaiah_56_66",
          TRUE ~ book
        ),

        macro_genre = dplyr::case_when(
          book %in% c("1 Samuel", "2 Samuel", "1 Kings", "2 Kings") ~ "Narrative",
          book %in% c("Psalms", "Job", "Proverbs") ~ "Poetic/Wisdom",
          TRUE ~ "Prophetic"
        )
      )

    kjv %>%
      dplyr::group_by(
        corpus_id,
        translation,
        book,
        chapter,
        section,
        macro_genre
      ) %>%
      dplyr::summarize(
        text = stringr::str_c(text, collapse = " "),
        n_verses = dplyr::n(),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        doc_id = paste(book, chapter, sep = "_")
      )
  }
}
```

## 5. Prepare chapter-level source documents

These are the documents being scored.

```{r prepare-chapters}
chapter_docs <- prepare_kjv_chapters(kjv)

chapter_docs %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  ) %>%
  dplyr::arrange(book, chapter)
```

## 6. Prepare section-level target documents

These are the reference targets.

```{r prepare-sections}
section_docs <- prepare_kjv_sections(kjv)

section_docs %>%
  dplyr::select(
    doc_id,
    section,
    macro_genre,
    n_verses,
    n_chapters,
    books
  )
```

## 7. Save prepared documents

```{r save-prepared}
dir.create(
  "./../data/prepared",
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  chapter_docs,
  "./../data/prepared/kjv_chapter_docs.csv"
)

readr::write_csv(
  section_docs,
  "./../data/prepared/kjv_section_docs.csv"
)
```

## 8. Run chapter-to-section ensemble

This is the main introductory example.

Each chapter is compared against each section/book target.

```{r run-chapter-to-section-ensemble}
config <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "punctuation",
    "sentence",
    "word_length",
    "lexical"
  )
)

result <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config
)
```



```{r display_results}

show_nearest_targets(result)


show_nearest_targets(
  result,
  doc_id = "Isaiah_40"
);

show_feature_votes(
  result,
  doc_id = "Isaiah_40"
);

show_feature_rankings(
  result,
  doc_id = "Isaiah_40"
);

show_top_feature_rankings(
  result,
  doc_id = "Isaiah_40"
);

show_weights(result);

explain_doc(
  result,
  "Isaiah_40"
);

explain_isaiah_chapter(
  result,
  40
);

```


## 9. Overall chapter-to-section similarity

Rows are source chapters. Columns are target sections.

```{r chapter-section-similarity-preview}
sim <- round(
  result$overall_similarity,
  3
)

sim[1:min(10, nrow(sim)), 1:min(10, ncol(sim))]
```

## 10. Nearest target sections

Depending on your local version of `run_stylometry_ensemble()`, this object may be named either `nearest_targets` or `nearest_neighbors`.

```{r nearest-target-sections}
nearest_tbl <- if("nearest_targets" %in% names(result)){
  result$nearest_targets
} else {
  result$nearest_neighbors
}

nearest_tbl %>%
  dplyr::arrange(doc_id, rank)
```

## 11. Isaiah example: target sections for Isaiah 40

```{r isaiah-40-target-sections}
nearest_tbl %>%
  dplyr::filter(doc_id == "Isaiah_40") %>%
  dplyr::arrange(rank)
```

## 12. Optional: chapter-to-chapter ensemble

This mode compares chapters directly to other chapters.

```{r run-chapter-to-chapter-ensemble}
chapter_result <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  config = config
)

round(
  chapter_result$overall_similarity[1:10, 1:10],
  3
)
```

## 13. Optional: nearest chapter neighbors

```{r nearest-chapter-neighbors}
chapter_neighbors <- if("nearest_targets" %in% names(chapter_result)){
  chapter_result$nearest_targets
} else {
  chapter_result$nearest_neighbors
}

chapter_neighbors %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        section,
        macro_genre
      ),
    by = "doc_id"
  ) %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        neighbor_doc_id = doc_id,
        neighbor_section = section,
        neighbor_macro_genre = macro_genre
      ),
    by = "neighbor_doc_id"
  ) %>%
  dplyr::arrange(doc_id, rank) %>%
  dplyr::select(
    doc_id,
    section,
    neighbor_doc_id,
    neighbor_section,
    similarity,
    rank
  )
```

## 14. Optional: section-to-section ensemble

This mode compares whole sections/books directly.

```{r run-section-to-section-ensemble}
section_result <- run_stylometry_ensemble(
  source_docs = section_docs,
  config = config
)

round(
  section_result$overall_similarity,
  3
)
```
