---
title: "02 - KJV Isaiah Example"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

## 1. Load prepared data if available

This notebook assumes `01_Introduction.Rmd` has already created:

```text
data/prepared/kjv_chapter_docs.csv
data/prepared/kjv_section_docs.csv
```

If those files do not exist, this notebook will prepare them from:

```text
data/raw/kjv.csv
```

```{r load-or-prepare}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(
    chapter_path,
    show_col_types = FALSE
  )

  section_docs <- readr::read_csv(
    section_path,
    show_col_types = FALSE
  )

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd after placing kjv.csv in data/raw/.")
  }

  kjv <- readr::read_csv(
    raw_path,
    show_col_types = FALSE
  )

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(
    file.path(STYLO_ROOT, "data", "prepared"),
    recursive = TRUE,
    showWarnings = FALSE
  )

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 2. Inspect Isaiah chapter documents

```{r inspect-isaiah-chapters}
chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )
```

## 3. Inspect target sections

```{r inspect-sections}
section_docs %>%
  dplyr::select(
    doc_id,
    section,
    macro_genre,
    n_verses,
    n_chapters,
    books
  )
```

## 4. Configure a style-oriented ensemble

This example uses a lightweight style-oriented ensemble and excludes POS/readability for speed.

```{r config}
config <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "punctuation",
    "sentence",
    "lexical"
  )
)

show_weights_placeholder <- config$weights
show_weights_placeholder
```

## 5. Run chapter-to-section ensemble

Rows are chapters. Columns are sections/books.

```{r run-chapter-to-section}
result <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config
)

show_result_summary(result)
show_weights(result)
```

## 6. Overall Isaiah chapter-to-section matrix

```{r isaiah-section-matrix}
isaiah_docs <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::pull(doc_id)

isaiah_targets <- c(
  "Isaiah_1_35",
  "Isaiah_36_39",
  "Isaiah_40_55",
  "Isaiah_56_66",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos"
)

show_overall_similarity(
  result,
  docs = isaiah_docs,
  targets = isaiah_targets,
  digits = 3
)
```

## 7. Nearest target sections for Isaiah 40

```{r isaiah-40-nearest}
show_nearest_targets(
  result,
  doc_id = "Isaiah_40",
  k = 10
)
```

## 8. Feature votes for Isaiah 40

```{r isaiah-40-votes}
show_feature_votes(
  result,
  doc_id = "Isaiah_40"
)
```

## 9. Top target by every feature variant for Isaiah 40

```{r isaiah-40-feature-winners}
show_top_feature_rankings(
  result,
  doc_id = "Isaiah_40"
)
```

## 10. Full feature rankings for Isaiah 40

```{r isaiah-40-all-rankings}
show_feature_rankings(
  result,
  doc_id = "Isaiah_40",
  rank_max = 5
)
```

## 11. Explain Isaiah 40 in one object

```{r explain-isaiah-40}
isa40 <- explain_doc(
  result,
  "Isaiah_40",
  k = 10
)

names(isa40)

isa40$nearest_targets
isa40$feature_votes
isa40$top_feature_rankings
```

## 12. Compare all chapters in Isaiah 40-55

```{r deutero-block-votes}
deutero_docs <- chapter_docs %>%
  dplyr::filter(section == "Isaiah_40_55") %>%
  dplyr::pull(doc_id)

result$feature_votes %>%
  dplyr::filter(doc_id %in% deutero_docs) %>%
  dplyr::arrange(doc_id, dplyr::desc(votes), dplyr::desc(avg_similarity))
```

## 13. Which target wins most often for Isaiah 40-55?

```{r deutero-section-vote-summary}
result$feature_votes %>%
  dplyr::filter(doc_id %in% deutero_docs) %>%
  dplyr::group_by(neighbor_doc_id) %>%
  dplyr::summarize(
    total_votes = sum(votes, na.rm = TRUE),
    mean_avg_similarity = mean(avg_similarity, na.rm = TRUE),
    n_chapters = dplyr::n_distinct(doc_id),
    .groups = "drop"
  ) %>%
  dplyr::arrange(
    dplyr::desc(total_votes),
    dplyr::desc(mean_avg_similarity)
  )
```

## 14. Proto-Isaiah vs Deutero-Isaiah diagnostic table

This table compares each Isaiah chapter against the four Isaiah target sections.

```{r isaiah-diagnostic-table}
diagnostic_targets <- c(
  "Isaiah_1_35",
  "Isaiah_36_39",
  "Isaiah_40_55",
  "Isaiah_56_66"
)

diag_mat <- show_overall_similarity(
  result,
  docs = isaiah_docs,
  targets = diagnostic_targets,
  digits = 4
)

diag_tbl <- as.data.frame(diag_mat) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, chapter, section),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, chapter, section)

diag_tbl
```


```{r show consensus}

show_consensus(
  result,
  "Isaiah_40"
)

```


## 15. Best Isaiah target per chapter

```{r best-isaiah-target-per-chapter}
diag_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(diagnostic_targets),
    names_to = "target_section",
    values_to = "similarity"
  ) %>%
  dplyr::group_by(doc_id, chapter, section) %>%
  dplyr::slice_max(
    order_by = similarity,
    n = 1,
    with_ties = FALSE
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(chapter)
```

## 16. Save Isaiah example outputs

```{r save-outputs}
out_dir <- file.path(STYLO_ROOT, "output", "summaries")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  show_nearest_targets(result, doc_id = "Isaiah_40", k = 10),
  file.path(out_dir, "isaiah_40_nearest_targets.csv")
)

readr::write_csv(
  show_feature_votes(result, doc_id = "Isaiah_40"),
  file.path(out_dir, "isaiah_40_feature_votes.csv")
)

readr::write_csv(
  show_top_feature_rankings(result, doc_id = "Isaiah_40"),
  file.path(out_dir, "isaiah_40_top_feature_rankings.csv")
)

readr::write_csv(
  diag_tbl,
  file.path(out_dir, "isaiah_diagnostic_table.csv")
)
```
