---
title: "03 - POS and Readability Features"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

## 1. Purpose

This notebook extends the earlier lightweight ensemble by adding:

- POS n-grams: `pos3`, `pos4`, `pos5`
- readability features

POS tagging is usually slower than character, word, function-word, punctuation, sentence, and lexical features. Use cache when possible.

## 2. Load prepared data

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(
    chapter_path,
    show_col_types = FALSE
  )

  section_docs <- readr::read_csv(
    section_path,
    show_col_types = FALSE
  )

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(
    raw_path,
    show_col_types = FALSE
  )

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(
    file.path(STYLO_ROOT, "data", "prepared"),
    recursive = TRUE,
    showWarnings = FALSE
  )

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}

chapter_docs %>%
  dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses) %>%
  dplyr::slice_head(n = 10)

section_docs %>%
  dplyr::select(doc_id, section, macro_genre, n_verses, n_chapters, books)
```

## 3. Check packages

The POS feature class normally requires `udpipe`.

The readability feature class may require `quanteda` and `quanteda.textstats`, depending on how your `features_readability.R` is written.

```{r check-packages}
required_optional <- c(
  "udpipe",
  "quanteda",
  "quanteda.textstats"
)

available_optional <- vapply(
  required_optional,
  requireNamespace,
  logical(1),
  quietly = TRUE
)

data.frame(
  package = required_optional,
  available = available_optional,
  row.names = NULL
)
```

## 4. POS model setup

If your POS functions require a loaded `ud_model`, use this chunk.

The model will be stored under:

```text
models/
```

```{r pos-model-setup}
model_dir <- file.path(STYLO_ROOT, "models")

dir.create(
  model_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

if(requireNamespace("udpipe", quietly = TRUE)){

  model_file <- file.path(
    model_dir,
    "english-ewt-ud-2.5-191206.udpipe"
  )

  if(!file.exists(model_file)){

    message("Downloading udpipe English model...")

    downloaded <- udpipe::udpipe_download_model(
      language = "english",
      model_dir = model_dir
    )

    model_file <- downloaded$file_model
  }

  ud_model <- udpipe::udpipe_load_model(model_file)

} else {

  ud_model <- NULL

  warning("Package 'udpipe' is not installed. POS features will not run.")
}
```

## 5. POS + readability configuration

This configuration includes the earlier features plus POS and readability.

```{r config-pos-readability}
config_pos_readability <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "lexical",
    "readability"
  )
)

config_pos_readability$pos_n <- c(3, 4, 5)

show_weights_placeholder <- config_pos_readability$weights
show_weights_placeholder
```

## 6. Ensure POS model is in config if needed

Different versions of the codebase may expect the model in different places. This chunk supports the common pattern:

```r
config$pos_model <- ud_model
```

```{r attach-pos-model}
if(!is.null(ud_model)){
  config_pos_readability$pos_model <- ud_model
}
```

## 7. Run chapter-to-section ensemble with POS/readability

This may take longer the first time.

```{r run-pos-readability}
result_pos_readability <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_pos_readability,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_pos_readability)
show_weights(result_pos_readability)
```

## 8. Overall Isaiah 40 result

```{r isaiah-40-nearest-pos-readability}
show_nearest_targets(
  result_pos_readability,
  doc_id = "Isaiah_40",
  k = 10
)
```

## 9. Consensus votes for Isaiah 40

```{r isaiah-40-votes-pos-readability}
show_feature_votes(
  result_pos_readability,
  doc_id = "Isaiah_40"
)
```

If your display helpers include `show_consensus()`, this gives vote percentages:

```{r isaiah-40-consensus-pos-readability, eval=exists("show_consensus")}
cons <- show_consensus(
  result_pos_readability,
  "Isaiah_40"
)

cons

attr(cons, "summary")

show_consensus_summary(cons)
```

## 10. Feature winners for Isaiah 40

```{r isaiah-40-feature-winners-pos-readability}
show_top_feature_rankings(
  result_pos_readability,
  doc_id = "Isaiah_40"
)
```

## 11. POS-only class similarity

This lets you inspect the POS class independent of the whole ensemble.

```{r pos-class-similarity}
if("pos" %in% names(result_pos_readability$class_similarity)){

  show_class_similarity(
    result_pos_readability,
    class_name = "pos",
    docs = "Isaiah_40",
    digits = 3
  )

} else {

  message("No POS class found in result_pos_readability$class_similarity.")
}
```

## 12. POS 3, POS 4, POS 5 variants for Isaiah 40

```{r pos-variant-similarity}
if("pos" %in% names(result_pos_readability$variant_similarity)){

  pos_variants <- names(result_pos_readability$variant_similarity$pos)

  pos_variants

  lapply(
    pos_variants,
    function(v){
      show_variant_similarity(
        result_pos_readability,
        class_name = "pos",
        variant_name = v,
        docs = "Isaiah_40",
        digits = 3
      )
    }
  )

} else {

  message("No POS variants found.")
}
```

## 13. Readability class similarity

```{r readability-class-similarity}
if("readability" %in% names(result_pos_readability$class_similarity)){

  show_class_similarity(
    result_pos_readability,
    class_name = "readability",
    docs = "Isaiah_40",
    digits = 3
  )

} else {

  message("No readability class found in result_pos_readability$class_similarity.")
}
```

## 14. Compare Isaiah 40-55 consensus

```{r deutero-consensus-pos-readability}
deutero_docs <- chapter_docs %>%
  dplyr::filter(section == "Isaiah_40_55") %>%
  dplyr::pull(doc_id)

result_pos_readability$feature_votes %>%
  dplyr::filter(doc_id %in% deutero_docs) %>%
  dplyr::group_by(doc_id) %>%
  dplyr::mutate(
    pct_votes = votes / sum(votes)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    doc_id,
    dplyr::desc(votes),
    dplyr::desc(avg_similarity)
  )
```

## 15. Section-level vote summary for Isaiah 40-55

```{r deutero-section-summary-pos-readability}
result_pos_readability$feature_votes %>%
  dplyr::filter(doc_id %in% deutero_docs) %>%
  dplyr::group_by(neighbor_doc_id) %>%
  dplyr::summarize(
    total_votes = sum(votes, na.rm = TRUE),
    mean_avg_similarity = mean(avg_similarity, na.rm = TRUE),
    n_source_chapters = dplyr::n_distinct(doc_id),
    .groups = "drop"
  ) %>%
  dplyr::arrange(
    dplyr::desc(total_votes),
    dplyr::desc(mean_avg_similarity)
  )
```

## 16. Diagnostic table: Isaiah chapters against Isaiah sections

```{r isaiah-diagnostic-pos-readability}
isaiah_docs <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::pull(doc_id)

diagnostic_targets <- c(
  "Isaiah_1_35",
  "Isaiah_36_39",
  "Isaiah_40_55",
  "Isaiah_56_66"
)

diag_mat <- show_overall_similarity(
  result_pos_readability,
  docs = isaiah_docs,
  targets = diagnostic_targets,
  digits = 4
)

diag_tbl <- as.data.frame(diag_mat) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, chapter, section),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, chapter, section)

diag_tbl
```

## 17. Best Isaiah section target per chapter

```{r best-isaiah-target-pos-readability}
diag_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(diagnostic_targets),
    names_to = "target_section",
    values_to = "similarity"
  ) %>%
  dplyr::group_by(doc_id, chapter, section) %>%
  dplyr::slice_max(
    order_by = similarity,
    n = 1,
    with_ties = FALSE
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(chapter)
```

## 18. Save POS/readability outputs

```{r save-pos-readability-outputs}
out_dir <- file.path(STYLO_ROOT, "output", "summaries")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  show_nearest_targets(result_pos_readability, doc_id = "Isaiah_40", k = 10),
  file.path(out_dir, "isaiah_40_nearest_targets_pos_readability.csv")
)

readr::write_csv(
  show_feature_votes(result_pos_readability, doc_id = "Isaiah_40"),
  file.path(out_dir, "isaiah_40_feature_votes_pos_readability.csv")
)

readr::write_csv(
  show_top_feature_rankings(result_pos_readability, doc_id = "Isaiah_40"),
  file.path(out_dir, "isaiah_40_top_feature_rankings_pos_readability.csv")
)

readr::write_csv(
  diag_tbl,
  file.path(out_dir, "isaiah_diagnostic_table_pos_readability.csv")
)
```
