---
title: "04 - Author Identification / Attribution Mode"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 100
)
```

## 1. Purpose

This notebook treats the stylometry ensemble as an attribution engine.

The core question is:

> For each Isaiah chapter, which section/book profile receives the strongest ensemble consensus?

This notebook runs:

```text
source_docs = chapter_docs
target_docs = section_docs
```

Then summarizes every Isaiah chapter using:

- consensus winner
- vote share
- weighted vote share
- vote margin
- weighted vote margin
- dominance ratio
- consensus strength
- entropy
- effective targets

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(
    chapter_path,
    show_col_types = FALSE
  )

  section_docs <- readr::read_csv(
    section_path,
    show_col_types = FALSE
  )

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(
    raw_path,
    show_col_types = FALSE
  )

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(
    file.path(STYLO_ROOT, "data", "prepared"),
    recursive = TRUE,
    showWarnings = FALSE
  )

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Isaiah chapter list

```{r isaiah-chapters}
isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_doc_ids <- isaiah_chapters$doc_id

isaiah_chapters %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    macro_genre,
    n_verses
  )
```

## 4. Target profiles

```{r targets}
section_docs %>%
  dplyr::select(
    doc_id,
    section,
    macro_genre,
    n_verses,
    n_chapters,
    books
  )
```

## 5. Configure attribution ensemble

This configuration includes POS and readability if your installation supports them.

If POS is too slow or unavailable, remove `"pos"` from `include`.

```{r config}
config_author_id <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "lexical",
    "readability"
  )
)

config_author_id$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_author_id$pos_model <- ud_model
}

config_author_id$weights
```

## 6. Run chapter-to-section attribution ensemble

```{r run-attribution}
result_author_id <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_author_id,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_author_id)
show_weights(result_author_id)
```

## 7. Build consensus summary for each Isaiah chapter

```{r all-isaiah-consensus}
isaiah_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_author_id,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_consensus
```

## 8. Clean attribution table

```{r clean-attribution-table}
isaiah_attribution <- isaiah_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,

    winner_pct_votes_pct =
      100 * winner_pct_votes,

    winner_weighted_pct_votes_pct =
      100 * winner_weighted_pct_votes,

    vote_margin_pct =
      100 * vote_margin,

    weighted_vote_margin_pct =
      100 * weighted_vote_margin
  ) %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_matches_actual_section,
    winner_pct_votes_pct,
    winner_weighted_pct_votes_pct,
    vote_margin_pct,
    weighted_vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    elbow_rank,
    doc_entropy,
    effective_targets,
    n_targets
  )

isaiah_attribution
```


```{r false matches}
show_false_matches(
  isaiah_attribution
)

summarize_findings(
    result_author_id,
    "Isaiah_12"
)

summarize_findings(
    result_author_id,
    "Isaiah_26"
)
```



## 9. Winner counts by actual Isaiah section

```{r winner-counts-by-section}
isaiah_attribution %>%
  dplyr::count(
    actual_section,
    winner,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(actual_section) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    actual_section,
    dplyr::desc(n_chapters)
  )
```

## 10. Overall winner counts across Isaiah

```{r overall-winner-counts}
isaiah_attribution %>%
  dplyr::count(
    winner,
    name = "n_chapters"
  ) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_chapters),
    winner
  )
```

## 11. Isaiah section self-identification rate

This measures how often a chapter's consensus winner is its own assigned Isaiah section.

```{r self-identification-rate}
isaiah_attribution %>%
  dplyr::group_by(actual_section) %>%
  dplyr::summarize(
    n_chapters = dplyr::n(),
    n_self_identified = sum(winner_matches_actual_section, na.rm = TRUE),
    pct_self_identified = n_self_identified / n_chapters,
    mean_winner_pct_votes = mean(winner_pct_votes_pct, na.rm = TRUE),
    mean_weighted_winner_pct_votes = mean(winner_weighted_pct_votes_pct, na.rm = TRUE),
    mean_vote_margin = mean(vote_margin_pct, na.rm = TRUE),
    mean_weighted_vote_margin = mean(weighted_vote_margin_pct, na.rm = TRUE),
    mean_entropy = mean(doc_entropy, na.rm = TRUE),
    mean_effective_targets = mean(effective_targets, na.rm = TRUE),
    .groups = "drop"
  )
```

## 12. Consensus strength by Isaiah chapter

```{r consensus-strength-by-chapter}
isaiah_attribution %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    doc_entropy,
    effective_targets
  ) %>%
  dplyr::arrange(chapter)
```

## 13. Chapters with strongest consensus

```{r strongest-consensus}
isaiah_attribution %>%
  dplyr::arrange(
    dplyr::desc(winner_pct_votes_pct),
    dplyr::desc(vote_margin_pct)
  ) %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    doc_entropy
  ) %>%
  dplyr::slice_head(n = 15)
```

## 14. Chapters with weakest consensus

```{r weakest-consensus}
isaiah_attribution %>%
  dplyr::arrange(
    winner_pct_votes_pct,
    vote_margin_pct
  ) %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    doc_entropy,
    effective_targets
  ) %>%
  dplyr::slice_head(n = 15)
```

## 15. Detailed findings for selected Isaiah chapters

```{r selected-chapter-findings}
summarize_findings(
  result_author_id,
  doc_id = "Isaiah_40",
  k = 10,
  plot = TRUE
)
```

Try additional chapters:

```{r selected-chapters-extra, eval=FALSE}
summarize_findings(result_author_id, "Isaiah_1",  k = 10, plot = TRUE)
summarize_findings(result_author_id, "Isaiah_35", k = 10, plot = TRUE)
summarize_findings(result_author_id, "Isaiah_53", k = 10, plot = TRUE)
summarize_findings(result_author_id, "Isaiah_66", k = 10, plot = TRUE)
```

## 16. Consensus scree plots for representative chapters

```{r representative-scree-plots}
cons_isa1 <- show_consensus(result_author_id, "Isaiah_1")
plot_consensus_scree(cons_isa1)

cons_isa40 <- show_consensus(result_author_id, "Isaiah_40")
plot_consensus_scree(cons_isa40)

cons_isa53 <- show_consensus(result_author_id, "Isaiah_53")
plot_consensus_scree(cons_isa53)

cons_isa66 <- show_consensus(result_author_id, "Isaiah_66")
plot_consensus_scree(cons_isa66)
```

## 17. Feature winners for all Isaiah chapters

```{r feature-winners-all-isaiah}
isaiah_feature_winners <- result_author_id$feature_rankings %>%
  dplyr::filter(
    doc_id %in% isaiah_doc_ids,
    rank == 1
  ) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    feature_class,
    feature_variant,
    neighbor_doc_id
  ) %>%
  dplyr::arrange(
    chapter,
    feature_class,
    feature_variant
  )

isaiah_feature_winners
```

## 18. Feature-class winner patterns

```{r feature-class-winner-patterns}
isaiah_feature_winners %>%
  dplyr::count(
    actual_section,
    feature_class,
    neighbor_doc_id,
    name = "wins"
  ) %>%
  dplyr::group_by(
    actual_section,
    feature_class
  ) %>%
  dplyr::mutate(
    pct_wins = wins / sum(wins)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    actual_section,
    feature_class,
    dplyr::desc(wins)
  )
```

## 19. Isaiah 40-55 attribution profile

```{r deutero-attribution-profile}
isaiah_attribution %>%
  dplyr::filter(actual_section == "Isaiah_40_55") %>%
  dplyr::select(
    chapter,
    doc_id,
    winner,
    winner_matches_actual_section,
    winner_pct_votes_pct,
    winner_weighted_pct_votes_pct,
    vote_margin_pct,
    weighted_vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    doc_entropy,
    effective_targets
  ) %>%
  dplyr::arrange(chapter)
```

## 20. Isaiah 1-35 attribution profile

```{r proto-attribution-profile}
isaiah_attribution %>%
  dplyr::filter(actual_section == "Isaiah_1_35") %>%
  dplyr::select(
    chapter,
    doc_id,
    winner,
    winner_matches_actual_section,
    winner_pct_votes_pct,
    winner_weighted_pct_votes_pct,
    vote_margin_pct,
    weighted_vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    doc_entropy,
    effective_targets
  ) %>%
  dplyr::arrange(chapter)
```

## 21. Export author-identification outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "author_identification")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  isaiah_consensus,
  file.path(out_dir, "isaiah_consensus_summary.csv")
)

readr::write_csv(
  isaiah_attribution,
  file.path(out_dir, "isaiah_attribution_table.csv")
)

readr::write_csv(
  isaiah_feature_winners,
  file.path(out_dir, "isaiah_feature_winners.csv")
)

readr::write_csv(
  result_author_id$feature_votes %>%
    dplyr::filter(doc_id %in% isaiah_doc_ids),
  file.path(out_dir, "isaiah_feature_votes_all.csv")
)

readr::write_csv(
  result_author_id$nearest_targets %>%
    dplyr::filter(doc_id %in% isaiah_doc_ids),
  file.path(out_dir, "isaiah_nearest_targets_all.csv")
)
```
