---
title: "05 - Genre Identification"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 100
)
```

## 1. Purpose

This notebook tests whether the ensemble can identify **macro-genre** rather than section/book authorship.

The target profiles are:

```text
Narrative
Prophetic
Poetic/Wisdom
```

This is useful because some Isaiah chapters, such as Isaiah 12 and Isaiah 26, behave like hymns/songs and may classify closer to Psalms or the broader Poetic/Wisdom profile.

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(
    chapter_path,
    show_col_types = FALSE
  )

  section_docs <- readr::read_csv(
    section_path,
    show_col_types = FALSE
  )

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(
    raw_path,
    show_col_types = FALSE
  )

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(
    file.path(STYLO_ROOT, "data", "prepared"),
    recursive = TRUE,
    showWarnings = FALSE
  )

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Build genre target profiles

`chapter_docs` are the source documents.

`genre_docs` are the target profiles, formed by concatenating all chapters within each macro genre.

```{r build-genre-docs}
genre_docs <- chapter_docs %>%
  dplyr::group_by(
    corpus_id,
    translation,
    macro_genre
  ) %>%
  dplyr::summarize(
    text = stringr::str_c(text, collapse = " "),
    n_verses = sum(n_verses, na.rm = TRUE),
    n_chapters = dplyr::n(),
    books = stringr::str_c(sort(unique(book)), collapse = "; "),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    doc_id = macro_genre,
    section = macro_genre
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    section,
    macro_genre,
    n_verses,
    n_chapters,
    books,
    text
  )

genre_docs
```

## 4. Isaiah chapter source documents

```{r isaiah-chapters}
isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_doc_ids <- isaiah_chapters$doc_id

isaiah_chapters %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    macro_genre,
    n_verses
  )
```

## 5. Configure genre ensemble

This configuration includes all major feature classes, including POS and readability, if available.

If POS is too slow or unavailable, remove `"pos"` from `include`.

```{r config}
config_genre <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_genre$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_genre$pos_model <- ud_model
}

config_genre$weights
```

## 6. Run chapter-to-genre ensemble

```{r run-genre}
result_genre <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = genre_docs,
  config = config_genre,
  k_neighbors = 3,
  use_cache = TRUE
)

show_result_summary(result_genre)
show_weights(result_genre)
```

## 7. Overall Isaiah chapter-to-genre matrix

```{r isaiah-genre-matrix}
show_overall_similarity(
  result_genre,
  docs = isaiah_doc_ids,
  targets = genre_docs$doc_id,
  digits = 3
)
```

## 8. Consensus summaries for every Isaiah chapter

```{r isaiah-genre-consensus}
isaiah_genre_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_genre,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        isaiah_section = section,
        assigned_macro_genre = macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    isaiah_section,
    assigned_macro_genre,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_genre_consensus
```

## 9. Clean genre attribution table

```{r clean-genre-attribution}
isaiah_genre_attribution <- isaiah_genre_consensus %>%
  dplyr::mutate(
    winner_matches_assigned_macro_genre =
      winner == assigned_macro_genre,

    winner_pct_votes_pct =
      100 * winner_pct_votes,

    winner_weighted_pct_votes_pct =
      100 * winner_weighted_pct_votes,

    vote_margin_pct =
      100 * vote_margin,

    weighted_vote_margin_pct =
      100 * weighted_vote_margin
  ) %>%
  dplyr::select(
    chapter,
    doc_id,
    isaiah_section,
    assigned_macro_genre,
    winner,
    winner_matches_assigned_macro_genre,
    winner_pct_votes_pct,
    winner_weighted_pct_votes_pct,
    vote_margin_pct,
    weighted_vote_margin_pct,
    winner_dominance_ratio,
    consensus_strength,
    elbow_rank,
    doc_entropy,
    effective_targets,
    n_targets
  )

isaiah_genre_attribution
```

## 10. Isaiah chapters not classified as Prophetic

```{r non-prophetic-isaiah}
isaiah_genre_attribution %>%
  dplyr::filter(winner != "Prophetic") %>%
  dplyr::arrange(chapter)
```

## 11. Genre winner counts across Isaiah

```{r genre-winner-counts}
isaiah_genre_attribution %>%
  dplyr::count(
    winner,
    name = "n_chapters"
  ) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_chapters)
  )
```

## 12. Genre winner counts by Isaiah section

```{r genre-winner-counts-by-section}
isaiah_genre_attribution %>%
  dplyr::count(
    isaiah_section,
    winner,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(isaiah_section) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    isaiah_section,
    dplyr::desc(n_chapters)
  )
```

## 13. Section-level genre summary

```{r section-level-genre-summary}
isaiah_genre_attribution %>%
  dplyr::group_by(isaiah_section) %>%
  dplyr::summarize(
    n_chapters = dplyr::n(),
    n_prophetic = sum(winner == "Prophetic", na.rm = TRUE),
    n_poetic_wisdom = sum(winner == "Poetic/Wisdom", na.rm = TRUE),
    n_narrative = sum(winner == "Narrative", na.rm = TRUE),
    pct_prophetic = n_prophetic / n_chapters,
    pct_poetic_wisdom = n_poetic_wisdom / n_chapters,
    pct_narrative = n_narrative / n_chapters,
    mean_winner_pct_votes = mean(winner_pct_votes_pct, na.rm = TRUE),
    mean_weighted_winner_pct_votes = mean(winner_weighted_pct_votes_pct, na.rm = TRUE),
    mean_vote_margin = mean(vote_margin_pct, na.rm = TRUE),
    mean_entropy = mean(doc_entropy, na.rm = TRUE),
    mean_effective_targets = mean(effective_targets, na.rm = TRUE),
    .groups = "drop"
  )
```

## 14. Detailed findings for Isaiah 12 and Isaiah 26

These chapters previously classified closer to Psalms in section/book attribution. Here we test whether they classify as Poetic/Wisdom at the broader genre level.

```{r detailed-isaiah-12}
summarize_findings(
  result_genre,
  doc_id = "Isaiah_12",
  k = 3,
  plot = TRUE
)
```

```{r detailed-isaiah-26}
summarize_findings(
  result_genre,
  doc_id = "Isaiah_26",
  k = 3,
  plot = TRUE
)
```

## 15. Compare representative Isaiah chapters

```{r representative-genre-scree}
cons_isa1 <- show_consensus(result_genre, "Isaiah_1")
plot_consensus_scree(cons_isa1)

cons_isa12 <- show_consensus(result_genre, "Isaiah_12")
plot_consensus_scree(cons_isa12)

cons_isa26 <- show_consensus(result_genre, "Isaiah_26")
plot_consensus_scree(cons_isa26)

cons_isa40 <- show_consensus(result_genre, "Isaiah_40")
plot_consensus_scree(cons_isa40)

cons_isa53 <- show_consensus(result_genre, "Isaiah_53")
plot_consensus_scree(cons_isa53)

cons_isa66 <- show_consensus(result_genre, "Isaiah_66")
plot_consensus_scree(cons_isa66)
```

## 16. Feature winners for Isaiah 12 and Isaiah 26

```{r feature-winners-isaiah-12-26}
result_genre$feature_rankings %>%
  dplyr::filter(
    doc_id %in% c("Isaiah_12", "Isaiah_26"),
    rank == 1
  ) %>%
  dplyr::arrange(
    doc_id,
    feature_class,
    feature_variant
  )
```

## 17. Feature-class genre patterns for all Isaiah chapters

```{r feature-class-genre-patterns}
isaiah_genre_feature_winners <- result_genre$feature_rankings %>%
  dplyr::filter(
    doc_id %in% isaiah_doc_ids,
    rank == 1
  ) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        isaiah_section = section
      ),
    by = "doc_id"
  )

isaiah_genre_feature_winners %>%
  dplyr::count(
    isaiah_section,
    feature_class,
    neighbor_doc_id,
    name = "wins"
  ) %>%
  dplyr::group_by(
    isaiah_section,
    feature_class
  ) %>%
  dplyr::mutate(
    pct_wins = wins / sum(wins)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    isaiah_section,
    feature_class,
    dplyr::desc(wins)
  )
```

## 18. Export genre-identification outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "genre_identification")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  genre_docs,
  file.path(out_dir, "genre_target_profiles.csv")
)

readr::write_csv(
  isaiah_genre_consensus,
  file.path(out_dir, "isaiah_genre_consensus_summary.csv")
)

readr::write_csv(
  isaiah_genre_attribution,
  file.path(out_dir, "isaiah_genre_attribution_table.csv")
)

readr::write_csv(
  result_genre$feature_votes %>%
    dplyr::filter(doc_id %in% isaiah_doc_ids),
  file.path(out_dir, "isaiah_genre_feature_votes_all.csv")
)

readr::write_csv(
  result_genre$feature_rankings %>%
    dplyr::filter(doc_id %in% isaiah_doc_ids),
  file.path(out_dir, "isaiah_genre_feature_rankings_all.csv")
)
```
