---
title: "06 - Featurized PCA"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 150
)
```

## 1. Purpose

This notebook performs PCA on the **raw weighted feature matrix**.

This is different from the ensemble similarity analysis.

Featurized PCA answers:

> What broad stylistic dimensions emerge directly from the combined feature space?

Because the feature matrix is very large and sparse, individual PCs may explain small percentages of total variance. That is expected.

The companion notebook `07_Ensemble_MDS.Rmd` analyzes the compressed ensemble similarity space using:

```r
result$overall_similarity
result$class_similarity
```

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Configure feature set

```{r config}
config_pca <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_pca$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_pca$pos_model <- ud_model
}

config_pca$weights
```

## 4. Build weighted feature matrix

```{r feature-matrix-functions}
scale_feature_block <- function(mat){

  mat <- as.matrix(mat)
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0

  if(ncol(mat) == 0){
    return(mat)
  }

  scaled <- scale(mat)

  scaled[is.na(scaled)] <- 0
  scaled[is.nan(scaled)] <- 0
  scaled[is.infinite(scaled)] <- 0

  as.matrix(scaled)
}


build_pca_feature_matrix <- function(
    docs,
    config,
    text_col = "text",
    doc_id_col = "doc_id"){

  feature_sets <- build_feature_sets(
    docs,
    config,
    text_col = text_col,
    doc_id_col = doc_id_col
  )

  active_classes <- names(feature_sets)

  weights <- config$weights[active_classes]
  weights <- weights / sum(weights, na.rm = TRUE)

  blocks <- list()

  for(class_name in active_classes){

    class_variants <- feature_sets[[class_name]]

    for(variant_name in names(class_variants)){

      mat <- class_variants[[variant_name]]
      mat <- scale_feature_block(mat)

      mat <- mat * sqrt(weights[[class_name]]) / sqrt(length(class_variants))

      colnames(mat) <- paste(
        class_name,
        variant_name,
        colnames(mat),
        sep = "__"
      )

      blocks[[paste(class_name, variant_name, sep = "__")]] <- mat
    }
  }

  feature_matrix <- do.call(cbind, blocks)

  feature_matrix[is.na(feature_matrix)] <- 0
  feature_matrix[is.nan(feature_matrix)] <- 0
  feature_matrix[is.infinite(feature_matrix)] <- 0

  list(
    feature_matrix = feature_matrix,
    feature_sets = feature_sets,
    weights_used = weights
  )
}
```

```{r build-feature-matrix}
pca_input <- build_pca_feature_matrix(
  chapter_docs,
  config = config_pca
)

X <- pca_input$feature_matrix

dim(X)

pca_input$weights_used
```

## 5. Run feature-space PCA

```{r run-pca}
pca_features <- stats::prcomp(
  X,
  center = FALSE,
  scale. = FALSE
)

 pca_features$sdev^2
```

## 6. Variance explained

```{r variance-table}
pca_var <- pca_features$sdev^2

pca_variance <- data.frame(
  pc = paste0("PC", seq_along(pca_var)),
  dimension = seq_along(pca_var),
  eigenvalue = pca_var,
  prop_variance = pca_var / sum(pca_var),
  pct_variance = 100 * pca_var / sum(pca_var),
  cumulative_prop_variance = cumsum(pca_var / sum(pca_var)),
  cumulative_pct_variance = 100 * cumsum(pca_var / sum(pca_var))
)

pca_variance %>%
  dplyr::slice_head(n = 20)
```

## 7. Feature-space scree plot

```{r pca-scree}
pca_variance %>%
  dplyr::slice_head(n = 25) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = dimension,
      y = pct_variance
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::geom_line(
    ggplot2::aes(
      y = cumulative_pct_variance,
      group = 1
    )
  ) +
  ggplot2::geom_point(
    ggplot2::aes(
      y = cumulative_pct_variance
    )
  ) +
  ggplot2::scale_x_continuous(
    breaks = 1:25
  ) +
  ggplot2::labs(
    title = "Feature-space PCA scree plot",
    subtitle = "Low individual variance is expected in a high-dimensional sparse feature matrix.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  ggplot2::theme_minimal()
```

## 8. PCA scores

```{r scores}
pca_scores <- as.data.frame(pca_features$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

isaiah_pca_scores <- pca_scores %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_pca_scores %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    PC1,
    PC2,
    PC3,
    PC4,
    PC5,
    PC6,
    PC7,
    PC8,
    PC9
  )
```

## 9. PC1 vs PC2

```{r pc1-pc2}
plotPCA(
  pca_scores,
  pca_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Feature-space PCA: Isaiah PC1 vs PC2"
)
```

## 10. PC2 vs PC3

```{r pc2-pc3}
plotPCA(
  pca_scores,
  pca_variance,
  x = 2,
  y = 3,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Feature-space PCA: Isaiah PC2 vs PC3"
)
```

## 11. PC3 vs PC4

```{r pc3-pc4}
plotPCA(
  pca_scores,
  pca_variance,
  x = 3,
  y = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Feature-space PCA: Isaiah PC3 vs PC4"
)
```

## 12. Generate PC1–PC9 adjacent pair plots

```{r all-pair-plots}
pc_pairs <- tibble::tibble(
  pc_x = 1:8,
  pc_y = 2:9
)

pca_pair_plots <- purrr::pmap(
  pc_pairs,
  function(pc_x, pc_y){

    plotPCA(
      pca_scores,
      pca_variance,
      x = pc_x,
      y = pc_y,
      book = "Isaiah",
      color = "section",
      label = "chapter",
      show_labels = TRUE,
      show_centroids = TRUE,
      show_hulls = TRUE,
      title = paste0(
        "Feature-space PCA: Isaiah PC",
        pc_x,
        " vs PC",
        pc_y
      )
    )
  }
)

pca_pair_plots[[1]]
```

```{r print-all-pairs, eval=FALSE}
for(i in seq_along(pca_pair_plots)){
  print(pca_pair_plots[[i]])
}
```

## 13. Interactive 3D feature-space PCA

These are disabled for knitting. Run interactively in RStudio.

```{r pca3d-123, eval=FALSE}
plotPCA3d(
  pca_scores,
  pca_variance,
  x = 1,
  y = 2,
  z = 3,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  title = "Feature-space PCA: Isaiah PC1 vs PC2 vs PC3"
)
```

```{r pca3d-234, eval=FALSE}
plotPCA3d(
  pca_scores,
  pca_variance,
  x = 2,
  y = 3,
  z = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  title = "Feature-space PCA: Isaiah PC2 vs PC3 vs PC4"
)
```

## 14. Section centroids and outliers

```{r centroids}
pc_cols <- paste0("PC", 1:9)

section_centroids <- isaiah_pca_scores %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(
      dplyr::all_of(pc_cols),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

section_centroids
```

```{r centroid-distance}
centroid_lookup <- section_centroids %>%
  dplyr::select(section, dplyr::all_of(pc_cols))

isaiah_pca_distances <- isaiah_pca_scores %>%
  dplyr::left_join(
    centroid_lookup,
    by = "section",
    suffix = c("", "_centroid")
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    distance_to_section_centroid = sqrt(
      sum(
        (
          c_across(dplyr::all_of(pc_cols)) -
          c_across(dplyr::all_of(paste0(pc_cols, "_centroid")))
        )^2,
        na.rm = TRUE
      )
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    distance_to_section_centroid,
    dplyr::all_of(pc_cols)
  ) %>%
  dplyr::arrange(
    dplyr::desc(distance_to_section_centroid)
  )

isaiah_pca_distances %>%
  dplyr::slice_head(n = 15)
```

## 15. Findings

```{r findings}
cat("\nFeature-space PCA Findings\n")
cat("==========================\n\n")

cat("Top nine components explain: ")
cat(
  round(
    pca_variance$cumulative_pct_variance[9],
    2
  ),
  "% cumulative variance.\n\n"
)

cat("This value may be low because the ensemble feature space is high-dimensional and sparse.\n")
cat("Use 07_Ensemble_MDS.Rmd for the compressed ensemble-similarity view.\n\n")

cat("Most PCA-distinct Isaiah chapters from assigned-section centroid:\n")
print(
  isaiah_pca_distances %>%
    dplyr::slice_head(n = 10) %>%
    dplyr::select(
      chapter,
      doc_id,
      section,
      distance_to_section_centroid
    ),
  width = 120
)
```

## 16. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "featurized_pca")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  pca_variance,
  file.path(out_dir, "feature_pca_variance.csv")
)

readr::write_csv(
  isaiah_pca_scores,
  file.path(out_dir, "isaiah_feature_pca_scores.csv")
)

readr::write_csv(
  section_centroids,
  file.path(out_dir, "isaiah_feature_pca_section_centroids.csv")
)

readr::write_csv(
  isaiah_pca_distances,
  file.path(out_dir, "isaiah_feature_pca_distance_to_centroid.csv")
)
```

## 17. Save plots

```{r save-plots}
plot_dir <- file.path(STYLO_ROOT, "output", "featurized_pca", "plots")

dir.create(
  plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

for(i in seq_along(pca_pair_plots)){

  ggplot2::ggsave(
    filename = file.path(
      plot_dir,
      paste0(
        "feature_pca_isaiah_pc",
        pc_pairs$pc_x[i],
        "_pc",
        pc_pairs$pc_y[i],
        ".png"
      )
    ),
    plot = pca_pair_plots[[i]],
    width = 9,
    height = 6,
    dpi = 300
  )
}
```
