---
title: "06 - PCA Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

options(tibble.width = Inf, tibble.print_max = 150)
```

## 1. Purpose

This notebook performs PCA exploratory analysis on the ensemble feature space.

It analyzes:

```text
PC1 vs PC2
PC2 vs PC3
PC3 vs PC4
...
PC8 vs PC9
```

It shows:

- variance explained by each dimension,
- cumulative variance through PC9,
- Isaiah chapters colored by assigned section,
- chapter labels,
- section centroids,
- PCA outliers,
- relationship to ensemble attribution findings.

PCA is exploratory. The ensemble consensus results remain the attribution model.

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Configure feature set

```{r config}
config_pca <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_pca$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_pca$pos_model <- ud_model
}

config_pca$weights
```

## 4. Build a PCA feature matrix

This builds all feature matrices from `chapter_docs`, scales each block, applies class weights, and combines the blocks.

```{r feature-matrix-functions}
scale_feature_block <- function(mat){

  mat <- as.matrix(mat)
  mat[is.na(mat)] <- 0
  mat[is.nan(mat)] <- 0
  mat[is.infinite(mat)] <- 0

  if(ncol(mat) == 0){
    return(mat)
  }

  scaled <- scale(mat)

  scaled[is.na(scaled)] <- 0
  scaled[is.nan(scaled)] <- 0
  scaled[is.infinite(scaled)] <- 0

  as.matrix(scaled)
}


build_pca_feature_matrix <- function(
    docs,
    config,
    text_col = "text",
    doc_id_col = "doc_id"){

  feature_sets <- build_feature_sets(
    docs,
    config,
    text_col = text_col,
    doc_id_col = doc_id_col
  )

  active_classes <- names(feature_sets)

  weights <- config$weights[active_classes]
  weights <- weights / sum(weights, na.rm = TRUE)

  blocks <- list()

  for(class_name in active_classes){

    class_variants <- feature_sets[[class_name]]

    for(variant_name in names(class_variants)){

      mat <- class_variants[[variant_name]]
      mat <- scale_feature_block(mat)

      mat <- mat * sqrt(weights[[class_name]]) / sqrt(length(class_variants))

      colnames(mat) <- paste(
        class_name,
        variant_name,
        colnames(mat),
        sep = "__"
      )

      blocks[[paste(class_name, variant_name, sep = "__")]] <- mat
    }
  }

  feature_matrix <- do.call(cbind, blocks)

  feature_matrix[is.na(feature_matrix)] <- 0
  feature_matrix[is.nan(feature_matrix)] <- 0
  feature_matrix[is.infinite(feature_matrix)] <- 0

  list(
    feature_matrix = feature_matrix,
    feature_sets = feature_sets,
    weights_used = weights
  )
}
```

## 5. Run PCA

```{r run-pca}
pca_input <- build_pca_feature_matrix(
  chapter_docs,
  config = config_pca
)

X <- pca_input$feature_matrix

dim(X)

pca <- stats::prcomp(
  X,
  center = FALSE,
  scale. = FALSE
)

pca
```

## 6. Variance explained

```{r variance}
pca_var <- pca$sdev^2

pca_variance <- data.frame(
  pc = paste0("PC", seq_along(pca_var)),
  dimension = seq_along(pca_var),
  eigenvalue = pca_var,
  prop_variance = pca_var / sum(pca_var),
  pct_variance = 100 * pca_var / sum(pca_var),
  cumulative_prop_variance = cumsum(pca_var / sum(pca_var)),
  cumulative_pct_variance = 100 * cumsum(pca_var / sum(pca_var))
)

pca_variance %>%
  dplyr::slice_head(n = 15)
```

## 7. PCA scree plot

```{r pca-scree}
pca_variance %>%
  dplyr::slice_head(n = 15) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = dimension, y = pct_variance)
  ) +
  ggplot2::geom_col() +
  ggplot2::geom_line(
    ggplot2::aes(y = cumulative_pct_variance, group = 1)
  ) +
  ggplot2::geom_point(
    ggplot2::aes(y = cumulative_pct_variance)
  ) +
  ggplot2::scale_x_continuous(breaks = 1:15) +
  ggplot2::labs(
    title = "PCA scree plot",
    subtitle = "Bars show individual variance; line shows cumulative variance.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  ggplot2::theme_minimal()
```

## 8. PCA score table

```{r scores}
pca_scores <- as.data.frame(pca$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

isaiah_pca_scores <- pca_scores %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_pca_scores %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    PC1,
    PC2,
    PC3,
    PC4,
    PC5,
    PC6,
    PC7,
    PC8,
    PC9
  )
```

## 9. PCA plotting helper

```{r plot-helper}
plot_pca_pair <- function(
    scores,
    variance_tbl,
    pc_x,
    pc_y,
    color_col = "section",
    label_col = "chapter",
    title = NULL,
    show_labels = TRUE,
    show_centroids = TRUE){

  x_name <- paste0("PC", pc_x)
  y_name <- paste0("PC", pc_y)

  x_var <- variance_tbl$pct_variance[variance_tbl$dimension == pc_x]
  y_var <- variance_tbl$pct_variance[variance_tbl$dimension == pc_y]

  if(is.null(title)){
    title <- paste0(x_name, " vs ", y_name)
  }

  plt <- ggplot2::ggplot(
    scores,
    ggplot2::aes(
      x = .data[[x_name]],
      y = .data[[y_name]],
      color = .data[[color_col]]
    )
  ) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.25, alpha = 0.4) +
    ggplot2::geom_vline(xintercept = 0, linewidth = 0.25, alpha = 0.4) +
    ggplot2::geom_point(size = 2.8, alpha = 0.85) +
    ggplot2::labs(
      title = title,
      subtitle = paste0(
        x_name, " = ", round(x_var, 2), "% variance; ",
        y_name, " = ", round(y_var, 2), "% variance"
      ),
      x = paste0(x_name, " (", round(x_var, 2), "%)"),
      y = paste0(y_name, " (", round(y_var, 2), "%)"),
      color = "Section"
    ) +
    ggplot2::theme_minimal()

  if(show_labels){

    if(requireNamespace("ggrepel", quietly = TRUE)){

      plt <- plt +
        ggrepel::geom_text_repel(
          ggplot2::aes(label = .data[[label_col]]),
          size = 3,
          max.overlaps = 100,
          show.legend = FALSE
        )

    } else {

      plt <- plt +
        ggplot2::geom_text(
          ggplot2::aes(label = .data[[label_col]]),
          size = 3,
          vjust = -0.8,
          show.legend = FALSE
        )
    }
  }

  if(show_centroids){

    centroids <- scores %>%
      dplyr::group_by(.data[[color_col]]) %>%
      dplyr::summarize(
        x = mean(.data[[x_name]], na.rm = TRUE),
        y = mean(.data[[y_name]], na.rm = TRUE),
        .groups = "drop"
      )

    plt <- plt +
      ggplot2::geom_point(
        data = centroids,
        ggplot2::aes(
          x = x,
          y = y,
          color = .data[[color_col]]
        ),
        inherit.aes = FALSE,
        shape = 4,
        size = 5,
        stroke = 1.2,
        show.legend = FALSE
      )
  }

  plt
}
```

## 10. PC1 vs PC2

```{r pc1-pc2}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 1,
  pc_y = 2,
  title = "Isaiah chapters: PC1 vs PC2"
)
```

## 11. PC2 vs PC3

```{r pc2-pc3}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 3,
  pc_y = 2,
  title = "Isaiah chapters: PC2 vs PC3"
)
```

## 12. PC3 vs PC4

```{r pc3-pc4}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 3,
  pc_y = 4,
  title = "Isaiah chapters: PC3 vs PC4"
)
```

## 13. PC4 vs PC5

```{r pc4-pc5}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 4,
  pc_y = 5,
  title = "Isaiah chapters: PC4 vs PC5"
)
```

## 14. PC5 vs PC6

```{r pc5-pc6}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 5,
  pc_y = 6,
  title = "Isaiah chapters: PC5 vs PC6"
)
```

## 15. PC6 vs PC7

```{r pc6-pc7}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 6,
  pc_y = 7,
  title = "Isaiah chapters: PC6 vs PC7"
)
```

## 16. PC7 vs PC8

```{r pc7-pc8}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 7,
  pc_y = 8,
  title = "Isaiah chapters: PC7 vs PC8"
)
```

## 17. PC8 vs PC9

```{r pc8-pc9}
plot_pca_pair(
  isaiah_pca_scores,
  pca_variance,
  pc_x = 8,
  pc_y = 9,
  title = "Isaiah chapters: PC8 vs PC9"
)
```

## 18. Generate all PC pair plots

```{r all-pairs}
pc_pairs <- tibble::tibble(
  pc_x = 1:8,
  pc_y = 2:9
)

pca_pair_plots <- purrr::pmap(
  pc_pairs,
  function(pc_x, pc_y){

    plot_pca_pair(
      isaiah_pca_scores,
      pca_variance,
      pc_x = pc_x,
      pc_y = pc_y,
      title = paste0("Isaiah chapters: PC", pc_x, " vs PC", pc_y)
    )
  }
)

pca_pair_plots[[1]]
```

## 19. Section centroids

```{r section-centroids}
section_centroids <- isaiah_pca_scores %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(
      PC1:PC9,
      ~ mean(.x, na.rm = TRUE)
    ),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

section_centroids
```

## 20. Distance to assigned-section centroid

```{r centroid-distance}
pc_cols <- paste0("PC", 1:9)

centroid_lookup <- section_centroids %>%
  dplyr::select(section, dplyr::all_of(pc_cols))

isaiah_pca_distances <- isaiah_pca_scores %>%
  dplyr::left_join(
    centroid_lookup,
    by = "section",
    suffix = c("", "_centroid")
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    distance_to_section_centroid = sqrt(
      sum(
        (
          c_across(dplyr::all_of(pc_cols)) -
          c_across(dplyr::all_of(paste0(pc_cols, "_centroid")))
        )^2,
        na.rm = TRUE
      )
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    distance_to_section_centroid,
    dplyr::all_of(pc_cols)
  ) %>%
  dplyr::arrange(dplyr::desc(distance_to_section_centroid))

isaiah_pca_distances
```

## 21. Most PCA-distinct Isaiah chapters

```{r distinct-chapters}
isaiah_pca_distances %>%
  dplyr::slice_head(n = 15)
```

## 22. Build or load author-identification findings

This reconstructs the key author-identification table if it is not already available.

```{r author-findings}
if(!exists("isaiah_attribution")){

  result_author_id <- run_stylometry_ensemble(
    source_docs = chapter_docs,
    target_docs = section_docs,
    config = config_pca,
    k_neighbors = 10,
    use_cache = TRUE
  )

  isaiah_doc_ids <- isaiah_pca_scores$doc_id

  isaiah_chapters <- chapter_docs %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::arrange(chapter)

  isaiah_consensus <- purrr::map_dfr(
    isaiah_doc_ids,
    function(id){

      cons <- show_consensus(result_author_id, doc_id = id)

      attr(cons, "summary")
    }
  ) %>%
    dplyr::left_join(
      isaiah_chapters %>%
        dplyr::select(
          doc_id,
          chapter,
          actual_section = section,
          n_verses
        ),
      by = "doc_id"
    ) %>%
    dplyr::relocate(
      doc_id,
      chapter,
      actual_section,
      winner
    ) %>%
    dplyr::arrange(chapter)

  isaiah_attribution <- isaiah_consensus %>%
    dplyr::mutate(
      winner_matches_actual_section = winner == actual_section,
      winner_pct_votes_pct = 100 * winner_pct_votes,
      winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
      vote_margin_pct = 100 * vote_margin,
      weighted_vote_margin_pct = 100 * weighted_vote_margin
    )
}
```

## 23. PCA versus attribution findings

```{r pca-vs-attribution}
pca_vs_attribution <- isaiah_pca_distances %>%
  dplyr::left_join(
    isaiah_attribution %>%
      dplyr::select(
        doc_id,
        winner,
        winner_matches_actual_section,
        winner_pct_votes_pct,
        vote_margin_pct,
        winner_dominance_ratio,
        consensus_strength,
        doc_entropy,
        effective_targets
      ),
    by = "doc_id"
  ) %>%
  dplyr::arrange(dplyr::desc(distance_to_section_centroid))

pca_vs_attribution
```

## 24. False matches in PCA context

```{r false-matches}
pca_vs_attribution %>%
  dplyr::filter(!winner_matches_actual_section)
```

Expected finding from prior author-identification experiments:

```text
Isaiah_12 → Psalms
Isaiah_26 → Psalms
```

If these chapters are also PCA-distinct from `Isaiah_1_35`, that supports the interpretation that the mismatches are genre-driven hymn/song chapters.

## 25. PCA findings summary

```{r findings-summary}
cat("\nPCA Findings Summary\n")
cat("====================\n\n")

cat("Top nine components explain: ")
cat(round(pca_variance$cumulative_pct_variance[9], 2), "% cumulative variance.\n\n")

cat("Most distant Isaiah chapters from assigned-section centroid:\n")
print(
  isaiah_pca_distances %>%
    dplyr::slice_head(n = 10) %>%
    dplyr::select(
      chapter,
      doc_id,
      section,
      distance_to_section_centroid
    ),
  width = 120
)

cat("\nFalse matches from author-identification model:\n")
print(
  pca_vs_attribution %>%
    dplyr::filter(!winner_matches_actual_section) %>%
    dplyr::select(
      chapter,
      doc_id,
      section,
      winner,
      winner_pct_votes_pct,
      vote_margin_pct,
      distance_to_section_centroid
    ),
  width = 120
)
```

## 26. Export PCA outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "pca_analysis")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(
  pca_variance,
  file.path(out_dir, "pca_variance.csv")
)

readr::write_csv(
  isaiah_pca_scores,
  file.path(out_dir, "isaiah_pca_scores.csv")
)

readr::write_csv(
  section_centroids,
  file.path(out_dir, "isaiah_section_centroids_pc1_pc9.csv")
)

readr::write_csv(
  isaiah_pca_distances,
  file.path(out_dir, "isaiah_pca_distance_to_section_centroid.csv")
)

readr::write_csv(
  pca_vs_attribution,
  file.path(out_dir, "isaiah_pca_vs_attribution.csv")
)
```

## 27. Save PCA plots

```{r save-plots}
plot_dir <- file.path(STYLO_ROOT, "output", "pca_analysis", "plots")

dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

for(i in seq_along(pca_pair_plots)){

  ggplot2::ggsave(
    filename = file.path(
      plot_dir,
      paste0("isaiah_pc", pc_pairs$pc_x[i], "_pc", pc_pairs$pc_y[i], ".png")
    ),
    plot = pca_pair_plots[[i]],
    width = 9,
    height = 6,
    dpi = 300
  )
}
```
