---
title: "07 - Ensemble MDS / Similarity Embeddings"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 150
)
```

## 1. Purpose

This notebook analyzes the **compressed ensemble similarity space** using:

```r
result$overall_similarity
result$class_similarity
```

Unlike raw feature-space PCA, this notebook starts from the output of `run_stylometry_ensemble()`.

This answers:

> How do documents arrange themselves after all feature classes have already been compressed into ensemble similarity?

This is closer to:

- multidimensional scaling,
- spectral embedding,
- kernel PCA,
- ensemble-space visualization.

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Configure ensemble

```{r config}
config_mds <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_mds$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_mds$pos_model <- ud_model
}

config_mds$weights
```

## 4. Run chapter-to-section ensemble

This reproduces author-identification mode:

```text
chapter_docs → section_docs
```

This result is rectangular:

```text
chapters × target sections/books
```

We can PCA or MDS the rows of the similarity profile.

```{r run-chapter-section}
result_chapter_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_mds,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_chapter_section)
show_weights(result_chapter_section)
```

## 5. PCA on chapter-to-section ensemble similarity profiles

Here each chapter is represented by its similarity to each section/book target.

```{r pca-on-rectangular-similarity}
S_rect <- as.matrix(result_chapter_section$overall_similarity)

S_rect[is.na(S_rect)] <- 0
S_rect[is.nan(S_rect)] <- 0
S_rect[is.infinite(S_rect)] <- 0

pca_similarity_profiles <- stats::prcomp(
  S_rect,
  center = TRUE,
  scale. = TRUE
)

sim_pca_var <- pca_similarity_profiles$sdev^2

sim_pca_variance <- data.frame(
  pc = paste0("PC", seq_along(sim_pca_var)),
  dimension = seq_along(sim_pca_var),
  eigenvalue = sim_pca_var,
  prop_variance = sim_pca_var / sum(sim_pca_var),
  pct_variance = 100 * sim_pca_var / sum(sim_pca_var),
  cumulative_prop_variance = cumsum(sim_pca_var / sum(sim_pca_var)),
  cumulative_pct_variance = 100 * cumsum(sim_pca_var / sum(sim_pca_var))
)

sim_pca_variance %>%
  dplyr::slice_head(n = 15)
```

## 6. Similarity-profile PCA scree plot

```{r profile-pca-scree}
sim_pca_variance %>%
  dplyr::slice_head(n = 15) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = dimension,
      y = pct_variance
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::geom_line(
    ggplot2::aes(
      y = cumulative_pct_variance,
      group = 1
    )
  ) +
  ggplot2::geom_point(
    ggplot2::aes(
      y = cumulative_pct_variance
    )
  ) +
  ggplot2::scale_x_continuous(
    breaks = 1:15
  ) +
  ggplot2::labs(
    title = "PCA on ensemble similarity profiles",
    subtitle = "Each chapter is represented by similarity to target sections/books.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  ggplot2::theme_minimal()
```

## 7. Similarity-profile PCA scores

```{r profile-pca-scores}
sim_pca_scores <- as.data.frame(pca_similarity_profiles$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

isaiah_sim_pca_scores <- sim_pca_scores %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_sim_pca_scores %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    PC1,
    PC2,
    PC3,
    PC4,
    PC5,
    PC6,
    PC7,
    PC8,
    PC9
  )
```

## 8. Similarity-profile PCA plots

```{r profile-pca-pc1-pc2}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Ensemble similarity-profile PCA: Isaiah PC1 vs PC2"
)
```

```{r profile-pca-pc2-pc3}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 3,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Ensemble similarity-profile PCA: Isaiah PC2 vs PC3"
)
```

```{r profile-pca-pc3-pc4}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 3,
  y = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Ensemble similarity-profile PCA: Isaiah PC3 vs PC4"
)
```

## 9. Classical MDS on chapter-to-section profiles

Because the similarity matrix is rectangular, we compute distances between chapters based on their target-similarity profiles.

```{r mds-profile}
D_profile <- stats::dist(
  S_rect,
  method = "euclidean"
)

mds_profile <- stats::cmdscale(
  D_profile,
  k = 9,
  eig = TRUE
)

mds_profile_scores <- as.data.frame(mds_profile$points)

colnames(mds_profile_scores) <- paste0("Dim", seq_len(ncol(mds_profile_scores)))

mds_profile_scores <- mds_profile_scores %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

mds_eig <- mds_profile$eig[mds_profile$eig > 0]

mds_variance <- data.frame(
  pc = paste0("PC", seq_along(mds_eig)),
  dimension = seq_along(mds_eig),
  eigenvalue = mds_eig,
  prop_variance = mds_eig / sum(mds_eig),
  pct_variance = 100 * mds_eig / sum(mds_eig),
  cumulative_prop_variance = cumsum(mds_eig / sum(mds_eig)),
  cumulative_pct_variance = 100 * cumsum(mds_eig / sum(mds_eig))
)

# Rename Dim columns to PC columns so plotPCA can be reused.
mds_plot_scores <- mds_profile_scores %>%
  dplyr::rename_with(
    ~ stringr::str_replace(.x, "^Dim", "PC"),
    dplyr::starts_with("Dim")
  )

mds_variance %>%
  dplyr::slice_head(n = 9)
```

## 10. MDS plots from similarity profiles

```{r mds-pc1-pc2}
plotPCA(
  mds_plot_scores,
  mds_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Ensemble MDS from similarity profiles: Dim1 vs Dim2"
)
```

```{r mds-pc2-pc3}
plotPCA(
  mds_plot_scores,
  mds_variance,
  x = 2,
  y = 3,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Ensemble MDS from similarity profiles: Dim2 vs Dim3"
)
```

## 11. Run chapter-to-chapter ensemble

For a true square similarity matrix, run:

```text
chapter_docs → chapter_docs
```

This makes `result$overall_similarity` square.

```{r run-chapter-chapter}
result_chapter_chapter <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = chapter_docs,
  config = config_mds,
  k_neighbors = 10,
  use_cache = TRUE
)

dim(result_chapter_chapter$overall_similarity)
```

## 12. Classical MDS on square ensemble similarity

```{r mds-square}
S_square <- as.matrix(result_chapter_chapter$overall_similarity)

S_square[is.na(S_square)] <- 0
S_square[is.nan(S_square)] <- 0
S_square[is.infinite(S_square)] <- 0

# Convert similarity to distance.
D_square <- 1 - S_square

D_square[D_square < 0] <- 0

# Symmetrize in case of numerical asymmetry.
D_square <- (D_square + t(D_square)) / 2

diag(D_square) <- 0

mds_square <- stats::cmdscale(
  as.dist(D_square),
  k = 9,
  eig = TRUE
)

mds_square_scores <- as.data.frame(mds_square$points)
colnames(mds_square_scores) <- paste0("PC", seq_len(ncol(mds_square_scores)))

mds_square_scores <- mds_square_scores %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

mds_square_eig <- mds_square$eig[mds_square$eig > 0]

mds_square_variance <- data.frame(
  pc = paste0("PC", seq_along(mds_square_eig)),
  dimension = seq_along(mds_square_eig),
  eigenvalue = mds_square_eig,
  prop_variance = mds_square_eig / sum(mds_square_eig),
  pct_variance = 100 * mds_square_eig / sum(mds_square_eig),
  cumulative_prop_variance = cumsum(mds_square_eig / sum(mds_square_eig)),
  cumulative_pct_variance = 100 * cumsum(mds_square_eig / sum(mds_square_eig))
)

mds_square_variance %>%
  dplyr::slice_head(n = 9)
```

## 13. Square ensemble MDS plots

```{r square-mds-pc1-pc2}
plotPCA(
  mds_square_scores,
  mds_square_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Square ensemble MDS: Isaiah Dim1 vs Dim2"
)


plotPCA(
  mds_square_scores,
  mds_square_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  include_books = c("Psalms", "Hosea", "Job", "Proverbs"),
  show_other_books = TRUE,
  label = "doc_id",
  title = "Isaiah with poetic comparison books: Dim2 vs Dim3"
)

plotPCA(
  mds_square_scores,
  mds_square_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  include_books = c("Psalms", "Job", "Proverbs", "Hosea"),
  aggregate_other_books = TRUE,
  show_other_books = TRUE,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Isaiah with poetic comparison book centroids: Dim2 vs Dim3"
)

```

```{r square-mds-pc2-pc3}
plotPCA(
  mds_square_scores,
  mds_square_variance,
  x = 3,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Square ensemble MDS: Isaiah Dim2 vs Dim3"
)
```

```{r square-mds-pc3-pc4}
plotPCA(
  mds_square_scores,
  mds_square_variance,
  x = 3,
  y = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Square ensemble MDS: Isaiah Dim3 vs Dim4"
)
```

## 14. Interactive 3D square ensemble MDS

Disabled by default for knitting.

```{r square-mds-3d, eval=FALSE}
plotPCA3d(
  mds_square_scores,
  mds_square_variance,
  x = 1,
  y = 2,
  z = 3,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  title = "Square ensemble MDS: Isaiah Dim1 vs Dim2 vs Dim3"
)
```

## 15. Class-specific MDS

This uses `result$class_similarity`.

Each class similarity matrix is square because this section uses `result_chapter_chapter`.

```{r class-specific-mds-functions}
mds_from_similarity <- function(sim, metadata, k = 9){

  S <- as.matrix(sim)

  S[is.na(S)] <- 0
  S[is.nan(S)] <- 0
  S[is.infinite(S)] <- 0

  D <- 1 - S
  D[D < 0] <- 0
  D <- (D + t(D)) / 2
  diag(D) <- 0

  fit <- stats::cmdscale(
    as.dist(D),
    k = k,
    eig = TRUE
  )

  scores <- as.data.frame(fit$points)
  colnames(scores) <- paste0("PC", seq_len(ncol(scores)))

  scores <- scores %>%
    tibble::rownames_to_column("doc_id") %>%
    dplyr::left_join(
      metadata,
      by = "doc_id"
    )

  eig <- fit$eig[fit$eig > 0]

  variance <- data.frame(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  list(
    fit = fit,
    scores = scores,
    variance = variance
  )
}
```

```{r class-specific-mds}
metadata <- chapter_docs %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

class_mds <- lapply(
  result_chapter_chapter$class_similarity,
  mds_from_similarity,
  metadata = metadata,
  k = 9
)

names(class_mds)
```

## 16. Plot class-specific MDS examples

```{r class-char-mds}
if("char" %in% names(class_mds)){

  plotPCA(
    class_mds$char$scores,
    class_mds$char$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific MDS: Character features"
  )
}
```

```{r class-pos-mds}
if("pos" %in% names(class_mds)){

  plotPCA(
    class_mds$pos$scores,
    class_mds$pos$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific MDS: POS features"
  )
}
```

```{r class-function-mds}
if("function_words" %in% names(class_mds)){

  plotPCA(
    class_mds$function_words$scores,
    class_mds$function_words$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific MDS: Function-word features"
  )
}
```

## 17. Compare MDS outliers with attribution findings

```{r attribution-summary}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_chapter_section,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_attribution <- isaiah_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  )

show_false_matches(isaiah_attribution)
```

## 18. MDS section centroids and distances

```{r mds-centroids}
pc_cols <- paste0("PC", 1:9)

isaiah_mds_square_scores <- mds_square_scores %>%
  dplyr::filter(book == "Isaiah")

mds_section_centroids <- isaiah_mds_square_scores %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(
      dplyr::all_of(pc_cols),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

mds_centroid_lookup <- mds_section_centroids %>%
  dplyr::select(
    section,
    dplyr::all_of(pc_cols)
  )

isaiah_mds_distances <- isaiah_mds_square_scores %>%
  dplyr::left_join(
    mds_centroid_lookup,
    by = "section",
    suffix = c("", "_centroid")
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    distance_to_section_centroid = sqrt(
      sum(
        (
          c_across(dplyr::all_of(pc_cols)) -
          c_across(dplyr::all_of(paste0(pc_cols, "_centroid")))
        )^2,
        na.rm = TRUE
      )
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    distance_to_section_centroid,
    dplyr::all_of(pc_cols)
  ) %>%
  dplyr::arrange(
    dplyr::desc(distance_to_section_centroid)
  )

isaiah_mds_distances %>%
  dplyr::slice_head(n = 15)
```

## 19. MDS findings summary

```{r findings}
cat("\nEnsemble MDS Findings\n")
cat("=====================\n\n")

cat("Chapter-to-section similarity-profile PCA: PC1-PC3 cumulative variance: ")
cat(round(sim_pca_variance$cumulative_pct_variance[3], 2), "%\n")

cat("Square chapter-to-chapter ensemble MDS: Dim1-Dim3 cumulative positive-eigen variance: ")
cat(round(mds_square_variance$cumulative_pct_variance[3], 2), "%\n\n")

cat("False matches from author-identification model:\n")
print(
  show_false_matches(isaiah_attribution),
  width = 120
)

cat("\nMost MDS-distinct Isaiah chapters from assigned-section centroid:\n")
print(
  isaiah_mds_distances %>%
    dplyr::slice_head(n = 10) %>%
    dplyr::select(
      chapter,
      doc_id,
      section,
      distance_to_section_centroid
    ),
  width = 120
)
```

## 20. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "ensemble_mds")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  sim_pca_variance,
  file.path(out_dir, "similarity_profile_pca_variance.csv")
)

readr::write_csv(
  isaiah_sim_pca_scores,
  file.path(out_dir, "isaiah_similarity_profile_pca_scores.csv")
)

readr::write_csv(
  mds_square_variance,
  file.path(out_dir, "square_ensemble_mds_variance.csv")
)

readr::write_csv(
  mds_square_scores,
  file.path(out_dir, "square_ensemble_mds_scores.csv")
)

readr::write_csv(
  isaiah_mds_distances,
  file.path(out_dir, "isaiah_square_ensemble_mds_distance_to_centroid.csv")
)

readr::write_csv(
  isaiah_attribution,
  file.path(out_dir, "isaiah_attribution_from_chapter_section_result.csv")
)
```

## 21. Save selected plots

```{r save-plots}
plot_dir <- file.path(STYLO_ROOT, "output", "ensemble_mds", "plots")

dir.create(
  plot_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

ggplot2::ggsave(
  filename = file.path(plot_dir, "similarity_profile_pca_pc1_pc2.png"),
  plot = plotPCA(
    sim_pca_scores,
    sim_pca_variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Ensemble similarity-profile PCA: Isaiah PC1 vs PC2"
  ),
  width = 9,
  height = 6,
  dpi = 300
)

ggplot2::ggsave(
  filename = file.path(plot_dir, "square_ensemble_mds_dim1_dim2.png"),
  plot = plotPCA(
    mds_square_scores,
    mds_square_variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Square ensemble MDS: Isaiah Dim1 vs Dim2"
  ),
  width = 9,
  height = 6,
  dpi = 300
)
```
