---
title: "08 - Latent Dimensions"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

options(tibble.width = Inf, tibble.print_max = 200)
```

## 1. Purpose

This notebook interprets the **latent dimensions** discovered in the ensemble similarity space.

It analyzes:

- `result$overall_similarity`
- `result$class_similarity`
- similarity-profile PCA loadings
- section centroid distances
- class-specific separation
- heatmaps and clustering
- author-identification false matches

The central question is whether the latent structure separates Isaiah by author-like section breaks, genre effects, or both.

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Configure ensemble

```{r config}
config_latent <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_latent$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_latent$pos_model <- ud_model
}

config_latent$weights
```

## 4. Run chapter-to-section ensemble

```{r run-chapter-section}
result_chapter_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_latent,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_chapter_section)
show_weights(result_chapter_section)
```

## 5. PCA on ensemble similarity profiles

Each chapter is represented by its similarity to each target profile.

```{r similarity-profile-pca}
S_rect <- as.matrix(result_chapter_section$overall_similarity)

S_rect[is.na(S_rect)] <- 0
S_rect[is.nan(S_rect)] <- 0
S_rect[is.infinite(S_rect)] <- 0

pca_similarity_profiles <- stats::prcomp(
  S_rect,
  center = TRUE,
  scale. = TRUE
)

sim_pca_var <- pca_similarity_profiles$sdev^2

sim_pca_variance <- data.frame(
  pc = paste0("PC", seq_along(sim_pca_var)),
  dimension = seq_along(sim_pca_var),
  eigenvalue = sim_pca_var,
  prop_variance = sim_pca_var / sum(sim_pca_var),
  pct_variance = 100 * sim_pca_var / sum(sim_pca_var),
  cumulative_prop_variance = cumsum(sim_pca_var / sum(sim_pca_var)),
  cumulative_pct_variance = 100 * cumsum(sim_pca_var / sum(sim_pca_var))
)

sim_pca_variance %>%
  dplyr::slice_head(n = 12)
```

## 6. Similarity-profile PCA loadings

```{r pca-loadings}
profile_loadings <- as.data.frame(pca_similarity_profiles$rotation) %>%
  tibble::rownames_to_column("target_profile")

profile_loadings %>%
  dplyr::select(target_profile, PC1, PC2, PC3, PC4, PC5) %>%
  dplyr::arrange(dplyr::desc(abs(PC1)))
```

## 7. Top positive and negative loadings by dimension

```{r loading-helper}
show_pc_loadings <- function(loadings_tbl, pc = 1, n = 8){

  pc_col <- paste0("PC", pc)

  positive <- loadings_tbl %>%
    dplyr::select(target_profile, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(dplyr::desc(loading)) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "positive")

  negative <- loadings_tbl %>%
    dplyr::select(target_profile, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(loading) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "negative")

  dplyr::bind_rows(positive, negative) %>%
    dplyr::arrange(direction, dplyr::desc(abs(loading)))
}

show_pc_loadings(profile_loadings, pc = 1, n = 8)
show_pc_loadings(profile_loadings, pc = 2, n = 8)
show_pc_loadings(profile_loadings, pc = 3, n = 8)
show_pc_loadings(profile_loadings, pc = 4, n = 8)
```

## 8. Interpret PC1-PC4 from loadings

```{r interpret-loadings}
for(pc in 1:4){

  cat("\n============================================================\n")
  cat("PC", pc, " loadings\n", sep = "")
  cat("Variance explained: ", round(sim_pca_variance$pct_variance[pc], 2), "%\n", sep = "")
  cat("============================================================\n\n")

  print(show_pc_loadings(profile_loadings, pc = pc, n = 6), width = 120)
}
```

## 9. Similarity-profile PCA scores

```{r pca-scores}
sim_pca_scores <- as.data.frame(pca_similarity_profiles$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, book, chapter, section, macro_genre, n_verses)

isaiah_sim_pca_scores <- sim_pca_scores %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_sim_pca_scores %>%
  dplyr::select(doc_id, chapter, section, PC1, PC2, PC3, PC4)
```

## 10. Plot interpreted dimensions

```{r pc1-pc2}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Latent dimensions: PC1 vs PC2"
)
```

```{r pc2-pc3}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 3,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Latent dimensions: PC2 vs PC3"
)
```

```{r pc3-pc4}
plotPCA(
  sim_pca_scores,
  sim_pca_variance,
  x = 3,
  y = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Latent dimensions: PC3 vs PC4"
)
```

## 11. Isaiah section centroids in latent space

```{r section-centroids}
pc_cols <- paste0("PC", 1:9)

section_centroids <- isaiah_sim_pca_scores %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(dplyr::all_of(pc_cols), ~ mean(.x, na.rm = TRUE)),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

section_centroids
```

## 12. Pairwise centroid distances

```{r centroid-distance-function}
centroid_distance_table <- function(centroids, dims = paste0("PC", 1:4)){

  mat <- centroids %>%
    dplyr::select(dplyr::all_of(dims)) %>%
    as.matrix()

  rownames(mat) <- centroids$section

  d <- as.matrix(stats::dist(mat))

  expand.grid(
    section_a = rownames(d),
    section_b = colnames(d),
    stringsAsFactors = FALSE
  ) %>%
    dplyr::filter(section_a < section_b) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(distance = d[section_a, section_b]) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(distance)
}

centroid_distances_pc1_pc4 <- centroid_distance_table(
  section_centroids,
  dims = paste0("PC", 1:4)
)

centroid_distances_pc1_pc4
```

## 13. Centroid distances by number of dimensions

```{r centroid-distances-by-dim}
centroid_distances_by_dim <- purrr::map_dfr(
  1:9,
  function(k){

    centroid_distance_table(section_centroids, dims = paste0("PC", 1:k)) %>%
      dplyr::mutate(n_dimensions = k)
  }
) %>%
  dplyr::relocate(n_dimensions)

centroid_distances_by_dim %>%
  dplyr::filter(
    section_a %in% c("Isaiah_40_55", "Isaiah_56_66") |
      section_b %in% c("Isaiah_40_55", "Isaiah_56_66")
  )
```

## 14. Plot centroid distance stability

```{r centroid-distance-plot}
centroid_distances_by_dim %>%
  dplyr::mutate(pair = paste(section_a, section_b, sep = " ↔ ")) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = n_dimensions,
      y = distance,
      color = pair
    )
  ) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::scale_x_continuous(breaks = 1:9) +
  ggplot2::labs(
    title = "Isaiah section centroid distances by latent dimensionality",
    subtitle = "Smaller distance indicates stronger latent-space proximity.",
    x = "Number of PCs used",
    y = "Centroid distance",
    color = "Section pair"
  ) +
  ggplot2::theme_minimal()
```

## 15. Closest section pairs by dimensionality

```{r closest-section-summary}
centroid_distances_by_dim %>%
  dplyr::group_by(n_dimensions) %>%
  dplyr::slice_min(order_by = distance, n = 1, with_ties = TRUE) %>%
  dplyr::ungroup()
```

## 16. Run chapter-to-chapter ensemble

```{r run-chapter-chapter}
result_chapter_chapter <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = chapter_docs,
  config = config_latent,
  k_neighbors = 10,
  use_cache = TRUE
)

dim(result_chapter_chapter$overall_similarity)
names(result_chapter_chapter$class_similarity)
```

## 17. MDS helper from similarity matrix

```{r mds-helper}
mds_from_similarity <- function(sim, metadata, k = 9){

  S <- as.matrix(sim)

  S[is.na(S)] <- 0
  S[is.nan(S)] <- 0
  S[is.infinite(S)] <- 0

  D <- 1 - S
  D[D < 0] <- 0
  D <- (D + t(D)) / 2
  diag(D) <- 0

  fit <- stats::cmdscale(as.dist(D), k = k, eig = TRUE)

  scores <- as.data.frame(fit$points)
  colnames(scores) <- paste0("PC", seq_len(ncol(scores)))

  scores <- scores %>%
    tibble::rownames_to_column("doc_id") %>%
    dplyr::left_join(metadata, by = "doc_id")

  eig <- fit$eig[fit$eig > 0]

  variance <- data.frame(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  list(fit = fit, scores = scores, variance = variance)
}
```

## 18. Overall square ensemble MDS

```{r overall-square-mds}
metadata <- chapter_docs %>%
  dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses)

overall_mds <- mds_from_similarity(
  result_chapter_chapter$overall_similarity,
  metadata = metadata,
  k = 9
)

overall_mds$variance %>%
  dplyr::slice_head(n = 9)
```

```{r overall-mds-plot}
plotPCA(
  overall_mds$scores,
  overall_mds$variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Overall ensemble MDS: Isaiah Dim1 vs Dim2"
)
```

## 19. Feature-class-specific MDS objects

```{r class-mds}
class_mds <- lapply(
  result_chapter_chapter$class_similarity,
  mds_from_similarity,
  metadata = metadata,
  k = 9
)

names(class_mds)
```

## 20. Class-specific centroid distances

```{r class-centroid-distance}
section_centroid_distances_for_scores <- function(scores, dims = paste0("PC", 1:4)){

  isaiah_scores <- scores %>%
    dplyr::filter(book == "Isaiah")

  centroids <- isaiah_scores %>%
    dplyr::group_by(section) %>%
    dplyr::summarize(
      dplyr::across(dplyr::all_of(dims), ~ mean(.x, na.rm = TRUE)),
      n_chapters = dplyr::n(),
      .groups = "drop"
    )

  centroid_distance_table(centroids, dims = dims)
}

class_centroid_distances <- purrr::imap_dfr(
  class_mds,
  function(obj, class_name){

    section_centroid_distances_for_scores(obj$scores, dims = paste0("PC", 1:4)) %>%
      dplyr::mutate(feature_class = class_name)
  }
) %>%
  dplyr::relocate(feature_class)

class_centroid_distances
```

## 21. Isaiah_40_55 vs other sections by feature class

```{r class-focus-4055}
class_centroid_distances %>%
  dplyr::filter(
    section_a == "Isaiah_40_55" |
      section_b == "Isaiah_40_55"
  ) %>%
  dplyr::arrange(feature_class, distance)
```

## 22. Isaiah_40_55 vs Isaiah_56_66 by feature class

```{r class-4055-5666}
dist_4055_5666_by_class <- class_centroid_distances %>%
  dplyr::filter(
    (section_a == "Isaiah_40_55" & section_b == "Isaiah_56_66") |
      (section_a == "Isaiah_56_66" & section_b == "Isaiah_40_55")
  ) %>%
  dplyr::arrange(distance)

dist_4055_5666_by_class
```

## 23. Plot 40-55 vs 56-66 distance by feature class

```{r class-distance-plot}
dist_4055_5666_by_class %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(feature_class, distance),
      y = distance
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "Latent-space distance: Isaiah_40_55 vs Isaiah_56_66",
    subtitle = "Computed separately for each feature class using class-specific MDS.",
    x = "Feature class",
    y = "Centroid distance"
  ) +
  ggplot2::theme_minimal()
```

## 24. Class-specific plots

```{r class-pos-plot}
if("pos" %in% names(class_mds)){

  plotPCA(
    class_mds$pos$scores,
    class_mds$pos$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific latent space: POS"
  )
  
  
  
  section_overlap(
    class_mds$pos$scores
)
  
}
```

```{r class-function-plot}
if("function_words" %in% names(class_mds)){

  plotPCA(
    class_mds$function_words$scores,
    class_mds$function_words$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific latent space: Function words"
  )
  
  section_overlap(
    class_mds$function_words$scores
)
  
}
```

```{r class-lexical-plot}
if("lexical" %in% names(class_mds)){

  plotPCA(
    class_mds$lexical$scores,
    class_mds$lexical$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific latent space: Lexical richness"
  )
}
```

```{r class-word-length-plot}
if("word_length" %in% names(class_mds)){

  plotPCA(
    class_mds$word_length$scores,
    class_mds$word_length$variance,
    x = 1,
    y = 2,
    book = "Isaiah",
    color = "section",
    label = "chapter",
    show_labels = TRUE,
    show_centroids = TRUE,
    show_hulls = TRUE,
    title = "Class-specific latent space: Word length"
  )
}
```

## 25. Overall similarity heatmap

```{r heatmap-data}
overall_sim <- as.matrix(result_chapter_chapter$overall_similarity)

isaiah_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::pull(doc_id)

isaiah_sim <- overall_sim[isaiah_ids, isaiah_ids, drop = FALSE]

hc <- stats::hclust(
  stats::as.dist(1 - isaiah_sim),
  method = "average"
)

ordered_ids <- hc$labels[hc$order]

isaiah_sim_long <- as.data.frame(isaiah_sim[ordered_ids, ordered_ids]) %>%
  tibble::rownames_to_column("doc_a") %>%
  tidyr::pivot_longer(
    cols = -doc_a,
    names_to = "doc_b",
    values_to = "similarity"
  ) %>%
  dplyr::mutate(
    doc_a = factor(doc_a, levels = ordered_ids),
    doc_b = factor(doc_b, levels = ordered_ids)
  )

isaiah_sim_long %>%
  dplyr::slice_head(n = 10)
```

```{r heatmap}
isaiah_sim_long %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = doc_b,
      y = doc_a,
      fill = similarity
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    title = "Overall ensemble similarity heatmap",
    subtitle = "Isaiah chapters reordered by hierarchical clustering.",
    x = "Chapter",
    y = "Chapter",
    fill = "Similarity"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5, size = 6),
    axis.text.y = ggplot2::element_text(size = 6)
  )
```

## 26. Dendrogram order

```{r dendrogram-order}
dendrogram_order <- data.frame(
  order = seq_along(ordered_ids),
  doc_id = ordered_ids
) %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, chapter, section),
    by = "doc_id"
  )

dendrogram_order
```

## 27. Author-identification false matches

```{r false-matches}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(result_chapter_section, doc_id = id)

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(doc_id, chapter, actual_section = section, n_verses),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, chapter, actual_section, winner) %>%
  dplyr::arrange(chapter)

isaiah_attribution <- isaiah_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  )

show_false_matches(isaiah_attribution)
```

## 28. Summarize latent findings

```{r final-summary}
cat("\nLatent Dimension Findings\n")
cat("=========================\n\n")

cat("Similarity-profile PCA cumulative variance PC1-PC3: ")
cat(round(sim_pca_variance$cumulative_pct_variance[3], 2), "%\n\n")

cat("Closest Isaiah section pair using PC1-PC4 centroids:\n")
print(
  centroid_distances_pc1_pc4 %>%
    dplyr::slice_head(n = 1),
  width = 120
)

cat("\nIsaiah_40_55 vs Isaiah_56_66 distance by feature class:\n")
print(dist_4055_5666_by_class, width = 120)

cat("\nAuthor-identification false matches:\n")
print(show_false_matches(isaiah_attribution), width = 120)
```

```{r sentence info}
#
#colnames(sentence_features)


#plot_sentence_features_by_section()

#show_sentence_summary()

#plot_sentence_distributions()

```



## 29. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "latent_dimensions")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(sim_pca_variance, file.path(out_dir, "similarity_profile_pca_variance.csv"))
readr::write_csv(profile_loadings, file.path(out_dir, "similarity_profile_pca_loadings.csv"))
readr::write_csv(isaiah_sim_pca_scores, file.path(out_dir, "isaiah_similarity_profile_pca_scores.csv"))
readr::write_csv(section_centroids, file.path(out_dir, "isaiah_section_centroids_similarity_pca.csv"))
readr::write_csv(centroid_distances_by_dim, file.path(out_dir, "isaiah_centroid_distances_by_dimension.csv"))
readr::write_csv(class_centroid_distances, file.path(out_dir, "class_specific_centroid_distances.csv"))
readr::write_csv(dist_4055_5666_by_class, file.path(out_dir, "isaiah_4055_vs_5666_by_feature_class.csv"))
readr::write_csv(dendrogram_order, file.path(out_dir, "isaiah_similarity_dendrogram_order.csv"))
readr::write_csv(isaiah_attribution, file.path(out_dir, "isaiah_attribution_summary.csv"))
```
