---
title: "09 - Sentence Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 200
)
```

## 1. Purpose

Sentence features emerged as the largest feature-class separation between:

```text
Isaiah_40_55
Isaiah_56_66
```

This notebook analyzes sentence features **through the ensemble framework**.

It does **not** call `build_sentence_features()` directly.

Primary objects:

```r
result_sentence_section$class_similarity$sentence
result_sentence_chapter$class_similarity$sentence
result_sentence_section$overall_similarity
result_sentence_chapter$overall_similarity
```

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. Configure sentence-only ensemble

```{r config-sentence-only}
config_sentence <- default_config(
  include = c("sentence")
)

config_sentence$weights <- c(
  sentence = 1
)

config_sentence$weights
```

## 4. Run sentence-only chapter-to-section ensemble

```{r run-sentence-section}
result_sentence_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_sentence,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_sentence_section)
show_weights(result_sentence_section)
```

## 5. Run sentence-only chapter-to-chapter ensemble

```{r run-sentence-chapter}
result_sentence_chapter <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = chapter_docs,
  config = config_sentence,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_sentence_chapter)
show_weights(result_sentence_chapter)
```

## 6. Inspect sentence similarity matrices

```{r inspect-sentence-matrices}
names(result_sentence_section$class_similarity)

dim(result_sentence_section$overall_similarity)
dim(result_sentence_chapter$overall_similarity)

sentence_section_similarity <- result_sentence_section$class_similarity$sentence
sentence_chapter_similarity <- result_sentence_chapter$class_similarity$sentence

dim(sentence_section_similarity)
dim(sentence_chapter_similarity)
```

## 7. Sentence-only attribution for Isaiah chapters

```{r sentence-attribution}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_sentence_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_sentence_section,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_sentence_attribution <- isaiah_sentence_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  )

isaiah_sentence_attribution
```

## 8. Sentence-only false matches

```{r sentence-false-matches}
show_false_matches(
  isaiah_sentence_attribution
)
```

## 9. Sentence-only winner counts

```{r sentence-winner-counts}
isaiah_sentence_attribution %>%
  dplyr::count(
    actual_section,
    winner,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(actual_section) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    actual_section,
    dplyr::desc(n_chapters)
  )
```

## 10. Detailed sentence-only findings

```{r sentence-detailed-findings-40}
summarize_findings(
  result_sentence_section,
  doc_id = "Isaiah_40",
  k = 10,
  plot = TRUE
)
```

```{r sentence-detailed-findings-56}
summarize_findings(
  result_sentence_section,
  doc_id = "Isaiah_56",
  k = 10,
  plot = TRUE
)
```

## 11. Helper: MDS from similarity matrix

```{r mds-helper}
mds_from_similarity <- function(sim, metadata, k = 9){

  S <- as.matrix(sim)

  S[is.na(S)] <- 0
  S[is.nan(S)] <- 0
  S[is.infinite(S)] <- 0

  D <- 1 - S
  D[D < 0] <- 0

  D <- (D + t(D)) / 2
  diag(D) <- 0

  fit <- stats::cmdscale(
    as.dist(D),
    k = k,
    eig = TRUE
  )

  scores <- as.data.frame(fit$points)
  colnames(scores) <- paste0("PC", seq_len(ncol(scores)))

  scores <- scores %>%
    tibble::rownames_to_column("doc_id") %>%
    dplyr::left_join(metadata, by = "doc_id")

  eig <- fit$eig[fit$eig > 0]

  variance <- data.frame(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  list(
    fit = fit,
    scores = scores,
    variance = variance
  )
}
```

## 12. Sentence-only MDS

```{r sentence-mds}
metadata <- chapter_docs %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

sentence_mds <- mds_from_similarity(
  result_sentence_chapter$overall_similarity,
  metadata = metadata,
  k = 9
)

sentence_mds$variance %>%
  dplyr::slice_head(n = 9)
```

## 13. Sentence-only MDS plots

```{r sentence-mds-12}
plotPCA(
  sentence_mds$scores,
  sentence_mds$variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Sentence-only MDS: Isaiah Dim1 vs Dim2"
)
```

```{r sentence-mds-23}
plotPCA(
  sentence_mds$scores,
  sentence_mds$variance,
  x = 3,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Sentence-only MDS: Isaiah Dim2 vs Dim3"
)
```

```{r sentence-mds-34}
plotPCA(
  sentence_mds$scores,
  sentence_mds$variance,
  x = 3,
  y = 4,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Sentence-only MDS: Isaiah Dim3 vs Dim4"
)
```

## 14. Sentence-only section overlap

```{r sentence-overlap}
section_overlap(
  sentence_mds$scores
)
```

## 15. Sentence-only centroid distances

```{r sentence-centroid-distances}
pc_cols <- paste0("PC", 1:9)

sentence_isaiah_scores <- sentence_mds$scores %>%
  dplyr::filter(book == "Isaiah")

sentence_section_centroids <- sentence_isaiah_scores %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(
      dplyr::all_of(pc_cols),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

centroid_distance_table <- function(centroids, dims = paste0("PC", 1:4)){

  mat <- centroids %>%
    dplyr::select(dplyr::all_of(dims)) %>%
    as.matrix()

  rownames(mat) <- centroids$section

  d <- as.matrix(stats::dist(mat))

  expand.grid(
    section_a = rownames(d),
    section_b = colnames(d),
    stringsAsFactors = FALSE
  ) %>%
    dplyr::filter(section_a < section_b) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(distance = d[section_a, section_b]) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(distance)
}

sentence_centroid_distances_pc1_pc4 <- centroid_distance_table(
  sentence_section_centroids,
  dims = paste0("PC", 1:4)
)

sentence_centroid_distances_pc1_pc4
```

## 16. Sentence centroid distance by dimensions

```{r sentence-centroid-distances-by-dim}
sentence_centroid_distances_by_dim <- purrr::map_dfr(
  1:9,
  function(k){

    centroid_distance_table(
      sentence_section_centroids,
      dims = paste0("PC", 1:k)
    ) %>%
      dplyr::mutate(
        n_dimensions = k
      )
  }
) %>%
  dplyr::relocate(n_dimensions)

sentence_centroid_distances_by_dim
```

```{r sentence-centroid-distance-plot}
sentence_centroid_distances_by_dim %>%
  dplyr::mutate(
    pair = paste(section_a, section_b, sep = " ↔ ")
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = n_dimensions,
      y = distance,
      color = pair
    )
  ) +
  ggplot2::geom_line() +
  ggplot2::geom_point() +
  ggplot2::scale_x_continuous(breaks = 1:9) +
  ggplot2::labs(
    title = "Sentence-only section centroid distances",
    subtitle = "Distances computed in sentence-only MDS space.",
    x = "Number of MDS dimensions used",
    y = "Centroid distance",
    color = "Section pair"
  ) +
  ggplot2::theme_minimal()
```

## 17. Closest section pair under sentence-only features

```{r sentence-closest-pairs}
sentence_centroid_distances_by_dim %>%
  dplyr::group_by(n_dimensions) %>%
  dplyr::slice_min(
    order_by = distance,
    n = 1,
    with_ties = TRUE
  ) %>%
  dplyr::ungroup()
```

## 18. Sentence-only similarity heatmap

```{r sentence-heatmap-data}
sentence_sim <- as.matrix(result_sentence_chapter$overall_similarity)

isaiah_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::pull(doc_id)

sentence_isaiah_sim <- sentence_sim[isaiah_ids, isaiah_ids, drop = FALSE]

sentence_heatmap_chapter_order <- as.data.frame(sentence_isaiah_sim[isaiah_ids, isaiah_ids]) %>%
  tibble::rownames_to_column("doc_a") %>%
  tidyr::pivot_longer(
    cols = -doc_a,
    names_to = "doc_b",
    values_to = "similarity"
  ) %>%
  dplyr::mutate(
    doc_a = factor(doc_a, levels = rev(isaiah_ids)),
    doc_b = factor(doc_b, levels = isaiah_ids)
  )

hc_sentence <- stats::hclust(
  stats::as.dist(1 - sentence_isaiah_sim),
  method = "average"
)

ordered_ids <- hc_sentence$labels[hc_sentence$order]

sentence_heatmap_cluster_order <- as.data.frame(sentence_isaiah_sim[ordered_ids, ordered_ids]) %>%
  tibble::rownames_to_column("doc_a") %>%
  tidyr::pivot_longer(
    cols = -doc_a,
    names_to = "doc_b",
    values_to = "similarity"
  ) %>%
  dplyr::mutate(
    doc_a = factor(doc_a, levels = rev(ordered_ids)),
    doc_b = factor(doc_b, levels = ordered_ids)
  )
```

```{r sentence-heatmap-chapter-order}
sentence_heatmap_chapter_order %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = doc_b,
      y = doc_a,
      fill = similarity
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    title = "Sentence-only similarity heatmap",
    subtitle = "Isaiah chapters in canonical chapter order.",
    x = "Chapter",
    y = "Chapter",
    fill = "Similarity"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5, size = 6),
    axis.text.y = ggplot2::element_text(size = 6)
  )
```

```{r sentence-heatmap-cluster-order}
sentence_heatmap_cluster_order %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = doc_b,
      y = doc_a,
      fill = similarity
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    title = "Sentence-only similarity heatmap",
    subtitle = "Isaiah chapters reordered by hierarchical clustering.",
    x = "Chapter",
    y = "Chapter",
    fill = "Similarity"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5, size = 6),
    axis.text.y = ggplot2::element_text(size = 6)
  )
```

## 19. Sentence-only nearest targets for late Isaiah

```{r late-isaiah-nearest}
result_sentence_section$nearest_targets %>%
  dplyr::filter(
    doc_id %in% paste0("Isaiah_", 40:66)
  ) %>%
  dplyr::arrange(
    doc_id,
    rank
  )
```

## 20. Compare sentence-only with full ensemble

```{r full-ensemble}
config_full <- default_config(
  include = c(
    "char",
    "word",
    "stem",
    "function_words",
    "pos",
    "punctuation",
    "sentence",
    "word_length",
    "lexical",
    "readability"
  )
)

config_full$pos_n <- c(3, 4, 5)

if(exists("ud_model")){
  config_full$pos_model <- ud_model
}

result_full_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_full,
  k_neighbors = 10,
  use_cache = TRUE
)
```

```{r full-vs-sentence-attribution}
isaiah_full_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_full_section,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

comparison_sentence_vs_full <- isaiah_sentence_attribution %>%
  dplyr::select(
    doc_id,
    chapter,
    actual_section,
    sentence_winner = winner,
    sentence_winner_pct = winner_pct_votes_pct,
    sentence_margin = vote_margin_pct
  ) %>%
  dplyr::left_join(
    isaiah_full_consensus %>%
      dplyr::mutate(
        full_winner_pct = 100 * winner_pct_votes,
        full_margin = 100 * vote_margin
      ) %>%
      dplyr::select(
        doc_id,
        full_winner = winner,
        full_winner_pct,
        full_margin
      ),
    by = "doc_id"
  ) %>%
  dplyr::mutate(
    same_winner = sentence_winner == full_winner
  ) %>%
  dplyr::arrange(chapter)

comparison_sentence_vs_full
```

## 21. Chapters where sentence-only disagrees with full ensemble

```{r sentence-full-disagreements}
comparison_sentence_vs_full %>%
  dplyr::filter(!same_winner)
```

## 22. Final sentence-analysis summary

```{r final-summary}
cat("\nSentence Analysis Findings\n")
cat("==========================\n\n")

cat("Sentence-only attribution false matches:\n")
print(show_false_matches(isaiah_sentence_attribution), width = 120)

cat("\nSentence-only section overlap:\n")
print(section_overlap(sentence_mds$scores), width = 120)

cat("\nSentence-only closest section pair by dimensionality:\n")
print(
  sentence_centroid_distances_by_dim %>%
    dplyr::group_by(n_dimensions) %>%
    dplyr::slice_min(order_by = distance, n = 1, with_ties = TRUE) %>%
    dplyr::ungroup(),
  width = 120
)

cat("\nSentence-only vs full-ensemble disagreements:\n")
print(
  comparison_sentence_vs_full %>%
    dplyr::filter(!same_winner),
  width = 120
)
```

## 23. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "sentence_analysis")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(
  isaiah_sentence_attribution,
  file.path(out_dir, "sentence_only_attribution.csv")
)

readr::write_csv(
  sentence_mds$variance,
  file.path(out_dir, "sentence_only_mds_variance.csv")
)

readr::write_csv(
  sentence_mds$scores,
  file.path(out_dir, "sentence_only_mds_scores.csv")
)

readr::write_csv(
  sentence_centroid_distances_pc1_pc4,
  file.path(out_dir, "sentence_centroid_distances_pc1_pc4.csv")
)

readr::write_csv(
  sentence_centroid_distances_by_dim,
  file.path(out_dir, "sentence_centroid_distances_by_dim.csv")
)

readr::write_csv(
  comparison_sentence_vs_full,
  file.path(out_dir, "sentence_vs_full_ensemble_comparison.csv")
)
```
