---
title: "09 - Sentence Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 200
)
```

## 1. Purpose

Sentence features emerged as the largest feature-class separation between:

```text
Isaiah_40_55
Isaiah_56_66
```

This notebook does two things:

1. It analyzes sentence features **through the ensemble framework**.
2. It builds **raw sentence-length profiles** by Isaiah section.

The raw profile analysis answers:

> What is the shape of each section's sentence-length distribution?

This includes:

- sentence-bin proportions,
- section profile tables,
- radar/star charts,
- chapter-level sentence-bin profiles,
- 40–55 vs 56–66 profile differences.

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)

  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

# Part A: Raw sentence-bin profiles

## 3. Sentence profile helpers

These functions are local to this notebook. They do not depend on `build_sentence_features()`.

```{r sentence-profile-helpers}
split_sentences_local <- function(text){

  text <- as.character(text)

  text <- stringr::str_replace_all(
    text,
    "\\s+",
    " "
  )

  text <- stringr::str_trim(text)

  if(is.na(text) || text == ""){
    return(character(0))
  }

  # KJV punctuation is fairly regular. Split on sentence-ending punctuation.
  sents <- unlist(
    stringr::str_split(
      text,
      "(?<=[.!?])\\s+"
    )
  )

  sents <- stringr::str_trim(sents)

  sents[nchar(sents) > 0]
}


count_sentence_words_local <- function(sentence){

  words <- stringr::str_extract_all(
    sentence,
    "\\b[A-Za-z']+\\b"
  )[[1]]

  length(words)
}


sentence_lengths_for_text <- function(text){

  sents <- split_sentences_local(text)

  if(length(sents) == 0){
    return(numeric(0))
  }

  vapply(
    sents,
    count_sentence_words_local,
    numeric(1)
  )
}


sentence_bin_label <- function(len){

  dplyr::case_when(
    len < 10 ~ "lt_10",
    len >= 10 & len < 20 ~ "10_19",
    len >= 20 & len < 30 ~ "20_29",
    len >= 30 & len < 40 ~ "30_39",
    len >= 40 & len < 50 ~ "40_49",
    len >= 50 ~ "50_plus",
    TRUE ~ NA_character_
  )
}


build_sentence_profile_table <- function(docs){

  rows <- lapply(
    seq_len(nrow(docs)),
    function(i){

      lens <- sentence_lengths_for_text(docs$text[i])

      if(length(lens) == 0){
        return(
          data.frame(
            doc_id = docs$doc_id[i],
            sentence_id = integer(0),
            sentence_length = numeric(0),
            sentence_bin = character(0)
          )
        )
      }

      data.frame(
        doc_id = docs$doc_id[i],
        sentence_id = seq_along(lens),
        sentence_length = lens,
        sentence_bin = sentence_bin_label(lens)
      )
    }
  )

  dplyr::bind_rows(rows) %>%
    dplyr::left_join(
      docs %>%
        dplyr::select(
          doc_id,
          book,
          chapter,
          section,
          macro_genre,
          n_verses
        ),
      by = "doc_id"
    )
}


sentence_bins <- c(
  "lt_10",
  "10_19",
  "20_29",
  "30_39",
  "40_49",
  "50_plus"
)
```

## 4. Build raw sentence profile table

```{r sentence-profile-table}
sentence_profile <- build_sentence_profile_table(
  chapter_docs
)

isaiah_sentence_profile <- sentence_profile %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::mutate(
    sentence_bin = factor(
      sentence_bin,
      levels = sentence_bins
    )
  )

isaiah_sentence_profile %>%
  dplyr::slice_head(n = 20)
```

## 5. Chapter-level sentence profile

```{r chapter-level-profile}
chapter_sentence_profile <- isaiah_sentence_profile %>%
  dplyr::count(
    doc_id,
    chapter,
    section,
    sentence_bin,
    name = "n_sentences"
  ) %>%
  tidyr::complete(
    doc_id,
    chapter,
    section,
    sentence_bin = factor(sentence_bins, levels = sentence_bins),
    fill = list(n_sentences = 0)
  ) %>%
  dplyr::group_by(
    doc_id,
    chapter,
    section
  ) %>%
  dplyr::mutate(
    total_sentences = sum(n_sentences),
    pct_sentences = dplyr::if_else(
      total_sentences > 0,
      n_sentences / total_sentences,
      0
    )
  ) %>%
  dplyr::ungroup()

chapter_sentence_profile
```

## 6. Section-level sentence profile

```{r section-level-profile}
section_sentence_profile <- isaiah_sentence_profile %>%
  dplyr::count(
    section,
    sentence_bin,
    name = "n_sentences"
  ) %>%
  tidyr::complete(
    section,
    sentence_bin = factor(sentence_bins, levels = sentence_bins),
    fill = list(n_sentences = 0)
  ) %>%
  dplyr::group_by(section) %>%
  dplyr::mutate(
    total_sentences = sum(n_sentences),
    pct_sentences = dplyr::if_else(
      total_sentences > 0,
      n_sentences / total_sentences,
      0
    )
  ) %>%
  dplyr::ungroup()

section_sentence_profile
```

## 7. Sentence-bin profile by section

```{r section-profile-bar}
section_sentence_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = sentence_bin,
      y = pct_sentences,
      fill = section
    )
  ) +
  ggplot2::geom_col(
    position = "dodge"
  ) +
  ggplot2::scale_y_continuous(
    labels = scales::percent_format()
  ) +
  ggplot2::labs(
    title = "Sentence-length bin profile by Isaiah section",
    subtitle = "Proportion of sentences falling into each word-count bin.",
    x = "Sentence length bin",
    y = "Percent of sentences",
    fill = "Section"
  ) +
  ggplot2::theme_minimal()
```

## 8. Sentence profile star chart / radar chart

This is a radar-style plot where each spoke is a sentence-length bin.

```{r radar-helper}
plot_sentence_profile_radar <- function(
    section_profile,
    sections = unique(section_profile$section),
    value_col = "pct_sentences",
    title = "Sentence-length profile radar chart"){

  bins <- sentence_bins
  n_bins <- length(bins)

  angle_tbl <- tibble::tibble(
    sentence_bin = factor(bins, levels = bins),
    angle = seq(0, 2 * pi, length.out = n_bins + 1)[1:n_bins],
    x_axis = sin(angle),
    y_axis = cos(angle)
  )

  dat <- section_profile %>%
    dplyr::filter(section %in% sections) %>%
    dplyr::mutate(
      sentence_bin = factor(sentence_bin, levels = bins)
    ) %>%
    dplyr::left_join(angle_tbl, by = "sentence_bin") %>%
    dplyr::mutate(
      r = .data[[value_col]],
      x = r * x_axis,
      y = r * y_axis
    ) %>%
    dplyr::arrange(section, sentence_bin)

  dat_closed <- dat %>%
    dplyr::group_by(section) %>%
    dplyr::arrange(sentence_bin) %>%
    dplyr::group_modify(~ dplyr::bind_rows(.x, .x[1, ])) %>%
    dplyr::ungroup()

  max_r <- max(dat[[value_col]], na.rm = TRUE)

  rings <- tibble::tibble(
    r = pretty(c(0, max_r), n = 4)
  ) %>%
    dplyr::filter(r > 0)

  ring_lines <- purrr::map_dfr(
    rings$r,
    function(rr){
      tibble::tibble(
        r = rr,
        angle = seq(0, 2 * pi, length.out = 200),
        x = rr * sin(angle),
        y = rr * cos(angle)
      )
    }
  )

  spokes <- angle_tbl %>%
    dplyr::mutate(
      x0 = 0,
      y0 = 0,
      x1 = max_r * 1.08 * x_axis,
      y1 = max_r * 1.08 * y_axis
    )

  labels <- angle_tbl %>%
    dplyr::mutate(
      x = max_r * 1.18 * x_axis,
      y = max_r * 1.18 * y_axis,
      label = as.character(sentence_bin)
    )

  ggplot2::ggplot() +
    ggplot2::geom_path(
      data = ring_lines,
      ggplot2::aes(x = x, y = y, group = r),
      color = "grey85"
    ) +
    ggplot2::geom_segment(
      data = spokes,
      ggplot2::aes(x = x0, y = y0, xend = x1, yend = y1),
      color = "grey80"
    ) +
    ggplot2::geom_polygon(
      data = dat_closed,
      ggplot2::aes(x = x, y = y, group = section, color = section, fill = section),
      alpha = 0.15,
      linewidth = 1
    ) +
    ggplot2::geom_point(
      data = dat,
      ggplot2::aes(x = x, y = y, color = section),
      size = 2.5
    ) +
    ggplot2::geom_text(
      data = labels,
      ggplot2::aes(x = x, y = y, label = label),
      size = 3
    ) +
    ggplot2::coord_equal() +
    ggplot2::labs(
      title = title,
      x = NULL,
      y = NULL,
      color = "Section",
      fill = "Section"
    ) +
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "right",
      plot.title = ggplot2::element_text(hjust = 0.5)
    )
}
```

```{r radar-all-sections}
plot_sentence_profile_radar(
  section_sentence_profile,
  title = "Isaiah section sentence-length profiles"
)
```

## 9. Focused star chart: Isaiah 40–55 vs 56–66

```{r radar-late-isaiah}
plot_sentence_profile_radar(
  section_sentence_profile,
  sections = c("Isaiah_40_55", "Isaiah_56_66"),
  title = "Sentence-length profile: Isaiah_40_55 vs Isaiah_56_66"
)
```

## 10. Section profile table, wide format

```{r section-profile-wide}
section_sentence_profile_wide <- section_sentence_profile %>%
  dplyr::select(
    section,
    sentence_bin,
    pct_sentences
  ) %>%
  tidyr::pivot_wider(
    names_from = sentence_bin,
    values_from = pct_sentences
  )

section_sentence_profile_wide
```

## 11. Difference profile: Isaiah 56–66 minus Isaiah 40–55

```{r difference-profile}
late_sentence_difference <- section_sentence_profile %>%
  dplyr::filter(
    section %in% c("Isaiah_40_55", "Isaiah_56_66")
  ) %>%
  dplyr::select(
    section,
    sentence_bin,
    pct_sentences
  ) %>%
  tidyr::pivot_wider(
    names_from = section,
    values_from = pct_sentences
  ) %>%
  dplyr::mutate(
    difference_56_66_minus_40_55 =
      Isaiah_56_66 - Isaiah_40_55,
    abs_difference =
      abs(difference_56_66_minus_40_55)
  ) %>%
  dplyr::arrange(
    dplyr::desc(abs_difference)
  )

late_sentence_difference
```

```{r difference-profile-plot}
late_sentence_difference %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = sentence_bin,
      y = difference_56_66_minus_40_55
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(
    labels = scales::percent_format()
  ) +
  ggplot2::labs(
    title = "Sentence profile difference",
    subtitle = "Positive values mean Isaiah_56_66 has a larger share than Isaiah_40_55.",
    x = "Sentence length bin",
    y = "Isaiah_56_66 minus Isaiah_40_55"
  ) +
  ggplot2::theme_minimal()
```

## 12. Chapter-by-chapter sentence-bin heatmap

```{r chapter-bin-heatmap}
chapter_sentence_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = sentence_bin,
      y = factor(chapter),
      fill = pct_sentences
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c(
    labels = scales::percent_format()
  ) +
  ggplot2::facet_grid(
    section ~ .,
    scales = "free_y",
    space = "free_y"
  ) +
  ggplot2::labs(
    title = "Chapter-level sentence-bin profiles",
    subtitle = "Each row is an Isaiah chapter; cells show sentence-bin proportions.",
    x = "Sentence length bin",
    y = "Chapter",
    fill = "Share"
  ) +
  ggplot2::theme_minimal()
```

## 13. Mean and dispersion of sentence lengths by section

```{r sentence-length-summary}
sentence_length_summary_by_section <- isaiah_sentence_profile %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    n_sentences = dplyr::n(),
    mean_sentence_length = mean(sentence_length, na.rm = TRUE),
    median_sentence_length = stats::median(sentence_length, na.rm = TRUE),
    sd_sentence_length = stats::sd(sentence_length, na.rm = TRUE),
    pct_lt_10 = mean(sentence_length < 10, na.rm = TRUE),
    pct_10_19 = mean(sentence_length >= 10 & sentence_length < 20, na.rm = TRUE),
    pct_20_29 = mean(sentence_length >= 20 & sentence_length < 30, na.rm = TRUE),
    pct_30_39 = mean(sentence_length >= 30 & sentence_length < 40, na.rm = TRUE),
    pct_40_49 = mean(sentence_length >= 40 & sentence_length < 50, na.rm = TRUE),
    pct_50_plus = mean(sentence_length >= 50, na.rm = TRUE),
    .groups = "drop"
  )

sentence_length_summary_by_section
```

```{r sentence-length-boxplot}
isaiah_sentence_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = sentence_length,
      fill = section
    )
  ) +
  ggplot2::geom_boxplot(
    alpha = 0.5,
    outlier.alpha = 0.5
  ) +
  ggplot2::coord_cartesian(
    ylim = c(
      0,
      stats::quantile(
        isaiah_sentence_profile$sentence_length,
        0.95,
        na.rm = TRUE
      )
    )
  ) +
  ggplot2::labs(
    title = "Sentence-length distribution by Isaiah section",
    subtitle = "Y-axis clipped at the 95th percentile for readability.",
    x = "Section",
    y = "Sentence length in words",
    fill = "Section"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 45,
      hjust = 1
    )
  )
```

# Part B: Ensemble-based sentence analysis

## 14. Configure sentence-only ensemble

```{r config-sentence-only}
config_sentence <- default_config(
  include = c("sentence")
)

config_sentence$weights <- c(
  sentence = 1
)

config_sentence$weights
```

## 15. Run sentence-only chapter-to-section ensemble

```{r run-sentence-section}
result_sentence_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_sentence,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_sentence_section)
show_weights(result_sentence_section)
```

## 16. Run sentence-only chapter-to-chapter ensemble

```{r run-sentence-chapter}
result_sentence_chapter <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = chapter_docs,
  config = config_sentence,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_sentence_chapter)
show_weights(result_sentence_chapter)
```

## 17. Sentence-only attribution for Isaiah chapters

```{r sentence-attribution}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_sentence_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_sentence_section,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_sentence_attribution <- isaiah_sentence_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  )

isaiah_sentence_attribution
```

## 18. Sentence-only false matches

```{r sentence-false-matches}
show_false_matches(
  isaiah_sentence_attribution
)
```

## 19. Helper: MDS from similarity matrix

```{r mds-helper}
mds_from_similarity <- function(sim, metadata, k = 9){

  S <- as.matrix(sim)

  S[is.na(S)] <- 0
  S[is.nan(S)] <- 0
  S[is.infinite(S)] <- 0

  D <- 1 - S
  D[D < 0] <- 0

  D <- (D + t(D)) / 2
  diag(D) <- 0

  fit <- stats::cmdscale(
    as.dist(D),
    k = k,
    eig = TRUE
  )

  scores <- as.data.frame(fit$points)
  colnames(scores) <- paste0("PC", seq_len(ncol(scores)))

  scores <- scores %>%
    tibble::rownames_to_column("doc_id") %>%
    dplyr::left_join(metadata, by = "doc_id")

  eig <- fit$eig[fit$eig > 0]

  variance <- data.frame(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  list(
    fit = fit,
    scores = scores,
    variance = variance
  )
}
```

## 20. Sentence-only MDS

```{r sentence-mds}
metadata <- chapter_docs %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

sentence_mds <- mds_from_similarity(
  result_sentence_chapter$overall_similarity,
  metadata = metadata,
  k = 9
)

sentence_mds$variance %>%
  dplyr::slice_head(n = 9)
```

## 21. Sentence-only MDS plots

```{r sentence-mds-12}
plotPCA(
  sentence_mds$scores,
  sentence_mds$variance,
  x = 1,
  y = 2,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Sentence-only MDS: Isaiah Dim1 vs Dim2"
)
```

```{r sentence-mds-23}
plotPCA(
  sentence_mds$scores,
  sentence_mds$variance,
  x = 2,
  y = 3,
  book = "Isaiah",
  color = "section",
  label = "chapter",
  show_labels = TRUE,
  show_centroids = TRUE,
  show_hulls = TRUE,
  title = "Sentence-only MDS: Isaiah Dim2 vs Dim3"
)
```

## 22. Sentence-only section overlap

```{r sentence-overlap}
section_overlap(
  sentence_mds$scores
)
```

## 23. Final sentence-analysis summary

```{r final-summary}
cat("\nSentence Analysis Findings\n")
cat("==========================\n\n")

cat("Raw sentence profile: Isaiah_56_66 minus Isaiah_40_55:\n")
print(late_sentence_difference, width = 120)

cat("\nSentence length summary by section:\n")
print(sentence_length_summary_by_section, width = 120)

cat("\nSentence-only attribution false matches:\n")
print(show_false_matches(isaiah_sentence_attribution), width = 120)

cat("\nSentence-only section overlap:\n")
print(section_overlap(sentence_mds$scores), width = 120)
```

## 24. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "sentence_analysis")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(
  isaiah_sentence_profile,
  file.path(out_dir, "isaiah_sentence_profile_raw.csv")
)

readr::write_csv(
  chapter_sentence_profile,
  file.path(out_dir, "chapter_sentence_bin_profile.csv")
)

readr::write_csv(
  section_sentence_profile,
  file.path(out_dir, "section_sentence_bin_profile.csv")
)

readr::write_csv(
  section_sentence_profile_wide,
  file.path(out_dir, "section_sentence_bin_profile_wide.csv")
)

readr::write_csv(
  late_sentence_difference,
  file.path(out_dir, "sentence_profile_5666_minus_4055.csv")
)

readr::write_csv(
  sentence_length_summary_by_section,
  file.path(out_dir, "sentence_length_summary_by_section.csv")
)

readr::write_csv(
  isaiah_sentence_attribution,
  file.path(out_dir, "sentence_only_attribution.csv")
)

readr::write_csv(
  sentence_mds$variance,
  file.path(out_dir, "sentence_only_mds_variance.csv")
)

readr::write_csv(
  sentence_mds$scores,
  file.path(out_dir, "sentence_only_mds_scores.csv")
)
```
