---
title: "10 - POS Radar Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 250
)




## User parameters

top_n <- 12

# alternative:
# top_n <- 12
# top_n <- 30
# top_n <- 20

```

## 1. Purpose

This notebook analyzes **POS-5 profiles** across Isaiah sections.

Goals:

1. Build POS-5 feature profiles.
2. Select the top 20 POS-5 bins.
3. Compare section-level POS-5 profiles using:
   - radar/star charts,
   - heatmaps,
   - difference charts.
4. Map selected POS-5 patterns back to original word phrases and count examples.

This bridges:

```text
abstract POS stylometry
        ↓
actual repeated phrase structures
```

## 2. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 3. POS configuration

This notebook focuses on POS-5 only.

```{r config-pos5}
config_pos5 <- default_config(
  include = c("pos")
)

config_pos5$pos_n <- c(5)

if(exists("ud_model")){
  config_pos5$pos_model <- ud_model
}

config_pos5$weights <- c(pos = 1)

config_pos5
```

## 4. Build POS-5 features

The code below uses `build_feature_sets()` so it follows the same pipeline as the ensemble.

```{r build-pos5}
pos5_feature_sets <- build_feature_sets(
  chapter_docs,
  config = config_pos5,
  text_col = "text",
  doc_id_col = "doc_id"
)

names(pos5_feature_sets)
names(pos5_feature_sets$pos)

pos5_mat <- pos5_feature_sets$pos[[1]]

dim(pos5_mat)

pos5_df <- as.data.frame(pos5_mat) %>%
  tibble::rownames_to_column("doc_id")

pos5_df %>%
  dplyr::slice_head(n = 5)
```

## 5. Join metadata

```{r join-metadata}
pos5_with_meta <- pos5_df %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(
        doc_id,
        book,
        chapter,
        section,
        macro_genre,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  )

isaiah_pos5 <- pos5_with_meta %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

pos5_cols <- setdiff(
  colnames(isaiah_pos5),
  c("doc_id", "book", "chapter", "section", "macro_genre", "n_verses")
)

length(pos5_cols)
```

## 6. Section-level POS-5 profiles

```{r section-pos5-profile}
section_pos5_profile <- isaiah_pos5 %>%
  dplyr::group_by(section) %>%
  dplyr::summarize(
    dplyr::across(
      dplyr::all_of(pos5_cols),
      ~ mean(.x, na.rm = TRUE)
    ),
    n_chapters = dplyr::n(),
    .groups = "drop"
  )

section_pos5_profile
```

## 7. Select top 20 POS-5 patterns

We rank POS-5 patterns by section-level variance and also compute the direct 40–55 vs 56–66 difference.

```{r top-pos5}
## 7. Select top-N POS-5 patterns

overall_freq <- isaiah_pos5 %>%

  summarise(

    across(

      all_of(pos5_cols),

      mean,

      na.rm = TRUE

    )

  ) %>%

  tidyr::pivot_longer(

    everything(),

    names_to = "pos5",

    values_to = "overall_freq"

  ) %>%

  arrange(

    desc(overall_freq)

  )


top_pos5 <- overall_freq %>%

  slice_head(

    n = top_n

  ) %>%

  pull(pos5)


overall_freq %>%

  slice_head(

    n = top_n

  )
```

## 8. Top-20 POS-5 section profile table

```{r top-profile-table}
top_pos5_profile <- pos5_section_long %>%
  dplyr::filter(pos5 %in% top_pos5) %>%
  dplyr::mutate(
    pos5 = factor(pos5, levels = top_pos5)
  ) %>%
  dplyr::arrange(pos5, section)

top_pos5_profile
```

## 9. Top-20 POS-5 heatmap

```{r top-heatmap}
top_pos5_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = pos5,
      fill = mean_freq
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    title = "Top 20 POS-5 profiles by Isaiah section",
    subtitle = "POS-5 patterns ranked by variance across Isaiah sections.",
    x = "Section",
    y = "POS-5 pattern",
    fill = "Mean frequency"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
    axis.text.y = ggplot2::element_text(size = 7)
  )
```

## 10. Top-20 POS-5 grouped bar chart

```{r top-bar}
top_pos5_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = pos5,
      y = mean_freq,
      fill = section
    )
  ) +
  ggplot2::geom_col(
    position = "dodge"
  ) +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "Top 20 POS-5 profiles by section",
    x = "POS-5 pattern",
    y = "Mean frequency",
    fill = "Section"
  ) +
  ggplot2::theme_minimal()
```

## 11. POS-5 radar/star helper

This creates a true spider/star chart using Cartesian coordinates.

```{r radar-helper}

isaiah_section_colors <- c(

  Isaiah_1_35  = "#F8766D",

  Isaiah_36_39 = "#7CAE00",

  Isaiah_40_55 = "#00BFC4",

  Isaiah_56_66 = "#C77CFF"

)


plot_pos5_radar <- function(
    profile_long,
    pos5_levels,
    sections = unique(profile_long$section),
    value_col = "mean_freq",
    colors = NULL,
    title = "POS-5 radar chart"){

  n_bins <- length(pos5_levels)

  angle_tbl <- tibble::tibble(
    pos5 = factor(pos5_levels, levels = pos5_levels),
    angle = seq(0, 2 * pi, length.out = n_bins + 1)[1:n_bins],
    x_axis = sin(angle),
    y_axis = cos(angle)
  )

  dat <- profile_long %>%
    dplyr::filter(section %in% sections) %>%
    dplyr::mutate(
      pos5 = factor(pos5, levels = pos5_levels)
    ) %>%
    dplyr::left_join(
      angle_tbl,
      by = "pos5"
    ) %>%
    dplyr::mutate(
      r = .data[[value_col]],
      x = r * x_axis,
      y = r * y_axis
    ) %>%
    dplyr::arrange(section, pos5)

  dat_closed <- dat %>%
    dplyr::group_by(section) %>%
    dplyr::arrange(pos5) %>%
    dplyr::group_modify(
      ~ dplyr::bind_rows(.x, .x[1, ])
    ) %>%
    dplyr::ungroup()

  max_r <- max(dat[[value_col]], na.rm = TRUE)

  rings <- tibble::tibble(
    r = pretty(c(0, max_r), n = 4)
  ) %>%
    dplyr::filter(r > 0)

  ring_lines <- purrr::map_dfr(
    rings$r,
    function(rr){
      tibble::tibble(
        r = rr,
        angle = seq(0, 2 * pi, length.out = 250),
        x = rr * sin(angle),
        y = rr * cos(angle)
      )
    }
  )

  spokes <- angle_tbl %>%
    dplyr::mutate(
      x0 = 0,
      y0 = 0,
      x1 = max_r * 1.10 * x_axis,
      y1 = max_r * 1.10 * y_axis
    )

  labels <- angle_tbl %>%
    dplyr::mutate(
      x = max_r * 1.25 * x_axis,
      y = max_r * 1.25 * y_axis,
      label = as.character(pos5)
    )

  ggplot2::ggplot() +
    ggplot2::geom_path(
      data = ring_lines,
      ggplot2::aes(x = x, y = y, group = r),
      color = "grey85"
    ) +
    ggplot2::geom_segment(
      data = spokes,
      ggplot2::aes(x = x0, y = y0, xend = x1, yend = y1),
      color = "grey80"
    ) +
    ggplot2::geom_polygon(
      data = dat_closed,
      ggplot2::aes(
        x = x,
        y = y,
        group = section,
        color = section,
        fill = section
      ),
      alpha = 0.12,
      linewidth = 0.8
    ) +
    ggplot2::geom_point(
      data = dat,
      ggplot2::aes(
        x = x,
        y = y,
        color = section
      ),
      size = 2
    ) +
    ggplot2::geom_text(
      data = labels,
      ggplot2::aes(
        x = x,
        y = y,
        label = label
      ),
      size = 2.2
    ) +
    ggplot2::coord_equal() +
    ggplot2::labs(
      title = title,
      x = NULL,
      y = NULL,
      color = "Section",
      fill = "Section"
    ) +
    
    scale_color_manual(

    values = colors

) +

scale_fill_manual(

    values = colors

) +
    
    ggplot2::theme_void() +
    ggplot2::theme(
      legend.position = "right",
      plot.title = ggplot2::element_text(hjust = 0.5)
    )
}
```

## 12. Radar chart: all Isaiah sections

```{r radar-all}
plot_pos5_radar(
  top_pos5_profile,
  pos5_levels = top_pos5,
  colors = isaiah_section_colors,
  title = "Top N=12 POS-5 profiles by Isaiah section"
)
```

## 13. Radar chart: Isaiah 40–55 vs 56–66

```{r radar-late}
plot_pos5_radar(
  top_pos5_profile,
  pos5_levels = top_pos5,
  sections = c("Isaiah_40_55", "Isaiah_56_66"),
  colors = isaiah_section_colors,
  title = "Top N=12 POS-5 profile: Isaiah_40_55 vs Isaiah_56_66"
)
```

## 14. Top POS-5 differences: Isaiah 56–66 minus 40–55

```{r late-diff-table}
top_pos5_late_diff <- pos5_late_diff %>%
  dplyr::arrange(
    dplyr::desc(abs_late_diff)
  ) %>%
  dplyr::slice_head(n = 30)

top_pos5_late_diff
```

```{r late-diff-plot}
top_pos5_late_diff %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(pos5, diff_56_66_minus_40_55),
      y = diff_56_66_minus_40_55
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "POS-5 difference: Isaiah_56_66 minus Isaiah_40_55",
    subtitle = "Positive values mean the pattern is more frequent in Isaiah_56_66.",
    x = "POS-5 pattern",
    y = "Mean frequency difference"
  ) +
  ggplot2::theme_minimal()
```

## 15. POS-5 phrase recovery

The next functions map POS-5 patterns back to original word phrases.

This depends on the POS tagger used in your `features_pos.R`. The helper below uses `udpipe` if `ud_model` exists.

```{r phrase-recovery-helpers}
tag_text_for_pos_examples <- function(text, doc_id = "doc"){

  if(!exists("ud_model")){
    stop("ud_model not found. Load the udpipe model in 00_source_project.R or setup.")
  }

  ann <- udpipe::udpipe_annotate(
    ud_model,
    x = text,
    doc_id = doc_id
  )

  as.data.frame(ann)
}


extract_pos_ngrams_with_phrases <- function(
    docs,
    n = 5,
    text_col = "text",
    doc_id_col = "doc_id"){

  rows <- list()
  k <- 1

  for(i in seq_len(nrow(docs))){

    doc_id <- docs[[doc_id_col]][i]
    text <- docs[[text_col]][i]

    ann <- tag_text_for_pos_examples(
      text,
      doc_id = doc_id
    )

    if(!all(c("sentence_id", "token", "upos") %in% names(ann))){
      stop("POS annotation must include sentence_id, token, and upos columns.")
    }

    ann <- ann %>%
      dplyr::filter(
        !is.na(token),
        !is.na(upos)
      ) %>%
      dplyr::arrange(
        sentence_id,
        token_id
      )

    sentence_ids <- unique(ann$sentence_id)

    for(sid in sentence_ids){

      sent <- ann %>%
        dplyr::filter(sentence_id == sid)

      if(nrow(sent) < n){
        next
      }

      for(j in seq_len(nrow(sent) - n + 1)){

        window <- sent[j:(j+n-1), ]

        pos_ngram <- paste(window$upos, collapse = "_")
        phrase <- paste(window$token, collapse = " ")

        rows[[k]] <- data.frame(
          doc_id = doc_id,
          sentence_id = sid,
          start_token = window$token_id[1],
          pos_ngram = pos_ngram,
          phrase = phrase,
          stringsAsFactors = FALSE
        )

        k <- k + 1
      }
    }
  }

  dplyr::bind_rows(rows) %>%
    dplyr::left_join(
      docs %>%
        dplyr::select(
          doc_id,
          book,
          chapter,
          section,
          macro_genre
        ),
      by = "doc_id"
    )
}


show_pos5_examples <- function(
    pos_phrase_table,
    pos5_pattern,
    sections = NULL,
    top_n = 20){

  dat <- pos_phrase_table %>%
    dplyr::filter(
      pos_ngram == pos5_pattern
    )

  if(!is.null(sections)){
    dat <- dat %>%
      dplyr::filter(section %in% sections)
  }

  dat %>%
    dplyr::count(
      section,
      phrase,
      name = "frequency"
    ) %>%
    dplyr::arrange(
      section,
      dplyr::desc(frequency),
      phrase
    ) %>%
    dplyr::group_by(section) %>%
    dplyr::slice_head(n = top_n) %>%
    dplyr::ungroup()
}


show_top_pos5_examples <- function(
    pos_phrase_table,
    pos5_patterns,
    sections = c("Isaiah_40_55", "Isaiah_56_66"),
    top_n = 10){

  purrr::map_dfr(
    pos5_patterns,
    function(pattern){

      show_pos5_examples(
        pos_phrase_table,
        pos5_pattern = pattern,
        sections = sections,
        top_n = top_n
      ) %>%
        dplyr::mutate(
          pos5_pattern = pattern
        ) %>%
        dplyr::relocate(pos5_pattern)
    }
  )
}
```

## 16. Build POS-5 phrase table for Isaiah only

This can take time. It is cached to `output/pos_radar/isaiah_pos5_phrase_table.csv`.

```{r build-phrase-table}
out_dir <- file.path(STYLO_ROOT, "output", "pos_radar")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

phrase_cache <- file.path(
  out_dir,
  "isaiah_pos5_phrase_table.csv"
)

if(file.exists(phrase_cache)){

  isaiah_pos5_phrases <- readr::read_csv(
    phrase_cache,
    show_col_types = FALSE
  )

} else {

  isaiah_docs <- chapter_docs %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::arrange(chapter)

  isaiah_pos5_phrases <- extract_pos_ngrams_with_phrases(
    isaiah_docs,
    n = 5,
    text_col = "text",
    doc_id_col = "doc_id"
  )

  readr::write_csv(
    isaiah_pos5_phrases,
    phrase_cache
  )
}

isaiah_pos5_phrases %>%
  dplyr::slice_head(n = 20)
```

## 17. Phrase examples for top late-difference POS-5 patterns

```{r phrase-examples-late-diff}
top_late_diff_patterns <- top_pos5_late_diff %>%
  dplyr::slice_head(n = 10) %>%
  dplyr::pull(pos5)

top_late_diff_patterns

top_late_diff_examples <- show_top_pos5_examples(
  isaiah_pos5_phrases,
  pos5_patterns = top_late_diff_patterns,
  sections = c("Isaiah_40_55", "Isaiah_56_66"),
  top_n = 8
)

top_late_diff_examples
```

## 18. Phrase examples for radar top-20 patterns

```{r phrase-examples-top}
top_examples <- show_top_pos5_examples(
  isaiah_pos5_phrases,
  pos5_patterns = top_pos5,
  sections = c("Isaiah_1_35", "Isaiah_36_39", "Isaiah_40_55", "Isaiah_56_66"),
  top_n = 5
)

top_examples



show_top <- function(
    pos5_pattern,
    sections = c(
      "Isaiah_1_35",
      "Isaiah_36_39",
      "Isaiah_40_55",
      "Isaiah_56_66"
    ),
    top_n = 10,
    phrase_table = isaiah_pos5_phrases,
    wide = TRUE,
    sort_by = "total"
){

  out <- phrase_table %>%

    dplyr::filter(

      pos_ngram == pos5_pattern,

      section %in% sections

    ) %>%

    dplyr::count(

      section,

      phrase,

      name = "frequency"

    )


  #
  # LONG FORMAT
  #

  if(!wide){

    out <- out %>%

      dplyr::arrange(

        section,

        dplyr::desc(frequency),

        phrase

      ) %>%

      dplyr::group_by(

        section

      ) %>%

      dplyr::slice_head(

        n = top_n

      ) %>%

      dplyr::ungroup() %>%

      dplyr::mutate(

        pos5_pattern = pos5_pattern

      ) %>%

      dplyr::select(

        pos5_pattern,

        section,

        phrase,

        frequency

      )

    return(out)

  }


  #
  # WIDE FORMAT
  #

  out <- out %>%

    tidyr::pivot_wider(

      names_from = section,

      values_from = frequency,

      values_fill = 0

    )


  #
  # total frequency
  #

  out <- out %>%

    dplyr::mutate(

      total_frequency =

        rowSums(

          dplyr::across(

            dplyr::all_of(sections)

          )

        )

    )


  #
  # sorting
  #

  if(sort_by == "total"){

    out <- out %>%

      dplyr::arrange(

        dplyr::desc(total_frequency),

        phrase

      )

  }

  else{

    if(!(sort_by %in% names(out))){

      stop(

        paste(

          "sort_by must be one of:",

          paste(

            c("total", sections),

            collapse=", "

          )

        )

      )

    }


    out <- out %>%

      dplyr::arrange(

        dplyr::desc(

          .data[[sort_by]]

        ),

        dplyr::desc(total_frequency),

        phrase

      )

  }


  out %>%

    dplyr::slice_head(

      n = top_n

    ) %>%

    dplyr::relocate(

      phrase,

      total_frequency

    )

}

show_top(

  "DET_NOUN_ADP_DET_NOUN"

)

show_top(
  "DET_NOUN_ADP_DET_NOUN",
  sort_by = "Isaiah_40_55"
)

show_top(
  "DET_NOUN_ADP_DET_NOUN",
  sort_by = "Isaiah_56_66"
)

```

## 19. POS-5 phrase count by section

```{r phrase-count-by-section}
pos5_phrase_counts_by_section <- isaiah_pos5_phrases %>%
  dplyr::filter(pos_ngram %in% top_pos5) %>%
  dplyr::count(
    section,
    pos_ngram,
    name = "n_occurrences"
  ) %>%
  dplyr::group_by(section) %>%
  dplyr::mutate(
    section_total_pos5 = sum(n_occurrences),
    pct_occurrences = n_occurrences / section_total_pos5
  ) %>%
  dplyr::ungroup()

pos5_phrase_counts_by_section
```

## 20. POS-5 phrase count heatmap

```{r phrase-count-heatmap}
pos5_phrase_counts_by_section %>%
  dplyr::mutate(
    pos_ngram = factor(pos_ngram, levels = top_pos5)
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = pos_ngram,
      fill = pct_occurrences
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c(
    labels = scales::percent_format()
  ) +
  ggplot2::labs(
    title = "Top POS-5 phrase occurrences by section",
    subtitle = "Computed from recovered POS-5 windows in Isaiah.",
    x = "Section",
    y = "POS-5 pattern",
    fill = "Share"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
    axis.text.y = ggplot2::element_text(size = 7)
  )
```

## 21. POS-5-only ensemble

```{r pos5-only-ensemble}
result_pos5_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_pos5,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_pos5_section)
show_weights(result_pos5_section)
```

## 22. POS-5 attribution for Isaiah chapters

```{r pos5-attribution}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter)

isaiah_pos5_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_pos5_section,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(
        doc_id,
        chapter,
        actual_section = section,
        n_verses
      ),
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::arrange(chapter)

isaiah_pos5_attribution <- isaiah_pos5_consensus %>%
  dplyr::mutate(
    winner_matches_actual_section = winner == actual_section,
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  )

isaiah_pos5_attribution
```

## 23. POS-5 false matches

```{r pos5-false-matches}
show_false_matches(
  isaiah_pos5_attribution
)
```

## 24. Final summary

```{r final-summary}
cat("\nPOS-5 Radar Analysis Findings\n")
cat("=============================\n\n")

cat("Top 20 POS-5 patterns by section-level variance:\n")
print(
  pos5_ranked %>%
    dplyr::slice_head(n = 20),
  width = 120
)

cat("\nTop POS-5 differences: Isaiah_56_66 minus Isaiah_40_55:\n")
print(
  top_pos5_late_diff,
  width = 120
)

cat("\nPOS-5-only attribution false matches:\n")
print(
  show_false_matches(isaiah_pos5_attribution),
  width = 120
)
```

## 25. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "pos_radar")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  pos5_ranked,
  file.path(out_dir, "pos5_ranked_patterns.csv")
)

readr::write_csv(
  top_pos5_profile,
  file.path(out_dir, "top_pos5_section_profile.csv")
)

readr::write_csv(
  top_pos5_late_diff,
  file.path(out_dir, "top_pos5_5666_minus_4055.csv")
)

readr::write_csv(
  top_late_diff_examples,
  file.path(out_dir, "top_late_diff_pos5_phrase_examples.csv")
)

readr::write_csv(
  top_examples,
  file.path(out_dir, "top_pos5_phrase_examples.csv")
)

readr::write_csv(
  pos5_phrase_counts_by_section,
  file.path(out_dir, "top_pos5_phrase_counts_by_section.csv")
)

readr::write_csv(
  isaiah_pos5_attribution,
  file.path(out_dir, "pos5_only_attribution.csv")
)
```
