---
title: "12 - PCA Loadings"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 250)
```

## 1. Purpose

This notebook interprets PCA/MDS axes using loadings and feature correlations.

It asks:

- What features push PC1 positive or negative?
- Is positive PC1 a poetic/formulaic direction?
- Which feature classes dominate each latent dimension?
- Which target profiles define ensemble similarity PCA axes?
- Which features correlate with square ensemble MDS axes?

## 2. User parameters

```{r parameters}
top_n_loadings <- 25
pcs_to_show <- 1:6
focus_book <- "Isaiah"
comparison_books <- c("Psalms", "Job", "Proverbs", "Hosea")

include_feature_classes <- c(
  "char", "word", "stem", "function_words", "pos",
  "punctuation", "sentence", "word_length", "lexical", "readability"
)

pos_n <- c(3, 4, 5)
```

## 3. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){
  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)
} else {
  if(!file.exists(raw_path)) stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)
  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs
  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)
  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 4. Configure analysis

```{r config}
config_loadings <- default_config(include = include_feature_classes)
config_loadings$pos_n <- pos_n
if(exists("ud_model")) config_loadings$pos_model <- ud_model
config_loadings$weights
```

# Part A: Feature-space PCA loadings

## 5. Build feature sets

```{r build-features}
feature_sets <- build_feature_sets(
  chapter_docs,
  config = config_loadings,
  text_col = "text",
  doc_id_col = "doc_id"
)

names(feature_sets)
```

## 6. Flatten feature sets

```{r flatten-feature-sets}
flatten_feature_sets <- function(feature_sets){
  parts <- list()

  for(class_name in names(feature_sets)){
    obj <- feature_sets[[class_name]]

    if(is.matrix(obj) || is.data.frame(obj)){
      mat <- as.matrix(obj)
      colnames(mat) <- paste(class_name, colnames(mat), sep = "__")
      parts[[class_name]] <- mat
    } else if(is.list(obj)){
      for(i in seq_along(obj)){
        mat <- as.matrix(obj[[i]])
        suffix <- names(obj)[i]
        if(is.null(suffix) || suffix == "") suffix <- paste0("variant", i)
        colnames(mat) <- paste(class_name, suffix, colnames(mat), sep = "__")
        parts[[paste(class_name, suffix, sep = "__")]] <- mat
      }
    }
  }

  common_docs <- Reduce(intersect, lapply(parts, rownames))

  parts <- lapply(parts, function(m){
    m[common_docs, , drop = FALSE]
  })

  X <- do.call(cbind, parts)
  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0
  X
}

X_features <- flatten_feature_sets(feature_sets)
dim(X_features)
```

## 7. Filter zero-variance features

```{r filter-features}
feature_variance <- apply(X_features, 2, stats::var, na.rm = TRUE)
keep_features <- is.finite(feature_variance) & feature_variance > 0
X_features_filtered <- X_features[, keep_features, drop = FALSE]
dim(X_features_filtered)
```

## 8. Feature-space PCA

```{r feature-pca}
feature_pca <- stats::prcomp(
  X_features_filtered,
  center = TRUE,
  scale. = TRUE
)

feature_pca_var <- feature_pca$sdev^2

feature_pca_variance <- data.frame(
  pc = paste0("PC", seq_along(feature_pca_var)),
  dimension = seq_along(feature_pca_var),
  eigenvalue = feature_pca_var,
  prop_variance = feature_pca_var / sum(feature_pca_var),
  pct_variance = 100 * feature_pca_var / sum(feature_pca_var),
  cumulative_prop_variance = cumsum(feature_pca_var / sum(feature_pca_var)),
  cumulative_pct_variance = 100 * cumsum(feature_pca_var / sum(feature_pca_var))
)

feature_pca_variance %>% dplyr::slice_head(n = 12)
```

## 9. Feature PCA scores

```{r feature-pca-scores}
feature_pca_scores <- as.data.frame(feature_pca$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, book, chapter, section, macro_genre, n_verses)

feature_pca_scores %>%
  dplyr::filter(book == focus_book) %>%
  dplyr::arrange(chapter) %>%
  dplyr::select(doc_id, chapter, section, PC1, PC2, PC3, PC4)
```

## 10. Feature PCA plots

```{r feature-pca-plot-12}
plotPCA(
  feature_pca_scores,
  feature_pca_variance,
  x = 1,
  y = 2,
  book = focus_book,
  include_books = comparison_books,
  aggregate_other_books = TRUE,
  show_other_books = TRUE,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Feature-space PCA with comparison book centroids: PC1 vs PC2"
)
```

```{r feature-pca-plot-23}
plotPCA(
  feature_pca_scores,
  feature_pca_variance,
  x = 3,
  y = 2,
  book = focus_book,
  include_books = comparison_books,
  aggregate_other_books = TRUE,
  show_other_books = TRUE,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Feature-space PCA with comparison book centroids: PC2 vs PC3"
)
```

## 11. Parse feature names

```{r parse-feature-names}
parse_feature_name <- function(feature){
  parts <- strsplit(feature, "__", fixed = TRUE)[[1]]
  feature_class <- parts[1]

  if(length(parts) >= 3){
    variant <- parts[2]
    feature_name <- paste(parts[-c(1,2)], collapse = "__")
  } else if(length(parts) == 2){
    variant <- NA_character_
    feature_name <- parts[2]
  } else {
    variant <- NA_character_
    feature_name <- feature
  }

  data.frame(
    feature = feature,
    feature_class = feature_class,
    variant = variant,
    feature_name = feature_name,
    stringsAsFactors = FALSE
  )
}

feature_lookup <- purrr::map_dfr(colnames(X_features_filtered), parse_feature_name)
feature_lookup %>% dplyr::slice_head(n = 20)
```

## 12. Feature PCA loadings table

```{r feature-loadings-table}
feature_loadings <- as.data.frame(feature_pca$rotation) %>%
  tibble::rownames_to_column("feature") %>%
  dplyr::left_join(feature_lookup, by = "feature") %>%
  dplyr::relocate(feature, feature_class, variant, feature_name)

feature_loadings %>%
  dplyr::select(feature, feature_class, variant, feature_name, PC1, PC2, PC3, PC4) %>%
  dplyr::slice_head(n = 20)
```

## 13. Show top positive and negative feature loadings

```{r feature-loading-helper}
show_feature_pc_loadings <- function(
    loadings_tbl,
    pc = 1,
    n = top_n_loadings,
    feature_class_filter = NULL){

  pc_col <- paste0("PC", pc)
  dat <- loadings_tbl

  if(!is.null(feature_class_filter)){
    dat <- dat %>% dplyr::filter(feature_class %in% feature_class_filter)
  }

  positive <- dat %>%
    dplyr::select(feature, feature_class, variant, feature_name, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(dplyr::desc(loading)) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "positive")

  negative <- dat %>%
    dplyr::select(feature, feature_class, variant, feature_name, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(loading) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "negative")

  dplyr::bind_rows(positive, negative) %>%
    dplyr::arrange(direction, dplyr::desc(abs(loading)))
}

show_feature_pc_loadings(feature_loadings, pc = 1, n = top_n_loadings)
```

## 14. Loadings by PC

```{r feature-loadings-by-pc}
for(pc in pcs_to_show){
  cat("\n============================================================\n")
  cat("FEATURE-SPACE PC", pc, "\n", sep = "")
  cat("Variance explained: ", round(feature_pca_variance$pct_variance[pc], 2), "%\n", sep = "")
  cat("============================================================\n\n")

  print(
    show_feature_pc_loadings(feature_loadings, pc = pc, n = top_n_loadings),
    width = 140
  )
}
```

## 15. Plot feature loadings

```{r loading-plot-helper}
plot_feature_pc_loadings <- function(
    loadings_tbl,
    pc = 1,
    n = 15,
    feature_class_filter = NULL,
    title = NULL){

  dat <- show_feature_pc_loadings(
    loadings_tbl,
    pc = pc,
    n = n,
    feature_class_filter = feature_class_filter
  ) %>%
    dplyr::mutate(
      feature_label = paste0(direction, ": ", feature_class, " / ", feature_name),
      feature_label = stringr::str_trunc(feature_label, width = 75)
    )

  if(is.null(title)) title <- paste0("Top feature loadings: PC", pc)

  ggplot2::ggplot(
    dat,
    ggplot2::aes(
      x = reorder(feature_label, loading),
      y = loading,
      fill = direction
    )
  ) +
    ggplot2::geom_col() +
    ggplot2::coord_flip() +
    ggplot2::labs(
      title = title,
      subtitle = "Positive and negative features define opposite ends of the latent axis.",
      x = "Feature",
      y = paste0("PC", pc, " loading"),
      fill = "Direction"
    ) +
    ggplot2::theme_minimal()
}
```

```{r feature-loading-plots}
plot_feature_pc_loadings(feature_loadings, pc = 1, n = 20, title = "Feature-space PC1 loadings")
plot_feature_pc_loadings(feature_loadings, pc = 2, n = 20, title = "Feature-space PC2 loadings")
plot_feature_pc_loadings(feature_loadings, pc = 3, n = 20, title = "Feature-space PC3 loadings")
```

## 16. Feature-class contribution to each PC

```{r feature-class-contribution}
feature_class_contribution <- purrr::map_dfr(
  pcs_to_show,
  function(pc){
    pc_col <- paste0("PC", pc)

    feature_loadings %>%
      dplyr::mutate(
        abs_loading = abs(.data[[pc_col]]),
        squared_loading = .data[[pc_col]]^2
      ) %>%
      dplyr::group_by(feature_class) %>%
      dplyr::summarize(
        sum_abs_loading = sum(abs_loading, na.rm = TRUE),
        mean_abs_loading = mean(abs_loading, na.rm = TRUE),
        sum_squared_loading = sum(squared_loading, na.rm = TRUE),
        n_features = dplyr::n(),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        pc = paste0("PC", pc),
        dimension = pc,
        pct_squared_loading = sum_squared_loading / sum(sum_squared_loading)
      )
  }
)

feature_class_contribution %>%
  dplyr::arrange(dimension, dplyr::desc(pct_squared_loading))
```

```{r feature-class-contribution-plot}
feature_class_contribution %>%
  dplyr::filter(dimension %in% paste0("PC", 1:4)) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = feature_class,
      y = pct_squared_loading,
      fill = feature_class
    )
  ) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~ dimension, scales = "free_y") +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "Feature-class contribution to PCA dimensions",
    subtitle = "Contribution computed as share of squared loadings by feature class.",
    x = "Feature class",
    y = "Share of squared loading"
  ) +
  ggplot2::theme_minimal()
```

# Part B: Similarity-profile PCA loadings

## 17. Run chapter-to-section ensemble

```{r run-chapter-section}
result_chapter_section <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_loadings,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_chapter_section)
show_weights(result_chapter_section)
```

## 18. Similarity-profile PCA

```{r similarity-profile-pca}
S_rect <- as.matrix(result_chapter_section$overall_similarity)
S_rect[is.na(S_rect)] <- 0
S_rect[is.nan(S_rect)] <- 0
S_rect[is.infinite(S_rect)] <- 0

similarity_pca <- stats::prcomp(S_rect, center = TRUE, scale. = TRUE)
similarity_pca_var <- similarity_pca$sdev^2

similarity_pca_variance <- data.frame(
  pc = paste0("PC", seq_along(similarity_pca_var)),
  dimension = seq_along(similarity_pca_var),
  eigenvalue = similarity_pca_var,
  prop_variance = similarity_pca_var / sum(similarity_pca_var),
  pct_variance = 100 * similarity_pca_var / sum(similarity_pca_var),
  cumulative_prop_variance = cumsum(similarity_pca_var / sum(similarity_pca_var)),
  cumulative_pct_variance = 100 * cumsum(similarity_pca_var / sum(similarity_pca_var))
)

similarity_pca_variance %>% dplyr::slice_head(n = 12)
```

## 19. Similarity-profile scores and plot

```{r similarity-scores}
similarity_pca_scores <- as.data.frame(similarity_pca$x) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    chapter_docs %>%
      dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses),
    by = "doc_id"
  ) %>%
  dplyr::relocate(doc_id, book, chapter, section, macro_genre, n_verses)
```

```{r similarity-pca-plot}
plotPCA(
  similarity_pca_scores,
  similarity_pca_variance,
  x = 1,
  y = 2,
  book = focus_book,
  include_books = comparison_books,
  aggregate_other_books = TRUE,
  show_other_books = TRUE,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Similarity-profile PCA with comparison book centroids: PC1 vs PC2"
)
```

## 20. Similarity-profile loadings

```{r similarity-loadings}
similarity_loadings <- as.data.frame(similarity_pca$rotation) %>%
  tibble::rownames_to_column("target_profile")

similarity_loadings %>%
  dplyr::select(target_profile, PC1, PC2, PC3, PC4) %>%
  dplyr::arrange(dplyr::desc(abs(PC1)))
```

## 21. Show target-profile loadings

```{r similarity-loading-helper}
show_similarity_pc_loadings <- function(loadings_tbl, pc = 1, n = top_n_loadings){
  pc_col <- paste0("PC", pc)

  positive <- loadings_tbl %>%
    dplyr::select(target_profile, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(dplyr::desc(loading)) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "positive")

  negative <- loadings_tbl %>%
    dplyr::select(target_profile, loading = dplyr::all_of(pc_col)) %>%
    dplyr::arrange(loading) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "negative")

  dplyr::bind_rows(positive, negative) %>%
    dplyr::arrange(direction, dplyr::desc(abs(loading)))
}

for(pc in pcs_to_show){
  cat("\n============================================================\n")
  cat("SIMILARITY-PROFILE PC", pc, "\n", sep = "")
  cat("Variance explained: ", round(similarity_pca_variance$pct_variance[pc], 2), "%\n", sep = "")
  cat("============================================================\n\n")

  print(show_similarity_pc_loadings(similarity_loadings, pc = pc, n = top_n_loadings), width = 120)
}
```

## 22. Plot target-profile loadings

```{r similarity-loading-plot-helper}
plot_similarity_pc_loadings <- function(loadings_tbl, pc = 1, n = 20, title = NULL){
  dat <- show_similarity_pc_loadings(loadings_tbl, pc = pc, n = n) %>%
    dplyr::mutate(target_label = paste0(direction, ": ", target_profile))

  if(is.null(title)) title <- paste0("Similarity-profile PC", pc, " loadings")

  ggplot2::ggplot(
    dat,
    ggplot2::aes(
      x = reorder(target_label, loading),
      y = loading,
      fill = direction
    )
  ) +
    ggplot2::geom_col() +
    ggplot2::coord_flip() +
    ggplot2::labs(
      title = title,
      subtitle = "Positive and negative target profiles define opposite ends of the latent axis.",
      x = "Target profile",
      y = paste0("PC", pc, " loading"),
      fill = "Direction"
    ) +
    ggplot2::theme_minimal()
}
```

```{r similarity-loading-plots}
plot_similarity_pc_loadings(similarity_loadings, pc = 1, n = 20, title = "Similarity-profile PC1 loadings")
plot_similarity_pc_loadings(similarity_loadings, pc = 2, n = 20, title = "Similarity-profile PC2 loadings")
plot_similarity_pc_loadings(similarity_loadings, pc = 3, n = 20, title = "Similarity-profile PC3 loadings")
```

# Part C: Square ensemble MDS axis interpretation

## 23. Run chapter-to-chapter ensemble

```{r run-chapter-chapter}
result_chapter_chapter <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = chapter_docs,
  config = config_loadings,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_chapter_chapter)
```

## 24. MDS helper

```{r mds-helper}
mds_from_similarity <- function(sim, metadata, k = 9){
  S <- as.matrix(sim)
  S[is.na(S)] <- 0
  S[is.nan(S)] <- 0
  S[is.infinite(S)] <- 0

  D <- 1 - S
  D[D < 0] <- 0
  D <- (D + t(D)) / 2
  diag(D) <- 0

  fit <- stats::cmdscale(as.dist(D), k = k, eig = TRUE)

  scores <- as.data.frame(fit$points)
  colnames(scores) <- paste0("PC", seq_len(ncol(scores)))

  scores <- scores %>%
    tibble::rownames_to_column("doc_id") %>%
    dplyr::left_join(metadata, by = "doc_id")

  eig <- fit$eig[fit$eig > 0]

  variance <- data.frame(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  list(fit = fit, scores = scores, variance = variance)
}
```

## 25. Square ensemble MDS

```{r square-mds}
metadata <- chapter_docs %>%
  dplyr::select(doc_id, book, chapter, section, macro_genre, n_verses)

overall_mds <- mds_from_similarity(
  result_chapter_chapter$overall_similarity,
  metadata = metadata,
  k = 9
)

overall_mds$variance %>% dplyr::slice_head(n = 9)
```

```{r square-mds-plot}
plotPCA(
  overall_mds$scores,
  overall_mds$variance,
  x = 2,
  y = 3,
  book = focus_book,
  include_books = comparison_books,
  aggregate_other_books = TRUE,
  show_other_books = TRUE,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Square ensemble MDS with comparison book centroids: Dim2 vs Dim3"
)
```

## 26. Interpret MDS axes by feature correlations

MDS axes do not have native feature loadings. We therefore correlate each MDS axis with the original feature values.

```{r mds-feature-correlations}
mds_scores_for_corr <- overall_mds$scores %>%
  dplyr::select(doc_id, dplyr::starts_with("PC"))

feature_df_for_corr <- as.data.frame(X_features_filtered) %>%
  tibble::rownames_to_column("doc_id")

mds_feature_data <- mds_scores_for_corr %>%
  dplyr::inner_join(feature_df_for_corr, by = "doc_id")

feature_cols_for_corr <- setdiff(colnames(feature_df_for_corr), "doc_id")

compute_axis_feature_correlations <- function(data, axis = "PC1", feature_cols = feature_cols_for_corr){
  purrr::map_dfr(
    feature_cols,
    function(f){
      x <- data[[axis]]
      y <- data[[f]]

      if(stats::sd(y, na.rm = TRUE) == 0 || stats::sd(x, na.rm = TRUE) == 0){
        r <- NA_real_
      } else {
        r <- stats::cor(x, y, use = "pairwise.complete.obs")
      }

      data.frame(axis = axis, feature = f, correlation = r, stringsAsFactors = FALSE)
    }
  ) %>%
    dplyr::left_join(feature_lookup, by = "feature") %>%
    dplyr::relocate(axis, feature, feature_class, variant, feature_name, correlation) %>%
    dplyr::arrange(dplyr::desc(abs(correlation)))
}

mds_axis_correlations <- purrr::map_dfr(
  paste0("PC", pcs_to_show),
  function(axis){
    compute_axis_feature_correlations(mds_feature_data, axis = axis)
  }
)

mds_axis_correlations %>% dplyr::slice_head(n = 30)
```

## 27. Show MDS axis feature correlations

```{r show-mds-axis-correlations}
show_mds_axis_correlations <- function(
    corr_tbl,
    axis = "PC1",
    n = top_n_loadings,
    feature_class_filter = NULL){

  dat <- corr_tbl %>% dplyr::filter(.data$axis == !!axis)

  if(!is.null(feature_class_filter)){
    dat <- dat %>% dplyr::filter(feature_class %in% feature_class_filter)
  }

  positive <- dat %>%
    dplyr::arrange(dplyr::desc(correlation)) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "positive")

  negative <- dat %>%
    dplyr::arrange(correlation) %>%
    dplyr::slice_head(n = n) %>%
    dplyr::mutate(direction = "negative")

  dplyr::bind_rows(positive, negative) %>%
    dplyr::arrange(direction, dplyr::desc(abs(correlation)))
}

for(pc in pcs_to_show){
  axis <- paste0("PC", pc)

  cat("\n============================================================\n")
  cat("MDS ", axis, " feature correlations\n", sep = "")
  cat("============================================================\n\n")

  print(
    show_mds_axis_correlations(mds_axis_correlations, axis = axis, n = top_n_loadings),
    width = 140
  )
}
```

## 28. Plot MDS axis correlations

```{r plot-mds-corr-helper}
plot_mds_axis_correlations <- function(
    corr_tbl,
    axis = "PC1",
    n = 20,
    feature_class_filter = NULL,
    title = NULL){

  dat <- show_mds_axis_correlations(
    corr_tbl,
    axis = axis,
    n = n,
    feature_class_filter = feature_class_filter
  ) %>%
    dplyr::mutate(
      feature_label = paste0(direction, ": ", feature_class, " / ", feature_name),
      feature_label = stringr::str_trunc(feature_label, width = 75)
    )

  if(is.null(title)) title <- paste0("MDS ", axis, " feature correlations")

  ggplot2::ggplot(
    dat,
    ggplot2::aes(
      x = reorder(feature_label, correlation),
      y = correlation,
      fill = direction
    )
  ) +
    ggplot2::geom_col() +
    ggplot2::coord_flip() +
    ggplot2::labs(
      title = title,
      subtitle = "Features most associated with the positive and negative ends of the axis.",
      x = "Feature",
      y = "Correlation",
      fill = "Direction"
    ) +
    ggplot2::theme_minimal()
}
```

```{r mds-corr-plots}
plot_mds_axis_correlations(mds_axis_correlations, axis = "PC1", n = 20, title = "MDS PC1 feature correlations")
plot_mds_axis_correlations(mds_axis_correlations, axis = "PC2", n = 20, title = "MDS PC2 feature correlations")
plot_mds_axis_correlations(mds_axis_correlations, axis = "PC3", n = 20, title = "MDS PC3 feature correlations")
```

# Part D: Interpretive summary

## 29. Candidate interpretation table

```{r interpretation-table}
candidate_interpretations <- data.frame(
  axis = c(
    "Feature PCA PC1", "Feature PCA PC2",
    "Similarity PCA PC1", "Similarity PCA PC2",
    "Square MDS PC1", "Square MDS PC2", "Square MDS PC3"
  ),
  tentative_interpretation = c(
    "Inspect top positive/negative feature loadings.",
    "Inspect top positive/negative feature loadings.",
    "Inspect target-profile loadings.",
    "Inspect target-profile loadings.",
    "Inspect feature correlations with MDS axis.",
    "Inspect feature correlations with MDS axis.",
    "Inspect feature correlations with MDS axis."
  ),
  notes = c(
    "May reflect broad feature-space variance; often sparse.",
    "May reflect secondary feature-space contrast.",
    "Often interpretable as similarity to target sections/books.",
    "May separate genre or section structure.",
    "May reflect poetic/formulaic direction if Psalms/Job/Proverbs lie positive.",
    "May reflect narrative vs prophecy or prophetic vs wisdom style.",
    "Often captures subtler section or genre effects."
  )
)

candidate_interpretations
```

## 30. Final summary

```{r final-summary}
cat("\nPCA Loadings Summary\n")
cat("====================\n\n")

cat("Feature-space variance explained:\n")
print(feature_pca_variance %>% dplyr::slice_head(n = 8), width = 120)

cat("\nSimilarity-profile variance explained:\n")
print(similarity_pca_variance %>% dplyr::slice_head(n = 8), width = 120)

cat("\nTop feature-space PC1 loadings:\n")
print(show_feature_pc_loadings(feature_loadings, pc = 1, n = 12), width = 140)

cat("\nTop similarity-profile PC1 loadings:\n")
print(show_similarity_pc_loadings(similarity_loadings, pc = 1, n = 12), width = 120)

cat("\nTop MDS PC1 feature correlations:\n")
print(show_mds_axis_correlations(mds_axis_correlations, axis = "PC1", n = 12), width = 140)
```

## 31. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "pca_loadings")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(feature_pca_variance, file.path(out_dir, "feature_pca_variance.csv"))
readr::write_csv(feature_pca_scores, file.path(out_dir, "feature_pca_scores.csv"))
readr::write_csv(feature_loadings, file.path(out_dir, "feature_pca_loadings.csv"))
readr::write_csv(feature_class_contribution, file.path(out_dir, "feature_class_contribution.csv"))
readr::write_csv(similarity_pca_variance, file.path(out_dir, "similarity_profile_pca_variance.csv"))
readr::write_csv(similarity_pca_scores, file.path(out_dir, "similarity_profile_pca_scores.csv"))
readr::write_csv(similarity_loadings, file.path(out_dir, "similarity_profile_pca_loadings.csv"))
readr::write_csv(overall_mds$variance, file.path(out_dir, "square_mds_variance.csv"))
readr::write_csv(overall_mds$scores, file.path(out_dir, "square_mds_scores.csv"))
readr::write_csv(mds_axis_correlations, file.path(out_dir, "mds_axis_feature_correlations.csv"))
```
