---
title: "13 - Control Effects"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 250
)
```

## 1. Purpose

This notebook tests whether Isaiah chapter attributions are better explained by:

1. **Specific book / section collections**, or
2. **Macro-genre control collections**.

The target set allows both collection types to compete directly:

```text
d = source document, such as Isaiah_12
c = target collection, such as Psalms, Isaiah_40_55, or GENRE_Poetic_Wisdom
```

If a genre collection wins, the result is likely a **genre effect**.

If a specific book or Isaiah section wins, the result may reflect a more specific **book/section/authorship-style effect**.

## 2. User parameters

```{r parameters}
target_books <- c(
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings",
  "Psalms",
  "Proverbs",
  "Job",
  "Song of Solomon",
  "Ecclesiastes",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos"
)

narrative_books <- c(
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings"
)

poetic_wisdom_books <- c(
  "Psalms",
  "Proverbs",
  "Job",
  "Song of Solomon",
  "Ecclesiastes"
)

prophetic_books <- c(
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos"
)

isaiah_sections_to_include <- c(
  "Isaiah_1_35",
  "Isaiah_36_39",
  "Isaiah_40_55",
  "Isaiah_56_66"
)

include_feature_classes <- c(
  "char",
  "word",
  "stem",
  "function_words",
  "pos",
  "punctuation",
  "sentence",
  "word_length",
  "lexical",
  "readability"
)

pos_n <- c(3, 4, 5)

k_neighbors <- 10
```

## 3. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}
```

## 4. Macro-genre lookup

```{r macro-genre-lookup}
macro_genre_lookup <- tibble::tibble(
  book = c(
    narrative_books,
    poetic_wisdom_books,
    prophetic_books
  ),
  macro_genre_control = c(
    rep("Narrative", length(narrative_books)),
    rep("Poetic_Wisdom", length(poetic_wisdom_books)),
    rep("Prophetic", length(prophetic_books))
  )
)

macro_genre_lookup
```

## 5. Attach macro genre to chapters

```{r attach-genre}
chapter_docs_genre <- chapter_docs %>%
  dplyr::left_join(
    macro_genre_lookup,
    by = "book"
  )

chapter_docs_genre %>%
  dplyr::count(book, macro_genre_control) %>%
  dplyr::arrange(macro_genre_control, book)
```

## 6. Build collection targets

We create three kinds of target collections:

1. Isaiah section collections.
2. Book collections.
3. Macro-genre control collections.

These are **concatenated collections**, not centroids.

```{r collection-builders}
make_collection_docs <- function(
    docs,
    group_cols,
    id_prefix = "",
    collection_type = "collection"){

  docs %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_cols))) %>%
    dplyr::summarize(
      text = paste(text, collapse = " "),
      n_source_docs = dplyr::n(),
      n_verses = sum(n_verses, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      doc_id = paste0(
        id_prefix,
        do.call(
          paste,
          c(
            dplyr::across(dplyr::all_of(group_cols)),
            sep = "_"
          )
        )
      ),
      book = doc_id,
      chapter = NA_integer_,
      section = doc_id,
      macro_genre = NA_character_,
      collection_type = collection_type
    ) %>%
    dplyr::relocate(
      doc_id,
      book,
      chapter,
      section,
      macro_genre,
      collection_type,
      n_source_docs,
      n_verses,
      text
    )
}
```

### 6.1 Isaiah section targets

```{r isaiah-section-targets}

if("books" %in% names(section_docs)){

  section_docs <- section_docs %>%
    dplyr::rename(
      book = books
    )

}


target_cols <- c(
  "corpus_id",
  "translation",
  "doc_id",
  "book",
  "chapter",
  "section",
  "macro_genre",
  "collection_type",
  "n_source_docs",
  "n_verses",
  "text"
)


isaiah_section_targets <- section_docs %>%
  dplyr::filter(doc_id %in% isaiah_sections_to_include) %>%
  dplyr::mutate(
    chapter = NA_integer_,
    collection_type = "isaiah_section",
    n_source_docs = n_chapters
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    collection_type,
    n_source_docs,
    n_verses,
    text
  )


isaiah_section_targets <- isaiah_section_targets %>%
  dplyr::select(
    dplyr::all_of(target_cols)
  )

isaiah_section_targets
```

### 6.2 Book-level targets

```{r book-targets}
book_targets <- chapter_docs_genre %>%
  dplyr::filter(book %in% target_books) %>%
  make_collection_docs(
    group_cols = c("book"),
    id_prefix = "BOOK_",
    collection_type = "book"
  )



book_targets <- book_targets %>%
  dplyr::mutate(
    corpus_id = "KJV",
    translation = "KJV"
  ) %>%
  dplyr::select(
    dplyr::all_of(target_cols)
  )





book_targets
```

### 6.3 Macro-genre control targets

Isaiah is excluded from the macro-genre controls to avoid circularity.

```{r genre-targets}
genre_targets <- chapter_docs_genre %>%
  dplyr::filter(
    book %in% target_books,
    !is.na(macro_genre_control)
  ) %>%
  make_collection_docs(
    group_cols = c("macro_genre_control"),
    id_prefix = "GENRE_",
    collection_type = "genre_control"
  )

genre_targets <- genre_targets %>%
  dplyr::mutate(
    corpus_id = "KJV",
    translation = "KJV",
    macro_genre = macro_genre_control
  ) %>%
  dplyr::select(
    dplyr::all_of(target_cols)
  )


genre_targets
```

### 6.4 Combined target collections

```{r combined-targets}
target_collections <- dplyr::bind_rows(
  isaiah_section_targets,
  book_targets,
  genre_targets
)

target_collections
```

## 7. Configure ensemble

```{r config}
config_control <- default_config(
  include = include_feature_classes
)

config_control$pos_n <- pos_n

if(exists("ud_model")){
  config_control$pos_model <- ud_model
}

config_control$weights



```

## 8. Run control-effect ensemble

```{r run-control-ensemble}

chapter_docs <- chapter_docs %>%
  dplyr::select(
    -dplyr::any_of("books")
  )

source_cols <- c(
  "corpus_id",
  "translation",
  "doc_id",
  "book",
  "chapter",
  "section",
  "macro_genre",
  "n_verses",
  "text"
)

chapter_docs_for_control <- chapter_docs %>%
  dplyr::select(
    dplyr::all_of(source_cols)
  )

target_collections_for_control <- target_collections %>%
  dplyr::select(
    dplyr::all_of(source_cols)
  )
```


```{r work}

result_control_effects <- run_stylometry_ensemble(
  source_docs = chapter_docs_for_control,
  target_docs = target_collections_for_control,
  config = config_control,
  k_neighbors = k_neighbors,
  use_cache = TRUE
)

show_result_summary(result_control_effects)
show_weights(result_control_effects)
```

## 9. Isaiah chapter consensus against competing collections

```{r isaiah-control-consensus}
isaiah_doc_ids <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::pull(doc_id)

isaiah_chapters <- chapter_docs %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::arrange(chapter) %>%
  dplyr::select(
    doc_id,
    chapter,
    actual_section = section,
    n_verses
  )


target_collections <- target_collections %>%
  dplyr::mutate(
    target_family = dplyr::case_when(
      collection_type == "genre_control"  ~ "GENRE_CONTROL",
      collection_type == "book"           ~ "BOOK_COLLECTION",
      collection_type == "isaiah_section" ~ "ISAIAH_SECTION",
      TRUE                                ~ "OTHER"
    )
  )

control_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_control_effects,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters,
    by = "doc_id"
  ) %>%
  dplyr::left_join(
    target_collections %>%
      dplyr::select(
        winner = doc_id,
        winner_collection_type = collection_type,
        winner_target_family = target_family
      ),
    by = "winner"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner,
    winner_collection_type,
    winner_target_family
  ) %>%
  dplyr::mutate(
    winner_pct_votes_pct = 100 * winner_pct_votes,
    winner_weighted_pct_votes_pct = 100 * winner_weighted_pct_votes,
    vote_margin_pct = 100 * vote_margin,
    weighted_vote_margin_pct = 100 * weighted_vote_margin
  ) %>%
  dplyr::arrange(chapter)

control_consensus
```

## 10. Winner-family summary

```{r winner-family-summary}
control_consensus %>%
  dplyr::count(
    actual_section,
    winner_target_family,
    winner_collection_type,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(actual_section) %>%
  dplyr::mutate(
    pct_chapters = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(
    actual_section,
    dplyr::desc(n_chapters)
  )
```

## 11. Genre-control wins

```{r genre-control-wins}
genre_control_wins <- control_consensus %>%
  dplyr::filter(winner_target_family == "GENRE_CONTROL") %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    doc_entropy,
    effective_targets
  )

genre_control_wins
```

## 12. Isaiah-section wins

```{r isaiah-section-wins}
isaiah_section_wins <- control_consensus %>%
  dplyr::filter(winner_target_family == "ISAIAH_SECTION") %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    doc_entropy,
    effective_targets
  )

isaiah_section_wins
```

## 13. Book-collection wins

```{r book-collection-wins}
book_collection_wins <- control_consensus %>%
  dplyr::filter(winner_target_family == "BOOK_COLLECTION") %>%
  dplyr::select(
    chapter,
    doc_id,
    actual_section,
    winner,
    winner_pct_votes_pct,
    vote_margin_pct,
    doc_entropy,
    effective_targets
  )

book_collection_wins
```

## 14. Nearest target table for selected chapters

```{r selected-nearest}
selected_docs <- c(
  "Isaiah_12",
  "Isaiah_14",
  "Isaiah_36",
  "Isaiah_37",
  "Isaiah_40",
  "Isaiah_53",
  "Isaiah_55",
  "Isaiah_56",
  "Isaiah_61",
  "Isaiah_66"
)

result_control_effects$nearest_targets %>%
  dplyr::filter(doc_id %in% selected_docs) %>%
  dplyr::left_join(
    target_collections %>%
      dplyr::select(
        neighbor_doc_id = doc_id,
        collection_type,
        target_family
      ),
    by = "neighbor_doc_id"
  ) %>%
  dplyr::arrange(doc_id, rank)
```


```{r similarity}
nearest_targets_aug <- result_control_effects$nearest_targets %>%

  dplyr::left_join(

    target_collections %>%

      dplyr::select(

        neighbor_doc_id = doc_id,

        collection_type,

        target_family

      ),

    by = "neighbor_doc_id"

  )

family_similarity <- nearest_targets_aug %>%

  dplyr::group_by(

    doc_id,

    target_family

  ) %>%

  dplyr::summarise(

    similarity = sum(similarity),

    .groups="drop"

  ) %>%

  dplyr::group_by(

    doc_id

  ) %>%

  dplyr::mutate(

    pct_similarity =

      similarity /

      sum(similarity)

  ) %>%

  dplyr::ungroup()


family_similarity


control_vote_table <- result_control_effects$nearest_targets %>%

  dplyr::left_join(

    target_collections %>%

      dplyr::select(

        neighbor_doc_id = doc_id,

        collection_type,

        target_family

      ),

    by="neighbor_doc_id"

  )

control_vote_table %>%

  group_by(

    doc_id,

    target_family

  )

control_vote_table


```


## 15. Detailed findings for selected chapters

```{r detailed-findings-selected}
for(id in selected_docs){

  if(id %in% isaiah_doc_ids){

    cat("\n\n============================================================\n")
    cat(id, "\n")
    cat("============================================================\n\n")

    print(
      summarize_findings(
        result_control_effects,
        doc_id = id,
        k = k_neighbors,
        plot = TRUE
      )
    )
  }
}
```

## 16. Genre-control profile table

For each Isaiah chapter, show vote shares going to target-family classes.

```{r target-family-vote-profile}
control_vote_table <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    show_consensus(
      result_control_effects,
      doc_id = id
    ) %>%
      dplyr::mutate(doc_id = id)
  }
) %>%
  dplyr::left_join(
    target_collections %>%
      dplyr::select(
        neighbor_doc_id = doc_id,
        collection_type,
        target_family
      ),
    by = "neighbor_doc_id"
  ) %>%
  dplyr::left_join(
    isaiah_chapters %>%
      dplyr::select(doc_id, chapter, actual_section),
    by = "doc_id"
  )

target_family_vote_profile <- control_vote_table %>%
  dplyr::group_by(
    doc_id,
    chapter,
    actual_section,
    target_family
  ) %>%
  dplyr::summarize(
    votes = sum(votes, na.rm = TRUE),
    weighted_votes = sum(weighted_votes, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(doc_id) %>%
  dplyr::mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(chapter, dplyr::desc(pct_votes))

target_family_vote_profile
```


```{r bar}

target_family_vote_profile <- target_family_vote_profile %>%

  dplyr::mutate(

    target_family = factor(

      target_family,

      levels = c(

        "ISAIAH_SECTION",

        "BOOK_COLLECTION",

        "GENRE_CONTROL"

      )

    )

  )



target_family_vote_profile %>%

  ggplot2::ggplot(

    ggplot2::aes(

      x = factor(chapter),

      y = pct_votes,

      fill = target_family

    )

  ) +

  ggplot2::geom_col(

    width = 0.9

  ) +

  ggplot2::facet_grid(

    actual_section ~ .,

    scales = "free_x",

    space = "free_x"

  ) +

  ggplot2::scale_y_continuous(

    labels = scales::percent_format()

  ) +

  ggplot2::labs(

    title =

      "Control-effect vote profile by Isaiah chapter",

    subtitle =

      "Genre controls, book collections, and Isaiah sections compete as targets.",

    x =

      "Isaiah chapter",

    y =

      "Vote share",

    fill =

      "Target family"

  ) +

  ggplot2::theme_minimal() +

  ggplot2::theme(

    legend.position = "top",

    panel.grid.minor = ggplot2::element_blank(),

    

  )



target_family_vote_profile %>%

group_by(

   actual_section,

   target_family

) %>%

summarise(

   avg_vote_share = mean(pct_votes),

   .groups="drop"

) %>%

mutate(

   avg_vote_share =

      scales::percent(

         avg_vote_share,

         accuracy=1

      )

)

```



## 17. Target-family vote heatmap

```{r target-family-heatmap}
target_family_vote_profile %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = target_family,
      y = factor(chapter),
      fill = pct_votes
    )
  ) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_viridis_c(
    labels = scales::percent_format()
  ) +
  ggplot2::facet_grid(
    actual_section ~ .,
    scales = "free_y",
    space = "free_y"
  ) +
  ggplot2::labs(
    title = "Control-effect vote profile by target family",
    subtitle = "Genre controls, Isaiah sections, and book collections compete as target collections.",
    x = "Target family",
    y = "Isaiah chapter",
    fill = "Vote share"
  ) +
  ggplot2::theme_minimal()
```

## 18. Winner trajectory across Isaiah

```{r winner-trajectory}
control_consensus %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = chapter,
      y = winner_pct_votes_pct,
      color = winner_target_family,
      shape = winner_collection_type
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line(alpha = 0.45) +
  ggplot2::facet_grid(
    actual_section ~ .,
    scales = "free_x",
    space = "free_x"
  ) +
  ggplot2::labs(
    title = "Winning control target across Isaiah",
    subtitle = "Point color indicates whether the winner is genre, book, or Isaiah section.",
    x = "Isaiah chapter",
    y = "Winner vote share (%)",
    color = "Target family",
    shape = "Collection type"
  ) +
  ggplot2::theme_minimal()
```

## 19. Compare control-effect result to Isaiah-section-only result

```{r section-only-comparison}
result_section_only <- run_stylometry_ensemble(
  source_docs = chapter_docs,
  target_docs = section_docs,
  config = config_control,
  k_neighbors = k_neighbors,
  use_cache = TRUE
)

section_only_consensus <- purrr::map_dfr(
  isaiah_doc_ids,
  function(id){

    cons <- show_consensus(
      result_section_only,
      doc_id = id
    )

    attr(cons, "summary")
  }
) %>%
  dplyr::left_join(
    isaiah_chapters,
    by = "doc_id"
  ) %>%
  dplyr::relocate(
    doc_id,
    chapter,
    actual_section,
    winner
  ) %>%
  dplyr::mutate(
    section_only_winner = winner,
    section_only_vote_pct = 100 * winner_pct_votes
  ) %>%
  dplyr::select(
    doc_id,
    chapter,
    actual_section,
    section_only_winner,
    section_only_vote_pct
  )

control_vs_section_only <- control_consensus %>%
  dplyr::select(
    doc_id,
    chapter,
    actual_section,
    control_winner = winner,
    control_winner_family = winner_target_family,
    control_winner_type = winner_collection_type,
    control_vote_pct = winner_pct_votes_pct
  ) %>%
  dplyr::left_join(
    section_only_consensus,
    by = c("doc_id", "chapter", "actual_section")
  ) %>%
  dplyr::arrange(chapter)

control_vs_section_only
```

## 20. Chapters where control changes interpretation

```{r control-changes}
control_vs_section_only %>%
  dplyr::filter(
    control_winner_family != "ISAIAH_SECTION"
  )
```

## 21. Final summary

```{r final-summary}
cat("\nControl Effects Summary\n")
cat("=======================\n\n")

cat("Target collections:\n")
print(
  target_collections %>%
    dplyr::select(doc_id, collection_type, target_family, n_source_docs, n_verses) %>%
    dplyr::arrange(target_family, doc_id),
  width = 120
)

cat("\nWinner-family summary:\n")
print(
  control_consensus %>%
    dplyr::count(actual_section, winner_target_family, winner_collection_type, name = "n_chapters") %>%
    dplyr::group_by(actual_section) %>%
    dplyr::mutate(pct_chapters = n_chapters / sum(n_chapters)) %>%
    dplyr::ungroup(),
  width = 120
)

cat("\nGenre-control wins:\n")
print(genre_control_wins, width = 120)

cat("\nBook-collection wins:\n")
print(book_collection_wins, width = 120)

cat("\nIsaiah-section wins:\n")
print(isaiah_section_wins, width = 120)
```

## 22. Export outputs

```{r export}
out_dir <- file.path(STYLO_CACHE, "output", "control_effects")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  target_collections,
  file.path(out_dir, "target_collections.csv")
)

readr::write_csv(
  control_consensus,
  file.path(out_dir, "control_consensus.csv")
)

readr::write_csv(
  control_vote_table,
  file.path(out_dir, "control_vote_table.csv")
)

readr::write_csv(
  target_family_vote_profile,
  file.path(out_dir, "target_family_vote_profile.csv")
)

readr::write_csv(
  control_vs_section_only,
  file.path(out_dir, "control_vs_section_only.csv")
)

readr::write_csv(
  genre_control_wins,
  file.path(out_dir, "genre_control_wins.csv")
)

readr::write_csv(
  book_collection_wins,
  file.path(out_dir, "book_collection_wins.csv")
)

readr::write_csv(
  isaiah_section_wins,
  file.path(out_dir, "isaiah_section_wins.csv")
)
```
