---
title: "14 - Mclust Analysis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 250
)
```

## 1. Purpose

This notebook uses `mclust` to test whether Isaiah chapters are better modeled as one stylistic population or as multiple latent stylistic populations.

The goal is not to prove single authorship. The goal is to ask whether the data naturally supports strong, discrete clusters corresponding to:

```text
Isaiah_1_35
Isaiah_36_39
Isaiah_40_55
Isaiah_56_66
```

If `mclust` repeatedly prefers `G = 1`, that supports stylistic unity. If it prefers multiple clusters that align closely with traditional Isaiah divisions, that supports a stronger multi-population interpretation. If it prefers multiple clusters but they do not align with Isaiah sections, then the latent structure may reflect genre, hymnic material, narrative material, chapter length, or other stylistic effects.

## 2. Parameters

```{r parameters}
include_feature_classes <- c(
  "function_words",
  "pos",
  "punctuation",
  "sentence",
  "word_length",
  "lexical",
  "readability"
)

pos_n <- c(3, 4, 5)

max_pcs_for_mclust <- 9

mclust_G <- 1:6

focus_book <- "Isaiah"

comparison_books <- c(
  "Psalms",
  "Job",
  "Proverbs",
  "Song of Solomon",
  "Ecclesiastes",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos",
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings"
)

preferred_n_pcs <- 4

set.seed(1234)
```

## 3. Packages

```{r packages}
if(!requireNamespace("mclust", quietly = TRUE)){
  install.packages("mclust")
}

if(!requireNamespace("cluster", quietly = TRUE)){
  install.packages("cluster")
}

library(mclust)
library(cluster)
```

## 4. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}

if("books" %in% names(chapter_docs)){
  chapter_docs <- chapter_docs %>%
    dplyr::select(-books)
}

if("books" %in% names(section_docs)){
  section_docs <- section_docs %>%
    dplyr::rename(book = books)
}
```

## 5. Configure ensemble

```{r config}
config_mclust <- default_config(
  include = include_feature_classes
)

config_mclust$pos_n <- pos_n

if(exists("ud_model")){
  config_mclust$pos_model <- ud_model
}

config_mclust$weights
```

## 6. Isaiah metadata

```{r metadata}
isaiah_docs <- chapter_docs %>%
  dplyr::filter(book == focus_book) %>%
  dplyr::arrange(chapter)

isaiah_meta <- isaiah_docs %>%
  dplyr::select(
    doc_id,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  ) %>%
  dplyr::rename(
    actual_section = section
  )

isaiah_meta
```

# Part A: Overall similarity-profile clustering

## 7. Run chapter-to-chapter ensemble

```{r run-ensemble}
analysis_docs <- chapter_docs %>%
  dplyr::filter(
    book == focus_book |
      book %in% comparison_books
  )

result_mclust <- run_stylometry_ensemble(
  source_docs = analysis_docs,
  target_docs = analysis_docs,
  config = config_mclust,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_mclust)
```

## 8. Build Isaiah similarity profile matrix

```{r similarity-profile-matrix}
S <- as.matrix(result_mclust$overall_similarity)

S[is.na(S)] <- 0
S[is.nan(S)] <- 0
S[is.infinite(S)] <- 0

X_similarity_isaiah <- S[isaiah_meta$doc_id, , drop = FALSE]

dim(X_similarity_isaiah)
```

## 9. PCA helper

```{r process-pca-helper}
processPCA <- function(
    X,
    metadata = NULL,
    center = TRUE,
    scale. = TRUE){

  X <- as.matrix(X)

  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0

  feature_var <- apply(X, 2, stats::var, na.rm = TRUE)
  keep <- is.finite(feature_var) & feature_var > 0

  X <- X[, keep, drop = FALSE]

  pca <- stats::prcomp(
    X,
    center = center,
    scale. = scale.
  )

  eig <- pca$sdev^2

  variance <- tibble::tibble(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  scores <- as.data.frame(pca$x) %>%
    tibble::rownames_to_column("doc_id")

  if(!is.null(metadata)){
    scores <- scores %>%
      dplyr::left_join(
        metadata,
        by = "doc_id"
      )
  }

  loadings <- as.data.frame(pca$rotation) %>%
    tibble::rownames_to_column("feature")

  list(
    X = X,
    pca = pca,
    scores = scores,
    loadings = loadings,
    variance = variance
  )
}
```

## 10. Similarity-profile PCA

```{r similarity-pca}
sim_pca <- processPCA(
  X_similarity_isaiah,
  metadata = isaiah_meta,
  center = TRUE,
  scale. = TRUE
)

sim_pca$variance %>%
  dplyr::slice_head(n = 12)
```

```{r similarity-pca-plot}
sim_pca$scores <- sim_pca$scores %>%
  dplyr::mutate(
    section = actual_section
  )

plotPCA(
  sim_pca$scores,
  sim_pca$variance,
  x = 1,
  y = 2,
  book = focus_book,
  show_hulls = TRUE,
  show_centroids = TRUE,
  show_labels = TRUE,
  title = "Isaiah similarity-profile PCA: PC1 vs PC2"
)
```

## 11. Mclust helper functions

```{r mclust-helper}
run_mclust_on_pcs <- function(
    pca_obj,
    metadata,
    n_pcs = 3,
    G = mclust_G,
    modelNames = NULL){

  pc_cols <- paste0("PC", seq_len(n_pcs))

  X <- pca_obj$scores %>%
    dplyr::select(dplyr::all_of(pc_cols)) %>%
    as.matrix()

  rownames(X) <- pca_obj$scores$doc_id

  fit <- mclust::Mclust(
    X,
    G = G,
    modelNames = modelNames
  )

  classified <- tibble::tibble(
    doc_id = rownames(X),
    mclust_cluster = fit$classification,
    uncertainty = fit$uncertainty
  ) %>%
    dplyr::left_join(
      metadata,
      by = "doc_id"
    ) %>%
    dplyr::arrange(chapter)

  list(
    fit = fit,
    classified = classified,
    n_pcs = n_pcs,
    pc_cols = pc_cols
  )
}

summarize_mclust_fit <- function(mc_obj){

  fit <- mc_obj$fit

  tibble::tibble(
    n_pcs = mc_obj$n_pcs,
    selected_G = fit$G,
    selected_model = fit$modelName,
    bic = fit$bic,
    loglik = fit$loglik,
    df = fit$df,
    n = fit$n,
    mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
    max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
  )
}
```

## 12. Mclust sweep by number of PCs

```{r mclust-sweep}
pc_max_available <- min(
  max_pcs_for_mclust,
  sum(grepl("^PC", names(sim_pca$scores)))
)

mclust_sweep <- purrr::map(
  1:pc_max_available,
  function(k){
    run_mclust_on_pcs(
      sim_pca,
      metadata = isaiah_meta,
      n_pcs = k,
      G = mclust_G
    )
  }
)

mclust_sweep_summary <- purrr::map_dfr(
  mclust_sweep,
  summarize_mclust_fit
)

mclust_sweep_summary
```

```{r plot-selected-g}
mclust_sweep_summary %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = n_pcs,
      y = selected_G
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = 1:pc_max_available) +
  ggplot2::scale_y_continuous(breaks = mclust_G) +
  ggplot2::labs(
    title = "Mclust selected clusters by PCA dimensionality",
    subtitle = "Repeated G = 1 supports weak evidence for discrete stylistic populations.",
    x = "Number of PCs used",
    y = "Selected G"
  ) +
  ggplot2::theme_minimal()
```

## 13. Preferred mclust model

```{r preferred-model}
preferred_n_pcs <- min(preferred_n_pcs, pc_max_available)

mc_sim <- run_mclust_on_pcs(
  sim_pca,
  metadata = isaiah_meta,
  n_pcs = preferred_n_pcs,
  G = mclust_G
)

summary(mc_sim$fit)

mc_sim$classified
```

## 14. Cluster vs Isaiah section

```{r cluster-section-table}
cluster_section_table <- mc_sim$classified %>%
  dplyr::count(
    actual_section,
    mclust_cluster,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(actual_section) %>%
  dplyr::mutate(
    pct_section = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup()

cluster_section_table
```

```{r cluster-section-plot}
cluster_section_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = actual_section,
      y = pct_section,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Mclust clusters by Isaiah section",
    subtitle = paste0("Similarity-profile PCA space, first ", preferred_n_pcs, " PCs."),
    x = "Isaiah section",
    y = "Section share",
    fill = "Mclust cluster"
  ) +
  ggplot2::theme_minimal()
```

## 15. PCA plot colored by mclust cluster

```{r pca-cluster-plot}
sim_scores_clustered <- sim_pca$scores %>%
  dplyr::left_join(
    mc_sim$classified %>%
      dplyr::select(doc_id, mclust_cluster, uncertainty),
    by = "doc_id"
  ) %>%
  dplyr::mutate(
    mclust_cluster = factor(mclust_cluster)
  )

ggplot2::ggplot(
  sim_scores_clustered,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = mclust_cluster,
    label = chapter
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(size = 3) +
  ggrepel::geom_text_repel(size = 2.8, max.overlaps = Inf) +
  ggplot2::labs(
    title = "Isaiah similarity-profile PCA colored by mclust cluster",
    subtitle = paste0(
      "Selected G = ",
      mc_sim$fit$G,
      "; model = ",
      mc_sim$fit$modelName,
      "; PCs used = ",
      preferred_n_pcs
    ),
    x = paste0("PC1 (", round(sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Cluster"
  ) +
  ggplot2::theme_minimal()
```

## 16. Classification uncertainty

```{r uncertainty-plot}
mc_sim$classified %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = chapter,
      y = uncertainty,
      color = actual_section
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line(alpha = 0.5) +
  ggplot2::labs(
    title = "Mclust classification uncertainty across Isaiah",
    subtitle = "Higher uncertainty indicates weaker assignment to a discrete cluster.",
    x = "Isaiah chapter",
    y = "Classification uncertainty",
    color = "Actual section"
  ) +
  ggplot2::theme_minimal()
```

# Part B: Forced-cluster diagnostics

## 17. Force G = 1 through G = 4

```{r forced-clusters}
forced_mclust <- purrr::map_dfr(
  1:4,
  function(g){

    pc_cols <- paste0("PC", seq_len(preferred_n_pcs))

    X <- sim_pca$scores %>%
      dplyr::select(dplyr::all_of(pc_cols)) %>%
      as.matrix()

    rownames(X) <- sim_pca$scores$doc_id

    fit <- mclust::Mclust(
      X,
      G = g
    )

    tibble::tibble(
      forced_G = g,
      bic = fit$bic,
      model = fit$modelName,
      loglik = fit$loglik,
      df = fit$df,
      mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
      max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
    )
  }
)

forced_mclust
```

```{r forced-bic-plot}
forced_mclust %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = forced_G,
      y = bic
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = 1:4) +
  ggplot2::labs(
    title = "Mclust BIC under forced cluster counts",
    subtitle = "Higher BIC is preferred.",
    x = "Forced number of clusters",
    y = "BIC"
  ) +
  ggplot2::theme_minimal()
```

## 18. Silhouette diagnostics

```{r silhouette}
silhouette_summary <- purrr::map_dfr(
  2:6,
  function(k){

    pc_cols <- paste0("PC", seq_len(preferred_n_pcs))

    X <- sim_pca$scores %>%
      dplyr::select(dplyr::all_of(pc_cols)) %>%
      as.matrix()

    km <- stats::kmeans(
      X,
      centers = k,
      nstart = 100
    )

    sil <- cluster::silhouette(
      km$cluster,
      dist(X)
    )

    tibble::tibble(
      k = k,
      avg_silhouette = mean(sil[, "sil_width"])
    )
  }
)

silhouette_summary
```

```{r silhouette-plot}
silhouette_summary %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = k,
      y = avg_silhouette
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = 2:6) +
  ggplot2::labs(
    title = "Forced k-means silhouette diagnostics",
    subtitle = "Low silhouette values suggest weak discrete separation.",
    x = "Number of clusters",
    y = "Average silhouette width"
  ) +
  ggplot2::theme_minimal()
```

# Part C: Feature-class similarity clustering

## 19. Class-specific mclust

```{r class-specific-mclust}
class_mclust_results <- purrr::imap(
  result_mclust$class_similarity,
  function(sim, class_name){

    S_class <- as.matrix(sim)

    S_class[is.na(S_class)] <- 0
    S_class[is.nan(S_class)] <- 0
    S_class[is.infinite(S_class)] <- 0

    if(!all(isaiah_meta$doc_id %in% rownames(S_class))){
      return(NULL)
    }

    X_class <- S_class[isaiah_meta$doc_id, , drop = FALSE]

    pca_class <- processPCA(
      X_class,
      metadata = isaiah_meta,
      center = TRUE,
      scale. = TRUE
    )

    pc_max <- min(
      preferred_n_pcs,
      sum(grepl("^PC", names(pca_class$scores)))
    )

    mc <- run_mclust_on_pcs(
      pca_class,
      metadata = isaiah_meta,
      n_pcs = pc_max,
      G = mclust_G
    )

    list(
      class_name = class_name,
      pca = pca_class,
      mclust = mc,
      summary = summarize_mclust_fit(mc) %>%
        dplyr::mutate(feature_class = class_name)
    )
  }
)

class_mclust_results <- class_mclust_results[
  !vapply(class_mclust_results, is.null, logical(1))
]

class_mclust_summary <- purrr::map_dfr(
  class_mclust_results,
  "summary"
)

class_mclust_summary %>%
  dplyr::arrange(feature_class)
```

```{r class-selected-g-plot}
class_mclust_summary %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(feature_class, selected_G),
      y = selected_G,
      fill = feature_class
    )
  ) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::coord_flip() +
  ggplot2::scale_y_continuous(breaks = mclust_G) +
  ggplot2::labs(
    title = "Selected cluster count by feature class",
    subtitle = "Class-specific similarity spaces reveal whether a feature class supports discrete clustering.",
    x = "Feature class",
    y = "Selected G"
  ) +
  ggplot2::theme_minimal()
```

## 20. Class-specific cluster tables

```{r class-cluster-tables}
for(class_name in names(class_mclust_results)){

  cat("\n============================================================\n")
  cat("Feature class: ", class_name, "\n", sep = "")
  cat("============================================================\n\n")

  mc <- class_mclust_results[[class_name]]$mclust

  print(
    summarize_mclust_fit(mc),
    width = 120
  )

  print(
    mc$classified %>%
      dplyr::count(actual_section, mclust_cluster, name = "n_chapters") %>%
      dplyr::group_by(actual_section) %>%
      dplyr::mutate(pct_section = n_chapters / sum(n_chapters)) %>%
      dplyr::ungroup(),
    width = 120
  )
}
```

# Part D: Alignment and interpretation

## 21. Adjusted Rand Index

```{r adjusted-rand-index}
ari_overall <- mclust::adjustedRandIndex(
  mc_sim$classified$actual_section,
  mc_sim$classified$mclust_cluster
)

ari_overall
```

```{r ari-by-class}
ari_by_class <- purrr::imap_dfr(
  class_mclust_results,
  function(obj, class_name){

    tibble::tibble(
      feature_class = class_name,
      adjusted_rand_index =
        mclust::adjustedRandIndex(
          obj$mclust$classified$actual_section,
          obj$mclust$classified$mclust_cluster
        ),
      selected_G = obj$mclust$fit$G,
      selected_model = obj$mclust$fit$modelName,
      mean_uncertainty = mean(obj$mclust$fit$uncertainty, na.rm = TRUE)
    )
  }
)

ari_by_class %>%
  dplyr::arrange(dplyr::desc(adjusted_rand_index))
```

## 22. Chapters by cluster

```{r chapters-by-cluster}
mc_sim$classified %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::summarise(
    chapters = paste(chapter, collapse = ", "),
    sections = paste(unique(actual_section), collapse = "; "),
    mean_uncertainty = mean(uncertainty, na.rm = TRUE),
    .groups = "drop"
  )
```

## 23. Interpretive guide

```{r interpretive-guide}
interpretive_guide <- tibble::tribble(
  ~finding, ~interpretation,
  "G = 1 repeatedly selected", "Supports a single broad stylistic population; weak evidence for discrete authorial partitions.",
  "G > 1 selected, but clusters do not match Isaiah sections", "Suggests latent structure, but not the traditional Deutero/Trito division; investigate genre, length, hymns, narrative, or outliers.",
  "G = 2 selected and clusters split near chapter 40", "Potential support for a Deutero-Isaiah-style partition, especially if stable across feature classes.",
  "G = 4 selected and clusters match Isaiah_1_35 / 36_39 / 40_55 / 56_66", "Strongest clustering support for traditional section-level differentiation.",
  "High mean uncertainty", "Even if multiple clusters are selected, assignments are weak or overlapping.",
  "Low silhouette", "Forced clusters are not well separated."
)

interpretive_guide
```

## 24. Final summary

```{r final-summary}
cat("\nMclust Summary\n")
cat("==============\n\n")

cat("Similarity-profile PCA variance:\n")
print(
  sim_pca$variance %>%
    dplyr::slice_head(n = 8),
  width = 120
)

cat("\nMclust sweep summary:\n")
print(
  mclust_sweep_summary,
  width = 120
)

cat("\nPreferred model:\n")
print(
  summarize_mclust_fit(mc_sim),
  width = 120
)

cat("\nAdjusted Rand Index vs Isaiah sections:\n")
print(ari_overall)

cat("\nClass-specific mclust summary:\n")
print(
  class_mclust_summary,
  width = 120
)

cat("\nARI by feature class:\n")
print(
  ari_by_class,
  width = 120
)
```

## 25. Export outputs

```{r export}
out_dir <- file.path(STYLO_CACHE, "output", "mclust")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  sim_pca$variance,
  file.path(out_dir, "similarity_pca_variance.csv")
)

readr::write_csv(
  sim_pca$scores,
  file.path(out_dir, "similarity_pca_scores.csv")
)

readr::write_csv(
  mclust_sweep_summary,
  file.path(out_dir, "mclust_sweep_summary.csv")
)

readr::write_csv(
  mc_sim$classified,
  file.path(out_dir, "mclust_preferred_classified.csv")
)

readr::write_csv(
  cluster_section_table,
  file.path(out_dir, "mclust_cluster_section_table.csv")
)

readr::write_csv(
  forced_mclust,
  file.path(out_dir, "mclust_forced_g.csv")
)

readr::write_csv(
  silhouette_summary,
  file.path(out_dir, "silhouette_summary.csv")
)

readr::write_csv(
  class_mclust_summary,
  file.path(out_dir, "class_mclust_summary.csv")
)

readr::write_csv(
  ari_by_class,
  file.path(out_dir, "ari_by_class.csv")
)
```
