---
title: "15 - Mclust All Chapters"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook extends the `mclust` experiment beyond Isaiah and examines **all available chapters** in the current analysis corpus.

The question is:

```text
Across 520+ biblical chapters, how many broad stylometric populations are naturally supported?
```

We test:

```text
G = 1 through G = 9
```

and evaluate whether clusters align with:

- macro genre,
- book,
- Isaiah sections,
- poetic/wisdom books,
- narrative books,
- prophetic books.

This provides a broader control context for the Isaiah-only result.

## 2. Parameters

```{r parameters}
include_feature_classes <- c(
  "function_words",
  #"pos",
  "punctuation",
  "sentence",
  "word_length",
  "lexical",
  "readability"
)

# where is char/word pos

pos_n <- c(3, 4, 5)

mclust_G <- 1:9

max_pcs_for_mclust <- 12

preferred_n_pcs <- 6

comparison_books <- c(
  "Isaiah",
  "Psalms",
  "Job",
  "Proverbs",
  "Song of Solomon",
  "Ecclesiastes",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos",
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings"
)

set.seed(1234)
mclust_G
```

## 3. Packages

```{r packages}
if(!requireNamespace("mclust", quietly = TRUE)){
  install.packages("mclust")
}

if(!requireNamespace("cluster", quietly = TRUE)){
  install.packages("cluster")
}

library(mclust)
library(cluster)
```

## 4. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}

if("books" %in% names(chapter_docs)){
  chapter_docs <- chapter_docs %>%
    dplyr::select(-books)
}

if("books" %in% names(section_docs)){
  section_docs <- section_docs %>%
    dplyr::rename(book = books)
}
```

## 5. Macro genre normalization

```{r macro-genre}
macro_genre_lookup <- tibble::tribble(
  ~book,              ~macro_genre_control,
  "1 Samuel",         "Narrative",
  "2 Samuel",         "Narrative",
  "1 Kings",          "Narrative",
  "2 Kings",          "Narrative",

  "Psalms",           "Poetic_Wisdom",
  "Job",              "Poetic_Wisdom",
  "Proverbs",         "Poetic_Wisdom",
  "Song of Solomon",  "Poetic_Wisdom",
  "Ecclesiastes",     "Poetic_Wisdom",

  "Isaiah",           "Prophetic",
  "Jeremiah",         "Prophetic",
  "Ezekiel",          "Prophetic",
  "Hosea",            "Prophetic",
  "Amos",             "Prophetic"
)

chapter_docs <- chapter_docs %>%
  dplyr::left_join(
    macro_genre_lookup,
    by = "book"
  ) %>%
  dplyr::mutate(
    macro_genre = dplyr::coalesce(
      macro_genre_control,
      macro_genre
    )
  ) %>%
  dplyr::select(-dplyr::any_of("macro_genre_control"))

chapter_docs %>%
  dplyr::count(book, macro_genre) %>%
  dplyr::arrange(macro_genre, book)
```

## 6. Analysis corpus

```{r analysis-corpus}
analysis_docs <- chapter_docs %>%
  dplyr::filter(book %in% comparison_books) %>%
  dplyr::arrange(book, chapter)

analysis_meta <- analysis_docs %>%
  dplyr::select(
    doc_id,
    corpus_id,
    translation,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  ) %>%
  dplyr::mutate(
    is_isaiah = book == "Isaiah",
    isaiah_section = dplyr::if_else(
      book == "Isaiah",
      section,
      NA_character_
    )
  )

analysis_meta %>%
  dplyr::count(book, macro_genre)

nrow(analysis_docs)
```

## 7. Configure ensemble

```{r config}
config_all_mclust <- default_config(
  include = include_feature_classes
)

config_all_mclust$pos_n <- pos_n

if(exists("ud_model")){
  config_all_mclust$pos_model <- ud_model
}

config_all_mclust$weights

config_equal <- config_all_mclust

config_equal$weights <- c(
  function_words = 1,
  punctuation    = 1,
  sentence       = 1,
  word_length    = 1,
  lexical        = 1,
  readability    = 1
)


config_equal$weights <- config_equal$weights / sum(config_equal$weights)
config_equal$weights


config_byfeature <- config_all_mclust
config_byfeature$weights <- c(
  function_words = 0.832,
  punctuation    = 0.071,
  sentence       = 0.031,
  word_length    = 0.031,
  lexical        = 0.019,
  readability    = 0.016
)

config_byfeature$weights <- config_byfeature$weights / sum(config_byfeature$weights)
config_byfeature$weights




config_freq_only <- config_all_mclust

config_freq_only$use_tfidf_function_words <- FALSE



```

# Part A: All-chapter similarity space

## 8. Run chapter-to-chapter ensemble

```{r count features}
count_features <- function(x){

  if(is.matrix(x) || is.data.frame(x)){

    return(ncol(x))

  }

  if(is.list(x)){

    return(

      sum(

        purrr::map_int(

          x,

          count_features

        )

      )

    )

  }

  0L

}


show_feature_counts <- function(feature_sets){

  out <- tibble::tibble(

    feature_class = names(feature_sets),

    n_features = purrr::map_int(

      feature_sets,

      count_features

    )

  ) %>%

  dplyr::mutate(

    pct_features =

      n_features /

      sum(n_features),

    pct_features =

      scales::percent(

        pct_features,

        accuracy = 0.1

      )

  ) %>%

  dplyr::arrange(

    dplyr::desc(n_features)

  )

  total_row <- tibble::tibble(

    feature_class = "TOTAL",

    n_features = sum(out$n_features),

    pct_features = "100.0%"

  )

  dplyr::bind_rows(

    out,

    total_row

  )

}
```

```{r run-ensemble}
result_all_mclust <- run_stylometry_ensemble(
  source_docs = analysis_docs,
  target_docs = analysis_docs,
  config = config_all_mclust,
  #config = config_equal,
  #config = config_byfeature,
  #config = config_freq_only,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_all_mclust)
show_feature_counts(
  result_all_mclust$feature_sets
)
```

## 9. Build similarity matrix

```{r similarity-matrix}
S_all <- as.matrix(result_all_mclust$overall_similarity)

S_all[is.na(S_all)] <- 0
S_all[is.nan(S_all)] <- 0
S_all[is.infinite(S_all)] <- 0

S_all <- S_all[analysis_meta$doc_id, analysis_meta$doc_id, drop = FALSE]

dim(S_all)
```

## 10. PCA helper

```{r pca-helper}
processPCA <- function(
    X,
    metadata = NULL,
    center = TRUE,
    scale. = TRUE){

  X <- as.matrix(X)

  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0

  feature_var <- apply(X, 2, stats::var, na.rm = TRUE)
  keep <- is.finite(feature_var) & feature_var > 0

  X <- X[, keep, drop = FALSE]

  pca <- stats::prcomp(
    X,
    center = center,
    scale. = scale.
  )

  eig <- pca$sdev^2

  variance <- tibble::tibble(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  scores <- as.data.frame(pca$x) %>%
    tibble::rownames_to_column("doc_id")

  if(!is.null(metadata)){
    scores <- scores %>%
      dplyr::left_join(
        metadata,
        by = "doc_id"
      )
  }

  loadings <- as.data.frame(pca$rotation) %>%
    tibble::rownames_to_column("feature")

  list(
    X = X,
    pca = pca,
    scores = scores,
    loadings = loadings,
    variance = variance
  )
}
```

## 11. PCA on all-chapter similarity profiles

```{r all-similarity-pca}
all_sim_pca <- processPCA(
  S_all,
  metadata = analysis_meta,
  center = TRUE,
  scale. = TRUE
)

all_sim_pca$variance %>%
  dplyr::slice_head(n = 15)
```

## 12. PCA plots by macro genre and book

```{r pca-by-genre}
ggplot2::ggplot(
  all_sim_pca$scores,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = macro_genre,
    label = ifelse(book == "Isaiah", chapter, "")
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(alpha = 0.75, size = 2.5) +
  ggrepel::geom_text_repel(size = 2.6, max.overlaps = Inf, show.legend = FALSE) +
  ggplot2::labs(
    title = "All-chapter similarity-profile PCA: PC1 vs PC2",
    subtitle = "Isaiah chapters are labeled; all chapters are colored by macro genre.",
    x = paste0("PC1 (", round(all_sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(all_sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Macro genre"
  ) +
  ggplot2::theme_minimal()
```

```{r pca-by-book}
ggplot2::ggplot(
  all_sim_pca$scores,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = book,
    label = ifelse(book == "Isaiah", chapter, "")
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(alpha = 0.7, size = 2.3) +
  ggrepel::geom_text_repel(size = 2.5, max.overlaps = Inf, show.legend = FALSE) +
  ggplot2::labs(
    title = "All-chapter similarity-profile PCA colored by book",
    subtitle = "Isaiah chapters are labeled.",
    x = paste0("PC1 (", round(all_sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(all_sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Book"
  ) +
  ggplot2::theme_minimal()
```

# Part B: Mclust on all chapters

## 13. Mclust helpers

```{r mclust-helpers}
run_mclust_on_pcs <- function(
    pca_obj,
    metadata,
    n_pcs = 3,
    G = mclust_G,
    modelNames = NULL){

  pc_cols <- paste0("PC", seq_len(n_pcs))

  X <- pca_obj$scores %>%
    dplyr::select(dplyr::all_of(pc_cols)) %>%
    as.matrix()

  rownames(X) <- pca_obj$scores$doc_id

  fit <- mclust::Mclust(
    X,
    G = G,
    modelNames = modelNames
  )

  classified <- tibble::tibble(
    doc_id = rownames(X),
    mclust_cluster = fit$classification,
    uncertainty = fit$uncertainty
  ) %>%
    dplyr::left_join(
      metadata,
      by = "doc_id"
    ) %>%
    dplyr::arrange(book, chapter)

  list(
    fit = fit,
    classified = classified,
    n_pcs = n_pcs,
    pc_cols = pc_cols
  )
}

summarize_mclust_fit <- function(mc_obj){

  fit <- mc_obj$fit

  tibble::tibble(
    n_pcs = mc_obj$n_pcs,
    selected_G = fit$G,
    selected_model = fit$modelName,
    bic = fit$bic,
    loglik = fit$loglik,
    df = fit$df,
    n = fit$n,
    mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
    max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
  )
}
```

## 14. Mclust sweep across PC dimensions

```{r mclust-sweep}
pc_max_available <- min(
  max_pcs_for_mclust,
  sum(grepl("^PC", names(all_sim_pca$scores)))
)

mclust_sweep_all <- purrr::map(
  1:pc_max_available,
  function(k){
    run_mclust_on_pcs(
      all_sim_pca,
      metadata = analysis_meta,
      n_pcs = k,
      G = mclust_G
    )
  }
)

mclust_sweep_all_summary <- purrr::map_dfr(
  mclust_sweep_all,
  summarize_mclust_fit
)

mclust_sweep_all_summary
```

```{r selected-g-plot}
mclust_sweep_all_summary %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = n_pcs,
      y = selected_G
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = 1:pc_max_available) +
  ggplot2::scale_y_continuous(breaks = mclust_G) +
  ggplot2::labs(
    title = "Selected mclust G by PCA dimensionality",
    subtitle = "All available chapters, tested over G = 1 through 6.",
    x = "Number of PCs used",
    y = "Selected G"
  ) +
  ggplot2::theme_minimal()
```

## 15. Preferred all-chapter model

```{r preferred-model}
preferred_n_pcs <- min(preferred_n_pcs, pc_max_available)

mc_all <- run_mclust_on_pcs(
  all_sim_pca,
  metadata = analysis_meta,
  n_pcs = preferred_n_pcs,
  G = mclust_G
)

summary(mc_all$fit)

mc_all$classified
```

## 16. Forced G = 1 through G = 9

```{r forced-g}
forced_mclust_all <- purrr::map_dfr(
  mclust_G,
  function(g){

    pc_cols <- paste0("PC", seq_len(preferred_n_pcs))

    X <- all_sim_pca$scores %>%
      dplyr::select(dplyr::all_of(pc_cols)) %>%
      as.matrix()

    rownames(X) <- all_sim_pca$scores$doc_id

    fit <- mclust::Mclust(
      X,
      G = g
    )

    tibble::tibble(
      forced_G = g,
      bic = fit$bic,
      model = fit$modelName,
      loglik = fit$loglik,
      df = fit$df,
      mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
      max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
    )
  }
)

forced_mclust_all
```

```{r forced-g-plot}
forced_mclust_all %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = forced_G,
      y = bic
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = mclust_G) +
  ggplot2::labs(
    title = "All-chapter mclust BIC under forced cluster counts",
    subtitle = "Higher BIC is preferred.",
    x = "Forced number of clusters",
    y = "BIC"
  ) +
  ggplot2::theme_minimal()
```

## 17. Silhouette diagnostics

```{r silhouette}
silhouette_summary_all <- purrr::map_dfr(
  2:6,
  function(k){

    pc_cols <- paste0("PC", seq_len(preferred_n_pcs))

    X <- all_sim_pca$scores %>%
      dplyr::select(dplyr::all_of(pc_cols)) %>%
      as.matrix()

    km <- stats::kmeans(
      X,
      centers = k,
      nstart = 100
    )

    sil <- cluster::silhouette(
      km$cluster,
      dist(X)
    )

    tibble::tibble(
      k = k,
      avg_silhouette = mean(sil[, "sil_width"])
    )
  }
)

silhouette_summary_all
```

```{r silhouette-plot}
silhouette_summary_all %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = k,
      y = avg_silhouette
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(breaks = 2:6) +
  ggplot2::labs(
    title = "All-chapter forced k-means silhouette diagnostics",
    subtitle = "Low silhouette values suggest weak discrete separation.",
    x = "Number of clusters",
    y = "Average silhouette width"
  ) +
  ggplot2::theme_minimal()
```

# Part C: Cluster interpretation

## 18. Cluster profile by macro genre

```{r cluster-by-genre}
cluster_genre_table <- mc_all$classified %>%
  dplyr::count(
    mclust_cluster,
    macro_genre,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(
    pct_cluster = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup()

cluster_genre_table
```



```{r cluster-by-genre-plot}
cluster_genre_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = factor(mclust_cluster),
      y = pct_cluster,
      fill = macro_genre
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "All-chapter mclust clusters by macro genre",
    subtitle = "Each bar shows the genre composition of a selected cluster.",
    x = "Mclust cluster",
    y = "Cluster share",
    fill = "Macro genre"
  ) +
  ggplot2::theme_minimal()
```

## 19. Cluster profile by book

```{r cluster-by-book}
cluster_book_table <- mc_all$classified %>%
  dplyr::count(
    mclust_cluster,
    book,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(
    pct_cluster = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_book_table
```

```{r cluster-by-book-plot}
cluster_book_table %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::slice_max(order_by = n_chapters, n = 8, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(book, n_chapters),
      y = n_chapters,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~ mclust_cluster, scales = "free_y") +
  ggplot2::labs(
    title = "Top books represented in each mclust cluster",
    subtitle = "Only the top eight books per cluster are shown.",
    x = "Book",
    y = "Chapter count"
  ) +
  ggplot2::theme_minimal()
```

## 20. Isaiah chapters by all-chapter cluster

```{r isaiah-cluster-table}
isaiah_cluster_table <- mc_all$classified %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::count(
    section,
    mclust_cluster,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(section) %>%
  dplyr::mutate(
    pct_section = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup()

isaiah_cluster_table
```

```{r isaiah-cluster-plot}
isaiah_cluster_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = pct_section,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Isaiah section distribution across all-chapter mclust clusters",
    subtitle = "Tests whether all-chapter clusters reproduce traditional Isaiah divisions.",
    x = "Isaiah section",
    y = "Section share",
    fill = "Mclust cluster"
  ) +
  ggplot2::theme_minimal()
```

## 21. PCA colored by mclust cluster

```{r pca-cluster-plot}
all_scores_clustered <- all_sim_pca$scores %>%
  dplyr::left_join(
    mc_all$classified %>%
      dplyr::select(doc_id, mclust_cluster, uncertainty),
    by = "doc_id"
  ) %>%
  dplyr::mutate(
    mclust_cluster = factor(mclust_cluster)
  )

ggplot2::ggplot(
  all_scores_clustered,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = mclust_cluster,
    label = ifelse(book == "Isaiah", chapter, "")
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(size = 2.5, alpha = 0.75) +
  ggrepel::geom_text_repel(size = 2.6, max.overlaps = Inf, show.legend = FALSE) +
  ggplot2::labs(
    title = "All-chapter PCA colored by mclust cluster",
    subtitle = paste0(
      "Selected G = ",
      mc_all$fit$G,
      "; model = ",
      mc_all$fit$modelName,
      "; PCs used = ",
      preferred_n_pcs
    ),
    x = paste0("PC1 (", round(all_sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(all_sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Cluster"
  ) +
  ggplot2::theme_minimal()
```

## 22. Classification uncertainty by book

```{r uncertainty-by-book}
mc_all$classified %>%
  dplyr::group_by(book, macro_genre) %>%
  dplyr::summarise(
    mean_uncertainty = mean(uncertainty, na.rm = TRUE),
    median_uncertainty = median(uncertainty, na.rm = TRUE),
    n_chapters = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_uncertainty))
```

```{r uncertainty-plot}
mc_all$classified %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(book, uncertainty, FUN = median),
      y = uncertainty,
      fill = macro_genre
    )
  ) +
  ggplot2::geom_boxplot(outlier.alpha = 0.35) +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "Mclust uncertainty by book",
    subtitle = "Higher values indicate weaker assignment to a discrete cluster.",
    x = "Book",
    y = "Classification uncertainty",
    fill = "Macro genre"
  ) +
  ggplot2::theme_minimal()
```

# Part D: Alignment metrics

## 23. Adjusted Rand Index against macro genre and book

```{r ari}
ari_macro_genre <- mclust::adjustedRandIndex(
  mc_all$classified$macro_genre,
  mc_all$classified$mclust_cluster
)

ari_book <- mclust::adjustedRandIndex(
  mc_all$classified$book,
  mc_all$classified$mclust_cluster
)

ari_isaiah_section <- mclust::adjustedRandIndex(
  mc_all$classified %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::pull(section),
  mc_all$classified %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::pull(mclust_cluster)
)

tibble::tibble(
  comparison = c(
    "cluster_vs_macro_genre",
    "cluster_vs_book",
    "cluster_vs_isaiah_section"
  ),
  adjusted_rand_index = c(
    ari_macro_genre,
    ari_book,
    ari_isaiah_section
  )
)
```

## 24. Interpretive guide

```{r interpretive-guide}
interpretive_guide <- tibble::tribble(
  ~finding, ~interpretation,
  "G = 1 selected", "The all-chapter corpus does not support broad discrete stylistic populations.",
  "G = 2 or 3 selected and clusters align with macro genre", "The model is largely detecting genre structure.",
  "G > 1 selected and clusters align with books", "The model is detecting book-level style or topic/register effects.",
  "Isaiah chapters scatter across many clusters", "Isaiah is internally diverse in the broader biblical corpus.",
  "Isaiah chapters remain concentrated in one cluster", "Isaiah remains stylistically cohesive even in the broader corpus.",
  "Low ARI with Isaiah section", "All-chapter clusters do not reproduce traditional Isaiah divisions.",
  "High uncertainty", "The selected clusters overlap substantially."
)

interpretive_guide
```

## 25. Final summary

```{r final-summary}
cat("\nAll-Chapter Mclust Summary\n")
cat("==========================\n\n")

cat("Number of chapters analyzed:\n")
print(nrow(analysis_docs))

cat("\nBooks included:\n")
print(sort(unique(analysis_docs$book)))

cat("\nPCA variance:\n")
print(
  all_sim_pca$variance %>%
    dplyr::slice_head(n = 12),
  width = 120
)

cat("\nMclust sweep summary:\n")
print(
  mclust_sweep_all_summary,
  width = 120
)

cat("\nPreferred model:\n")
print(
  summarize_mclust_fit(mc_all),
  width = 120
)

cat("\nForced G results:\n")
print(
  forced_mclust_all,
  width = 120
)

cat("\nSilhouette results:\n")
print(
  silhouette_summary_all,
  width = 120
)

cat("\nARI metrics:\n")
print(
  tibble::tibble(
    comparison = c(
      "cluster_vs_macro_genre",
      "cluster_vs_book",
      "cluster_vs_isaiah_section"
    ),
    adjusted_rand_index = c(
      ari_macro_genre,
      ari_book,
      ari_isaiah_section
    )
  ),
  width = 120
)
```

## 26. Export outputs

```{r export}
out_dir <- file.path(STYLO_CACHE, "output", "mclust_all_nopos")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  analysis_meta,
  file.path(out_dir, "analysis_meta.csv")
)

readr::write_csv(
  all_sim_pca$variance,
  file.path(out_dir, "all_similarity_pca_variance.csv")
)

readr::write_csv(
  all_sim_pca$scores,
  file.path(out_dir, "all_similarity_pca_scores.csv")
)

readr::write_csv(
  mclust_sweep_all_summary,
  file.path(out_dir, "mclust_sweep_all_summary.csv")
)

readr::write_csv(
  mc_all$classified,
  file.path(out_dir, "mclust_all_classified.csv")
)

readr::write_csv(
  forced_mclust_all,
  file.path(out_dir, "forced_mclust_all.csv")
)

readr::write_csv(
  silhouette_summary_all,
  file.path(out_dir, "silhouette_summary_all.csv")
)

readr::write_csv(
  cluster_genre_table,
  file.path(out_dir, "cluster_genre_table.csv")
)

readr::write_csv(
  cluster_book_table,
  file.path(out_dir, "cluster_book_table.csv")
)

readr::write_csv(
  isaiah_cluster_table,
  file.path(out_dir, "isaiah_cluster_table.csv")
)
```
