---
title: "16 - Mclust All Chapters, G = 4 Review"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

Notebook 15 showed that the all-chapter corpus strongly supports a small number of broad clusters.  
The best model was approximately:

```text
G = 3, with G = 4 nearly competitive
```

The `G = 3` solution corresponded closely to:

```text
Narrative
Prophetic
Poetic/Wisdom
```

This notebook forces and reviews the **G = 4** solution to determine what the fourth cluster represents.

Key questions:

1. Does G = 4 split one of the macro genres?
2. Does the fourth cluster isolate Psalms, Wisdom, Ezekiel, Kings, or another book-level effect?
3. Does G = 4 reveal any meaningful Isaiah section structure?
4. Does the G = 4 solution strengthen or weaken the Deutero-Isaiah interpretation?

## 2. Parameters

```{r parameters}
include_feature_classes <- c(
  "function_words",
  "pos",
  "punctuation",
  "sentence",
  "word_length",
  "lexical",
  "readability"
)

pos_n <- c(3, 4, 5)

forced_G <- 4

preferred_n_pcs <- 6

comparison_books <- c(
  "Isaiah",
  "Psalms",
  "Job",
  "Proverbs",
  "Song of Solomon",
  "Ecclesiastes",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos",
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings"
)

set.seed(1234)
```

## 3. Packages

```{r packages}
if(!requireNamespace("mclust", quietly = TRUE)){
  install.packages("mclust")
}

if(!requireNamespace("cluster", quietly = TRUE)){
  install.packages("cluster")
}

library(mclust)
library(cluster)
```

## 4. Load prepared documents

```{r load-prepared}
chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
section_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_section_docs.csv")
raw_path     <- file.path(STYLO_ROOT, "data", "raw", "kjv.csv")

if(file.exists(chapter_path) && file.exists(section_path)){

  chapter_docs <- readr::read_csv(chapter_path, show_col_types = FALSE)
  section_docs <- readr::read_csv(section_path, show_col_types = FALSE)

} else {

  if(!file.exists(raw_path)){
    stop("Could not find data/raw/kjv.csv. Run 01_Introduction.Rmd first.")
  }

  kjv <- readr::read_csv(raw_path, show_col_types = FALSE)
  prepared <- prepare_kjv_docs(kjv)

  chapter_docs <- prepared$chapter_docs
  section_docs <- prepared$section_docs

  dir.create(file.path(STYLO_ROOT, "data", "prepared"), recursive = TRUE, showWarnings = FALSE)

  readr::write_csv(chapter_docs, chapter_path)
  readr::write_csv(section_docs, section_path)
}

if("books" %in% names(chapter_docs)){
  chapter_docs <- chapter_docs %>%
    dplyr::select(-books)
}

if("books" %in% names(section_docs)){
  section_docs <- section_docs %>%
    dplyr::rename(book = books)
}
```

## 5. Macro genre normalization

```{r macro-genre}
macro_genre_lookup <- tibble::tribble(
  ~book,              ~macro_genre_control,
  "1 Samuel",         "Narrative",
  "2 Samuel",         "Narrative",
  "1 Kings",          "Narrative",
  "2 Kings",          "Narrative",

  "Psalms",           "Poetic_Wisdom",
  "Job",              "Poetic_Wisdom",
  "Proverbs",         "Poetic_Wisdom",
  "Song of Solomon",  "Poetic_Wisdom",
  "Ecclesiastes",     "Poetic_Wisdom",

  "Isaiah",           "Prophetic",
  "Jeremiah",         "Prophetic",
  "Ezekiel",          "Prophetic",
  "Hosea",            "Prophetic",
  "Amos",             "Prophetic"
)

chapter_docs <- chapter_docs %>%
  dplyr::left_join(
    macro_genre_lookup,
    by = "book"
  ) %>%
  dplyr::mutate(
    macro_genre = dplyr::coalesce(
      macro_genre_control,
      macro_genre
    )
  ) %>%
  dplyr::select(-dplyr::any_of("macro_genre_control"))
```

## 6. Analysis corpus

```{r analysis-corpus}
analysis_docs <- chapter_docs %>%
  dplyr::filter(book %in% comparison_books) %>%
  dplyr::arrange(book, chapter)

analysis_meta <- analysis_docs %>%
  dplyr::select(
    doc_id,
    corpus_id,
    translation,
    book,
    chapter,
    section,
    macro_genre,
    n_verses
  ) %>%
  dplyr::mutate(
    is_isaiah = book == "Isaiah",
    isaiah_section = dplyr::if_else(
      book == "Isaiah",
      section,
      NA_character_
    )
  )

analysis_meta %>%
  dplyr::count(book, macro_genre)

nrow(analysis_docs)
```

## 7. Configure ensemble

```{r config}
config_all_mclust_4 <- default_config(
  include = include_feature_classes
)

config_all_mclust_4$pos_n <- pos_n

if(exists("ud_model")){
  config_all_mclust_4$pos_model <- ud_model
}

config_all_mclust_4$weights
```

# Part A: Rebuild all-chapter similarity PCA

## 8. Run all-chapter ensemble

```{r run-ensemble}
result_all_mclust_4 <- run_stylometry_ensemble(
  source_docs = analysis_docs,
  target_docs = analysis_docs,
  config = config_all_mclust_4,
  k_neighbors = 10,
  use_cache = TRUE
)

show_result_summary(result_all_mclust_4)
```

## 9. Build similarity matrix

```{r similarity-matrix}
S_all <- as.matrix(result_all_mclust_4$overall_similarity)

S_all[is.na(S_all)] <- 0
S_all[is.nan(S_all)] <- 0
S_all[is.infinite(S_all)] <- 0

S_all <- S_all[analysis_meta$doc_id, analysis_meta$doc_id, drop = FALSE]

dim(S_all)
```


```{r feature count}
count_features <- function(x){

  if(is.matrix(x) || is.data.frame(x)){

    return(ncol(x))

  }

  if(is.list(x)){

    return(

      sum(

        purrr::map_int(

          x,

          count_features

        )

      )

    )

  }

  0

}


purrr::map_int(

  result_all_mclust_4$feature_sets,

  count_features

)
```


## 10. PCA helper

```{r pca-helper}
processPCA <- function(
    X,
    metadata = NULL,
    center = TRUE,
    scale. = TRUE){

  X <- as.matrix(X)

  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0

  feature_var <- apply(X, 2, stats::var, na.rm = TRUE)
  keep <- is.finite(feature_var) & feature_var > 0

  X <- X[, keep, drop = FALSE]

  pca <- stats::prcomp(
    X,
    center = center,
    scale. = scale.
  )

  eig <- pca$sdev^2

  variance <- tibble::tibble(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  scores <- as.data.frame(pca$x) %>%
    tibble::rownames_to_column("doc_id")

  if(!is.null(metadata)){
    scores <- scores %>%
      dplyr::left_join(
        metadata,
        by = "doc_id"
      )
  }

  loadings <- as.data.frame(pca$rotation) %>%
    tibble::rownames_to_column("feature")

  list(
    X = X,
    pca = pca,
    scores = scores,
    loadings = loadings,
    variance = variance
  )
}
```

## 11. PCA on all-chapter similarity profiles

```{r all-similarity-pca}
all_sim_pca <- processPCA(
  S_all,
  metadata = analysis_meta,
  center = TRUE,
  scale. = TRUE
)

all_sim_pca$variance %>%
  dplyr::slice_head(n = 15)
```

## 12. Scree plot

```{r scree-plot}
all_sim_pca$variance %>%
  dplyr::slice_head(n = 20) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = dimension,
      y = pct_variance
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::geom_line(
    ggplot2::aes(y = cumulative_pct_variance),
    linewidth = 0.8
  ) +
  ggplot2::geom_point(
    ggplot2::aes(y = cumulative_pct_variance),
    size = 2
  ) +
  ggplot2::scale_x_continuous(breaks = 1:20) +
  ggplot2::labs(
    title = "All-chapter similarity PCA variance",
    subtitle = "Bars show individual PC variance; line shows cumulative variance.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  ggplot2::theme_minimal()
```

# Part B: Force G = 4

## 13. Mclust helper

```{r mclust-helper}
run_mclust_on_pcs <- function(
    pca_obj,
    metadata,
    n_pcs = 6,
    G = forced_G,
    modelNames = NULL){

  pc_cols <- paste0("PC", seq_len(n_pcs))

  X <- pca_obj$scores %>%
    dplyr::select(dplyr::all_of(pc_cols)) %>%
    as.matrix()

  rownames(X) <- pca_obj$scores$doc_id

  fit <- mclust::Mclust(
    X,
    G = G,
    modelNames = modelNames
  )

  classified <- tibble::tibble(
    doc_id = rownames(X),
    mclust_cluster = fit$classification,
    uncertainty = fit$uncertainty
  ) %>%
    dplyr::left_join(
      metadata,
      by = "doc_id"
    ) %>%
    dplyr::arrange(mclust_cluster, book, chapter)

  list(
    fit = fit,
    classified = classified,
    n_pcs = n_pcs,
    pc_cols = pc_cols
  )
}

summarize_mclust_fit <- function(mc_obj){

  fit <- mc_obj$fit

  tibble::tibble(
    n_pcs = mc_obj$n_pcs,
    selected_G = fit$G,
    selected_model = fit$modelName,
    bic = fit$bic,
    loglik = fit$loglik,
    df = fit$df,
    n = fit$n,
    mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
    max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
  )
}
```

## 14. Fit forced G = 4 model

```{r fit-g4}
preferred_n_pcs <- min(
  preferred_n_pcs,
  sum(grepl("^PC", names(all_sim_pca$scores)))
)

mc_all_g4 <- run_mclust_on_pcs(
  all_sim_pca,
  metadata = analysis_meta,
  n_pcs = preferred_n_pcs,
  G = forced_G
)

summary(mc_all_g4$fit)

summarize_mclust_fit(mc_all_g4)

mc_all_g4$classified
```

## 15. Classification counts

```{r classification-counts}
mc_all_g4$classified %>%
  dplyr::count(
    mclust_cluster,
    name = "n_chapters"
  ) %>%
  dplyr::arrange(mclust_cluster)
```

# Part C: What does the fourth cluster represent?

## 16. Cluster composition by macro genre

```{r cluster-genre-table}
cluster_genre_table_g4 <- mc_all_g4$classified %>%
  dplyr::count(
    mclust_cluster,
    macro_genre,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(
    pct_cluster = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_genre_table_g4
```

```{r cluster-genre-plot}
cluster_genre_table_g4 %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = factor(mclust_cluster),
      y = pct_cluster,
      fill = macro_genre
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Forced G = 4 clusters by macro genre",
    subtitle = "Each bar shows genre composition within a cluster.",
    x = "Mclust cluster",
    y = "Cluster share",
    fill = "Macro genre"
  ) +
  ggplot2::theme_minimal()
```

## 17. Cluster composition by book

```{r cluster-book-table}
cluster_book_table_g4 <- mc_all_g4$classified %>%
  dplyr::count(
    mclust_cluster,
    book,
    macro_genre,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(
    pct_cluster = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_book_table_g4
```

```{r cluster-book-plot}
cluster_book_table_g4 %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::slice_max(order_by = n_chapters, n = 10, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(book, n_chapters),
      y = n_chapters,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~ mclust_cluster, scales = "free_y") +
  ggplot2::labs(
    title = "Top books represented in each forced G = 4 cluster",
    subtitle = "Only the top ten books per cluster are shown.",
    x = "Book",
    y = "Chapter count"
  ) +
  ggplot2::theme_minimal()
```

## 18. Cluster labels by dominant genre/book

```{r cluster-labels}
cluster_labels_g4 <- cluster_genre_table_g4 %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::slice_max(order_by = pct_cluster, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    mclust_cluster,
    dominant_genre = macro_genre,
    dominant_genre_pct = pct_cluster
  ) %>%
  dplyr::left_join(
    cluster_book_table_g4 %>%
      dplyr::group_by(mclust_cluster) %>%
      dplyr::slice_max(order_by = pct_cluster, n = 1, with_ties = FALSE) %>%
      dplyr::ungroup() %>%
      dplyr::select(
        mclust_cluster,
        dominant_book = book,
        dominant_book_pct = pct_cluster
      ),
    by = "mclust_cluster"
  )

cluster_labels_g4
```

# Part D: Isaiah under G = 4

## 19. Isaiah cluster distribution

```{r isaiah-cluster-table}
isaiah_cluster_table_g4 <- mc_all_g4$classified %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::count(
    section,
    mclust_cluster,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(section) %>%
  dplyr::mutate(
    pct_section = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(section, mclust_cluster)

isaiah_cluster_table_g4
```

```{r isaiah-cluster-plot}
isaiah_cluster_table_g4 %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = pct_section,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Isaiah section distribution across forced G = 4 clusters",
    subtitle = "Tests whether the fourth cluster reproduces traditional Isaiah divisions.",
    x = "Isaiah section",
    y = "Section share",
    fill = "Mclust cluster"
  ) +
  ggplot2::theme_minimal()
```

## 20. Isaiah chapters by cluster

```{r isaiah-chapters-by-cluster}
isaiah_chapters_g4 <- mc_all_g4$classified %>%
  dplyr::filter(book == "Isaiah") %>%
  dplyr::select(
    doc_id,
    chapter,
    section,
    mclust_cluster,
    uncertainty,
    macro_genre
  ) %>%
  dplyr::arrange(mclust_cluster, chapter)

isaiah_chapters_g4
```

```{r isaiah-chapter-list}
isaiah_chapters_g4 %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::summarise(
    chapters = paste(chapter, collapse = ", "),
    sections = paste(unique(section), collapse = "; "),
    n_chapters = dplyr::n(),
    mean_uncertainty = mean(uncertainty, na.rm = TRUE),
    .groups = "drop"
  )
```

## 21. Isaiah uncertainty by section

```{r isaiah-uncertainty}
mc_all_g4$classified %>%
  dplyr::filter(book == "Isaiah") %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = chapter,
      y = uncertainty,
      color = section
    )
  ) +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_line(alpha = 0.45) +
  ggplot2::labs(
    title = "Isaiah classification uncertainty under forced G = 4",
    subtitle = "Higher uncertainty indicates weaker assignment to a discrete cluster.",
    x = "Isaiah chapter",
    y = "Classification uncertainty",
    color = "Isaiah section"
  ) +
  ggplot2::theme_minimal()
```

# Part E: Visual review in PCA space

## 22. PC1 vs PC2 colored by G = 4 cluster

```{r pc12-g4}
scores_g4 <- all_sim_pca$scores %>%
  dplyr::left_join(
    mc_all_g4$classified %>%
      dplyr::select(doc_id, mclust_cluster, uncertainty),
    by = "doc_id"
  ) %>%
  dplyr::mutate(
    mclust_cluster = factor(mclust_cluster),
    isaiah_label = dplyr::if_else(book == "Isaiah", as.character(chapter), "")
  )

ggplot2::ggplot(
  scores_g4,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = mclust_cluster,
    label = isaiah_label
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(size = 2.5, alpha = 0.75) +
  ggrepel::geom_text_repel(size = 2.6, max.overlaps = Inf, show.legend = FALSE) +
  ggplot2::labs(
    title = "All-chapter PCA colored by forced G = 4 cluster",
    subtitle = paste0(
      "Model = ",
      mc_all_g4$fit$modelName,
      "; PCs used = ",
      preferred_n_pcs,
      "; Isaiah chapters labeled"
    ),
    x = paste0("PC1 (", round(all_sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(all_sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "G = 4 cluster"
  ) +
  ggplot2::theme_minimal()
```

## 23. PC-pair grid for first six PCs

```{r pc-pair-grid, fig.width=11, fig.height=14}
pcs_to_plot <- 1:min(6, sum(grepl("^PC", names(scores_g4))))

pc_pairs <- expand.grid(
  x = pcs_to_plot,
  y = pcs_to_plot
) %>%
  dplyr::filter(x < y)

pc_pair_plots <- purrr::pmap(
  pc_pairs,
  function(x, y){

    x_col <- paste0("PC", x)
    y_col <- paste0("PC", y)

    ggplot2::ggplot(
      scores_g4,
      ggplot2::aes(
        x = .data[[x_col]],
        y = .data[[y_col]],
        color = mclust_cluster
      )
    ) +
      ggplot2::geom_hline(yintercept = 0, color = "grey80", linewidth = 0.2) +
      ggplot2::geom_vline(xintercept = 0, color = "grey80", linewidth = 0.2) +
      ggplot2::geom_point(alpha = 0.65, size = 1.6) +
      ggplot2::labs(
        title = paste0("PC", x, " vs PC", y),
        x = paste0(
          "PC",
          x,
          " (",
          round(all_sim_pca$variance$pct_variance[x], 1),
          "%)"
        ),
        y = paste0(
          "PC",
          y,
          " (",
          round(all_sim_pca$variance$pct_variance[y], 1),
          "%)"
        ),
        color = "Cluster"
      ) +
      ggplot2::theme_minimal() +
      ggplot2::theme(
        legend.position = "none",
        plot.title = ggplot2::element_text(size = 10)
      )
  }
)

if(requireNamespace("patchwork", quietly = TRUE)){
  patchwork::wrap_plots(pc_pair_plots, ncol = 3)
} else {
  for(p in pc_pair_plots){
    print(p)
  }
}
```

## 24. Prophetic-only view under G = 4

```{r prophetic-view}
scores_g4 %>%
  dplyr::filter(macro_genre == "Prophetic") %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = PC1,
      y = PC2,
      color = book,
      shape = mclust_cluster,
      label = dplyr::if_else(book == "Isaiah", as.character(chapter), "")
    )
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey60", linewidth = 0.3) +
  ggplot2::geom_point(size = 2.7, alpha = 0.75) +
  ggrepel::geom_text_repel(size = 2.6, max.overlaps = Inf, show.legend = FALSE) +
  ggplot2::labs(
    title = "Prophetic chapters under forced G = 4",
    subtitle = "Isaiah chapters are labeled; color shows book; shape shows cluster.",
    x = paste0("PC1 (", round(all_sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(all_sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Book",
    shape = "Cluster"
  ) +
  ggplot2::theme_minimal()
```

# Part F: Compare G = 3 and G = 4

## 25. Fit G = 3 for comparison

```{r fit-g3}
mc_all_g3 <- run_mclust_on_pcs(
  all_sim_pca,
  metadata = analysis_meta,
  n_pcs = preferred_n_pcs,
  G = 3
)

summarize_mclust_fit(mc_all_g3)
summarize_mclust_fit(mc_all_g4)
```

## 26. Crosswalk from G = 3 to G = 4

```{r g3-g4-crosswalk}
g3_g4_crosswalk <- mc_all_g3$classified %>%
  dplyr::select(
    doc_id,
    cluster_g3 = mclust_cluster
  ) %>%
  dplyr::left_join(
    mc_all_g4$classified %>%
      dplyr::select(
        doc_id,
        cluster_g4 = mclust_cluster
      ),
    by = "doc_id"
  ) %>%
  dplyr::count(
    cluster_g3,
    cluster_g4,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(cluster_g3) %>%
  dplyr::mutate(
    pct_g3 = n_chapters / sum(n_chapters)
  ) %>%
  dplyr::ungroup()

g3_g4_crosswalk
```

```{r g3-g4-crosswalk-plot}
g3_g4_crosswalk %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = factor(cluster_g3),
      y = pct_g3,
      fill = factor(cluster_g4)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Crosswalk from G = 3 clusters to G = 4 clusters",
    subtitle = "Shows which G = 3 cluster is split by the fourth component.",
    x = "G = 3 cluster",
    y = "Share of G = 3 cluster",
    fill = "G = 4 cluster"
  ) +
  ggplot2::theme_minimal()
```

## 27. BIC comparison

```{r bic-comparison}
bic_comparison <- dplyr::bind_rows(
  summarize_mclust_fit(mc_all_g3) %>%
    dplyr::mutate(forced_G = 3),
  summarize_mclust_fit(mc_all_g4) %>%
    dplyr::mutate(forced_G = 4)
) %>%
  dplyr::select(
    forced_G,
    n_pcs,
    selected_model,
    bic,
    loglik,
    df,
    mean_uncertainty,
    max_uncertainty
  )

bic_comparison
```

# Part G: Alignment metrics

## 28. ARI for G = 4

```{r ari-g4}
ari_g4_macro_genre <- mclust::adjustedRandIndex(
  mc_all_g4$classified$macro_genre,
  mc_all_g4$classified$mclust_cluster
)

ari_g4_book <- mclust::adjustedRandIndex(
  mc_all_g4$classified$book,
  mc_all_g4$classified$mclust_cluster
)

ari_g4_isaiah_section <- mclust::adjustedRandIndex(
  mc_all_g4$classified %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::pull(section),
  mc_all_g4$classified %>%
    dplyr::filter(book == "Isaiah") %>%
    dplyr::pull(mclust_cluster)
)

ari_g4 <- tibble::tibble(
  comparison = c(
    "G4_cluster_vs_macro_genre",
    "G4_cluster_vs_book",
    "G4_cluster_vs_isaiah_section"
  ),
  adjusted_rand_index = c(
    ari_g4_macro_genre,
    ari_g4_book,
    ari_g4_isaiah_section
  )
)

ari_g4
```

## 29. Compare ARI G = 3 vs G = 4

```{r ari-g3-g4}
ari_g3 <- tibble::tibble(
  comparison = c(
    "G3_cluster_vs_macro_genre",
    "G3_cluster_vs_book",
    "G3_cluster_vs_isaiah_section"
  ),
  adjusted_rand_index = c(
    mclust::adjustedRandIndex(
      mc_all_g3$classified$macro_genre,
      mc_all_g3$classified$mclust_cluster
    ),
    mclust::adjustedRandIndex(
      mc_all_g3$classified$book,
      mc_all_g3$classified$mclust_cluster
    ),
    mclust::adjustedRandIndex(
      mc_all_g3$classified %>%
        dplyr::filter(book == "Isaiah") %>%
        dplyr::pull(section),
      mc_all_g3$classified %>%
        dplyr::filter(book == "Isaiah") %>%
        dplyr::pull(mclust_cluster)
    )
  )
)

ari_g3_g4 <- dplyr::bind_rows(
  ari_g3 %>%
    dplyr::mutate(G = 3),
  ari_g4 %>%
    dplyr::mutate(G = 4)
) %>%
  dplyr::mutate(
    comparison_type = dplyr::case_when(
      grepl("macro_genre", comparison) ~ "Macro genre",
      grepl("book", comparison) ~ "Book",
      grepl("isaiah_section", comparison) ~ "Isaiah section",
      TRUE ~ comparison
    )
  )

ari_g3_g4
```

```{r ari-g3-g4-plot}
ari_g3_g4 %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = comparison_type,
      y = adjusted_rand_index,
      fill = factor(G)
    )
  ) +
  ggplot2::geom_col(position = "dodge") +
  ggplot2::labs(
    title = "ARI comparison: G = 3 vs G = 4",
    subtitle = "Does the fourth cluster improve alignment with book or Isaiah section?",
    x = "Comparison",
    y = "Adjusted Rand Index",
    fill = "G"
  ) +
  ggplot2::theme_minimal()
```

# Part H: Interpretive guide

## 30. Interpretation table

```{r interpretation-table}
interpretation_table <- tibble::tribble(
  ~finding, ~interpretation,
  "G = 4 splits the Poetic/Wisdom cluster", "The fourth component likely distinguishes Psalms from wisdom/love poetry, or separates Job/Proverbs from Psalms.",
  "G = 4 splits the Prophetic cluster", "The fourth component may reflect book-level prophetic style, such as Ezekiel vs Isaiah/Jeremiah, rather than Deutero-Isaiah.",
  "G = 4 splits the Narrative cluster", "The fourth component may distinguish Kings from Samuel or court narrative from battle narrative.",
  "G = 4 increases ARI with macro genre", "The fourth cluster still mainly refines genre structure.",
  "G = 4 increases ARI with book", "The fourth cluster captures a book-level effect.",
  "G = 4 does not increase ARI with Isaiah section", "The fourth cluster does not support traditional Isaiah partitioning.",
  "Isaiah 40-55 remains mostly in the same cluster as Isaiah 1-35", "The G = 4 solution does not support a sharp Deutero-Isaiah division.",
  "Isaiah 36-39 shifts toward narrative clusters", "This is expected and indicates a genre/narrative effect."
)

interpretation_table
```

## 31. Final summary

```{r final-summary}
cat("\nForced G = 4 Mclust Review\n")
cat("==========================\n\n")

cat("G = 4 model summary:\n")
print(
  summarize_mclust_fit(mc_all_g4),
  width = 120
)

cat("\nCluster labels:\n")
print(
  cluster_labels_g4,
  width = 120
)

cat("\nCluster by genre:\n")
print(
  cluster_genre_table_g4,
  width = 120
)

cat("\nIsaiah section distribution:\n")
print(
  isaiah_cluster_table_g4,
  width = 120
)

cat("\nIsaiah chapters by cluster:\n")
print(
  isaiah_chapters_g4 %>%
    dplyr::group_by(mclust_cluster) %>%
    dplyr::summarise(
      chapters = paste(chapter, collapse = ", "),
      sections = paste(unique(section), collapse = "; "),
      n_chapters = dplyr::n(),
      mean_uncertainty = mean(uncertainty, na.rm = TRUE),
      .groups = "drop"
    ),
  width = 120
)

cat("\nG = 3 vs G = 4 BIC comparison:\n")
print(
  bic_comparison,
  width = 120
)

cat("\nARI G = 3 vs G = 4:\n")
print(
  ari_g3_g4,
  width = 120
)
```

## 32. Export outputs

```{r export}
out_dir <- file.path(STYLO_CACHE, "output", "mclust_all_g4")

dir.create(
  out_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

readr::write_csv(
  all_sim_pca$variance,
  file.path(out_dir, "all_similarity_pca_variance.csv")
)

readr::write_csv(
  all_sim_pca$scores,
  file.path(out_dir, "all_similarity_pca_scores.csv")
)

readr::write_csv(
  mc_all_g4$classified,
  file.path(out_dir, "mclust_all_g4_classified.csv")
)

readr::write_csv(
  cluster_genre_table_g4,
  file.path(out_dir, "cluster_genre_table_g4.csv")
)

readr::write_csv(
  cluster_book_table_g4,
  file.path(out_dir, "cluster_book_table_g4.csv")
)

readr::write_csv(
  cluster_labels_g4,
  file.path(out_dir, "cluster_labels_g4.csv")
)

readr::write_csv(
  isaiah_cluster_table_g4,
  file.path(out_dir, "isaiah_cluster_table_g4.csv")
)

readr::write_csv(
  isaiah_chapters_g4,
  file.path(out_dir, "isaiah_chapters_g4.csv")
)

readr::write_csv(
  g3_g4_crosswalk,
  file.path(out_dir, "g3_g4_crosswalk.csv")
)

readr::write_csv(
  bic_comparison,
  file.path(out_dir, "bic_comparison_g3_g4.csv")
)

readr::write_csv(
  ari_g3_g4,
  file.path(out_dir, "ari_g3_g4.csv")
)
```
