---
title: "30 - Prepare Book of Mormon TSV Editions"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook downloads and parses the BYU ODH OpenScripture Book of Mormon TSV files.

The source files are token-level TSVs, with one row per word/punctuation token and one column per Book of Mormon edition:

```text
Citation    wID    1830    1837    1840    1841    1879    1920    1981    2013
1 Nephi 1:1 1      I       I       I       I       I       I       I       I
1 Nephi 1:1 1.01   ,       ,       ,       ,       ,       ,       ,       ,
1 Nephi 1:1 1.02   ⌴       ⌴       ⌴       ⌴       ⌴       ⌴       ⌴       ⌴
1 Nephi 1:1 2      Nephi   Nephi   Nephi   Nephi   Nephi   Nephi   Nephi   Nephi
```

The notebook collapses these token rows into verse-level CSV files:

```text
data/raw/bom_1830.csv
data/raw/bom_1837.csv
data/raw/bom_1840.csv
data/raw/bom_1841.csv
data/raw/bom_1879.csv
data/raw/bom_1920.csv
data/raw/bom_1981.csv
data/raw/bom_2013.csv
```

Each output file has normalized columns:

```text
corpus_id
translation
book
chapter
verse
text
```

These outputs can then be consumed by `31_Introduction_BOM.Rmd`.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
```

## 3. Parameters

```{r parameters}
openscripture_base_url <- "https://raw.githubusercontent.com/BYU-ODH/OpenScripture/refs/heads/main/book-of-mormon"

bom_tsv_dir <- file.path(
  STYLO_ROOT,
  "data",
  "raw",
  "bom_tsv"
)

raw_dir <- file.path(
  STYLO_ROOT,
  "data",
  "raw"
)

dir.create(
  bom_tsv_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  raw_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

edition_cols <- c(
  "1830",
  "1837",
  "1840",
  "1841",
  "1879",
  "1920",
  "1981",
  "2013"
)

download_tsvs <- TRUE
overwrite_tsvs <- FALSE
```

## 4. Canonical Book of Mormon order

```{r bom-book-order}
bom_book_order <- tibble::tribble(
  ~book,               ~book_order, ~file_name,
  "1 Nephi",            1L,          "1 Nephi.tsv",
  "2 Nephi",            2L,          "2 Nephi.tsv",
  "Jacob",              3L,          "Jacob.tsv",
  "Enos",               4L,          "Enos.tsv",
  "Jarom",              5L,          "Jarom.tsv",
  "Omni",               6L,          "Omni.tsv",
  "Words of Mormon",    7L,          "Words of Mormon.tsv",
  "Mosiah",             8L,          "Mosiah.tsv",
  "Alma",               9L,          "Alma.tsv",
  "Helaman",           10L,          "Helaman.tsv",
  "3 Nephi",           11L,          "3 Nephi.tsv",
  "4 Nephi",           12L,          "4 Nephi.tsv",
  "Mormon",            13L,          "Mormon.tsv",
  "Ether",             14L,          "Ether.tsv",
  "Moroni",            15L,          "Moroni.tsv"
)

bom_book_order
```

## 5. Download TSV files

```{r download-tsv-files}
bom_tsv_sources <- bom_book_order %>%
  dplyr::mutate(
    url = paste0(
      openscripture_base_url,
      "/",
      utils::URLencode(file_name, reserved = TRUE)
    ),
    local_path = file.path(
      bom_tsv_dir,
      file_name
    )
  )

bom_tsv_sources

if(isTRUE(download_tsvs)){

  purrr::pwalk(
    bom_tsv_sources %>%
      dplyr::select(url, local_path),
    function(url, local_path){

      if(!file.exists(local_path) || isTRUE(overwrite_tsvs)){

        message("Downloading: ", url)

        download.file(
          url = url,
          destfile = local_path,
          mode = "wb",
          quiet = FALSE
        )

      } else {

        message("Already exists: ", local_path)

      }

    }
  )

}

missing_tsvs <- bom_tsv_sources %>%
  dplyr::filter(!file.exists(local_path))

missing_tsvs

if(nrow(missing_tsvs) > 0){
  stop(
    "Some TSV files are missing. Check missing_tsvs."
  )
}
```

## 6. Helper functions

```{r helper-functions}
normalize_bom_book <- function(x){

  x <- as.character(x)
  x <- stringr::str_squish(x)
  x <- stringr::str_replace_all(x, "_", " ")

  y <- stringr::str_to_lower(x)

  dplyr::case_when(
    stringr::str_detect(y, "1\\s*nephi|first\\s+nephi") ~ "1 Nephi",
    stringr::str_detect(y, "2\\s*nephi|second\\s+nephi") ~ "2 Nephi",
    stringr::str_detect(y, "3\\s*nephi|third\\s+nephi") ~ "3 Nephi",
    stringr::str_detect(y, "4\\s*nephi|fourth\\s+nephi") ~ "4 Nephi",
    stringr::str_detect(y, "words\\s+of\\s+mormon") ~ "Words of Mormon",
    stringr::str_detect(y, "mosiah") ~ "Mosiah",
    stringr::str_detect(y, "alma") ~ "Alma",
    stringr::str_detect(y, "helaman") ~ "Helaman",
    stringr::str_detect(y, "mormon") ~ "Mormon",
    stringr::str_detect(y, "ether") ~ "Ether",
    stringr::str_detect(y, "moroni") ~ "Moroni",
    stringr::str_detect(y, "jacob") ~ "Jacob",
    stringr::str_detect(y, "enos") ~ "Enos",
    stringr::str_detect(y, "jarom") ~ "Jarom",
    stringr::str_detect(y, "omni") ~ "Omni",
    TRUE ~ stringr::str_to_title(x)
  )
}


clean_bom_token <- function(x){

  x <- as.character(x)

  x <- dplyr::na_if(x, "")
  x <- dplyr::na_if(x, "NA")
  x <- dplyr::na_if(x, "⌴")
  x <- dplyr::na_if(x, "∅")

  x
}


collapse_tokens_to_text <- function(tokens){

  text <- paste(tokens, collapse = " ")

  text <- stringr::str_replace_all(text, "\\s+([,.;:!?])", "\\1")
  text <- stringr::str_replace_all(text, "\\s+([\\)\\]\\}])", "\\1")
  text <- stringr::str_replace_all(text, "([\\(\\[\\{])\\s+", "\\1")
  text <- stringr::str_replace_all(text, "\\s+'", "'")
  text <- stringr::str_replace_all(text, "\\s+", " ")
  text <- stringr::str_squish(text)

  text
}


parse_bom_word_tsv <- function(path, expected_book = NULL){

  x <- readr::read_tsv(
    path,
    show_col_types = FALSE,
    col_types = readr::cols(.default = readr::col_character())
  )

  required <- c("Citation", "wID")

  missing_required <- setdiff(
    required,
    names(x)
  )

  if(length(missing_required) > 0){
    stop(
      "TSV file is missing required columns: ",
      paste(missing_required, collapse = ", "),
      "\nFile: ",
      path
    )
  }

  editions_found <- intersect(
    edition_cols,
    names(x)
  )

  if(length(editions_found) == 0){
    stop(
      "No edition columns found in TSV file: ",
      path
    )
  }

  x_long <- x %>%
    tidyr::pivot_longer(
      cols = dplyr::all_of(editions_found),
      names_to = "translation",
      values_to = "token"
    ) %>%
    dplyr::mutate(
      token = clean_bom_token(token)
    ) %>%
    dplyr::filter(!is.na(token)) %>%
    tidyr::extract(
      Citation,
      into = c("book", "chapter", "verse"),
      regex = "^(.*)\\s+(\\d+):(\\d+)$",
      remove = FALSE,
      convert = TRUE
    ) %>%
    dplyr::mutate(
      book = normalize_bom_book(book),
      chapter = as.integer(chapter),
      verse = as.integer(verse),
      wID_num = suppressWarnings(as.numeric(wID))
    )

  if(!is.null(expected_book)){

    unexpected_books <- x_long %>%
      dplyr::filter(book != expected_book) %>%
      dplyr::distinct(book)

    if(nrow(unexpected_books) > 0){
      warning(
        "Unexpected book labels in ",
        path,
        ": ",
        paste(unexpected_books$book, collapse = ", ")
      )
    }

  }

  x_long %>%
    dplyr::arrange(
      translation,
      book,
      chapter,
      verse,
      wID_num,
      wID
    ) %>%
    dplyr::group_by(
      translation,
      book,
      chapter,
      verse
    ) %>%
    dplyr::summarise(
      text = collapse_tokens_to_text(token),
      n_tokens = dplyr::n(),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      corpus_id = "BOM"
    ) %>%
    dplyr::select(
      corpus_id,
      translation,
      book,
      chapter,
      verse,
      text,
      n_tokens
    )
}
```

## 7. Parse all TSV files

```{r parse-all-tsv}
bom_all_versions <- purrr::map2_dfr(
  bom_tsv_sources$local_path,
  bom_tsv_sources$book,
  parse_bom_word_tsv
) %>%
  dplyr::left_join(
    bom_book_order %>%
      dplyr::select(book, book_order),
    by = "book"
  ) %>%
  dplyr::arrange(
    translation,
    book_order,
    chapter,
    verse
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    book,
    chapter,
    verse,
    text
  )

bom_all_versions %>%
  dplyr::slice_head(n = 30)

bom_all_versions %>%
  dplyr::count(translation, name = "n_verses")
```

## 8. Sanity checks

```{r sanity-checks}
expected_chapters <- tibble::tribble(
  ~book,              ~expected_chapters,
  "1 Nephi",           22L,
  "2 Nephi",           33L,
  "Jacob",              7L,
  "Enos",               1L,
  "Jarom",              1L,
  "Omni",               1L,
  "Words of Mormon",    1L,
  "Mosiah",            29L,
  "Alma",              63L,
  "Helaman",           16L,
  "3 Nephi",           30L,
  "4 Nephi",            1L,
  "Mormon",             9L,
  "Ether",             15L,
  "Moroni",            10L
)

chapter_check <- bom_all_versions %>%
  dplyr::distinct(translation, book, chapter) %>%
  dplyr::count(translation, book, name = "observed_chapters") %>%
  dplyr::left_join(
    expected_chapters,
    by = "book"
  ) %>%
  dplyr::mutate(
    matches_expected = observed_chapters == expected_chapters
  ) %>%
  dplyr::left_join(
    bom_book_order %>%
      dplyr::select(book, book_order),
    by = "book"
  ) %>%
  dplyr::arrange(
    translation,
    book_order
  )

chapter_check

if(any(!chapter_check$matches_expected, na.rm = TRUE)){
  warning(
    "Some BOM edition/book chapter counts do not match expected LDS chapter counts. Inspect chapter_check."
  )
}

empty_verses <- bom_all_versions %>%
  dplyr::filter(is.na(text) | text == "")

empty_verses

verse_counts <- bom_all_versions %>%
  dplyr::count(
    translation,
    book,
    chapter,
    name = "n_verses"
  ) %>%
  dplyr::left_join(
    bom_book_order %>%
      dplyr::select(book, book_order),
    by = "book"
  ) %>%
  dplyr::arrange(
    translation,
    book_order,
    chapter
  )

verse_counts %>%
  dplyr::slice_head(n = 40)
```

## 9. Preview selected verses

```{r preview-verses}
bom_all_versions %>%
  dplyr::filter(
    book == "1 Nephi",
    chapter == 1,
    verse == 1
  ) %>%
  dplyr::arrange(translation)

bom_all_versions %>%
  dplyr::filter(
    book == "Moroni",
    chapter == 10,
    verse == 34
  ) %>%
  dplyr::arrange(translation)
```

## 10. Write one CSV per edition

```{r write-edition-csvs}
bom_all_versions %>%
  dplyr::group_split(translation) %>%
  purrr::walk(function(df){

    ed <- unique(df$translation)

    out_path <- file.path(
      raw_dir,
      paste0("bom_", ed, ".csv")
    )

    readr::write_csv(
      df %>%
        dplyr::arrange(
          book,
          chapter,
          verse
        ) %>%
        dplyr::select(
          corpus_id,
          translation,
          book,
          chapter,
          verse,
          text
        ),
      out_path
    )

    message("Wrote: ", out_path)

  })
```

## 11. Write combined edition CSV

```{r write-combined}
combined_path <- file.path(
  raw_dir,
  "bom_all_editions.csv"
)

readr::write_csv(
  bom_all_versions,
  combined_path
)

cat("Wrote combined BOM edition file to:\n")
cat(combined_path, "\n")
```

## 12. Create default bom.csv alias

By default, downstream notebooks such as `31_Introduction_BOM.Rmd` look for:

```text
data/raw/bom.csv
```

This chunk writes the 2013 edition to `bom.csv` as the default modern edition.

Change `default_bom_edition` if you want another default.

```{r write-default-bom}
default_bom_edition <- "1830"

default_bom <- bom_all_versions %>%
  dplyr::filter(translation == default_bom_edition) %>%
  dplyr::arrange(
    book,
    chapter,
    verse
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    book,
    chapter,
    verse,
    text
  )

readr::write_csv(
  default_bom,
  file.path(raw_dir, "bom.csv")
)

cat("Wrote default BOM edition to:\n")
cat(file.path(raw_dir, "bom.csv"), "\n")
```

## 13. Output summary

```{r output-summary}
bom_summary_by_edition <- bom_all_versions %>%
  dplyr::group_by(
    translation
  ) %>%
  dplyr::summarise(
    n_books = dplyr::n_distinct(book),
    n_chapters = dplyr::n_distinct(paste(book, chapter)),
    n_verses = dplyr::n(),
    total_words = sum(stringr::str_count(text, "\\S+")),
    .groups = "drop"
  ) %>%
  dplyr::arrange(translation)

bom_summary_by_edition

bom_summary_by_book_edition <- bom_all_versions %>%
  dplyr::group_by(
    translation,
    book
  ) %>%
  dplyr::summarise(
    n_chapters = dplyr::n_distinct(chapter),
    n_verses = dplyr::n(),
    total_words = sum(stringr::str_count(text, "\\S+")),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    bom_book_order %>%
      dplyr::select(book, book_order),
    by = "book"
  ) %>%
  dplyr::arrange(
    translation,
    book_order
  )

bom_summary_by_book_edition
```

## 14. Next notebook

After this notebook succeeds, run:

```text
31_Introduction_BOM.Rmd
```

That notebook consumes:

```text
data/raw/bom.csv
```

or one of the edition-specific files:

```text
data/raw/bom_1830.csv
data/raw/bom_2013.csv
```

and produces:

```text
data/prepared/bom_chapter_docs.csv
data/prepared/bom_book_docs.csv
data/prepared/bom_mosiah_priority_docs.csv
data/prepared/bom_target_collections.csv
```
