---
title: "30 - Prepare Book of Mormon from Project Gutenberg"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook prepares a raw Book of Mormon CSV from a Project Gutenberg plain-text file.

It creates:

```text
data/raw/bom.csv
```

with columns:

```text
book
chapter
verse
text
```

This file is then consumed by:

```text
31_Introduction_BOM.Rmd
```

which builds chapter documents, book collections, and Mosiah Priority collections.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tibble)
library(purrr)
```

## 3. Parameters

```{r parameters}
bom_gutenberg_url <- "https://www.gutenberg.org/cache/epub/17/pg17.txt"

raw_dir <- file.path(
  STYLO_ROOT,
  "data",
  "raw"
)

dir.create(
  raw_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

bom_txt_path <- file.path(
  raw_dir,
  "bom_gutenberg_pg17.txt"
)

bom_csv_path <- file.path(
  raw_dir,
  "bom.csv"
)
```

## 4. Download Project Gutenberg text

```{r download-gutenberg}
if(!file.exists(bom_txt_path)){

  download.file(
    url = bom_gutenberg_url,
    destfile = bom_txt_path,
    mode = "wb",
    quiet = FALSE
  )

}

bom_lines_raw <- readr::read_lines(
  bom_txt_path,
  locale = readr::locale(encoding = "UTF-8")
)

length(bom_lines_raw)

bom_lines_raw[1:40]
```

## 5. Canonical Book of Mormon order

```{r bom-book-order}
bom_book_order <- tibble::tribble(
  ~book,               ~book_order,
  "1 Nephi",            1L,
  "2 Nephi",            2L,
  "Jacob",              3L,
  "Enos",               4L,
  "Jarom",              5L,
  "Omni",               6L,
  "Words of Mormon",    7L,
  "Mosiah",             8L,
  "Alma",               9L,
  "Helaman",           10L,
  "3 Nephi",           11L,
  "4 Nephi",           12L,
  "Mormon",            13L,
  "Ether",             14L,
  "Moroni",            15L
)

bom_book_order
```

## 6. Gutenberg boundary detection

Project Gutenberg files contain header and footer material that should not be analyzed.

This chunk removes the Gutenberg wrapper and keeps only the Book of Mormon body.

```{r strip-gutenberg-wrapper}
start_candidates <- which(
  stringr::str_detect(
    bom_lines_raw,
    regex("THE FIRST BOOK OF NEPHI|FIRST BOOK OF NEPHI|1 Nephi", ignore_case = TRUE)
  )
)

end_candidates <- which(
  stringr::str_detect(
    bom_lines_raw,
    regex("END OF THE PROJECT GUTENBERG|End of Project Gutenberg", ignore_case = TRUE)
  )
)

if(length(start_candidates) == 0){
  stop("Could not find the beginning of 1 Nephi in the Gutenberg text.")
}

body_start <- start_candidates[1]

body_end <- if(length(end_candidates) > 0){
  end_candidates[1] - 1L
} else {
  length(bom_lines_raw)
}

bom_lines <- bom_lines_raw[body_start:body_end]

bom_lines <- bom_lines %>%
  stringr::str_replace_all("\uFEFF", "") %>%
  stringr::str_replace_all("\\r", "") %>%
  stringr::str_squish()

bom_lines <- bom_lines[bom_lines != ""]

length(bom_lines)

bom_lines[1:80]
```

## 7. Book and chapter recognizers

```{r recognizers}
canonical_book_from_line <- function(line){

  x <- stringr::str_squish(line)
  y <- stringr::str_to_upper(x)

  dplyr::case_when(
    stringr::str_detect(y, "^THE FIRST BOOK OF NEPHI|^FIRST BOOK OF NEPHI|^1 NEPHI$") ~ "1 Nephi",
    stringr::str_detect(y, "^THE SECOND BOOK OF NEPHI|^SECOND BOOK OF NEPHI|^2 NEPHI$") ~ "2 Nephi",
    stringr::str_detect(y, "^THE BOOK OF JACOB|^JACOB$") ~ "Jacob",
    stringr::str_detect(y, "^THE BOOK OF ENOS|^ENOS$") ~ "Enos",
    stringr::str_detect(y, "^THE BOOK OF JAROM|^JAROM$") ~ "Jarom",
    stringr::str_detect(y, "^THE BOOK OF OMNI|^OMNI$") ~ "Omni",
    stringr::str_detect(y, "^WORDS OF MORMON|^THE WORDS OF MORMON") ~ "Words of Mormon",
    stringr::str_detect(y, "^THE BOOK OF MOSIAH|^MOSIAH$") ~ "Mosiah",
    stringr::str_detect(y, "^THE BOOK OF ALMA|^ALMA$") ~ "Alma",
    stringr::str_detect(y, "^THE BOOK OF HELAMAN|^HELAMAN$") ~ "Helaman",
    stringr::str_detect(y, "^THIRD NEPHI|^THE THIRD BOOK OF NEPHI|^3 NEPHI$") ~ "3 Nephi",
    stringr::str_detect(y, "^FOURTH NEPHI|^THE FOURTH BOOK OF NEPHI|^4 NEPHI$") ~ "4 Nephi",
    stringr::str_detect(y, "^THE BOOK OF MORMON$|^MORMON$") ~ "Mormon",
    stringr::str_detect(y, "^THE BOOK OF ETHER|^ETHER$") ~ "Ether",
    stringr::str_detect(y, "^THE BOOK OF MORONI|^MORONI$") ~ "Moroni",
    TRUE ~ NA_character_
  )
}


is_chapter_line <- function(line){

  stringr::str_detect(
    line,
    regex(
      paste(
        "^CHAPTER\\s+[0-9]+\\b",
        "^[1-4]?\\s*Nephi\\s+Chapter\\s+[0-9]+\\b",
        "^[A-Za-z ]+\\s+Chapter\\s+[0-9]+\\b",
        sep = "|"
      ),
      ignore_case = TRUE
    )
  )

}


chapter_from_line <- function(line){

  m <- stringr::str_match(
    line,
    regex("CHAPTER\\s+([0-9]+)", ignore_case = TRUE)
  )

  if(!is.na(m[1, 2])){
    return(as.integer(m[1, 2]))
  }

  NA_integer_
}
```

## 8. Parse into book/chapter blocks

```{r parse-blocks}
parsed_lines <- tibble::tibble(
  line_no = seq_along(bom_lines),
  line = bom_lines
) %>%
  dplyr::mutate(
    book_detected = purrr::map_chr(
      line,
      canonical_book_from_line
    ),
    chapter_detected = dplyr::if_else(
      is_chapter_line(line),
      purrr::map_int(line, chapter_from_line),
      NA_integer_
    )
  )

parsed_lines %>%
  dplyr::filter(!is.na(book_detected) | !is.na(chapter_detected)) %>%
  dplyr::slice_head(n = 80)
```

## 9. Main parser

This parser walks through the text line by line.

It detects:

- book headers,
- chapter headers,
- verse starts.

Project Gutenberg editions may format verses as either:

```text
1 I, Nephi, having been born...
```

or:

```text
1. I, Nephi, having been born...
```

The parser handles both.

```{r main-parser}
verse_start_regex <- "^([0-9]+)[\\.|\\)]?\\s+(.*)$"

current_book <- NA_character_
current_chapter <- NA_integer_
current_verse <- NA_integer_
current_text <- character(0)

records <- list()

flush_verse <- function(){

  if(!is.na(current_book) &&
     !is.na(current_chapter) &&
     !is.na(current_verse) &&
     length(current_text) > 0){

    tibble::tibble(
      book = current_book,
      chapter = current_chapter,
      verse = current_verse,
      text = stringr::str_squish(
        paste(current_text, collapse = " ")
      )
    )

  } else {

    NULL

  }

}

for(i in seq_along(bom_lines)){

  line <- bom_lines[[i]]

  detected_book <- canonical_book_from_line(line)

  if(!is.na(detected_book)){

    flushed <- flush_verse()

    if(!is.null(flushed)){
      records[[length(records) + 1L]] <- flushed
    }

    current_book <- detected_book
    current_chapter <- NA_integer_
    current_verse <- NA_integer_
    current_text <- character(0)

    next

  }

  if(is_chapter_line(line)){

    flushed <- flush_verse()

    if(!is.null(flushed)){
      records[[length(records) + 1L]] <- flushed
    }

    current_chapter <- chapter_from_line(line)
    current_verse <- NA_integer_
    current_text <- character(0)

    next

  }

  if(is.na(current_book) || is.na(current_chapter)){
    next
  }

  verse_match <- stringr::str_match(line, verse_start_regex)

  if(!is.na(verse_match[1, 2])){

    flushed <- flush_verse()

    if(!is.null(flushed)){
      records[[length(records) + 1L]] <- flushed
    }

    current_verse <- as.integer(verse_match[1, 2])
    current_text <- verse_match[1, 3]

  } else {

    if(!is.na(current_verse)){
      current_text <- c(current_text, line)
    }

  }

}

flushed <- flush_verse()

if(!is.null(flushed)){
  records[[length(records) + 1L]] <- flushed
}

bom <- dplyr::bind_rows(records)

bom <- bom %>%
  dplyr::left_join(
    bom_book_order,
    by = "book"
  ) %>%
  dplyr::arrange(
    book_order,
    chapter,
    verse
  ) %>%
  dplyr::select(
    book,
    chapter,
    verse,
    text
  )

bom %>%
  dplyr::slice_head(n = 30)

nrow(bom)
```

## 10. Sanity checks

```{r sanity-checks}
expected_chapters <- tibble::tribble(
  ~book,              ~expected_chapters,
  "1 Nephi",           22L,
  "2 Nephi",           33L,
  "Jacob",              7L,
  "Enos",               1L,
  "Jarom",              1L,
  "Omni",               1L,
  "Words of Mormon",    1L,
  "Mosiah",            29L,
  "Alma",              63L,
  "Helaman",           16L,
  "3 Nephi",           30L,
  "4 Nephi",            1L,
  "Mormon",             9L,
  "Ether",             15L,
  "Moroni",            10L
)

chapter_check <- bom %>%
  dplyr::distinct(book, chapter) %>%
  dplyr::count(book, name = "observed_chapters") %>%
  dplyr::left_join(
    expected_chapters,
    by = "book"
  ) %>%
  dplyr::left_join(
    bom_book_order,
    by = "book"
  ) %>%
  dplyr::mutate(
    matches_expected = observed_chapters == expected_chapters
  ) %>%
  dplyr::arrange(book_order)

chapter_check

verse_summary <- bom %>%
  dplyr::group_by(book, chapter) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    min_verse = min(verse),
    max_verse = max(verse),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    bom_book_order,
    by = "book"
  ) %>%
  dplyr::arrange(book_order, chapter)

verse_summary %>%
  dplyr::slice_head(n = 40)

if(any(!chapter_check$matches_expected, na.rm = TRUE)){
  warning("Some books do not match expected BOM chapter counts. Inspect chapter_check.")
}

if(any(is.na(bom$text)) || any(bom$text == "")){
  warning("Some verses have missing or empty text.")
}
```

## 11. Preview parsed verses

```{r preview}
bom %>%
  dplyr::filter(book == "1 Nephi", chapter == 1) %>%
  dplyr::slice_head(n = 10)

bom %>%
  dplyr::filter(book == "Moroni", chapter == 10) %>%
  dplyr::slice_tail(n = 10)
```

## 12. Write raw CSV

```{r write-csv}
readr::write_csv(
  bom,
  bom_csv_path
)

cat("Wrote BOM raw CSV to:\\n")
cat(bom_csv_path, "\\n")

readr::read_csv(
  bom_csv_path,
  show_col_types = FALSE
) %>%
  dplyr::slice_head(n = 10)
```

## 13. Output summary

```{r output-summary}
bom_summary <- bom %>%
  dplyr::group_by(book) %>%
  dplyr::summarise(
    n_chapters = dplyr::n_distinct(chapter),
    n_verses = dplyr::n(),
    total_words = sum(stringr::str_count(text, "\\\\S+")),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    bom_book_order,
    by = "book"
  ) %>%
  dplyr::arrange(book_order)

bom_summary
```

## 14. Next notebook

After this notebook succeeds, run:

```text
31_Introduction_BOM.Rmd
```

That notebook consumes:

```text
data/raw/bom.csv
```

and produces:

```text
data/prepared/bom_chapter_docs.csv
data/prepared/bom_book_docs.csv
data/prepared/bom_mosiah_priority_docs.csv
data/prepared/bom_target_collections.csv
```
