---
title: "31 - Introduction to Book of Mormon Stylometry"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook prepares Book of Mormon documents in a structure parallel to the KJV preparation notebooks.

It creates:

1. verse-level BOM data,
2. chapter-level BOM documents,
3. book-level BOM documents,
4. Mosiah Priority section documents,
5. combined target collections for later stylometry.

The Mosiah Priority grouping used here is:

```text
MP-1 = Mosiah, Alma, Helaman
MP-2 = 3 Nephi through Moroni
MP-3 = 1 Nephi through Words of Mormon
```

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
```

## 3. Parameters

```{r parameters}
BOM_TRANSLATION <- "1830"

bom_raw_candidates <- c(
  file.path(STYLO_ROOT, "data", "raw", "bom.csv"),
  file.path(STYLO_ROOT, "data", "raw", "book_of_mormon.csv"),
  file.path(STYLO_ROOT, "data", "raw", "Book_of_Mormon.csv")
)

bom_prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(
  bom_prepared_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

bom_verse_path   <- file.path(bom_prepared_dir, "bom_verse_docs.csv")
bom_chapter_path <- file.path(bom_prepared_dir, "bom_chapter_docs.csv")
bom_book_path    <- file.path(bom_prepared_dir, "bom_book_docs.csv")
bom_mp_path      <- file.path(bom_prepared_dir, "bom_mosiah_priority_docs.csv")
bom_targets_path <- file.path(bom_prepared_dir, "bom_target_collections.csv")
```

## 4. Canonical Book of Mormon order

```{r bom-book-order}
bom_book_order <- tibble::tribble(
  ~book,               ~book_order,
  "1 Nephi",            1L,
  "2 Nephi",            2L,
  "Jacob",              3L,
  "Enos",               4L,
  "Jarom",              5L,
  "Omni",               6L,
  "Words of Mormon",    7L,
  "Mosiah",             8L,
  "Alma",               9L,
  "Helaman",           10L,
  "3 Nephi",           11L,
  "4 Nephi",           12L,
  "Mormon",            13L,
  "Ether",             14L,
  "Moroni",            15L
)

bom_book_order
```

## 5. Mosiah Priority mapping

```{r mosiah-priority-map}
mosiah_priority_map <- bom_book_order %>%
  dplyr::mutate(
    mosiah_priority = dplyr::case_when(
      book %in% c("Mosiah", "Alma", "Helaman") ~ "MP-1",
      book %in% c("3 Nephi", "4 Nephi", "Mormon", "Ether", "Moroni") ~ "MP-2",
      book %in% c("1 Nephi", "2 Nephi", "Jacob", "Enos", "Jarom", "Omni", "Words of Mormon") ~ "MP-3",
      TRUE ~ NA_character_
    ),
    mosiah_priority_label = dplyr::case_when(
      mosiah_priority == "MP-1" ~ "MP-1_Mosiah_Alma_Helaman",
      mosiah_priority == "MP-2" ~ "MP-2_3Nephi_to_Moroni",
      mosiah_priority == "MP-3" ~ "MP-3_1Nephi_to_WordsOfMormon",
      TRUE ~ NA_character_
    )
  )

mosiah_priority_map
```

## 6. Helper functions

```{r helper-functions}


make_doc_id <- function(book, chapter = NULL, prefix = NULL){

  clean_book <- book %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_|_$", "")

  if(is.null(chapter)){
    id <- clean_book
  } else {
    id <- paste0(clean_book, "_", chapter)
  }

  if(!is.null(prefix)){
    id <- paste0(prefix, "_", id)
  }

  id
}


collapse_text <- function(x){

  x %>%
    as.character() %>%
    stringr::str_squish() %>%
    paste(collapse = " ") %>%
    stringr::str_squish()
}
```

## 7. Read raw BOM data

The raw CSV should contain these conceptual columns:

```text
book
chapter
verse
text
```

Common alternatives such as `book_name`, `chapter_number`, `verse_number`, and `verse_text` are normalized.

```{r read-raw-bom}
bom_raw_path <- bom_raw_candidates[file.exists(bom_raw_candidates)][1]

if(is.na(bom_raw_path)){
  stop(
    "Could not find a BOM raw CSV. Expected one of: ",
    paste(bom_raw_candidates, collapse = ", ")
  )
}

bom_raw <- readr::read_csv(
  bom_raw_path,
  show_col_types = FALSE
)

names(bom_raw)
```

## 8. Normalize raw columns

```{r normalize-columns}
bom <- bom_raw

names(bom) <- names(bom) %>%
  stringr::str_replace_all("\\s+", "_") %>%
  stringr::str_to_lower()

if(!"book" %in% names(bom) && "book_title" %in% names(bom)){
  bom <- bom %>% dplyr::rename(book = book_title)
}

if(!"book" %in% names(bom) && "book_name" %in% names(bom)){
  bom <- bom %>% dplyr::rename(book = book_name)
}

if(!"chapter" %in% names(bom) && "chapter_number" %in% names(bom)){
  bom <- bom %>% dplyr::rename(chapter = chapter_number)
}

if(!"verse" %in% names(bom) && "verse_number" %in% names(bom)){
  bom <- bom %>% dplyr::rename(verse = verse_number)
}

if(!"text" %in% names(bom) && "verse_text" %in% names(bom)){
  bom <- bom %>% dplyr::rename(text = verse_text)
}

if(!"text" %in% names(bom) && "scripture_text" %in% names(bom)){
  bom <- bom %>% dplyr::rename(text = scripture_text)
}

required_cols <- c("book", "chapter", "verse", "text")

missing_cols <- setdiff(required_cols, names(bom))

if(length(missing_cols) > 0){
  stop(
    "BOM raw data is missing required columns: ",
    paste(missing_cols, collapse = ", ")
  )
}

bom <- bom %>%
  dplyr::mutate(
    corpus_id = "BOM",
    translation = BOM_TRANSLATION,
    book = stringr::str_squish(book),
    chapter = as.integer(chapter),
    verse = as.integer(verse),
    text = as.character(text)
  ) %>%
  dplyr::left_join(
    mosiah_priority_map,
    by = "book"
  ) %>%
  dplyr::arrange(
    book_order,
    chapter,
    verse
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    book,
    book_order,
    chapter,
    verse,
    mosiah_priority,
    mosiah_priority_label,
    text
  )

bom %>%
  dplyr::count(book, book_order, mosiah_priority) %>%
  dplyr::arrange(book_order)
```

## 9. Verse-level documents

```{r verse-docs}
bom_verse_docs <- bom %>%
  dplyr::mutate(
    doc_id = paste0(make_doc_id(book, chapter), "_", verse),
    section = mosiah_priority,
    collection_type = "verse",
    n_verses = 1L,
    n_chapters = 1L
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    book,
    book_order,
    chapter,
    verse,
    section,
    mosiah_priority,
    mosiah_priority_label,
    collection_type,
    n_verses,
    n_chapters,
    text
  )

bom_verse_docs %>%
  dplyr::slice_head(n = 20)
```

## 10. Chapter-level documents

```{r chapter-docs}
bom_chapter_docs <- bom %>%
  dplyr::group_by(
    corpus_id,
    translation,
    book,
    book_order,
    chapter,
    mosiah_priority,
    mosiah_priority_label
  ) %>%
  dplyr::summarise(
    text = collapse_text(text),
    n_verses = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    doc_id = make_doc_id(book, chapter),
    section = mosiah_priority,
    collection_type = "chapter",
    n_chapters = 1L
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    book,
    book_order,
    chapter,
    section,
    mosiah_priority,
    mosiah_priority_label,
    collection_type,
    n_verses,
    n_chapters,
    text
  ) %>%
  dplyr::arrange(book_order, chapter)

bom_chapter_docs %>%
  dplyr::count(book, book_order, mosiah_priority, name = "n_chapters") %>%
  dplyr::arrange(book_order)
```

## 11. Book-level documents

These are analogous to using Psalms, Job, Proverbs, Jeremiah, and other books as target collections in the KJV analysis.

```{r book-docs}
bom_book_docs <- bom_chapter_docs %>%
  dplyr::group_by(
    corpus_id,
    translation,
    book,
    book_order,
    mosiah_priority,
    mosiah_priority_label
  ) %>%
  dplyr::summarise(
    text = collapse_text(text),
    n_verses = sum(n_verses),
    n_chapters = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    doc_id = paste0("BOOK_", make_doc_id(book)),
    chapter = NA_integer_,
    section = mosiah_priority,
    collection_type = "book"
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    book,
    book_order,
    chapter,
    section,
    mosiah_priority,
    mosiah_priority_label,
    collection_type,
    n_verses,
    n_chapters,
    text
  ) %>%
  dplyr::arrange(book_order)

bom_book_docs
```

## 12. Mosiah Priority documents

```{r mosiah-priority-docs}
bom_mp_docs <- bom_chapter_docs %>%
  dplyr::group_by(
    corpus_id,
    translation,
    mosiah_priority,
    mosiah_priority_label
  ) %>%
  dplyr::summarise(
    text = collapse_text(text),
    n_verses = sum(n_verses),
    n_chapters = dplyr::n(),
    books = paste(unique(book), collapse = "; "),
    min_book_order = min(book_order),
    max_book_order = max(book_order),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    doc_id = mosiah_priority_label,
    book = books,
    book_order = min_book_order,
    chapter = NA_integer_,
    section = mosiah_priority,
    collection_type = "mosiah_priority"
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    doc_id,
    book,
    book_order,
    chapter,
    section,
    mosiah_priority,
    mosiah_priority_label,
    collection_type,
    n_verses,
    n_chapters,
    books,
    text
  ) %>%
  dplyr::arrange(book_order)

bom_mp_docs
```

## 13. Combined target collections

```{r target-collections}
bom_target_collections <- dplyr::bind_rows(
  bom_book_docs,
  bom_mp_docs %>%
    dplyr::select(-dplyr::any_of("books"))
) %>%
  dplyr::mutate(
    target_family = dplyr::case_when(
      collection_type == "book" ~ "BOOK_COLLECTION",
      collection_type == "mosiah_priority" ~ "MOSIAH_PRIORITY",
      TRUE ~ stringr::str_to_upper(collection_type)
    )
  )

bom_target_collections %>%
  dplyr::select(
    doc_id,
    collection_type,
    target_family,
    book,
    section,
    n_chapters,
    n_verses
  )
```

## 14. Basic summaries

```{r summaries}
bom_summary_by_book <- bom_chapter_docs %>%
  dplyr::group_by(
    book,
    book_order,
    mosiah_priority
  ) %>%
  dplyr::summarise(
    n_chapters = dplyr::n(),
    n_verses = sum(n_verses),
    total_chars = sum(nchar(text)),
    total_words = sum(stringr::str_count(text, "\\S+")),
    mean_words_per_chapter = mean(stringr::str_count(text, "\\S+")),
    .groups = "drop"
  ) %>%
  dplyr::arrange(book_order)

bom_summary_by_book

bom_summary_by_mp <- bom_chapter_docs %>%
  dplyr::group_by(
    mosiah_priority,
    mosiah_priority_label
  ) %>%
  dplyr::summarise(
    books = paste(unique(book), collapse = "; "),
    n_books = dplyr::n_distinct(book),
    n_chapters = dplyr::n(),
    n_verses = sum(n_verses),
    total_chars = sum(nchar(text)),
    total_words = sum(stringr::str_count(text, "\\S+")),
    mean_words_per_chapter = mean(stringr::str_count(text, "\\S+")),
    .groups = "drop"
  ) %>%
  dplyr::arrange(mosiah_priority)

bom_summary_by_mp
```

## 15. Sanity checks

```{r sanity-checks}
expected_chapters <- tibble::tribble(
  ~book,              ~expected_chapters,
  "1 Nephi",           22L,
  "2 Nephi",           33L,
  "Jacob",              7L,
  "Enos",               1L,
  "Jarom",              1L,
  "Omni",               1L,
  "Words of Mormon",    1L,
  "Mosiah",            29L,
  "Alma",              63L,
  "Helaman",           16L,
  "3 Nephi",           30L,
  "4 Nephi",            1L,
  "Mormon",             9L,
  "Ether",             15L,
  "Moroni",            10L
)

chapter_check <- bom_chapter_docs %>%
  dplyr::count(book, name = "observed_chapters") %>%
  dplyr::left_join(
    expected_chapters,
    by = "book"
  ) %>%
  dplyr::mutate(
    matches_expected = observed_chapters == expected_chapters
  ) %>%
  dplyr::left_join(
    bom_book_order,
    by = "book"
  ) %>%
  dplyr::arrange(book_order)

chapter_check

if(any(!chapter_check$matches_expected, na.rm = TRUE)){
  warning("Some BOM books do not match expected LDS chapter counts. Inspect chapter_check.")
}

missing_mp <- bom_chapter_docs %>%
  dplyr::filter(is.na(mosiah_priority)) %>%
  dplyr::distinct(book)

missing_mp
```

## 16. Write prepared files

```{r write-prepared}
readr::write_csv(bom_verse_docs, bom_verse_path)
readr::write_csv(bom_chapter_docs, bom_chapter_path)
readr::write_csv(bom_book_docs, bom_book_path)
readr::write_csv(bom_mp_docs, bom_mp_path)
readr::write_csv(bom_target_collections, bom_targets_path)

readr::write_csv(
  bom_summary_by_book,
  file.path(bom_prepared_dir, "bom_summary_by_book.csv")
)

readr::write_csv(
  bom_summary_by_mp,
  file.path(bom_prepared_dir, "bom_summary_by_mosiah_priority.csv")
)

cat("Prepared BOM files written to:\n")
cat(bom_prepared_dir, "\n")
```

## 17. Objects created

```{r objects-created}
objects_created <- tibble::tribble(
  ~object, ~description,
  "bom", "Normalized verse-level Book of Mormon data.",
  "bom_verse_docs", "One row per verse.",
  "bom_chapter_docs", "One row per chapter.",
  "bom_book_docs", "One row per BOM book.",
  "bom_mp_docs", "One row per Mosiah Priority section.",
  "bom_target_collections", "Book and Mosiah Priority collections for ensemble target comparisons.",
  "bom_summary_by_book", "Book-level counts and word totals.",
  "bom_summary_by_mp", "Mosiah Priority group summaries."
)

objects_created
```
