---
title: "31 - Mclust All Chapters with Book of Mormon"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook extends the previous KJV all-chapter mclust experiment by adding Book of Mormon chapters.

The combined corpus includes:

```text
KJV chapters    ≈ 500+
BOM chapters    ≈ 200+
Total           ≈ 700+
```

The analysis:

1. loads prepared KJV chapter documents,
2. loads prepared BOM chapter documents,
3. treats Isaiah as ordinary chapter documents only,
4. builds an all-document similarity matrix,
5. performs similarity-profile PCA,
6. runs model-based clustering for `G = 1:9`,
7. evaluates cluster alignment with corpus identity, macro genre, book, Mosiah Priority, and Isaiah section diagnostics.

## 2. Packages

```{r packages}
if(!requireNamespace("mclust", quietly = TRUE)){
  install.packages("mclust")
}

library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
library(ggplot2)
library(mclust)

if(requireNamespace("ggrepel", quietly = TRUE)){
  library(ggrepel)
}
```

## 3. Parameters

```{r parameters}
set.seed(1234)

kjv_books_to_include <- c(
  "Isaiah",
  "Psalms",
  "Job",
  "Proverbs",
  "Song of Solomon",
  "Ecclesiastes",
  "Jeremiah",
  "Ezekiel",
  "Hosea",
  "Amos",
  "1 Samuel",
  "2 Samuel",
  "1 Kings",
  "2 Kings"
)

# NULL means use whatever edition is in data/prepared/bom_chapter_docs.csv.
# Set to "1830" or "2013" if the prepared file contains multiple BOM editions.
bom_translation_to_include <- NULL

include_feature_classes <- c(
  "function_words",
  "punctuation",
  "sentence",
  "word_length",
  "lexical",
  "readability"
)

use_pos_features <- FALSE

if(isTRUE(use_pos_features)){
  include_feature_classes <- c(include_feature_classes, "pos")
}

pos_n <- c(3, 4, 5)

k_neighbors <- 10
preferred_n_pcs <- 6
G_values <- 1:9
mclust_model_names <- NULL
```

## 4. Load KJV prepared chapters

```{r load-kjv}
kjv_chapter_path <- file.path(
  STYLO_ROOT,
  "data",
  "prepared",
  "kjv_chapter_docs.csv"
)

if(!file.exists(kjv_chapter_path)){
  stop(
    "Could not find prepared KJV chapter docs: ",
    kjv_chapter_path,
    "\nRun the KJV preparation notebook first."
  )
}

kjv_chapter_docs_raw <- readr::read_csv(
  kjv_chapter_path,
  show_col_types = FALSE
)

names(kjv_chapter_docs_raw)
```

## 5. Normalize KJV chapters

Isaiah is not collapsed into sections here. Each Isaiah chapter remains an ordinary document.

```{r normalize-kjv}
kjv_macro_genre_lookup <- tibble::tribble(
  ~book,              ~macro_genre_control,
  "1 Samuel",         "Narrative",
  "2 Samuel",         "Narrative",
  "1 Kings",          "Narrative",
  "2 Kings",          "Narrative",

  "Psalms",           "Poetic_Wisdom",
  "Job",              "Poetic_Wisdom",
  "Proverbs",         "Poetic_Wisdom",
  "Song of Solomon",  "Poetic_Wisdom",
  "Ecclesiastes",     "Poetic_Wisdom",

  "Isaiah",           "Prophetic",
  "Jeremiah",         "Prophetic",
  "Ezekiel",          "Prophetic",
  "Hosea",            "Prophetic",
  "Amos",             "Prophetic"
)

kjv_chapter_docs <- kjv_chapter_docs_raw %>%
  dplyr::filter(book %in% kjv_books_to_include) %>%
  dplyr::left_join(kjv_macro_genre_lookup, by = "book") %>%
  dplyr::mutate(
    corpus_id = dplyr::coalesce(corpus_id, "KJV"),
    translation = dplyr::coalesce(translation, "KJV"),
    corpus_family = "KJV",
    macro_genre = dplyr::coalesce(macro_genre_control, macro_genre),
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    source_family = "KJV",
    chapter = as.integer(chapter),
    doc_id_original = doc_id,
    doc_id = paste0("KJV__", doc_id)
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    n_verses,
    text
  )

kjv_chapter_docs %>%
  dplyr::count(book, macro_genre, name = "n_chapters") %>%
  dplyr::arrange(macro_genre, book)
```

## 6. Load BOM prepared chapters

This expects that you have run:

```text
30_Prep_BOM_TSV.Rmd
31_Introduction_BOM.Rmd
```

```{r load-bom}
bom_chapter_path <- file.path(
  STYLO_ROOT,
  "data",
  "prepared",
  "bom_chapter_docs.csv"
)

if(!file.exists(bom_chapter_path)){
  stop(
    "Could not find prepared BOM chapter docs: ",
    bom_chapter_path,
    "\nRun 30_Prep_BOM_TSV.Rmd and 31_Introduction_BOM.Rmd first."
  )
}

bom_chapter_docs_raw <- readr::read_csv(
  bom_chapter_path,
  show_col_types = FALSE
)

names(bom_chapter_docs_raw)
```

## 7. Normalize BOM chapters

```{r normalize-bom}
bom_chapter_docs <- bom_chapter_docs_raw

if(!is.null(bom_translation_to_include)){
  bom_chapter_docs <- bom_chapter_docs %>%
    dplyr::filter(translation == bom_translation_to_include)
}

bom_chapter_docs <- bom_chapter_docs %>%
  dplyr::mutate(
    translation = as.character(translation),

    corpus_id = "BOM",
    corpus_family = "BOM",

    source_family = paste0("BOM_", translation),

    macro_genre = "BOM",

    section = dplyr::coalesce(
      section,
      mosiah_priority
    ),

    doc_id_original = doc_id,

    doc_id = paste0(
      "BOM_",
      translation,
      "__",
      doc_id
    ),

    chapter = as.integer(chapter)
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    n_verses,
    text
  )

bom_chapter_docs %>%
  dplyr::count(translation, book, mosiah_priority, name = "n_chapters") %>%
  dplyr::arrange(translation, mosiah_priority, book)
```

## 8. Combined analysis corpus

```{r combined-corpus}
analysis_docs <- dplyr::bind_rows(
  kjv_chapter_docs,
  bom_chapter_docs
) %>%
  dplyr::mutate(
    text = as.character(text),
    chapter = as.integer(chapter),
    n_verses = as.integer(n_verses)
  ) %>%
  dplyr::filter(!is.na(text), text != "") %>%
  dplyr::arrange(corpus_family, source_family, book, chapter)

analysis_meta <- analysis_docs %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    n_verses
  ) %>%
  dplyr::mutate(
    is_kjv = corpus_family == "KJV",
    is_bom = corpus_family == "BOM",
    is_isaiah = corpus_family == "KJV" & book == "Isaiah",
    is_bom_mp1 = corpus_family == "BOM" & mosiah_priority == "MP-1",
    is_bom_mp2 = corpus_family == "BOM" & mosiah_priority == "MP-2",
    is_bom_mp3 = corpus_family == "BOM" & mosiah_priority == "MP-3"
  )

analysis_docs %>%
  dplyr::count(corpus_family, source_family, name = "n_docs")

analysis_docs %>%
  dplyr::count(corpus_family, macro_genre, name = "n_docs")

analysis_docs %>%
  dplyr::count(corpus_family, book, name = "n_docs") %>%
  dplyr::arrange(corpus_family, book)
```

## 9. Configure stylometry ensemble

```{r config}
config_all_bom <- default_config(
  include = include_feature_classes
)

config_all_bom$pos_n <- pos_n

if(!isTRUE(use_pos_features)){

  config_all_bom$weights <- c(
    function_words = 0.500,
    punctuation    = 0.075,
    sentence       = 0.175,
    word_length    = 0.125,
    lexical        = 0.075,
    readability    = 0.050
  )

} else {

  config_all_bom$weights <- c(
    function_words = 0.300,
    pos            = 0.300,
    punctuation    = 0.050,
    sentence       = 0.150,
    word_length    = 0.100,
    lexical        = 0.050,
    readability    = 0.050
  )

}

config_all_bom$use_tfidf_function_words <- TRUE

if(exists("ud_model")){
  config_all_bom$pos_model <- ud_model
}

config_all_bom$include
config_all_bom$weights
```

## 10. Feature-count helpers

```{r feature-count-helpers}
count_features <- function(x){

  if(is.matrix(x) || is.data.frame(x)){
    return(ncol(x))
  }

  if(is.list(x)){
    return(sum(purrr::map_int(x, count_features)))
  }

  0L
}


show_feature_counts <- function(feature_sets){

  out <- tibble::tibble(
    feature_class = names(feature_sets),
    n_features = purrr::map_int(feature_sets, count_features)
  ) %>%
    dplyr::mutate(
      pct_features_raw = n_features / sum(n_features),
      pct_features = scales::percent(pct_features_raw, accuracy = 0.1)
    ) %>%
    dplyr::arrange(dplyr::desc(n_features))

  total_row <- tibble::tibble(
    feature_class = "TOTAL",
    n_features = sum(out$n_features),
    pct_features_raw = 1,
    pct_features = "100.0%"
  )

  dplyr::bind_rows(out, total_row)
}


show_feature_counts_recursive <- function(x, prefix = ""){

  if(is.matrix(x) || is.data.frame(x)){
    return(tibble::tibble(feature = prefix, n_features = ncol(x)))
  }

  if(is.list(x)){
    return(
      purrr::imap_dfr(
        x,
        ~show_feature_counts_recursive(
          .x,
          if(prefix == "") .y else paste0(prefix, "::", .y)
        )
      )
    )
  }

  tibble::tibble(feature = prefix, n_features = 0L)
}
```

## 11. Run ensemble and build similarity matrix

```{r run-ensemble}
result_all_bom <- run_stylometry_ensemble(
  source_docs = analysis_docs,
  target_docs = analysis_docs,
  config = config_all_bom,
  k_neighbors = k_neighbors,
  use_cache = TRUE
)

result_all_bom$weights_used

feature_counts <- show_feature_counts(result_all_bom$feature_sets)
feature_counts

feature_counts_recursive <- show_feature_counts_recursive(result_all_bom$feature_sets)
feature_counts_recursive
```

```{r similarity-matrix}
S_all <- as.matrix(result_all_bom$overall_similarity)

S_all[is.na(S_all)] <- 0
S_all[is.nan(S_all)] <- 0
S_all[is.infinite(S_all)] <- 0

S_all <- S_all[
  analysis_meta$doc_id,
  analysis_meta$doc_id,
  drop = FALSE
]

dim(S_all)

S_all[1:5, 1:5]
```

## 12. PCA helper

```{r pca-helper}
processPCA <- function(
    X,
    metadata = NULL,
    center = TRUE,
    scale. = TRUE){

  X <- as.matrix(X)

  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0

  feature_var <- apply(X, 2, stats::var, na.rm = TRUE)
  keep <- is.finite(feature_var) & feature_var > 0

  X <- X[, keep, drop = FALSE]

  pca <- stats::prcomp(
    X,
    center = center,
    scale. = scale.
  )

  eig <- pca$sdev^2

  variance <- tibble::tibble(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  scores <- as.data.frame(pca$x) %>%
    tibble::rownames_to_column("doc_id")

  if(!is.null(metadata)){
    scores <- scores %>%
      dplyr::left_join(metadata, by = "doc_id")
  }

  loadings <- as.data.frame(pca$rotation) %>%
    tibble::rownames_to_column("feature")

  list(
    X = X,
    pca = pca,
    scores = scores,
    loadings = loadings,
    variance = variance
  )
}
```

## 13. Similarity-profile PCA

```{r pca}
sim_pca <- processPCA(
  S_all,
  metadata = analysis_meta,
  center = TRUE,
  scale. = TRUE
)

sim_pca$variance %>%
  dplyr::slice_head(n = 20)
```

```{r scree-plot}
sim_pca$variance %>%
  dplyr::slice_head(n = 20) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = dimension, y = pct_variance)
  ) +
  ggplot2::geom_col() +
  ggplot2::geom_line(
    ggplot2::aes(y = cumulative_pct_variance),
    linewidth = 0.8
  ) +
  ggplot2::geom_point(
    ggplot2::aes(y = cumulative_pct_variance),
    size = 2
  ) +
  ggplot2::scale_x_continuous(breaks = 1:20) +
  ggplot2::labs(
    title = "KJV + BOM similarity-profile PCA variance",
    subtitle = "Bars show individual PC variance; line shows cumulative variance.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  ggplot2::theme_minimal()
```

## 14. Mclust helpers

```{r mclust-helpers}
run_mclust_on_pcs <- function(
    pca_obj,
    metadata,
    n_pcs = 6,
    G = G_values,
    modelNames = NULL){

  pc_cols <- paste0("PC", seq_len(n_pcs))

  X <- pca_obj$scores %>%
    dplyr::select(dplyr::all_of(pc_cols)) %>%
    as.matrix()

  rownames(X) <- pca_obj$scores$doc_id

  fit <- mclust::Mclust(
    X,
    G = G,
    modelNames = modelNames
  )

  classified <- tibble::tibble(
    doc_id = rownames(X),
    mclust_cluster = fit$classification,
    uncertainty = fit$uncertainty
  ) %>%
    dplyr::left_join(metadata, by = "doc_id") %>%
    dplyr::arrange(mclust_cluster, corpus_family, book, chapter)

  list(
    fit = fit,
    classified = classified,
    n_pcs = n_pcs,
    pc_cols = pc_cols
  )
}


summarize_mclust_fit <- function(mc_obj){

  fit <- mc_obj$fit

  tibble::tibble(
    n_pcs = mc_obj$n_pcs,
    selected_G = fit$G,
    selected_model = fit$modelName,
    bic = fit$bic,
    loglik = fit$loglik,
    df = fit$df,
    n = fit$n,
    mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
    max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
  )
}


safe_ari <- function(a, b){

  keep <- !is.na(a) & !is.na(b)

  if(sum(keep) < 2){
    return(NA_real_)
  }

  mclust::adjustedRandIndex(a[keep], b[keep])
}
```

## 15. Fit G = 1:9 over increasing PC counts

```{r fit-over-pcs}
max_pcs_available <- sum(grepl("^PC", names(sim_pca$scores)))

n_pcs_grid <- seq_len(min(10, max_pcs_available))

mclust_by_pcs <- purrr::map(
  n_pcs_grid,
  ~run_mclust_on_pcs(
    sim_pca,
    metadata = analysis_meta,
    n_pcs = .x,
    G = G_values,
    modelNames = mclust_model_names
  )
)

mclust_fit_summary <- purrr::map_dfr(
  mclust_by_pcs,
  summarize_mclust_fit
)

mclust_fit_summary
```

```{r selected-g-plot}
mclust_fit_summary %>%
  ggplot2::ggplot(
    ggplot2::aes(x = n_pcs, y = selected_G)
  ) +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2.5) +
  ggplot2::scale_x_continuous(breaks = n_pcs_grid) +
  ggplot2::scale_y_continuous(breaks = G_values) +
  ggplot2::labs(
    title = "Selected G by number of PCs",
    subtitle = "KJV + BOM corpus; mclust tested G = 1:6.",
    x = "Number of PCs used",
    y = "Selected G"
  ) +
  ggplot2::theme_minimal()
```

## 16. Preferred model

```{r preferred-model}
preferred_n_pcs <- min(preferred_n_pcs, max_pcs_available)

mc_all_bom <- run_mclust_on_pcs(
  sim_pca,
  metadata = analysis_meta,
  n_pcs = preferred_n_pcs,
  G = G_values,
  modelNames = mclust_model_names
)

summary(mc_all_bom$fit)

summarize_mclust_fit(mc_all_bom)

classified <- mc_all_bom$classified

classified
```

## 17. BIC values for preferred model

```{r bic-preferred}
bic_matrix <- mc_all_bom$fit$BIC

bic_df <- as.data.frame(as.table(bic_matrix)) %>%
  dplyr::rename(
    model = Var1,
    G = Var2,
    BIC = Freq
  ) %>%
  dplyr::mutate(
    G = as.integer(as.character(G))
  ) %>%
  dplyr::filter(!is.na(BIC)) %>%
  dplyr::arrange(G, dplyr::desc(BIC))

bic_df %>%
  dplyr::group_by(G) %>%
  dplyr::slice_max(order_by = BIC, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup()
```

```{r bic-plot}
bic_df %>%
  dplyr::group_by(G) %>%
  dplyr::slice_max(order_by = BIC, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  ggplot2::ggplot(
    ggplot2::aes(x = G, y = BIC)
  ) +
  ggplot2::geom_line() +
  ggplot2::geom_point(size = 2.5) +
  ggplot2::scale_x_continuous(breaks = G_values) +
  ggplot2::labs(
    title = paste0("Best BIC by G, using first ", preferred_n_pcs, " PCs"),
    x = "G",
    y = "BIC"
  ) +
  ggplot2::theme_minimal()
```

## 18. Cluster composition tables

```{r cluster-corpus}
cluster_corpus_table <- classified %>%
  dplyr::count(mclust_cluster, corpus_family, name = "n_chapters") %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_corpus_table
```

```{r cluster-corpus-plot}
cluster_corpus_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = factor(mclust_cluster),
      y = pct_cluster,
      fill = corpus_family
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Cluster composition by corpus",
    x = "Mclust cluster",
    y = "Cluster share",
    fill = "Corpus"
  ) +
  ggplot2::theme_minimal()
```

```{r cluster-genre}
cluster_genre_table <- classified %>%
  dplyr::count(mclust_cluster, macro_genre, name = "n_chapters") %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_genre_table
```

```{r cluster-genre-plot}
cluster_genre_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = factor(mclust_cluster),
      y = pct_cluster,
      fill = macro_genre
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Cluster composition by macro genre",
    x = "Mclust cluster",
    y = "Cluster share",
    fill = "Macro genre"
  ) +
  ggplot2::theme_minimal()
```

```{r cluster-book}
cluster_book_table <- classified %>%
  dplyr::count(
    mclust_cluster,
    corpus_family,
    book,
    macro_genre,
    mosiah_priority,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mclust_cluster, dplyr::desc(n_chapters))

cluster_book_table
```

```{r cluster-book-plot, fig.height=10}
cluster_book_table %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::slice_max(order_by = n_chapters, n = 12, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(book_label = paste0(corpus_family, ": ", book)) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(book_label, n_chapters),
      y = n_chapters,
      fill = corpus_family
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~ mclust_cluster, scales = "free_y") +
  ggplot2::labs(
    title = "Top books represented in each cluster",
    subtitle = "Top 12 books per cluster.",
    x = "Book",
    y = "Chapter count",
    fill = "Corpus"
  ) +
  ggplot2::theme_minimal()
```

## 19. BOM Mosiah Priority distribution

```{r bom-mp-table}
bom_mp_cluster_table <- classified %>%
  dplyr::filter(corpus_family == "BOM") %>%
  dplyr::count(
    mosiah_priority,
    mosiah_priority_label,
    mclust_cluster,
    name = "n_chapters"
  ) %>%
  dplyr::group_by(mosiah_priority, mosiah_priority_label) %>%
  dplyr::mutate(pct_mp = n_chapters / sum(n_chapters)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(mosiah_priority, mclust_cluster)

bom_mp_cluster_table
```

```{r bom-mp-plot}
bom_mp_cluster_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = mosiah_priority,
      y = pct_mp,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "BOM Mosiah Priority groups across clusters",
    x = "Mosiah Priority group",
    y = "Group share",
    fill = "Mclust cluster"
  ) +
  ggplot2::theme_minimal()
```

## 20. Isaiah distribution across clusters

```{r isaiah-cluster-table}
isaiah_cluster_table <- classified %>%
  dplyr::filter(is_isaiah) %>%
  dplyr::count(section, mclust_cluster, name = "n_chapters") %>%
  dplyr::group_by(section) %>%
  dplyr::mutate(pct_section = n_chapters / sum(n_chapters)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(section, mclust_cluster)

isaiah_cluster_table
```

```{r isaiah-cluster-plot}
isaiah_cluster_table %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = section,
      y = pct_section,
      fill = factor(mclust_cluster)
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::percent_format()) +
  ggplot2::labs(
    title = "Isaiah chapters across KJV+BOM clusters",
    subtitle = "Isaiah is treated as ordinary chapter documents; sections are diagnostic only.",
    x = "Isaiah section",
    y = "Section share",
    fill = "Mclust cluster"
  ) +
  ggplot2::theme_minimal()
```

## 21. PCA plots

```{r pc12-cluster-corpus}
scores_clustered <- sim_pca$scores %>%
  dplyr::left_join(
    classified %>%
      dplyr::select(doc_id, mclust_cluster, uncertainty),
    by = "doc_id"
  ) %>%
  dplyr::mutate(
    cluster = factor(mclust_cluster),
    label = dplyr::case_when(
      is_isaiah ~ paste0("Isa ", chapter),
      corpus_family == "BOM" & book %in% c("Mosiah", "Alma", "Helaman") ~ paste0(book, " ", chapter),
      TRUE ~ ""
    )
  )

ggplot2::ggplot(
  scores_clustered,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = cluster,
    shape = corpus_family,
    label = label
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_point(alpha = 0.75, size = 2.4) +
  {
    if(requireNamespace("ggrepel", quietly = TRUE)){
      ggrepel::geom_text_repel(size = 2.3, max.overlaps = 120, show.legend = FALSE)
    }
  } +
  ggplot2::labs(
    title = "KJV + BOM similarity-profile PCA",
    subtitle = paste0("Colored by cluster; shaped by corpus; PCs used = ", preferred_n_pcs),
    x = paste0("PC1 (", round(sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Cluster",
    shape = "Corpus"
  ) +
  ggplot2::theme_minimal()
```


```{r take 2}
ggplot(
  scores_clustered,
  aes(
    x = PC1,
    y = PC2,
    color = cluster,
    shape = corpus_family
  )
) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.3
  ) +
  geom_vline(
    xintercept = 0,
    color = "grey70",
    linewidth = 0.3
  ) +
  geom_point(
    alpha = 0.75,
    size = 2.6
  ) +
  labs(
    title = "KJV + BOM similarity-profile PCA",
    subtitle = paste0(
      "Colored by mclust cluster; shaped by corpus; PCs used = ",
      preferred_n_pcs
    ),
    x = paste0(
      "PC1 (",
      round(sim_pca$variance$pct_variance[1], 2),
      "%)"
    ),
    y = paste0(
      "PC2 (",
      round(sim_pca$variance$pct_variance[2], 2),
      "%)"
    ),
    color = "Cluster",
    shape = "Corpus"
  ) +
  theme_minimal()











ggplot(
  scores_clustered,
  aes(
    x = PC1,
    y = PC2,
    color = cluster
  )
) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.3
  ) +
  geom_vline(
    xintercept = 0,
    color = "grey70",
    linewidth = 0.3
  ) +
  geom_point(
    alpha = 0.7,
    size = 2.5
  ) +
  facet_wrap(
    ~ corpus_family
  ) +
  labs(
    title = "KJV + BOM similarity-profile PCA",
    subtitle = "Panels separate KJV and BOM while preserving the same PCA space.",
    x = paste0(
      "PC1 (",
      round(sim_pca$variance$pct_variance[1], 2),
      "%)"
    ),
    y = paste0(
      "PC2 (",
      round(sim_pca$variance$pct_variance[2], 2),
      "%)"
    ),
    color = "Cluster"
  ) +
  theme_minimal()

```

```{r more PCA plots}
plotPCABoth <- function(
    scores,
    variance,
    x = 1,
    y = 2,
    color_var = "cluster",
    facet_var = "corpus_family",
    point_size = 2.5,
    alpha = 0.7,
    title = "KJV + BOM similarity-profile PCA",
    subtitle = "Panels separate corpora while preserving the same PCA space."
){

  x_col <- paste0("PC", x)
  y_col <- paste0("PC", y)

  stopifnot(x_col %in% names(scores))
  stopifnot(y_col %in% names(scores))
  stopifnot(color_var %in% names(scores))
  stopifnot(facet_var %in% names(scores))

  x_lab <- paste0(
    x_col,
    " (",
    round(variance$pct_variance[variance$dimension == x], 2),
    "%)"
  )

  y_lab <- paste0(
    y_col,
    " (",
    round(variance$pct_variance[variance$dimension == y], 2),
    "%)"
  )

  ggplot2::ggplot(
    scores,
    ggplot2::aes(
      x = .data[[x_col]],
      y = .data[[y_col]],
      color = .data[[color_var]]
    )
  ) +
    ggplot2::geom_hline(
      yintercept = 0,
      color = "grey70",
      linewidth = 0.3
    ) +
    ggplot2::geom_vline(
      xintercept = 0,
      color = "grey70",
      linewidth = 0.3
    ) +
    ggplot2::geom_point(
      alpha = alpha,
      size = point_size
    ) +
    ggplot2::facet_wrap(
      stats::as.formula(paste("~", facet_var))
    ) +
    ggplot2::labs(
      title = title,
      subtitle = subtitle,
      x = x_lab,
      y = y_lab,
      color = stringr::str_to_title(
        stringr::str_replace_all(color_var, "_", " ")
      )
    ) +
    ggplot2::theme_minimal()
}

```


```{r more plots}
plotPCABoth(scores_clustered, sim_pca$variance, x = 1, y = 2)

plotPCABoth(scores_clustered, sim_pca$variance, x = 3, y = 2)

plotPCABoth(scores_clustered, sim_pca$variance, x = 3, y = 4)

plotPCABoth(scores_clustered, sim_pca$variance, x = 5, y = 4)

plotPCABoth(scores_clustered, sim_pca$variance, x = 5, y = 6)




plotPCABoth(
  scores_clustered %>% dplyr::filter(corpus_family == "BOM"),
  sim_pca$variance,
  x = 1,
  y = 2,
  color_var = "mosiah_priority",
  facet_var = "corpus_family",
  title = "BOM chapters in KJV+BOM PCA space"
)
```
```{r}
scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC3 = mean(PC3),
    n=n()
  ) %>%
  arrange(desc(mean_PC3))

scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC3 = mean(PC3),
    n=n()
  ) %>%
  arrange(mean_PC3)

```




```{r pc12-genre}
ggplot2::ggplot(
  scores_clustered,
  ggplot2::aes(
    x = PC1,
    y = PC2,
    color = macro_genre,
    shape = corpus_family
  )
) +
  ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_point(alpha = 0.75, size = 2.4) +
  ggplot2::labs(
    title = "KJV + BOM PCA by macro genre",
    subtitle = "BOM chapters are marked as macro_genre = BOM.",
    x = paste0("PC1 (", round(sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Macro genre",
    shape = "Corpus"
  ) +
  ggplot2::theme_minimal()
```

```{r pc12-bom-mp}
scores_clustered %>%
  dplyr::filter(corpus_family == "BOM") %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = PC1,
      y = PC2,
      color = mosiah_priority,
      label = paste0(book, " ", chapter)
    )
  ) +
  ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
  ggplot2::geom_point(alpha = 0.8, size = 2.4) +
  {
    if(requireNamespace("ggrepel", quietly = TRUE)){
      ggrepel::geom_text_repel(size = 2.2, max.overlaps = 80, show.legend = FALSE)
    }
  } +
  ggplot2::labs(
    title = "BOM chapters in KJV+BOM PCA space",
    subtitle = "Colored by Mosiah Priority group.",
    x = paste0("PC1 (", round(sim_pca$variance$pct_variance[1], 2), "%)"),
    y = paste0("PC2 (", round(sim_pca$variance$pct_variance[2], 2), "%)"),
    color = "Mosiah Priority"
  ) +
  ggplot2::theme_minimal()


ggplot(
  scores_clustered %>%
    filter(corpus_family=="BOM"),
  aes(
    x = PC1,
    y = PC2,
    color = mosiah_priority
  )
) +
geom_hline(
  yintercept = 0,
  color = "grey70",
  linewidth = 0.3
) +
geom_vline(
  xintercept = 0,
  color = "grey70",
  linewidth = 0.3
) +
geom_point(
  alpha = 0.80,
  size = 2.7
) +
labs(
  title = "BOM chapters in KJV+BOM PCA space",
  subtitle = "Colored by Mosiah Priority group.",
  x = paste0(
    "PC1 (",
    round(sim_pca$variance$pct_variance[1], 2),
    "%)"
  ),
  y = paste0(
    "PC2 (",
    round(sim_pca$variance$pct_variance[2], 2),
    "%)"
  ),
  color = "Mosiah Priority"
) +
theme_minimal()


ggplot(
  scores_clustered %>%
    filter(corpus_family=="BOM"),
  aes(
    PC1,
    PC2,
    color = mosiah_priority
  )
) +
geom_hline(
  yintercept = 0,
  color = "grey70",
  linewidth = 0.3
) +
geom_vline(
  xintercept = 0,
  color = "grey70",
  linewidth = 0.3
) +
stat_ellipse(
  aes(group = mosiah_priority),
  linewidth = 1,
  alpha = 0.5
) +
geom_point(
  size = 2.7,
  alpha = 0.8
) +
labs(
  title = "BOM chapters in KJV+BOM PCA space",
  subtitle = "Ellipses summarize the Mosiah Priority groups.",
  x = paste0(
    "PC1 (",
    round(sim_pca$variance$pct_variance[1], 2),
    "%)"
  ),
  y = paste0(
    "PC2 (",
    round(sim_pca$variance$pct_variance[2], 2),
    "%)"
  ),
  color = "Mosiah Priority"
) +
theme_minimal()

```

## 22. Alignment metrics

```{r ari-metrics}
ari_metrics <- tibble::tibble(
  comparison = c(
    "cluster_vs_corpus_family",
    "cluster_vs_macro_genre",
    "cluster_vs_book",
    "cluster_vs_source_family",
    "cluster_vs_bom_mosiah_priority",
    "cluster_vs_isaiah_section"
  ),
  adjusted_rand_index = c(
    safe_ari(classified$corpus_family, classified$mclust_cluster),
    safe_ari(classified$macro_genre, classified$mclust_cluster),
    safe_ari(paste(classified$corpus_family, classified$book, sep = "__"), classified$mclust_cluster),
    safe_ari(classified$source_family, classified$mclust_cluster),
    safe_ari(
      classified %>% dplyr::filter(corpus_family == "BOM") %>% dplyr::pull(mosiah_priority),
      classified %>% dplyr::filter(corpus_family == "BOM") %>% dplyr::pull(mclust_cluster)
    ),
    safe_ari(
      classified %>% dplyr::filter(is_isaiah) %>% dplyr::pull(section),
      classified %>% dplyr::filter(is_isaiah) %>% dplyr::pull(mclust_cluster)
    )
  )
)

ari_metrics
```
```{r pca meanings}
scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC1 = mean(PC1),
    mean_PC2 = mean(PC2),
    n=n()
  ) %>%
  arrange(desc(mean_PC1))

scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC1 = mean(PC1),
    mean_PC2 = mean(PC2),
    n=n()
  ) %>%
  arrange((mean_PC1))



scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC1 = mean(PC1),
    mean_PC2 = mean(PC2),
    n=n()
  ) %>%
  arrange(desc(mean_PC2))


scores_clustered %>%
  group_by(book) %>%
  summarize(
    mean_PC1 = mean(PC1),
    mean_PC2 = mean(PC2),
    n=n()
  ) %>%
  arrange((mean_PC2))
```


```{r ari-plot}
ari_metrics %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = reorder(comparison, adjusted_rand_index),
      y = adjusted_rand_index
    )
  ) +
  ggplot2::geom_col() +
  ggplot2::coord_flip() +
  ggplot2::labs(
    title = "Adjusted Rand Index diagnostics",
    subtitle = "Which known labels best align with the latent clusters?",
    x = "Comparison",
    y = "Adjusted Rand Index"
  ) +
  ggplot2::theme_minimal()
```

## 23. Uncertainty diagnostics

```{r uncertainty}
classified %>%
  dplyr::group_by(corpus_family, macro_genre) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_uncertainty = mean(uncertainty, na.rm = TRUE),
    median_uncertainty = median(uncertainty, na.rm = TRUE),
    max_uncertainty = max(uncertainty, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_uncertainty))
```

```{r uncertainty-plot}
classified %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = corpus_family,
      y = uncertainty,
      fill = corpus_family
    )
  ) +
  ggplot2::geom_boxplot(alpha = 0.75, show.legend = FALSE) +
  ggplot2::labs(
    title = "Mclust classification uncertainty by corpus",
    x = "Corpus",
    y = "Uncertainty"
  ) +
  ggplot2::theme_minimal()
```


```{r isiah bom}
isaiah_ids <- analysis_meta %>%
  filter(
    corpus_family=="KJV",
    book=="Isaiah"
  ) %>%
  pull(doc_id)

bom_ids <- analysis_meta %>%
  filter(
    corpus_family=="BOM"
  ) %>%
  pull(doc_id)

S_bom_isaiah <- S_all[
    bom_ids,
    isaiah_ids,
    drop=FALSE
]

nearest_isaiah <- tibble(
    bom_doc = rownames(S_bom_isaiah),

    isaiah_doc =
      colnames(S_bom_isaiah)[
        max.col(S_bom_isaiah)
      ],

    similarity =
      apply(
        S_bom_isaiah,
        1,
        max
      )
)

nearest_isaiah



library(tidyr)

top5 <- as.data.frame(as.table(S_bom_isaiah)) %>%
  rename(
    bom_doc = Var1,
    isaiah_doc = Var2,
    similarity = Freq
  ) %>%
  group_by(bom_doc) %>%
  slice_max(
      similarity,
      n=5
  ) %>%
  arrange(
      bom_doc,
      desc(similarity)
  )

top5


pairs <- as.data.frame(as.table(S_bom_isaiah)) %>%
  rename(
    bom_doc=Var1,
    isaiah_doc=Var2,
    similarity=Freq
  )

summary(pairs$similarity)
quantile(
    pairs$similarity,
    c(.95,.99,.995,.999)
)

pairs %>%
    filter(
        similarity > .95
    ) %>%
    arrange(
        desc(similarity)
    )



pairs <- pairs %>%
  dplyr::mutate(

    isaiah_chapter =
      as.integer(
        str_extract(
          isaiah_doc,
          "(?<=Isaiah_)[0-9]+$"
        )
      ),

    bom_book =
      str_extract(
        bom_doc,
        "(1 Nephi|2 Nephi|Jacob|Enos|Jarom|Omni|Words of Mormon|Mosiah|Alma|Helaman|3 Nephi|4 Nephi|Mormon|Ether|Moroni)"
      ),

    bom_chapter =
      as.integer(
        str_extract(
          bom_doc,
          "[0-9]+$"
        )
      )
  )


ggplot(
    pairs,
    aes(
        isaiah_chapter,
        interaction(bom_book, bom_chapter),
        fill = similarity
    )
)+
geom_tile()


```
```{r more isiaha}
bom_meta <- analysis_docs %>%
  dplyr::filter(corpus_family == "BOM") %>%
  dplyr::select(
    doc_id,
    book,
    chapter
  ) %>%
  dplyr::rename(
    bom_doc = doc_id,
    bom_book = book,
    bom_chapter = chapter
  )

isa_meta <- analysis_docs %>%
  dplyr::filter(
    corpus_family == "KJV",
    book == "Isaiah"
  ) %>%
  dplyr::select(
    doc_id,
    chapter
  ) %>%
  dplyr::rename(
    isaiah_doc = doc_id,
    isaiah_chapter = chapter
  )


S_bom_isaiah <- S_all[
  bom_meta$bom_doc,
  isa_meta$isaiah_doc,
  drop = FALSE
]

S_bom_isaiah


nearest_isaiah <- tibble::tibble(
  bom_doc = rownames(S_bom_isaiah),

  isaiah_doc =
    colnames(S_bom_isaiah)[
      max.col(
        S_bom_isaiah,
        ties.method = "first"
      )
    ],

  cosine_similarity =
    apply(
      S_bom_isaiah,
      1,
      max
    )
)


nearest_isaiah

nearest_isaiah <- nearest_isaiah %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  )

nearest_isaiah


nearest_isaiah <- nearest_isaiah %>%
  dplyr::mutate(
    likely_direct_isaiah_match =
      cosine_similarity >= 0.96
  )

nearest_isaiah %>%
  dplyr::filter(likely_direct_isaiah_match) %>%
  dplyr::arrange(dplyr::desc(cosine_similarity))

nearest_isaiah %>%
  dplyr::filter(
    cosine_similarity < 0.96,
    cosine_similarity >= 0.93
  ) %>%
  dplyr::arrange(dplyr::desc(cosine_similarity))


nearest_isaiah <- nearest_isaiah %>%
  dplyr::mutate(
    isaiah_match_class = dplyr::case_when(
      cosine_similarity >= 0.96 ~ "Likely direct Isaiah match",
      cosine_similarity >= 0.93 ~ "High Isaiah affinity",
      cosine_similarity >= 0.90 ~ "Moderate Isaiah affinity",
      TRUE ~ "Low Isaiah affinity"
    )
  )


nearest_isaiah



known_full_isaiah_chapter_quotes <- tibble::tribble(
  ~bom_book, ~bom_chapter, ~isaiah_chapter,
  "1 Nephi", 20L, 48L,
  "1 Nephi", 21L, 49L,
  "2 Nephi", 7L, 50L,
  "2 Nephi", 8L, 51L,
  "2 Nephi", 12L, 2L,
  "2 Nephi", 13L, 3L,
  "2 Nephi", 14L, 4L,
  "2 Nephi", 15L, 5L,
  "2 Nephi", 16L, 6L,
  "2 Nephi", 17L, 7L,
  "2 Nephi", 18L, 8L,
  "2 Nephi", 19L, 9L,
  "2 Nephi", 20L, 10L,
  "2 Nephi", 21L, 11L,
  "2 Nephi", 22L, 12L,
  "2 Nephi", 23L, 13L,
  "2 Nephi", 24L, 14L,
  "Mosiah", 14L, 53L,
  "3 Nephi", 22L, 54L
)

known_partial_isaiah_matches <- tibble::tribble(
  ~bom_book, ~bom_chapter, ~isaiah_chapter,
  "2 Nephi", 27L, 29L
)


known_partial_isaiah_matches



nearest_isaiah_unique <- nearest_isaiah %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  ) %>%
  dplyr::distinct(
    isaiah_doc,
    .keep_all = TRUE
  ) %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  )

nearest_isaiah_unique

nearest_isaiah_unique_direct <- nearest_isaiah %>%
  dplyr::filter(
    likely_direct_isaiah_match
  ) %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  ) %>%
  dplyr::distinct(
    isaiah_doc,
    .keep_all = TRUE
  )

nearest_isaiah_unique_direct

nrow(nearest_isaiah_unique_direct)


```
```{r more iska}
isaiah_nearest_bom <- as.data.frame(as.table(S_bom_isaiah)) %>%
  dplyr::rename(
    bom_doc = Var1,
    isaiah_doc = Var2,
    cosine_similarity = Freq
  ) %>%
  dplyr::group_by(isaiah_doc) %>%
  dplyr::slice_max(
    order_by = cosine_similarity,
    n = 1,
    with_ties = FALSE
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(isaiah_doc)

isaiah_nearest_bom <- isaiah_nearest_bom %>%
  dplyr::left_join(
    bom_meta,
    by = "bom_doc"
  ) %>%
  dplyr::left_join(
    isa_meta,
    by = "isaiah_doc"
  ) %>%
  dplyr::select(
    isaiah_doc,
    isaiah_chapter,
    bom_doc,
    bom_book,
    bom_chapter,
    cosine_similarity
  ) %>%
  dplyr::arrange(isaiah_chapter)

nrow(isaiah_nearest_bom)
# 66

isaiah_nearest_bom




```


```{r isaiah clusters}
isaiah_bom_quotes <- tibble::tribble(
  ~book,      ~chapter,
  "1 Nephi", 20L,
  "1 Nephi", 21L,

  "2 Nephi",  7L,
  "2 Nephi",  8L,

  "2 Nephi", 12L,
  "2 Nephi", 13L,
  "2 Nephi", 14L,
  "2 Nephi", 15L,
  "2 Nephi", 16L,
  "2 Nephi", 17L,
  "2 Nephi", 18L,
  "2 Nephi", 19L,
  "2 Nephi", 20L,
  "2 Nephi", 21L,
  "2 Nephi", 22L,
  "2 Nephi", 23L,
  "2 Nephi", 24L,

  "Mosiah", 14L,

  "3 Nephi", 22L
)

scores_clustered %>%
  dplyr::filter(corpus_family=="BOM") %>%
  dplyr::semi_join(
      isaiah_bom_quotes,
      by=c("book","chapter")
  ) %>%
  dplyr::count(cluster, sort=TRUE)



scores_clustered %>%
  dplyr::filter(corpus_family=="BOM") %>%
  dplyr::semi_join(
      isaiah_bom_quotes,
      by=c("book","chapter")
  ) %>%
  dplyr::count(
      book,
      chapter,
      cluster
  ) %>%
  dplyr::arrange(book, chapter)


scores_clustered %>%
  filter(corpus_family=="BOM") %>%
  semi_join(
      isaiah_bom_quotes,
      by=c("book","chapter")
  ) %>%
  filter(cluster==2)

S_all[
  "BOM_1830__Mosiah_14",
  ]

sort(
  S_all[
    "BOM_1830__Mosiah_14",
  ],
  decreasing=TRUE
)[1:20]


scores_clustered %>%
  dplyr::semi_join(
    isaiah_bom_quotes,
    by = c("book", "chapter")
  ) %>%
  dplyr::filter(cluster == 4) %>%
  dplyr::count(book, chapter)


scores_clustered %>%
  filter(
    corpus_family=="KJV",
    book=="Isaiah",
    chapter==12
  ) %>%
  select(
    book,
    chapter,
    cluster
  )


scores_clustered %>%
  filter(
    book=="Isaiah",
    chapter==48
  ) %>%
  select(
    book,
    chapter,
    cluster
  )


sort(
  S_all["KJV__Isaiah_48",],
  decreasing=TRUE
)[1:20]


```

```{r compare KJV to KJV+BOM}
# had to also run 15_Mclust_all.Rmd first 

kjv_only_clusters <- all_scores_clustered %>%
  dplyr::transmute(
    corpus_family = "KJV",
    chapter_key = paste("KJV", book, chapter, sep = "__"),
    book,
    chapter = as.integer(chapter),
    cluster_KJV = mclust_cluster
  )

kjv_bom_clusters <- scores_clustered %>%
  dplyr::transmute(
    corpus_family,
    chapter_key = paste(corpus_family, book, chapter, sep = "__"),
    book,
    chapter = as.integer(chapter),
    cluster_KJV_BOM = cluster
  )

cluster_compare <- dplyr::full_join(
  kjv_only_clusters,
  kjv_bom_clusters,
  by = c("chapter_key", "book", "chapter", "corpus_family")
) %>%
  dplyr::arrange(
    corpus_family,
    book,
    chapter
  )




cluster_compare



cluster_compare %>%
  dplyr::filter(
    !is.na(cluster_KJV),
    !is.na(cluster_KJV_BOM),
    as.integer(as.character(cluster_KJV)) != as.integer(as.character(cluster_KJV_BOM))
  ) %>%
  dplyr::count(
    book,
    cluster_KJV,
    cluster_KJV_BOM,
    sort = TRUE
  )

cluster_compare



```

```{r get neighbors}
cluster_compare <- cluster_compare %>%
  dplyr::left_join(
    scores_clustered %>%
      dplyr::select(doc_id, corpus_family, book, chapter) %>%
      dplyr::mutate(chapter = as.integer(chapter)),
    by = c("corpus_family", "book", "chapter")
  )

neighbor_tbl <- purrr::map_dfr(
  cluster_compare$doc_id,
  ~{
    if(.x %in% rownames(S_all)){
      get_neighbors(.x, S_all, k = 10)
    } else {
      tibble::as_tibble(
        setNames(
          rep(list(NA), 20),
          as.vector(rbind(
            paste0("neighbor", 1:10),
            paste0("neighbor", 1:10, "_score")
          ))
        )
      )
    }
  }
)

cluster_compare_neighbors <- dplyr::bind_cols(
  cluster_compare,
  neighbor_tbl
)

cluster_compare_neighbors

sum(cluster_compare$doc_id %in% rownames(S_all))
nrow(cluster_compare)


```



## 24. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "mclust_all_bom")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(analysis_meta, file.path(out_dir, "analysis_meta.csv"))
readr::write_csv(sim_pca$variance, file.path(out_dir, "similarity_pca_variance.csv"))
readr::write_csv(sim_pca$scores, file.path(out_dir, "similarity_pca_scores.csv"))
readr::write_csv(classified, file.path(out_dir, "mclust_classified.csv"))
readr::write_csv(mclust_fit_summary, file.path(out_dir, "mclust_fit_summary_by_pcs.csv"))
readr::write_csv(bic_df, file.path(out_dir, "mclust_bic_preferred_model.csv"))
readr::write_csv(cluster_corpus_table, file.path(out_dir, "cluster_corpus_table.csv"))
readr::write_csv(cluster_genre_table, file.path(out_dir, "cluster_genre_table.csv"))
readr::write_csv(cluster_book_table, file.path(out_dir, "cluster_book_table.csv"))
readr::write_csv(bom_mp_cluster_table, file.path(out_dir, "bom_mosiah_priority_cluster_table.csv"))
readr::write_csv(isaiah_cluster_table, file.path(out_dir, "isaiah_cluster_table.csv"))
readr::write_csv(ari_metrics, file.path(out_dir, "ari_metrics.csv"))
readr::write_csv(feature_counts, file.path(out_dir, "feature_counts.csv"))
readr::write_csv(feature_counts_recursive, file.path(out_dir, "feature_counts_recursive.csv"))

cat("Exported outputs to:\n")
cat(out_dir, "\n")
```

## 25. Interpretation notes

```{r interpretation-notes}
interpretation_notes <- tibble::tribble(
  ~question, ~what_to_check,
  "Does the model separate BOM from KJV?", "Inspect cluster_corpus_table and cluster_vs_corpus_family ARI.",
  "Does BOM align more with KJV narrative, prophetic, or poetic chapters?", "Inspect PC plots and cluster_book_table.",
  "Does Mosiah Priority emerge as a latent structure?", "Inspect bom_mp_cluster_table and cluster_vs_bom_mosiah_priority ARI.",
  "Does adding BOM disturb the prior KJV genre structure?", "Inspect cluster_genre_table and cluster_vs_macro_genre ARI.",
  "Does Isaiah still fail to split into traditional sections?", "Inspect isaiah_cluster_table and cluster_vs_isaiah_section ARI.",
  "Does G=1:6 prefer a simple or complex model?", "Inspect mclust_fit_summary and the BIC plot."
)

interpretation_notes
```
