---
title: "41 - Prepare Book of Commandments 1833 from Wikisource"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook downloads and parses the **Book of Commandments (1833)** from Wikisource.

The Wikisource structure is approximately:

```text
Ch 3, February 1829

A Revelation given to Joseph, the father of Joseph,
in Harmony, Pennsylvania, February 1829. saying:

3:1 Now, behold, a marvelous work is about to come forth...

3:2 Remember temperance, patience, humility...
```

The notebook produces:

```text
data/raw/boc_1833_verses.csv
data/prepared/boc_1833_chapter_docs.csv
```

The chapter-level file is designed to be appended to the KJV/BOM chapter corpus in later PCA and mclust notebooks.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
library(rvest)
library(xml2)
```

## 3. Parameters

```{r parameters}
boc_url <- "https://en.wikisource.org/wiki/Book_of_Commandments"

raw_dir <- file.path(STYLO_ROOT, "data", "raw")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(raw_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

boc_html_path <- file.path(raw_dir, "boc_1833_wikisource.html")
boc_raw_text_path <- file.path(raw_dir, "boc_1833_wikisource_raw.txt")
boc_verses_path <- file.path(raw_dir, "boc_1833_verses.csv")
boc_chapters_path <- file.path(prepared_dir, "boc_1833_chapter_docs.csv")
```

## 4. Download Wikisource page

```{r download}
if(!file.exists(boc_html_path)){

  download.file(
    url = boc_url,
    destfile = boc_html_path,
    mode = "wb",
    quiet = FALSE
  )

}

boc_html <- xml2::read_html(boc_html_path)

boc_text_raw <- boc_html %>%
  rvest::html_element(".mw-parser-output") %>%
  rvest::html_text2()

readr::write_lines(boc_text_raw, boc_raw_text_path)

cat("Wrote raw Wikisource text to:\n")
cat(boc_raw_text_path, "\n")
```

## 5. Initial text cleanup
```{r debug}
## Diagnostic: detect chapter headings

boc_lines_debug <- readr::read_lines(
  boc_raw_text_path,
  locale = readr::locale(encoding = "UTF-8")
) %>%
  stringr::str_replace_all("\uFEFF", "") %>%
  stringr::str_replace_all("\r", "") %>%
  stringr::str_replace_all("\u00A0", " ") %>%
  stringr::str_squish()

boc_lines_debug <- boc_lines_debug[boc_lines_debug != ""]

chapter_heading_debug <- tibble::tibble(
  line_no = seq_along(boc_lines_debug),
  line = boc_lines_debug
) %>%
  dplyr::filter(
    stringr::str_detect(
      line,
      regex("^\\s*ch(?:apter)?\\.?\\s*[0-9]+", ignore_case = TRUE)
    )
  ) %>%
  dplyr::mutate(
    detected_chapter = as.integer(
      stringr::str_match(
        line,
        regex("^\\s*ch(?:apter)?\\.?\\s*([0-9]+)", ignore_case = TRUE)
      )[, 2]
    )
  )

chapter_heading_debug %>%
  print(n = 100)

chapter_heading_debug %>%
  dplyr::summarise(
    n_headings = dplyr::n(),
    first_chapter = min(detected_chapter, na.rm = TRUE),
    last_chapter = max(detected_chapter, na.rm = TRUE),
    missing_chapters = paste(
      setdiff(1:65, detected_chapter),
      collapse = ", "
    )
  )


```


```{r cleanup-text}
boc_lines <- readr::read_lines(
  boc_raw_text_path,
  locale = readr::locale(encoding = "UTF-8")
)

boc_lines <- boc_lines %>%
  stringr::str_replace_all("\uFEFF", "") %>%
  stringr::str_replace_all("\r", "") %>%
  stringr::str_replace_all("\u00A0", " ") %>%
  stringr::str_squish()

boc_lines <- boc_lines[boc_lines != ""]

length(boc_lines)

boc_lines[1:80]
```

## 6. Detect useful text boundaries

```{r boundaries}
chapter_heading_debug <- tibble::tibble(
  line_no = seq_along(boc_lines_debug),
  line = boc_lines_debug
) %>%
  dplyr::filter(
    stringr::str_detect(
      line,
      regex("^\\s*ch(?:apter)?\\.?\\s*[0-9]+", ignore_case = TRUE)
    )
  ) %>%
  dplyr::mutate(
    detected_chapter = as.integer(
      stringr::str_match(
        line,
        regex("^\\s*ch(?:apter)?\\.?\\s*([0-9]+)", ignore_case = TRUE)
      )[, 2]
    )
  )

body_start <- min(chapter_heading_debug$line_no)
body_end <- max(chapter_heading_debug$line_no)

# include everything after chapter 1 through the end of the page text
boc_body <- boc_lines_debug[body_start:length(boc_lines_debug)]
```

## 7. Helper functions

```{r helpers}


detect_chapter <- function(line){

  line <- stringr::str_squish(line)

  m <- stringr::str_match(
    line,
    regex("^\\s*ch(?:apter)?\\.?\\s*([0-9]+)\\s*,?\\s*(.*)$",
          ignore_case = TRUE)
  )

  if(!is.na(m[1, 2])){
    return(list(
      chapter = as.integer(m[1, 2]),
      chapter_date = stringr::str_squish(m[1, 3])
    ))
  }

  NULL
}




detect_verse <- function(line){

  line <- stringr::str_squish(line)

  m <- stringr::str_match(
    line,
    regex("^([0-9]+):([0-9]+)\\s+(.*)$", ignore_case = TRUE)
  )

  if(!is.na(m[1, 2])){
    return(
      list(
        chapter = as.integer(m[1, 2]),
        verse = as.integer(m[1, 3]),
        text = stringr::str_squish(m[1, 4])
      )
    )
  }

  # Fallback for sources using only verse numbers after a chapter heading.
  m2 <- stringr::str_match(
    line,
    regex("^([0-9]+)\\s+(.*)$", ignore_case = TRUE)
  )

  if(!is.na(m2[1, 2])){
    return(
      list(
        chapter = NA_integer_,
        verse = as.integer(m2[1, 2]),
        text = stringr::str_squish(m2[1, 3])
      )
    )
  }

  NULL
}


clean_revelation_text <- function(x){

  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()
}


make_doc_id <- function(corpus_id, translation, book, chapter){

  paste0(
    corpus_id,
    "_",
    translation,
    "__",
    stringr::str_replace_all(book, "\\s+", "_"),
    "_",
    chapter
  )

}
```

## 8. Parse chapters, headers, and verses

Important parsing rule:

- `chapter_date` is taken from the `Ch X, Date` line.
- `header` is the prose line(s) after the chapter heading but before the first verse.
- `text` excludes the header.
- `text_raw` preserves header + verse text for inspection.




```{r parse}
chapter_blocks <- chapter_heading_debug %>%
  dplyr::arrange(line_no) %>%
  dplyr::mutate(
    start_line = line_no,
    end_line = dplyr::lead(line_no, default = length(boc_lines_debug) + 1L) - 1L,
    chapter = detected_chapter,
    chapter_date = stringr::str_squish(
      stringr::str_match(
        line,
        regex("^\\s*ch(?:apter)?\\.?\\s*[0-9]+\\s*,?\\s*(.*)$",
              ignore_case = TRUE)
      )[, 2]
    )
  )

parse_chapter_block <- function(block){

  lines <- boc_lines_debug[block$start_line:block$end_line]

  # Remove chapter heading
  lines <- lines[-1]
  lines <- lines[lines != ""]

  block_text <- clean_revelation_text(
    paste(lines, collapse = " ")
  )

  marker_pattern <- paste0("\\b", block$chapter, ":([0-9]+)\\b")

  marker_locs <- stringr::str_locate_all(
    block_text,
    marker_pattern
  )[[1]]

  if(nrow(marker_locs) == 0){
    return(tibble::tibble())
  }

  first_marker_start <- marker_locs[1, "start"]

  header_text <- stringr::str_sub(
    block_text,
    1,
    first_marker_start - 1
  ) %>%
    clean_revelation_text()

  verse_starts <- marker_locs[, "start"]
  verse_ends <- c(
    marker_locs[-1, "start"] - 1L,
    stringr::str_length(block_text)
  )

  purrr::map2_dfr(
    verse_starts,
    verse_ends,
    function(s, e){

      verse_chunk <- stringr::str_sub(
        block_text,
        s,
        e
      ) %>%
        clean_revelation_text()

      m <- stringr::str_match(
        verse_chunk,
        paste0("^", block$chapter, ":([0-9]+)\\s*(.*)$")
      )

      tibble::tibble(
        corpus_id = "BOC",
        translation = "1833",
        book = "Book of Commandments",
        chapter = block$chapter,
        chapter_date = block$chapter_date,
        header = header_text,
        verse = as.integer(m[1, 2]),
        text = clean_revelation_text(m[1, 3]),
        text_raw = clean_revelation_text(
          paste(header_text, m[1, 3])
        )
      )
    }
  )
}

boc_verses <- purrr::map_dfr(
  split(chapter_blocks, chapter_blocks$chapter),
  parse_chapter_block
) %>%
  dplyr::arrange(chapter, verse)





nrow(boc_verses)
setdiff(1:65, unique(boc_verses$chapter))
```

## 9. Sanity checks

```{r sanity-checks}
boc_verses %>%
  dplyr::count(chapter, name = "n_verses") %>%
  dplyr::arrange(chapter)

boc_verses %>%
  dplyr::summarise(
    n_chapters = dplyr::n_distinct(chapter),
    n_verses = dplyr::n(),
    first_chapter = min(chapter, na.rm = TRUE),
    last_chapter = max(chapter, na.rm = TRUE),
    empty_text = sum(is.na(text) | text == ""),
    empty_header = sum(is.na(header) | header == "")
  )

verse_gaps <- boc_verses %>%
  dplyr::group_by(chapter) %>%
  dplyr::summarise(
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_verses = dplyr::n_distinct(verse),
    expected_n = max_verse - min_verse + 1L,
    has_gap = n_verses != expected_n,
    .groups = "drop"
  ) %>%
  dplyr::filter(has_gap)

verse_gaps
```
```{r debug}
boc_verses %>%
  dplyr::group_by(chapter) %>%
  dplyr::summarise(
    missing_verses = paste(
      setdiff(min(verse):max(verse), verse),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  dplyr::filter(missing_verses != "")

inspect_block <- function(ch){

  block <- chapter_blocks %>%
    dplyr::filter(chapter == ch)

  boc_lines_debug[block$start_line:block$end_line] %>%
    tibble::tibble(
      line_no = block$start_line:block$end_line,
      line = .
    )
}

inspect_block(16)
inspect_block(44)
inspect_block(54)



boc_verses %>%
  dplyr::count(chapter, verse) %>%
  dplyr::filter(n > 1)

boc_duplicate_verses <- boc_verses %>%
  dplyr::count(chapter, verse, name = "n") %>%
  dplyr::filter(n > 1)

boc_duplicate_verses


boc_missing_verses <- boc_verses %>%
  dplyr::group_by(chapter) %>%
  dplyr::summarise(
    missing_verses = paste(
      setdiff(min(verse):max(verse), verse),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  dplyr::filter(missing_verses != "")

boc_missing_verses

readr::write_csv(
  boc_missing_verses,
  file.path(raw_dir, "boc_1833_missing_verse_numbers.csv")
)

readr::write_csv(
  boc_duplicate_verses,
  file.path(raw_dir, "boc_1833_duplicate_verse_numbers.csv")
)



```

## 10. Build chapter documents

```{r chapter-docs}
boc_chapter_docs <- boc_verses %>%
  dplyr::group_by(
    corpus_id,
    translation,
    book,
    chapter,
    chapter_date,
    header
  ) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    text = clean_revelation_text(
      paste(text, collapse = " ")
    ),
    text_raw = clean_revelation_text(
      paste(unique(header), paste(text, collapse = " "), collapse = " ")
    ),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    corpus_family = "BOC",
    source_family = "BOC_1833",
    macro_genre = "Modern_Revelation",
    section = paste0("BOC_", chapter),
    doc_id = make_doc_id(corpus_id, translation, book, chapter),
    doc_id_original = doc_id
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    chapter_date,
    header,
    n_verses,
    text,
    text_raw
  ) %>%
  dplyr::arrange(chapter)

boc_chapter_docs %>%
  dplyr::slice_head(n = 20)

nrow(boc_chapter_docs)
```

## 11. Preview selected chapter

```{r preview}
boc_verses %>%
  dplyr::filter(chapter == 3)

boc_chapter_docs %>%
  dplyr::filter(chapter == 3) %>%
  dplyr::select(doc_id, chapter_date, header, n_verses, text)
```

## 12. Write CSV outputs

```{r write-csvs}
readr::write_csv(boc_verses, boc_verses_path)
readr::write_csv(boc_chapter_docs, boc_chapters_path)

cat("Wrote verse-level BOC CSV to:\n")
cat(boc_verses_path, "\n\n")

cat("Wrote chapter-level BOC document CSV to:\n")
cat(boc_chapters_path, "\n")
```

## 13. Output summary

```{r summary}
boc_summary <- boc_chapter_docs %>%
  dplyr::summarise(
    corpus_id = dplyr::first(corpus_id),
    translation = dplyr::first(translation),
    n_chapters = dplyr::n(),
    total_verses = sum(n_verses),
    total_words = sum(stringr::str_count(text, "\\S+")),
    mean_words_per_chapter = mean(stringr::str_count(text, "\\S+")),
    median_words_per_chapter = median(stringr::str_count(text, "\\S+"))
  )

boc_summary

boc_chapter_docs %>%
  dplyr::mutate(
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::select(
    chapter,
    chapter_date,
    n_verses,
    n_words,
    header
  ) %>%
  dplyr::arrange(chapter)
```

## 14. Next step

After this notebook succeeds, add `boc_1833_chapter_docs.csv` to your combined analysis corpus.

A later notebook can bind:

```text
data/prepared/kjv_chapter_docs.csv
data/prepared/bom_chapter_docs.csv
data/prepared/boc_1833_chapter_docs.csv
```

and then rerun:

```text
similarity matrix
PCA
mclust G = 1:9
nearest-neighbor diagnostics
```
