---
title: "42 - Mclust All Chapters with Book of Mormon and Book of Commandments"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.
source("./00_source_project.R")
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 300)
```

## 1. Purpose

This notebook extends the KJV + Book of Mormon mclust experiment by adding the **Book of Commandments (1833)**.

The analysis loads prepared KJV, BOM, and BOC chapter documents, builds an all-document similarity matrix, performs similarity-profile PCA, and runs model-based clustering with `G = 1:9`.

## 2. Packages

```{r packages}
if(!requireNamespace("mclust", quietly = TRUE)) install.packages("mclust")

library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
library(ggplot2)
library(mclust)

if(requireNamespace("ggrepel", quietly = TRUE)) library(ggrepel)
```

## 3. Parameters

```{r parameters}
set.seed(1234)

kjv_books_to_include <- c(
  "Isaiah", "Psalms", "Job", "Proverbs", "Song of Solomon", "Ecclesiastes",
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings"
)

bom_translation_to_include <- NULL

include_feature_classes <- c(
  "function_words", "punctuation", "sentence",
  "word_length", "lexical", "readability"
)

use_pos_features <- FALSE
if(isTRUE(use_pos_features)) include_feature_classes <- c(include_feature_classes, "pos")

pos_n <- c(3, 4, 5)
k_neighbors <- 10
preferred_n_pcs <- 6
G_values <- 1:9
mclust_model_names <- NULL
```

## 4. Load and normalize KJV chapters

```{r kjv}
kjv_chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "kjv_chapter_docs.csv")
if(!file.exists(kjv_chapter_path)) stop("Could not find: ", kjv_chapter_path)

kjv_chapter_docs_raw <- readr::read_csv(kjv_chapter_path, show_col_types = FALSE)

kjv_macro_genre_lookup <- tibble::tribble(
  ~book,              ~macro_genre_control,
  "1 Samuel",         "Narrative",
  "2 Samuel",         "Narrative",
  "1 Kings",          "Narrative",
  "2 Kings",          "Narrative",
  "Psalms",           "Poetic_Wisdom",
  "Job",              "Poetic_Wisdom",
  "Proverbs",         "Poetic_Wisdom",
  "Song of Solomon",  "Poetic_Wisdom",
  "Ecclesiastes",     "Poetic_Wisdom",
  "Isaiah",           "Prophetic",
  "Jeremiah",         "Prophetic",
  "Ezekiel",          "Prophetic",
  "Hosea",            "Prophetic",
  "Amos",             "Prophetic"
)

kjv_chapter_docs <- kjv_chapter_docs_raw %>%
  filter(book %in% kjv_books_to_include) %>%
  left_join(kjv_macro_genre_lookup, by = "book") %>%
  mutate(
    corpus_id = coalesce(corpus_id, "KJV"),
    translation = coalesce(translation, "KJV"),
    corpus_family = "KJV",
    source_family = "KJV",
    macro_genre = coalesce(macro_genre_control, macro_genre),
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    chapter_date = NA_character_,
    header = NA_character_,
    chapter = as.integer(chapter),
    doc_id_original = doc_id,
    doc_id = paste0("KJV__", doc_id)
  ) %>%
  select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original,
         book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label,
         chapter_date, header, n_verses, text)

dplyr::count(
  kjv_chapter_docs,
  book,
  macro_genre,
  name = "n_chapters"
) %>%
dplyr::arrange(
  macro_genre,
  book
)
```

## 5. Load and normalize BOM chapters

```{r bom}
bom_chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "bom_chapter_docs.csv")
if(!file.exists(bom_chapter_path)) stop("Could not find: ", bom_chapter_path)

bom_chapter_docs_raw <- readr::read_csv(bom_chapter_path, show_col_types = FALSE)

bom_chapter_docs <- bom_chapter_docs_raw %>%
  mutate(translation = as.character(translation))

if(!is.null(bom_translation_to_include)){
  bom_chapter_docs <- bom_chapter_docs %>% filter(translation == bom_translation_to_include)
}

bom_chapter_docs <- bom_chapter_docs %>%
  mutate(
    corpus_id = "BOM",
    corpus_family = "BOM",
    source_family = paste0("BOM_", translation),
    macro_genre = "BOM",
    section = coalesce(section, mosiah_priority),
    chapter_date = NA_character_,
    header = NA_character_,
    doc_id_original = doc_id,
    doc_id = paste0("BOM_", translation, "__", doc_id),
    chapter = as.integer(chapter)
  ) %>%
  select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original,
         book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label,
         chapter_date, header, n_verses, text)

bom_chapter_docs %>% dplyr::count(translation, book, mosiah_priority, name = "n_chapters") %>% arrange(translation, mosiah_priority, book)
```

## 6. Load and normalize Book of Commandments chapters

```{r boc}
boc_chapter_path <- file.path(STYLO_ROOT, "data", "prepared", "boc_1833_chapter_docs.csv")
if(!file.exists(boc_chapter_path)) stop("Could not find: ", boc_chapter_path, "\nRun 41_prepare_BC_Wiki.Rmd first.")

boc_chapter_docs_raw <- readr::read_csv(boc_chapter_path, show_col_types = FALSE)

boc_chapter_docs <- boc_chapter_docs_raw %>%
  mutate(
    translation = as.character(translation),
    corpus_id = "BOC",
    corpus_family = "BOC",
    source_family = "BOC_1833",
    macro_genre = "Modern_Revelation",
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    section = coalesce(section, paste0("BOC_", chapter)),
    chapter = as.integer(chapter),
    chapter_date = as.character(chapter_date),
    header = as.character(header),
    doc_id_original = doc_id,
    doc_id = if_else(str_detect(doc_id, "^BOC_1833__"), doc_id, paste0("BOC_1833__", doc_id))
  ) %>%
  select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original,
         book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label,
         chapter_date, header, n_verses, text)

boc_chapter_docs %>% dplyr::count(corpus_family, source_family, book, name = "n_chapters")
boc_chapter_docs %>% summarise(n_chapters = n_distinct(chapter), first_chapter = min(chapter), last_chapter = max(chapter))
```

## 7. Combined corpus

```{r combined-corpus}
analysis_docs <- bind_rows(kjv_chapter_docs, bom_chapter_docs, boc_chapter_docs) %>%
  mutate(text = as.character(text), chapter = as.integer(chapter), n_verses = as.integer(n_verses)) %>%
  filter(!is.na(text), text != "") %>%
  arrange(corpus_family, source_family, book, chapter)

analysis_meta <- analysis_docs %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    chapter_date,
    header,
    n_verses,
    text
  ) %>%
  dplyr::mutate(
    is_kjv = corpus_family == "KJV",
    is_bom = corpus_family == "BOM",
    is_boc = corpus_family == "BOC",
    is_isaiah = corpus_family == "KJV" & book == "Isaiah",
    is_bom_mp1 = corpus_family == "BOM" & mosiah_priority == "MP-1",
    is_bom_mp2 = corpus_family == "BOM" & mosiah_priority == "MP-2",
    is_bom_mp3 = corpus_family == "BOM" & mosiah_priority == "MP-3"
  )

analysis_docs %>% dplyr::count(corpus_family, source_family, name = "n_docs")
analysis_docs %>% dplyr::count(corpus_family, macro_genre, name = "n_docs")
analysis_docs %>% dplyr::count(corpus_family, book, name = "n_docs") %>% arrange(corpus_family, book)
nrow(analysis_docs)
```

## 8. Configure stylometry ensemble

```{r config}
config_all_boc <- default_config(include = include_feature_classes)
config_all_boc$pos_n <- pos_n

if(!isTRUE(use_pos_features)){
  config_all_boc$weights <- c(
    function_words = 0.500,
    punctuation    = 0.075,
    sentence       = 0.175,
    word_length    = 0.125,
    lexical        = 0.075,
    readability    = 0.050
  )
} else {
  config_all_boc$weights <- c(
    function_words = 0.300,
    pos            = 0.300,
    punctuation    = 0.050,
    sentence       = 0.150,
    word_length    = 0.100,
    lexical        = 0.050,
    readability    = 0.050
  )
}

config_all_boc$use_tfidf_function_words <- TRUE
if(exists("ud_model")) config_all_boc$pos_model <- ud_model

config_all_boc$include
config_all_boc$weights
```

## 9. Feature count helpers

```{r feature-count-helpers}
count_features <- function(x){
  if(is.matrix(x) || is.data.frame(x)) return(ncol(x))
  if(is.list(x)) return(sum(purrr::map_int(x, count_features)))
  0L
}

show_feature_counts <- function(feature_sets){
  out <- tibble(
    feature_class = names(feature_sets),
    n_features = purrr::map_int(feature_sets, count_features)
  ) %>%
    mutate(
      pct_features_raw = n_features / sum(n_features),
      pct_features = scales::percent(pct_features_raw, accuracy = 0.1)
    ) %>%
    arrange(desc(n_features))

  bind_rows(
    out,
    tibble(feature_class = "TOTAL", n_features = sum(out$n_features), pct_features_raw = 1, pct_features = "100.0%")
  )
}

show_feature_counts_recursive <- function(x, prefix = ""){
  if(is.matrix(x) || is.data.frame(x)) return(tibble(feature = prefix, n_features = ncol(x)))
  if(is.list(x)){
    return(purrr::imap_dfr(x, ~show_feature_counts_recursive(.x, if(prefix == "") .y else paste0(prefix, "::", .y))))
  }
  tibble(feature = prefix, n_features = 0L)
}
```

## 10. Run ensemble and build similarity matrix

```{r run-ensemble}
result_all_boc <- run_stylometry_ensemble(
  source_docs = analysis_docs,
  target_docs = analysis_docs,
  config = config_all_boc,
  k_neighbors = k_neighbors,
  use_cache = TRUE
)

result_all_boc$weights_used

feature_counts <- show_feature_counts(result_all_boc$feature_sets)
feature_counts

feature_counts_recursive <- show_feature_counts_recursive(result_all_boc$feature_sets)
feature_counts_recursive
```

```{r similarity-matrix}
S_all <- as.matrix(result_all_boc$overall_similarity)
S_all[is.na(S_all)] <- 0
S_all[is.nan(S_all)] <- 0
S_all[is.infinite(S_all)] <- 0

S_all <- S_all[analysis_meta$doc_id, analysis_meta$doc_id, drop = FALSE]

dim(S_all)
S_all[1:5, 1:5]
```

## 11. PCA helper and PCA

```{r pca-helper}
processPCA <- function(X, metadata = NULL, center = TRUE, scale. = TRUE){
  X <- as.matrix(X)
  X[is.na(X)] <- 0
  X[is.nan(X)] <- 0
  X[is.infinite(X)] <- 0

  feature_var <- apply(X, 2, stats::var, na.rm = TRUE)
  keep <- is.finite(feature_var) & feature_var > 0
  X <- X[, keep, drop = FALSE]

  pca <- stats::prcomp(X, center = center, scale. = scale.)
  eig <- pca$sdev^2

  variance <- tibble(
    pc = paste0("PC", seq_along(eig)),
    dimension = seq_along(eig),
    eigenvalue = eig,
    prop_variance = eig / sum(eig),
    pct_variance = 100 * eig / sum(eig),
    cumulative_prop_variance = cumsum(eig / sum(eig)),
    cumulative_pct_variance = 100 * cumsum(eig / sum(eig))
  )

  scores <- as.data.frame(pca$x) %>% rownames_to_column("doc_id")
  if(!is.null(metadata)) scores <- scores %>% left_join(metadata, by = "doc_id")

  loadings <- as.data.frame(pca$rotation) %>% rownames_to_column("feature")

  list(X = X, pca = pca, scores = scores, loadings = loadings, variance = variance)
}
```

```{r pca}
sim_pca <- processPCA(S_all, metadata = analysis_meta, center = TRUE, scale. = TRUE)
sim_pca$variance %>% slice_head(n = 20)
```

```{r scree-plot}
sim_pca$variance %>%
  slice_head(n = 20) %>%
  ggplot(aes(x = dimension, y = pct_variance)) +
  geom_col() +
  geom_line(aes(y = cumulative_pct_variance), linewidth = 0.8) +
  geom_point(aes(y = cumulative_pct_variance), size = 2) +
  scale_x_continuous(breaks = 1:20) +
  labs(
    title = "KJV + BOM + BOC similarity-profile PCA variance",
    subtitle = "Bars show individual PC variance; line shows cumulative variance.",
    x = "Principal component",
    y = "Percent variance"
  ) +
  theme_minimal()
```

## 12. Mclust helpers

```{r mclust-helpers}
run_mclust_on_pcs <- function(pca_obj, metadata, n_pcs = 6, G = 1:9, modelNames = NULL){
  pc_cols <- paste0("PC", seq_len(n_pcs))

  X <- pca_obj$scores %>% select(all_of(pc_cols)) %>% as.matrix()
  rownames(X) <- pca_obj$scores$doc_id

  fit <- mclust::Mclust(X, G = G, modelNames = modelNames)

  classified <- tibble(
    doc_id = rownames(X),
    mclust_cluster = fit$classification,
    uncertainty = fit$uncertainty
  ) %>%
    left_join(metadata, by = "doc_id") %>%
    arrange(mclust_cluster, corpus_family, book, chapter)

  list(fit = fit, classified = classified, n_pcs = n_pcs, pc_cols = pc_cols)
}

summarize_mclust_fit <- function(mc_obj){
  fit <- mc_obj$fit
  tibble(
    n_pcs = mc_obj$n_pcs,
    selected_G = fit$G,
    selected_model = fit$modelName,
    bic = fit$bic,
    loglik = fit$loglik,
    df = fit$df,
    n = fit$n,
    mean_uncertainty = mean(fit$uncertainty, na.rm = TRUE),
    max_uncertainty = max(fit$uncertainty, na.rm = TRUE)
  )
}

safe_ari <- function(a, b){
  keep <- !is.na(a) & !is.na(b)
  if(sum(keep) < 2) return(NA_real_)
  mclust::adjustedRandIndex(a[keep], b[keep])
}
```

## 13. Fit G = 1:9 over increasing PC counts

```{r fit-over-pcs}
max_pcs_available <- sum(grepl("^PC", names(sim_pca$scores)))
n_pcs_grid <- seq_len(min(10, max_pcs_available))

mclust_by_pcs <- purrr::map(
  n_pcs_grid,
  ~run_mclust_on_pcs(
    sim_pca,
    metadata = analysis_meta,
    n_pcs = .x,
    G = G_values,
    modelNames = mclust_model_names
  )
)

mclust_fit_summary <- purrr::map_dfr(mclust_by_pcs, summarize_mclust_fit)
mclust_fit_summary
```

```{r selected-g-plot}
mclust_fit_summary %>%
  ggplot(aes(x = n_pcs, y = selected_G)) +
  geom_line() +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = n_pcs_grid) +
  scale_y_continuous(breaks = G_values) +
  labs(
    title = "Selected G by number of PCs",
    subtitle = "KJV + BOM + BOC corpus; mclust tested G = 1:9.",
    x = "Number of PCs used",
    y = "Selected G"
  ) +
  theme_minimal()
```

## 14. Preferred model

```{r preferred-model}
preferred_n_pcs <- min(preferred_n_pcs, max_pcs_available)

mc_all_boc <- run_mclust_on_pcs(
  sim_pca,
  metadata = analysis_meta,
  n_pcs = preferred_n_pcs,
  G = G_values,
  modelNames = mclust_model_names
)

summary(mc_all_boc$fit)
summarize_mclust_fit(mc_all_boc)

classified <- mc_all_boc$classified
classified
```

## 15. BIC values for preferred model

```{r bic-preferred}
bic_matrix <- mc_all_boc$fit$BIC

bic_df <- as.data.frame(
  as.table(bic_matrix)
) %>%

dplyr::rename(

  G     = Var1,
  model = Var2,
  BIC   = Freq

) %>%

dplyr::mutate(

  G = as.integer(as.character(G))

) %>%

dplyr::filter(

  !is.na(BIC)

) %>%

dplyr::arrange(

  G,
  dplyr::desc(BIC)

)

bic_df


best_bic_by_G <- bic_df %>%

dplyr::group_by(
  G
) %>%

dplyr::slice_max(

  order_by = BIC,

  n = 1,

  with_ties = FALSE

) %>%

dplyr::ungroup()

best_bic_by_G

```

```{r bic-plot}
best_bic_by_G %>%
  ggplot(aes(x = G, y = BIC)) +
  geom_line() +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = G_values) +
  labs(title = paste0("Best BIC by G, using first ", preferred_n_pcs, " PCs"), x = "G", y = "BIC") +
  theme_minimal()
```

## 16. Cluster composition

```{r cluster-corpus}
cluster_corpus_table <- classified %>%
  dplyr::count(mclust_cluster, corpus_family, name = "n_chapters") %>%
  group_by(mclust_cluster) %>%
  mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  ungroup() %>%
  arrange(mclust_cluster, desc(n_chapters))

cluster_corpus_table
```

```{r cluster-corpus-plot}
cluster_corpus_table %>%
  ggplot(aes(x = factor(mclust_cluster), y = pct_cluster, fill = corpus_family)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Cluster composition by corpus", x = "Mclust cluster", y = "Cluster share", fill = "Corpus") +
  theme_minimal()
```

```{r cluster-genre}
cluster_genre_table <- classified %>%
  dplyr::count(mclust_cluster, macro_genre, name = "n_chapters") %>%
  group_by(mclust_cluster) %>%
  mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  ungroup() %>%
  arrange(mclust_cluster, desc(n_chapters))

cluster_genre_table
```

```{r cluster-book}
cluster_book_table <- classified %>%
  dplyr::count(mclust_cluster, corpus_family, book, macro_genre, mosiah_priority, name = "n_chapters") %>%
  group_by(mclust_cluster) %>%
  mutate(pct_cluster = n_chapters / sum(n_chapters)) %>%
  ungroup() %>%
  arrange(mclust_cluster, desc(n_chapters))

cluster_book_table
```

```{r cluster-book-plot, fig.height=11}
cluster_book_table %>%
  group_by(mclust_cluster) %>%
  slice_max(order_by = n_chapters, n = 12, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(book_label = paste0(corpus_family, ": ", book)) %>%
  ggplot(aes(x = reorder(book_label, n_chapters), y = n_chapters, fill = corpus_family)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ mclust_cluster, scales = "free_y") +
  labs(title = "Top books represented in each cluster", subtitle = "Top 12 books per cluster.", x = "Book", y = "Chapter count", fill = "Corpus") +
  theme_minimal()
```
```{r more}
classified %>%
dplyr::count(
  mclust_cluster,
  corpus_family,
  book,
  sort=TRUE
)

classified %>%
filter(
  corpus_family=="BOC"
) %>%
dplyr::count(
  mclust_cluster,
  chapter_date,
  sort=TRUE
)


classified %>%
filter(corpus_family=="BOC") %>%
dplyr::count(
  mclust_cluster,
  chapter_date,
  sort=TRUE
)




classified %>%
  dplyr::filter(corpus_family == "BOC") %>%
  dplyr::mutate(
    n_words = stringr::str_count(.data$text, "\\S+")
  ) %>%
  dplyr::group_by(mclust_cluster) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_words = mean(n_words),
    median_words = median(n_words),
    min_words = min(n_words),
    max_words = max(n_words),
    .groups = "drop"
  )


classified %>%
  dplyr::filter(
    corpus_family == "BOC",
    mclust_cluster == 4
  ) %>%
  dplyr::mutate(
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::select(
    chapter,
    chapter_date,
    mclust_cluster,
    n_verses,
    n_words,
    uncertainty,
    header
  ) %>%
  dplyr::arrange(chapter)


doc_id <- "BOC_1833__Book_of_Commandments_9"

tibble::tibble(
  neighbor = colnames(S_all),
  cosine_similarity = as.numeric(S_all[doc_id, ])
) %>%
  dplyr::filter(
    neighbor != doc_id
  ) %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  ) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::left_join(

    analysis_meta %>%

      dplyr::select(
        doc_id,
        corpus_family,
        book,
        chapter,
        macro_genre
      ),

    by = c(
      "neighbor" = "doc_id"
    )

  )


doc_id <- "BOC_1833__Book_of_Commandments_24"

tibble::tibble(
  neighbor = colnames(S_all),
  cosine_similarity = as.numeric(S_all[doc_id, ])
) %>%
  dplyr::filter(
    neighbor != doc_id
  ) %>%
  dplyr::arrange(
    dplyr::desc(cosine_similarity)
  ) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::left_join(

    analysis_meta %>%

      dplyr::select(
        doc_id,
        corpus_family,
        book,
        chapter,
        macro_genre
      ),

    by = c(
      "neighbor" = "doc_id"
    )

  )


```


## 17. BOC cluster distribution

```{r boc-cluster-table}
boc_cluster_table <- classified %>%
  filter(corpus_family == "BOC") %>%
  dplyr::count(mclust_cluster, name = "n_chapters") %>%
  mutate(pct_boc = n_chapters / sum(n_chapters)) %>%
  arrange(mclust_cluster)

boc_cluster_table
```

```{r boc-cluster-plot}
boc_cluster_table %>%
  ggplot(aes(x = factor(mclust_cluster), y = pct_boc, fill = factor(mclust_cluster))) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "BOC 1833 chapters across clusters", x = "Mclust cluster", y = "BOC share") +
  theme_minimal()
```

```{r boc-chapters-by-cluster}
classified %>%
  filter(corpus_family == "BOC") %>%
  select(mclust_cluster, chapter, chapter_date, n_verses, uncertainty, header) %>%
  arrange(mclust_cluster, chapter)
```

## 18. BOM Mosiah Priority distribution

```{r bom-mp-table}
bom_mp_cluster_table <- classified %>%
  filter(corpus_family == "BOM") %>%
  dplyr::count(mosiah_priority, mosiah_priority_label, mclust_cluster, name = "n_chapters") %>%
  group_by(mosiah_priority, mosiah_priority_label) %>%
  mutate(pct_mp = n_chapters / sum(n_chapters)) %>%
  ungroup() %>%
  arrange(mosiah_priority, mclust_cluster)

bom_mp_cluster_table
```

```{r bom-mp-plot}
bom_mp_cluster_table %>%
  ggplot(aes(x = mosiah_priority, y = pct_mp, fill = factor(mclust_cluster))) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "BOM Mosiah Priority groups across clusters", x = "Mosiah Priority group", y = "Group share", fill = "Mclust cluster") +
  theme_minimal()
```

## 19. Isaiah distribution across clusters

```{r isaiah-cluster-table}
isaiah_cluster_table <- classified %>%
  filter(is_isaiah) %>%
  dplyr::count(section, mclust_cluster, name = "n_chapters") %>%
  group_by(section) %>%
  mutate(pct_section = n_chapters / sum(n_chapters)) %>%
  ungroup() %>%
  arrange(section, mclust_cluster)

isaiah_cluster_table
```

```{r isaiah-cluster-plot}
isaiah_cluster_table %>%
  ggplot(aes(x = section, y = pct_section, fill = factor(mclust_cluster))) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Isaiah chapters across KJV+BOM+BOC clusters", subtitle = "Isaiah is treated as ordinary chapter documents; sections are diagnostic only.", x = "Isaiah section", y = "Section share", fill = "Mclust cluster") +
  theme_minimal()
```

## 20. PCA plotting helper and plots

```{r pca-plot-helper}
plotPCABoth <- function(scores, variance, x = 1, y = 2, color_var = "cluster", facet_var = "corpus_family", point_size = 2.5, alpha = 0.7, title = "KJV + BOM + BOC similarity-profile PCA", subtitle = "Panels separate corpora while preserving the same PCA space."){
  x_col <- paste0("PC", x)
  y_col <- paste0("PC", y)

  stopifnot(x_col %in% names(scores), y_col %in% names(scores), color_var %in% names(scores), facet_var %in% names(scores))

  x_lab <- paste0(x_col, " (", round(variance$pct_variance[variance$dimension == x], 2), "%)")
  y_lab <- paste0(y_col, " (", round(variance$pct_variance[variance$dimension == y], 2), "%)")

  ggplot(scores, aes(x = .data[[x_col]], y = .data[[y_col]], color = .data[[color_var]])) +
    geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
    geom_vline(xintercept = 0, color = "grey70", linewidth = 0.3) +
    geom_point(alpha = alpha, size = point_size) +
    facet_wrap(stats::as.formula(paste("~", facet_var))) +
    labs(title = title, subtitle = subtitle, x = x_lab, y = y_lab, color = str_to_title(str_replace_all(color_var, "_", " "))) +
    theme_minimal()
}
```

```{r pc-plots}
scores_clustered <- sim_pca$scores %>%
  left_join(classified %>% select(doc_id, mclust_cluster, uncertainty), by = "doc_id") %>%
  mutate(cluster = factor(mclust_cluster))

plotPCABoth(scores_clustered, sim_pca$variance, x = 1, y = 2, color_var = "cluster", facet_var = "corpus_family")
plotPCABoth(scores_clustered, sim_pca$variance, x = 2, y = 3, color_var = "cluster", facet_var = "corpus_family")
plotPCABoth(scores_clustered, sim_pca$variance, x = 3, y = 4, color_var = "cluster", facet_var = "corpus_family")
```

```{r pc-bom-mp}
plotPCABoth(
  scores_clustered %>% filter(corpus_family == "BOM"),
  sim_pca$variance,
  x = 1,
  y = 2,
  color_var = "mosiah_priority",
  facet_var = "corpus_family",
  title = "BOM chapters in KJV+BOM+BOC PCA space",
  subtitle = "Colored by Mosiah Priority group."
)
```

```{r pc-boc}
plotPCABoth(
  scores_clustered %>% filter(corpus_family == "BOC"),
  sim_pca$variance,
  x = 1,
  y = 2,
  color_var = "cluster",
  facet_var = "corpus_family",
  title = "BOC 1833 chapters in KJV+BOM+BOC PCA space",
  subtitle = "Colored by mclust cluster."
)
```

## 21. Book means for PCA interpretation

```{r book-means}
book_pc_means <- scores_clustered %>%
  group_by(corpus_family, book) %>%
  summarise(
    mean_PC1 = mean(PC1, na.rm = TRUE),
    mean_PC2 = mean(PC2, na.rm = TRUE),
    mean_PC3 = mean(PC3, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

book_pc_means %>% arrange(desc(mean_PC1))
book_pc_means %>% arrange(mean_PC1)
book_pc_means %>% arrange(desc(mean_PC2))
book_pc_means %>% arrange(mean_PC2)
book_pc_means %>% arrange(desc(mean_PC3))
book_pc_means %>% arrange(mean_PC3)
```

## 22. ARI metrics

```{r ari-metrics}
ari_metrics <- tibble(
  comparison = c(
    "cluster_vs_corpus_family",
    "cluster_vs_macro_genre",
    "cluster_vs_book",
    "cluster_vs_source_family",
    "cluster_vs_bom_mosiah_priority",
    "cluster_vs_isaiah_section",
    "cluster_vs_boc_chapter_date"
  ),
  adjusted_rand_index = c(
    safe_ari(classified$corpus_family, classified$mclust_cluster),
    safe_ari(classified$macro_genre, classified$mclust_cluster),
    safe_ari(paste(classified$corpus_family, classified$book, sep = "__"), classified$mclust_cluster),
    safe_ari(classified$source_family, classified$mclust_cluster),
    safe_ari(classified %>% filter(corpus_family == "BOM") %>% pull(mosiah_priority), classified %>% filter(corpus_family == "BOM") %>% pull(mclust_cluster)),
    safe_ari(classified %>% filter(is_isaiah) %>% pull(section), classified %>% filter(is_isaiah) %>% pull(mclust_cluster)),
    safe_ari(classified %>% filter(corpus_family == "BOC") %>% pull(chapter_date), classified %>% filter(corpus_family == "BOC") %>% pull(mclust_cluster))
  )
)

ari_metrics
```

```{r ari-plot}
ari_metrics %>%
  ggplot(aes(x = reorder(comparison, adjusted_rand_index), y = adjusted_rand_index)) +
  geom_col() +
  coord_flip() +
  labs(title = "Adjusted Rand Index diagnostics", subtitle = "Which known labels best align with the latent clusters?", x = "Comparison", y = "Adjusted Rand Index") +
  theme_minimal()
```

## 23. Nearest-neighbor helpers and BOC neighbors

```{r nearest-neighbor-helpers}
get_neighbors <- function(doc_id, S, k = 20){
  sims <- S[doc_id, ]

  tb <- tibble(neighbor = names(sims), score = as.numeric(sims)) %>%
    filter(neighbor != doc_id) %>%
    arrange(desc(score)) %>%
    slice_head(n = k)

  out <- list()
  for(i in seq_len(nrow(tb))){
    out[[paste0("neighbor", i)]] <- tb$neighbor[i]
    out[[paste0("neighbor", i, "_score")]] <- tb$score[i]
  }
  as_tibble(out)
}

nearest_neighbors_long <- function(doc_ids, S, k = 20){
  purrr::map_dfr(doc_ids, function(id){
    sims <- S[id, ]
    tibble(doc_id = id, neighbor = names(sims), cosine_similarity = as.numeric(sims)) %>%
      filter(neighbor != id) %>%
      arrange(desc(cosine_similarity)) %>%
      slice_head(n = k) %>%
      mutate(rank = row_number())
  })
}
```

```{r boc-nearest-neighbors}
boc_ids <- analysis_meta %>% filter(corpus_family == "BOC") %>% pull(doc_id)

boc_neighbors_long <- nearest_neighbors_long(boc_ids, S_all, k = 20)

boc_neighbors_annotated <- boc_neighbors_long %>%
  left_join(
    analysis_meta %>% select(neighbor_doc_id = doc_id, neighbor_corpus = corpus_family, neighbor_book = book, neighbor_chapter = chapter, neighbor_section = section, neighbor_macro_genre = macro_genre),
    by = c("neighbor" = "neighbor_doc_id")
  ) %>%
  arrange(doc_id, rank)

boc_neighbors_annotated %>% slice_head(n = 200)
```

```{r boc-nearest-summary}
boc_neighbors_annotated %>%
  filter(rank <= 10) %>%
  dplyr::count(neighbor_corpus, neighbor_macro_genre, name = "n_top10_neighbors") %>%
  arrange(desc(n_top10_neighbors))
```

## 24. Uncertainty diagnostics

```{r uncertainty}
classified %>%
  group_by(corpus_family, macro_genre) %>%
  summarise(
    n = n(),
    mean_uncertainty = mean(uncertainty, na.rm = TRUE),
    median_uncertainty = median(uncertainty, na.rm = TRUE),
    max_uncertainty = max(uncertainty, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_uncertainty))
```

```{r uncertainty-plot}
classified %>%
  ggplot(aes(x = corpus_family, y = uncertainty, fill = corpus_family)) +
  geom_boxplot(alpha = 0.75, show.legend = FALSE) +
  labs(title = "Mclust classification uncertainty by corpus", x = "Corpus", y = "Uncertainty") +
  theme_minimal()
```

## 25. Export outputs

```{r export}
out_dir <- file.path(STYLO_ROOT, "output", "mclust_all_bom_boc")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(analysis_meta, file.path(out_dir, "analysis_meta.csv"))
write_csv(sim_pca$variance, file.path(out_dir, "similarity_pca_variance.csv"))
write_csv(sim_pca$scores, file.path(out_dir, "similarity_pca_scores.csv"))
write_csv(classified, file.path(out_dir, "mclust_classified.csv"))
write_csv(mclust_fit_summary, file.path(out_dir, "mclust_fit_summary_by_pcs.csv"))
write_csv(bic_df, file.path(out_dir, "mclust_bic_preferred_model.csv"))
write_csv(best_bic_by_G, file.path(out_dir, "mclust_best_bic_by_G.csv"))
write_csv(cluster_corpus_table, file.path(out_dir, "cluster_corpus_table.csv"))
write_csv(cluster_genre_table, file.path(out_dir, "cluster_genre_table.csv"))
write_csv(cluster_book_table, file.path(out_dir, "cluster_book_table.csv"))
write_csv(boc_cluster_table, file.path(out_dir, "boc_cluster_table.csv"))
write_csv(bom_mp_cluster_table, file.path(out_dir, "bom_mosiah_priority_cluster_table.csv"))
write_csv(isaiah_cluster_table, file.path(out_dir, "isaiah_cluster_table.csv"))
write_csv(ari_metrics, file.path(out_dir, "ari_metrics.csv"))
write_csv(feature_counts, file.path(out_dir, "feature_counts.csv"))
write_csv(feature_counts_recursive, file.path(out_dir, "feature_counts_recursive.csv"))
write_csv(book_pc_means, file.path(out_dir, "book_pc_means.csv"))
write_csv(boc_neighbors_long, file.path(out_dir, "boc_nearest_neighbors_long.csv"))
write_csv(boc_neighbors_annotated, file.path(out_dir, "boc_nearest_neighbors_annotated.csv"))

cat("Exported outputs to:\n")
cat(out_dir, "\n")
```

## 26. Interpretation notes

```{r interpretation-notes}
interpretation_notes <- tibble::tribble(
  ~question, ~what_to_check,
  "Does BOC form its own cluster?", "Inspect boc_cluster_table and cluster_corpus_table.",
  "Does BOC align with BOM distinctive clusters?", "Inspect BOC rows in cluster_book_table and BOC nearest neighbors.",
  "Does BOC align with Isaiah / prophetic KJV?", "Inspect nearest-neighbor tables and PC3 positions.",
  "Does adding BOC disturb the KJV+BOM six-cluster solution?", "Compare selected G and cluster composition to 31_Mclust_all_BOM.Rmd.",
  "Does Isaiah still fail to split into traditional sections?", "Inspect isaiah_cluster_table and cluster_vs_isaiah_section ARI.",
  "Does BOC date correspond to cluster structure?", "Inspect cluster_vs_boc_chapter_date ARI and BOC chapter-date means."
)

interpretation_notes
```
