---
title: "43 - Mclust All BOM + BoC + DC 1835"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 300)

library(dplyr)
library(stringr)
library(readr)
library(tibble)
library(tidyr)
library(purrr)
library(ggplot2)
library(mclust)
library(quanteda)
library(Matrix)
```

## 1. Purpose

This notebook extends `42_Mclust_all_BOM_BoC.Rmd` by adding the 1835 Doctrine and Covenants chapter documents:

```text
data/prepared/dc_1835_chapter_docs.csv
```

It compares:

```text
BOM 1830 chapters
BOC 1833 chapters
DC 1835 chapters
```

and runs `mclust` with:

```text
G = 1:12
```

## 2. Input and output paths

```{r paths}
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
export_dir <- file.path(STYLO_ROOT, "export")

dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)

bom_path <- file.path(prepared_dir, "bom_1830_chapter_docs.csv")
boc_path <- file.path(prepared_dir, "boc_1833_chapter_docs.csv")
dc1835_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")

out_docs_path <- file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_docs.csv")
out_features_path <- file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_features.csv")
out_assignments_path <- file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_assignments.csv")
out_model_summary_path <- file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_model_summary.csv")
```

## 3. Read prepared chapter documents

```{r read-docs}
read_required_csv <- function(path){
  if(!file.exists(path)){
    stop("Missing required file: ", path)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

bom_docs_raw <- read_required_csv(bom_path)
boc_docs_raw <- read_required_csv(boc_path)
dc1835_docs_raw <- read_required_csv(dc1835_path)

list(
  BOM = dim(bom_docs_raw),
  BOC = dim(boc_docs_raw),
  DC1835 = dim(dc1835_docs_raw)
)
```

## 4. Standardize metadata

```{r standardize-docs}
standardize_chapter_docs <- function(df, corpus_family_default, source_family_default){
  df %>%
    mutate(
      corpus_family = if("corpus_family" %in% names(.)) corpus_family else corpus_family_default,
      source_family = if("source_family" %in% names(.)) source_family else source_family_default,
      book = if("book" %in% names(.)) book else if("work" %in% names(.)) work else corpus_family_default,
      chapter = if("chapter" %in% names(.)) as.integer(chapter) else row_number(),
      chapter_title = if("chapter_title" %in% names(.)) chapter_title else paste0(book, " ", chapter),
      doc_id = if("doc_id" %in% names(.)) doc_id else paste0(source_family_default, "__", stringr::str_pad(row_number(), 3, pad = "0")),
      text = text %>%
        stringr::str_replace_all("\r", " ") %>%
        stringr::str_replace_all("\n", " ") %>%
        stringr::str_squish(),
      n_words = stringr::str_count(text, "\\S+")
    ) %>%
    select(
      doc_id,
      corpus_family,
      source_family,
      book,
      chapter,
      chapter_title,
      n_words,
      text,
      everything()
    )
}

bom_docs <- standardize_chapter_docs(bom_docs_raw, "BOM", "BOM_1830")
boc_docs <- standardize_chapter_docs(boc_docs_raw, "BOC", "BOC_1833")
dc1835_docs <- standardize_chapter_docs(dc1835_docs_raw, "DC", "DC_1835")

all_docs <- bind_rows(
  bom_docs,
  boc_docs,
  dc1835_docs
) %>%
  mutate(
    analysis_id = paste0(corpus_family, "__", make.unique(doc_id)),
    text = text %>%
      str_replace_all("\r", " ") %>%
      str_replace_all("\n", " ") %>%
      str_squish(),
    n_words = str_count(text, "\\S+")
  ) %>%
  filter(!is.na(text), text != "", n_words >= 50) %>%
  distinct(analysis_id, .keep_all = TRUE)

all_docs %>%
  count(corpus_family, source_family, name = "n_docs") %>%
  arrange(corpus_family, source_family)

readr::write_csv(all_docs, out_docs_path)
```

## 5. Sanity checks

```{r sanity-docs}
all_docs %>%
  summarise(
    n_docs = n(),
    n_families = n_distinct(corpus_family),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )

all_docs %>%
  arrange(n_words) %>%
  select(analysis_id, corpus_family, source_family, book, chapter, chapter_title, n_words) %>%
  print(n = 25)

all_docs %>%
  arrange(desc(n_words)) %>%
  select(analysis_id, corpus_family, source_family, book, chapter, chapter_title, n_words) %>%
  print(n = 25)

if(!all(c("BOM", "BOC", "DC") %in% unique(all_docs$corpus_family))){
  warning("Expected BOM, BOC, and DC corpus families are not all present.")
}
```

## 6. Build stylometric features

This version uses character 3-grams and 4-grams with tf-idf, then z-scales the feature matrix before mclust.

```{r feature-build}
tokens_to_dfm_char <- function(texts, n){
  quanteda::tokens(
    texts,
    what = "character",
    remove_punct = FALSE,
    remove_symbols = FALSE,
    remove_numbers = FALSE,
    remove_separators = TRUE
  ) %>%
    quanteda::tokens_ngrams(n = n) %>%
    quanteda::dfm()
}

corp <- quanteda::corpus(
  all_docs,
  text_field = "text",
  docid_field = "analysis_id"
)

dfm_char3 <- tokens_to_dfm_char(quanteda::texts(corp), 3)
dfm_char4 <- tokens_to_dfm_char(quanteda::texts(corp), 4)

dfm_all <- dfm_char3 + dfm_char4

dfm_trimmed <- quanteda::dfm_trim(
  dfm_all,
  min_termfreq = 5,
  min_docfreq = 3
)

dfm_tfidf <- quanteda::dfm_tfidf(
  dfm_trimmed,
  scheme_tf = "prop"
)

X_sparse <- as(dfm_tfidf, "dgCMatrix")
X <- as.matrix(X_sparse)

feature_sd <- apply(X, 2, sd, na.rm = TRUE)
X <- X[, feature_sd > 0, drop = FALSE]

X_scaled <- scale(X)
X_scaled <- as.matrix(X_scaled)

feature_df <- as_tibble(X_scaled) %>%
  mutate(analysis_id = rownames(X_scaled), .before = 1)

readr::write_csv(feature_df, out_features_path)

dim(X_scaled)
```

## 7. Sanity check feature matrix

```{r sanity-features}
feature_summary <- tibble(
  n_docs = nrow(X_scaled),
  n_features = ncol(X_scaled),
  any_na = any(is.na(X_scaled)),
  any_infinite = any(is.infinite(X_scaled))
)

feature_summary

if(any(is.na(X_scaled))) stop("Feature matrix contains NA.")
if(any(is.infinite(X_scaled))) stop("Feature matrix contains infinite values.")
if(nrow(X_scaled) != nrow(all_docs)) stop("Feature matrix document count does not match all_docs.")
```

## 8. Run mclust G = 1:12

```{r run-mclust}
set.seed(1830)

mclust_fit <- mclust::Mclust(
  data = X_scaled,
  G = 1:12
)

mclust_fit
summary(mclust_fit)
```

## 9. Model summary

```{r model-summary}
bic_df <- as.data.frame(mclust_fit$BIC) %>%
  tibble::rownames_to_column("model_name") %>%
  tidyr::pivot_longer(
    cols = -model_name,
    names_to = "G",
    values_to = "BIC"
  ) %>%
  mutate(G = as.integer(G)) %>%
  filter(!is.na(BIC)) %>%
  arrange(desc(BIC))

best_model_summary <- tibble(
  selected_G = mclust_fit$G,
  selected_model = mclust_fit$modelName,
  loglik = mclust_fit$loglik,
  bic = mclust_fit$bic,
  icl = mclust_fit$icl
)

best_model_summary

readr::write_csv(best_model_summary, out_model_summary_path)

bic_df %>% slice_head(n = 20)
```

## 10. Plot BIC by G

```{r plot-bic, fig.width=9, fig.height=5}
bic_by_g <- bic_df %>%
  group_by(G) %>%
  summarise(best_BIC = max(BIC, na.rm = TRUE), .groups = "drop")

ggplot(bic_by_g, aes(x = G, y = best_BIC)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:12) +
  labs(
    title = "Mclust BIC by Number of Clusters",
    subtitle = "BOM 1830 + BOC 1833 + DC 1835",
    x = "Number of clusters G",
    y = "Best BIC"
  )
```

## 11. Cluster assignments

```{r assignments}
cluster_assignments <- all_docs %>%
  mutate(
    mclust_cluster = mclust_fit$classification,
    uncertainty = mclust_fit$uncertainty
  ) %>%
  select(
    analysis_id,
    doc_id,
    corpus_family,
    source_family,
    book,
    chapter,
    chapter_title,
    n_words,
    mclust_cluster,
    uncertainty,
    text
  ) %>%
  arrange(mclust_cluster, corpus_family, source_family, book, chapter)

cluster_assignments %>%
  count(mclust_cluster, corpus_family, source_family, name = "n_docs") %>%
  arrange(mclust_cluster, corpus_family, source_family) %>%
  print(n = 200)

readr::write_csv(cluster_assignments, out_assignments_path)
```

## 12. Cluster composition

```{r cluster-composition}
cluster_composition <- cluster_assignments %>%
  count(mclust_cluster, corpus_family, name = "n") %>%
  group_by(mclust_cluster) %>%
  mutate(
    cluster_total = sum(n),
    pct_cluster = n / cluster_total
  ) %>%
  ungroup() %>%
  arrange(mclust_cluster, desc(n))

cluster_composition %>% print(n = 200)

cluster_composition_wide <- cluster_composition %>%
  select(mclust_cluster, corpus_family, n) %>%
  tidyr::pivot_wider(
    names_from = corpus_family,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(mclust_cluster)

cluster_composition_wide
```

## 13. Plot cluster composition

```{r plot-cluster-composition, fig.width=10, fig.height=6}
ggplot(
  cluster_composition,
  aes(x = factor(mclust_cluster), y = n, fill = corpus_family)
) +
  geom_col() +
  labs(
    title = "Mclust Cluster Composition",
    subtitle = "BOM 1830 + BOC 1833 + DC 1835",
    x = "Mclust cluster",
    y = "Number of chapter documents",
    fill = "Corpus"
  )
```

## 14. Uncertainty diagnostics

```{r uncertainty-diagnostics}
cluster_assignments %>%
  arrange(desc(uncertainty)) %>%
  select(corpus_family, source_family, book, chapter, chapter_title, mclust_cluster, uncertainty, n_words) %>%
  print(n = 50)

cluster_assignments %>%
  group_by(corpus_family, source_family) %>%
  summarise(
    n_docs = n(),
    mean_uncertainty = mean(uncertainty),
    median_uncertainty = median(uncertainty),
    max_uncertainty = max(uncertainty),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_uncertainty))
```

## 15. Chapter chronology views

```{r chronology-plots, fig.width=11, fig.height=5}
cluster_assignments %>%
  filter(corpus_family %in% c("BOC", "DC")) %>%
  mutate(textual_order = chapter) %>%
  ggplot(aes(x = textual_order, y = mclust_cluster, color = corpus_family)) +
  geom_point(alpha = 0.8) +
  geom_line(alpha = 0.3) +
  scale_y_continuous(breaks = sort(unique(cluster_assignments$mclust_cluster))) +
  labs(
    title = "Cluster Assignment by Section/Chapter Order",
    subtitle = "BOC 1833 and DC 1835",
    x = "Section / chapter order",
    y = "Mclust cluster",
    color = "Corpus"
  )
```

## 16. Cross-tab comparisons

```{r cross-tabs}
table_cluster_family <- table(
  cluster_assignments$mclust_cluster,
  cluster_assignments$corpus_family
)

table_cluster_family
prop.table(table_cluster_family, margin = 1)

table_cluster_source <- table(
  cluster_assignments$mclust_cluster,
  cluster_assignments$source_family
)

table_cluster_source
```

## 17. Adjusted Rand Index comparisons

```{r ari}
ari_safe <- function(x, y){
  ok <- !is.na(x) & !is.na(y)
  mclust::adjustedRandIndex(x[ok], y[ok])
}

ari_results <- tibble(
  comparison = c(
    "cluster_vs_corpus_family",
    "cluster_vs_source_family",
    "cluster_vs_book"
  ),
  adjusted_rand_index = c(
    ari_safe(cluster_assignments$mclust_cluster, cluster_assignments$corpus_family),
    ari_safe(cluster_assignments$mclust_cluster, cluster_assignments$source_family),
    ari_safe(cluster_assignments$mclust_cluster, cluster_assignments$book)
  )
)

ari_results
```

## 18. Save additional exports

```{r save-extra}
readr::write_csv(
  cluster_composition,
  file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_cluster_composition.csv")
)

readr::write_csv(
  bic_df,
  file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_bic_grid.csv")
)

readr::write_csv(
  ari_results,
  file.path(prepared_dir, "mclust_all_BOM_BoC_DC1835_ari_results.csv")
)
```

## 19. Notes

Interpretation guardrails:

1. `mclust` assumes Gaussian mixture structure in the feature space.
2. The tf-idf character n-gram matrix is z-scaled before clustering.
3. Clusters are not authors by default; they are feature-space components.
4. Compare cluster assignments with corpus family, source family, book, and chronology.
5. Use this notebook as exploratory clustering before authorship claims.
