---
title: "45 - JSO Analytics: Nearest Neighbors and Ensemble Votes"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run from notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 500
)

library(dplyr)
library(stringr)
library(readr)
library(tibble)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)
```

## 1. Purpose

This notebook builds from `44_Mclust_all_BOM_BoC_DC1835_JSO.Rmd`.

The goal here is not merely clustering. The goal is to ask:

```text
For Joseph, Sidney, and Oliver writings, especially Sidney Rigdon,
which documents are nearest neighbors?
Which corpora/books receive the most votes?
What does the ensemble say?
```

Primary focus:

```text
Sidney Rigdon nearest neighbors
Sidney Rigdon vote counts
Sidney Rigdon consensus
```

This notebook assumes the project library already defines:

```r
run_stylometry_ensemble()
show_feature_votes()
show_consensus()
```

from the package/project source files.

## 2. Paths

```{r paths}
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
analytics_dir <- file.path(STYLO_ROOT, "export", "jso_analytics")

dir.create(analytics_dir, recursive = TRUE, showWarnings = FALSE)

kjv_path <- file.path(prepared_dir, "kjv_chapter_docs.csv")
bom_path <- file.path(prepared_dir, "bom_chapter_docs.csv")
boc_path <- file.path(prepared_dir, "boc_1833_chapter_docs.csv")
dc1835_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")
jso_path <- file.path(prepared_dir, "jso_manual_chapter_docs.csv")

out_sidney_neighbors <- file.path(analytics_dir, "sidney_nearest_neighbors.csv")
out_sidney_corpus_votes <- file.path(analytics_dir, "sidney_corpus_votes.csv")
out_sidney_book_votes <- file.path(analytics_dir, "sidney_book_votes.csv")
out_jso_author_votes <- file.path(analytics_dir, "jso_author_corpus_votes.csv")
```

## 3. Read prepared corpora

```{r read-corpora}
read_required <- function(path){
  if(!file.exists(path)){
    stop("Missing required file: ", path)
  }

  readr::read_csv(path, show_col_types = FALSE)
}

kjv_raw <- read_required(kjv_path)
bom_raw <- read_required(bom_path)
boc_raw <- read_required(boc_path)
dc1835_raw <- read_required(dc1835_path)
jso_raw <- read_required(jso_path)

list(
  kjv = dim(kjv_raw),
  bom = dim(bom_raw),
  boc = dim(boc_raw),
  dc1835 = dim(dc1835_raw),
  jso = dim(jso_raw)
)
```

## 4. Normalize document schemas

```{r normalize-docs}
clean_text_for_analysis <- function(x){
  x %>%
    as.character() %>%
    stringr::str_replace_all("\r", " ") %>%
    stringr::str_replace_all("\n", " ") %>%
    stringr::str_squish()
}

safe_chr <- function(df, col, default = NA_character_){
  if(col %in% names(df)) as.character(df[[col]]) else rep(default, nrow(df))
}

safe_int <- function(df, col, default = NA_integer_){
  if(col %in% names(df)) suppressWarnings(as.integer(df[[col]])) else rep(default, nrow(df))
}

normalize_general_docs <- function(df,
                                   corpus_id_default,
                                   translation_default,
                                   corpus_family_default,
                                   source_family_default,
                                   book_default,
                                   macro_genre_default){

  df %>%
    mutate(
      corpus_id = if("corpus_id" %in% names(.)) as.character(corpus_id) else corpus_id_default,
      translation = if("translation" %in% names(.)) as.character(translation) else translation_default,
      corpus_family = if("corpus_family" %in% names(.)) as.character(corpus_family) else corpus_family_default,
      source_family = if("source_family" %in% names(.)) as.character(source_family) else source_family_default,
      doc_id = if("doc_id" %in% names(.)) as.character(doc_id) else paste0(source_family_default, "__", stringr::str_pad(row_number(), 3, pad = "0")),
      doc_id_original = if("doc_id_original" %in% names(.)) as.character(doc_id_original) else doc_id,
      book = if("book" %in% names(.)) as.character(book) else if("work" %in% names(.)) as.character(work) else book_default,
      chapter = if("chapter" %in% names(.)) suppressWarnings(as.integer(chapter)) else row_number(),
      section = if("section" %in% names(.)) as.character(section) else if("section_label" %in% names(.)) as.character(section_label) else paste0(book, "_", chapter),
      macro_genre = if("macro_genre" %in% names(.)) as.character(macro_genre) else macro_genre_default,
      mosiah_priority = if("mosiah_priority" %in% names(.)) as.character(mosiah_priority) else NA_character_,
      mosiah_priority_label = if("mosiah_priority_label" %in% names(.)) as.character(mosiah_priority_label) else NA_character_,
      chapter_date = if("chapter_date" %in% names(.)) as.character(chapter_date) else NA_character_,
      header = if("header" %in% names(.)) as.character(header) else if("chapter_title" %in% names(.)) as.character(chapter_title) else paste0(book, " ", chapter),
      n_verses = if("n_verses" %in% names(.)) suppressWarnings(as.integer(n_verses)) else NA_integer_,
      text = clean_text_for_analysis(text)
    ) %>%
    select(
      corpus_id,
      translation,
      corpus_family,
      source_family,
      doc_id,
      doc_id_original,
      book,
      chapter,
      section,
      macro_genre,
      mosiah_priority,
      mosiah_priority_label,
      chapter_date,
      header,
      n_verses,
      text
    )
}

kjv_docs <- normalize_general_docs(
  kjv_raw,
  corpus_id_default = "KJV",
  translation_default = "KJV",
  corpus_family_default = "KJV",
  source_family_default = "KJV",
  book_default = "KJV",
  macro_genre_default = "Biblical"
)

bom_docs <- normalize_general_docs(
  bom_raw,
  corpus_id_default = "BOM",
  translation_default = "1830",
  corpus_family_default = "BOM",
  source_family_default = "BOM_1830",
  book_default = "Book of Mormon",
  macro_genre_default = "Modern_Scripture"
)

boc_docs <- normalize_general_docs(
  boc_raw,
  corpus_id_default = "BOC",
  translation_default = "1833",
  corpus_family_default = "BOC",
  source_family_default = "BOC_1833",
  book_default = "Book of Commandments",
  macro_genre_default = "Modern_Revelation"
)

dc1835_docs <- normalize_general_docs(
  dc1835_raw,
  corpus_id_default = "DC",
  translation_default = "1835",
  corpus_family_default = "DC",
  source_family_default = "DC_1835",
  book_default = "Doctrine and Covenants",
  macro_genre_default = "Modern_Revelation"
)

jso_docs <- jso_raw %>%
  mutate(
    author = as.character(author),
    author_code = as.character(author_code),
    corpus_id = "JSO",
    translation = "JSO",
    corpus_family = "JSO",
    source_family = "JSO",
    doc_id_original = as.character(doc_id),
    doc_id = paste0(author_code, "-", chapter),
    book = author,
    section = paste0(author_code, "-", chapter_original),
    macro_genre = "Early_LDS_Writing",
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    chapter_date = NA_character_,
    header = as.character(chapter_title),
    n_verses = NA_integer_,
    text = clean_text_for_analysis(text)
  ) %>%
  select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    doc_id,
    doc_id_original,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    chapter_date,
    header,
    n_verses,
    text,
    author,
    author_code,
    short_work,
    source_file,
    chapter_original,
    chapter_title
  )

analysis_docs <- bind_rows(
  kjv_docs,
  bom_docs,
  boc_docs,
  dc1835_docs,
  jso_docs %>%
    select(
      corpus_id,
      translation,
      corpus_family,
      source_family,
      doc_id,
      doc_id_original,
      book,
      chapter,
      section,
      macro_genre,
      mosiah_priority,
      mosiah_priority_label,
      chapter_date,
      header,
      n_verses,
      text
    )
) %>%
  mutate(
    text = clean_text_for_analysis(text),
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  filter(!is.na(text), text != "", n_words >= 50)

analysis_docs %>%
  dplyr::count(corpus_family, source_family, name = "n_docs") %>%
  arrange(corpus_family, source_family)

jso_docs %>%
  dplyr::count(author, author_code, name = "n_docs")
```

## 5. JSO author map

```{r jso-author-map}
jso_author_map <- jso_docs %>%
  transmute(
    doc_id,
    author,
    author_code,
    short_name = paste0(author_code, "-", chapter),
    chapter,
    chapter_original,
    chapter_title,
    source_file
  )

jso_author_map %>%
  arrange(author, chapter) %>%
  print(n = 100)
```

## 6. Create JSO ensemble

The source documents are JSO chapters. The target documents are all non-JSO documents:

```text
KJV + BOM + BOC + DC 1835
```

This prevents a JSO chapter from voting for another JSO chapter and focuses the comparison on the larger target corpus.

```{r ensemble-config}
source_docs_jso <- jso_docs %>%
  select(
    doc_id,
    corpus_id,
    translation,
    corpus_family,
    source_family,
    book,
    chapter,
    section,
    macro_genre,
    header,
    text
  )

target_docs_all_non_jso <- analysis_docs %>%
  filter(corpus_family != "JSO") %>%
  select(
    doc_id,
    corpus_id,
    translation,
    corpus_family,
    source_family,
    book,
    chapter,
    section,
    macro_genre,
    header,
    text
  )

source_docs_sidney <- source_docs_jso %>%
  semi_join(
    jso_author_map %>% filter(author == "Sidney Rigdon"),
    by = "doc_id"
  )

target_docs_all_non_jso %>%
  dplyr::count(corpus_family, source_family, name = "n_targets") %>%
  arrange(corpus_family, source_family)

source_docs_jso %>%
  dplyr::count(book, name = "n_sources")

source_docs_sidney %>%
  summarise(n_sidney_sources = n())
```

The configuration below is intentionally conservative and notebook-local. If your project already has a preferred ensemble config object, replace `jso_config` with that object.

```{r jso-config}
include_feature_classes <- c(
  "function_words", "punctuation", "sentence",
  "word_length", "lexical", "readability"
)


jso_config  <- default_config(include = include_feature_classes)


jso_config$weights <- c(
    function_words = 0.500,
    punctuation    = 0.075,
    sentence       = 0.175,
    word_length    = 0.125,
    lexical        = 0.075,
    readability    = 0.050
  )



```

```{r run-jso-ensemble}
result_jso <- run_stylometry_ensemble(
  source_docs = source_docs_jso,
  target_docs = target_docs_all_non_jso,
  config = jso_config,
  text_col = "text",
  doc_id_col = "doc_id",
  k_neighbors = 20,
  use_cache = TRUE,
  cache_dir = file.path(STYLO_CACHE, "ensemble")
)

names(result_jso)
```

```{r run-sidney-ensemble}
result_sidney <- run_stylometry_ensemble(
  source_docs = source_docs_sidney,
  target_docs = target_docs_all_non_jso,
  config = jso_config,
  text_col = "text",
  doc_id_col = "doc_id",
  k_neighbors = 20,
  use_cache = TRUE,
  cache_dir = file.path(STYLO_CACHE, "ensemble")
)

names(result_sidney)
```

## 7. Built-in display functions

These use your package/reporting layer.

```{r show-feature-votes-jso}
show_feature_votes(result_jso)
```

```{r show-consensus-jso}
show_consensus(result_jso)
```

```{r show-feature-votes-sidney}
show_feature_votes(result_sidney)
```

```{r show-consensus-sidney}
show_consensus(result_sidney)
```

## 8. Extract nearest neighbors from ensemble result

Different ensemble versions sometimes name the nearest-neighbor table slightly differently. This helper tries the common names.

```{r extract-nearest-helper}
extract_nearest_table <- function(result){

  candidates <- c(
    "nearest_neighbors",
    "nearest_targets",
    "feature_rankings",
    "feature_votes"
  )

  found <- candidates[candidates %in% names(result)]

  if(length(found) == 0){
    stop(
      "No nearest-neighbor-like table found. Available result names: ",
      paste(names(result), collapse = ", ")
    )
  }

  result[[found[[1]]]]
}

nn_jso_raw <- extract_nearest_table(result_jso)
nn_sidney_raw <- extract_nearest_table(result_sidney)

names(nn_jso_raw)
names(nn_sidney_raw)
```

## 9. Normalize nearest-neighbor table

This helper standardizes likely column names.

```{r normalize-nearest-table}
normalize_nearest_table <- function(nn){

  nm <- names(nn)

  source_col <- dplyr::case_when(
    "source_doc_id" %in% nm ~ "source_doc_id",
    "source_doc" %in% nm ~ "source_doc",
    "doc_id" %in% nm ~ "doc_id",
    "source_id" %in% nm ~ "source_id",
    TRUE ~ NA_character_
  )

  target_col <- dplyr::case_when(
    "target_doc_id" %in% nm ~ "target_doc_id",
    "target_doc" %in% nm ~ "target_doc",
    "neighbor_doc_id" %in% nm ~ "neighbor_doc_id",
    "neighbor_doc" %in% nm ~ "neighbor_doc",
    "nearest_doc_id" %in% nm ~ "nearest_doc_id",
    TRUE ~ NA_character_
  )

  sim_col <- dplyr::case_when(
    "similarity" %in% nm ~ "similarity",
    "score" %in% nm ~ "score",
    "avg_similarity" %in% nm ~ "avg_similarity",
    TRUE ~ NA_character_
  )

  rank_col <- dplyr::case_when(
    "rank" %in% nm ~ "rank",
    "neighbor_rank" %in% nm ~ "neighbor_rank",
    "target_rank" %in% nm ~ "target_rank",
    TRUE ~ NA_character_
  )

  if(is.na(source_col) || is.na(target_col)){
    stop(
      "Could not identify source/target columns. Columns are: ",
      paste(nm, collapse = ", ")
    )
  }

  out <- nn %>%
    transmute(
      source_doc = as.character(.data[[source_col]]),
      target_doc = as.character(.data[[target_col]]),
      similarity = if(!is.na(sim_col)) as.numeric(.data[[sim_col]]) else NA_real_,
      rank = if(!is.na(rank_col)) as.integer(.data[[rank_col]]) else NA_integer_
    )

  if(all(is.na(out$rank))){
    out <- out %>%
      group_by(source_doc) %>%
      arrange(desc(similarity), .by_group = TRUE) %>%
      mutate(rank = row_number()) %>%
      ungroup()
  }

  out
}

nn_jso <- normalize_nearest_table(nn_jso_raw)
nn_sidney <- normalize_nearest_table(nn_sidney_raw)

nn_jso %>% slice_head(n = 10)
nn_sidney %>% slice_head(n = 10)
```

## 10. Add metadata to nearest neighbors

```{r nearest-metadata}
target_meta <- target_docs_all_non_jso %>%
  transmute(
    target_doc = doc_id,
    target_corpus = corpus_family,
    target_source = source_family,
    target_book = book,
    target_chapter = chapter,
    target_section = section,
    target_genre = macro_genre,
    target_header = header
  )

source_meta_jso <- source_docs_jso %>%
  transmute(
    source_doc = doc_id,
    source_corpus = corpus_family,
    source_book = book,
    source_chapter = chapter,
    source_header = header
  ) %>%
  left_join(
    jso_author_map %>% select(source_doc = doc_id, author, author_code, short_name, chapter_original, chapter_title),
    by = "source_doc"
  )

nn_jso_labeled <- nn_jso %>%
  left_join(source_meta_jso, by = "source_doc") %>%
  left_join(target_meta, by = "target_doc") %>%
  arrange(author, source_doc, rank)

nn_sidney_labeled <- nn_sidney %>%
  left_join(source_meta_jso, by = "source_doc") %>%
  left_join(target_meta, by = "target_doc") %>%
  arrange(source_doc, rank)

nn_sidney_labeled %>%
  select(
    short_name,
    source_header,
    rank,
    target_corpus,
    target_book,
    target_chapter,
    target_header,
    similarity
  )

dplyr::glimpse(nn_sidney_labeled)

readr::write_csv(nn_sidney_labeled, out_sidney_neighbors)
```

## 11. Sidney vote counts by corpus

```{r sidney-corpus-votes}
sidney_corpus_votes <- nn_sidney_labeled %>%
  group_by(source_doc, short_name, source_header, target_corpus) %>%
  summarise(
    votes = n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    max_similarity = max(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(source_doc, short_name, source_header) %>%
  mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes)
  ) %>%
  ungroup() %>%
  arrange(short_name, desc(weighted_votes))

sidney_corpus_votes %>%
  print(n = 300)

readr::write_csv(sidney_corpus_votes, out_sidney_corpus_votes)
```

## 12. Sidney vote counts by book

```{r sidney-book-votes}
sidney_book_votes <- nn_sidney_labeled %>%
  group_by(source_doc, short_name, source_header, target_corpus, target_book) %>%
  summarise(
    votes = n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    max_similarity = max(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(source_doc, short_name, source_header) %>%
  mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes)
  ) %>%
  ungroup() %>%
  arrange(short_name, desc(weighted_votes))

sidney_book_votes %>%
  print(n = 500)

readr::write_csv(sidney_book_votes, out_sidney_book_votes)
```

## 13. Aggregate JSO author vote profile

```{r author-corpus-votes}
jso_author_corpus_votes <- nn_jso_labeled %>%
  group_by(author, author_code, target_corpus) %>%
  summarise(
    votes = n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    max_similarity = max(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(author, author_code) %>%
  mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes)
  ) %>%
  ungroup() %>%
  arrange(author, desc(weighted_votes))

jso_author_corpus_votes

readr::write_csv(jso_author_corpus_votes, out_jso_author_votes)
```

## 14. Plots: JSO author vote profile

```{r plot-author-corpus-votes, fig.width=9, fig.height=5}
ggplot(
  jso_author_corpus_votes,
  aes(
    x = author,
    y = weighted_pct_votes,
    fill = target_corpus
  )
) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(
    title = "Nearest-neighbor vote profile by JSO author",
    subtitle = "Top nearest targets from KJV, BOM, BOC, and DC 1835; weighted by similarity",
    x = "Author",
    y = "Weighted vote share",
    fill = "Neighbor corpus"
  )
```

## 15. Plots: Sidney vote profile by chapter

```{r plot-sidney-corpus-by-chapter, fig.width=12, fig.height=6}

sidney_corpus_votes <- sidney_corpus_votes %>%
  mutate(
    chapter_num = as.integer(stringr::str_extract(short_name, "\\d+$"))
  ) %>%
  mutate(
    short_name = factor(
      short_name,
      levels = short_name[order(chapter_num)] %>% unique()
    )
  )

sidney_corpus_votes <- sidney_corpus_votes %>%
  mutate(
    target_corpus = factor(
      target_corpus,
      levels = c("BOC", "DC", "KJV", "BOM")
    )
  )


ggplot(
  sidney_corpus_votes,
  aes(
    x = short_name,
    y = weighted_pct_votes,
    fill = target_corpus
  )
) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ) +
  labs(
    title = "Sidney Rigdon nearest-neighbor corpus votes by chapter",
    subtitle = "Each Sidney chapter's nearest-neighbor votes, weighted by similarity",
    x = "Sidney chapter",
    y = "Weighted vote share",
    fill = "Neighbor corpus"
  )



ggplot(
  sidney_corpus_votes,
  aes(
    x = short_name,
    y = weighted_pct_votes,
    fill = target_corpus
  )
) +
  geom_col() +
  scale_fill_manual(
    breaks = c("BOM", "KJV", "DC", "BOC"),
    values = c(
      "BOM" = "#7CAE00",   # green
      "KJV" = "#C77CFF",   # purple
      "DC"  = "#00BFC4",   # cyan
      "BOC" = "#F8766D"    # salmon/red
    ),
    drop = FALSE
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ) +
  labs(
    title = "Sidney Rigdon nearest-neighbor corpus votes by chapter",
    subtitle = "Each Sidney chapter's nearest-neighbor votes, weighted by similarity",
    x = "Sidney Chapter",
    y = "Weighted vote share",
    fill = "Neighbor corpus"
  )
```

## 16. Top Sidney target books

```{r plot-sidney-book-summary, fig.width=10, fig.height=7}
sidney_book_summary <- sidney_book_votes %>%
  group_by(target_corpus, target_book) %>%
  summarise(
    votes = sum(votes),
    weighted_votes = sum(weighted_votes),
    .groups = "drop"
  ) %>%
  arrange(desc(weighted_votes)) %>%
  slice_head(n = 25)

sidney_book_summary %>%
  print(n = 30)

ggplot(
  sidney_book_summary,
  aes(
    x = reorder(paste(target_corpus, target_book, sep = " / "), weighted_votes),
    y = weighted_votes,
    fill = target_corpus
  )
) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  labs(
    title = "Top target books for Sidney Rigdon nearest-neighbor votes",
    subtitle = "Weighted votes aggregated across Sidney chapters",
    x = "Target corpus / book",
    y = "Weighted votes",
    fill = "Corpus"
  )
```

## 17. Top individual Sidney nearest neighbors

```{r top-sidney-neighbors}
nn_sidney_labeled %>%
  group_by(source_doc, short_name, source_header) %>%
  arrange(rank, .by_group = TRUE) %>%
  slice_head(n = 10) %>%
  ungroup() %>%
  select(
    short_name,
    source_header,
    rank,
    target_corpus,
    target_book,
    target_chapter,
    target_header,
    similarity
  ) %>%
  print(n = 500)
```

## 18. Notes for interpretation

The ensemble vote tables are better evidence than the Mclust cluster labels alone because they ask a more direct question:

```text
For each JSO chapter, which known texts are closest?
```

Recommended interpretation order:

1. Read `show_feature_votes(result_sidney)`.
2. Read `show_consensus(result_sidney)`.
3. Inspect `sidney_corpus_votes`.
4. Inspect `sidney_book_votes`.
5. Inspect top individual nearest neighbors.
6. Compare Sidney's author-level vote profile with Joseph and Oliver.


```{r sidney chapter target book}
## ------------------------------------------------------------------
## Sidney Chapter × Target Book Matrix
## ------------------------------------------------------------------

sidney_book_matrix <- sidney_book_votes %>%
  dplyr::group_by(
    short_name,
    target_book
  ) %>%
  dplyr::summarise(
    weighted_votes = sum(weighted_votes),
    votes = sum(votes),
    .groups = "drop"
  )


book_order <- sidney_book_matrix %>%
  group_by(target_book) %>%
  summarise(
    total = sum(weighted_votes)
  ) %>%
  arrange(desc(total)) %>%
  pull(target_book)

sidney_book_matrix <- sidney_book_matrix %>%
  mutate(
    target_book = factor(
      target_book,
      levels = book_order
    )
  )

sidney_book_matrix

library(tidyr)

sidney_book_heat <- sidney_book_matrix %>%
  tidyr::pivot_wider(
    names_from = target_book,
    values_from = weighted_votes,
    values_fill = 0
  )

sidney_book_heat

ggplot(
  sidney_book_matrix,
  aes(
    x = target_book,
    y = short_name,
    fill = weighted_votes
  )
) +
  geom_tile(color = "white") +
  scale_fill_gradient(
    low = "white",
    high = "darkgreen"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  ) +
  labs(
    title = "Sidney Rigdon Chapter × Target Book",
    subtitle = "Weighted nearest-neighbor votes",
    x = "Target Book",
    y = "Sidney Chapter",
    fill = "Weighted\nVotes"
  )


```

```{r jacob 5}
analysis_docs %>%
  filter(
    book == "Jacob",
    chapter == 5
  ) %>%
  select(doc_id, book, chapter, header)


show_target_neighbors <- function(result, target_doc){

  nn <- result$nearest_neighbors

  nn %>%
    dplyr::filter(target_doc == !!target_doc) %>%
    dplyr::arrange(desc(similarity))
}


show_target_neighbors(
    result_jso,
    "Jacob_5"
)


show_target_votes <- function(result, target_doc, author_map){

  result$nearest_neighbors %>%
    dplyr::filter(neighbor_doc_id == target_doc) %>%
    dplyr::group_by(doc_id) %>%
    dplyr::summarise(
      votes = dplyr::n(),
      weighted_votes = sum(similarity),
      avg_similarity = mean(similarity),
      max_similarity = max(similarity),
      best_rank = min(rank),
      .groups = "drop"
    ) %>%
    dplyr::left_join(author_map, by = "doc_id") %>%
    dplyr::select(
      short_name,
      author,
      chapter_original,
      chapter_title,
      votes,
      weighted_votes,
      avg_similarity,
      best_rank
    ) %>%
    dplyr::arrange(dplyr::desc(weighted_votes))
}



show_target_votes(
    result_jso,
    "Jacob_5",
    jso_author_map
)

```




