---
title: "46 - Sidney Rigdon and the Mosiah Priority Hypothesis"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 500)

library(dplyr)
library(stringr)
library(readr)
library(tibble)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)
library(forcats)
library(mclust)
```

## 1. Purpose

This notebook tests a specific, falsifiable version of the **Sidney Rigdon contribution/editor hypothesis**:

```text
Sidney Rigdon's stylometric affinity should be strongest with MP-1
(Mosiah / Alma), moderate with MP-2 (Helaman through Moroni),
and weaker with MP-3 (1 Nephi through Words of Mormon).
```

Expected trend:

```text
MP-1 > MP-2 > MP-3
```

The notebook uses two frameworks:

1. **Mclust overlap** from the all-corpus analysis.
2. **Nearest-neighbor / ensemble votes** using `run_stylometry_ensemble()`, `show_feature_votes()`, and `show_consensus()`.

## 2. Paths

```{r paths}
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
export_dir <- file.path(STYLO_ROOT, "export", "sidney_mp_hypothesis")

dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)

kjv_path <- file.path(prepared_dir, "kjv_chapter_docs.csv")
bom_path <- file.path(prepared_dir, "bom_chapter_docs.csv")
boc_path <- file.path(prepared_dir, "boc_1833_chapter_docs.csv")
dc1835_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")
jso_path <- file.path(prepared_dir, "jso_manual_chapter_docs.csv")
corpus_1830_chapter_path <- file.path(prepared_dir, "corpus_1830_manual_chapter_docs.csv")
corpus_1830_book_path <- file.path(prepared_dir, "corpus_1830_manual_book_docs.csv")

out_mp_votes <- file.path(export_dir, "sidney_mp_votes.csv")
out_mp_author_votes <- file.path(export_dir, "jso_author_mp_votes.csv")
out_sidney_bom_neighbors <- file.path(export_dir, "sidney_bom_nearest_neighbors.csv")
out_cluster_mp <- file.path(export_dir, "mclust_cluster_mp_composition.csv")
```

## 3. Read corpora

```{r read-corpora}
read_required <- function(path){
  if(!file.exists(path)) stop("Missing required file: ", path)
  readr::read_csv(path, show_col_types = FALSE)
}

kjv_raw <- read_required(kjv_path)
bom_raw <- read_required(bom_path)
boc_raw <- read_required(boc_path)
dc1835_raw <- read_required(dc1835_path)
jso_raw <- read_required(jso_path)
corpus_1830_chapter_raw <- read_required(corpus_1830_chapter_path)
corpus_1830_book_raw <- read_required(corpus_1830_book_path)

list(
  kjv = dim(kjv_raw),
  bom = dim(bom_raw),
  boc = dim(boc_raw),
  dc1835 = dim(dc1835_raw),
  jso = dim(jso_raw),
  corpus_1830_chapter = dim(corpus_1830_chapter_raw),
  corpus_1830_book = dim(corpus_1830_book_raw)
)
```

## 4. Helpers

```{r helpers}
clean_text_for_analysis <- function(x){
  x %>%
    as.character() %>%
    stringr::str_replace_all("\r", " ") %>%
    stringr::str_replace_all("\n", " ") %>%
    stringr::str_squish()
}

make_safe_id <- function(x){
  x %>%
    as.character() %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")
}

assign_mosiah_priority <- function(book){
  dplyr::case_when(
    book %in% c("Mosiah", "Alma") ~ "MP-1",
    book %in% c("Helaman", "3 Nephi", "4 Nephi", "Mormon", "Ether", "Moroni") ~ "MP-2",
    book %in% c("1 Nephi", "2 Nephi", "Jacob", "Enos", "Jarom", "Omni", "Words of Mormon") ~ "MP-3",
    TRUE ~ NA_character_
  )
}

assign_mosiah_priority_label <- function(mp){
  dplyr::case_when(
    mp == "MP-1" ~ "MP-1: Mosiah-Alma",
    mp == "MP-2" ~ "MP-2: Helaman-Moroni",
    mp == "MP-3" ~ "MP-3: 1 Nephi-Words of Mormon",
    TRUE ~ NA_character_
  )
}

normalize_general_docs <- function(df, corpus_id_default, translation_default,
                                   corpus_family_default, source_family_default,
                                   book_default, macro_genre_default,
                                   doc_level_default = "chapter") {
  df %>%
    dplyr::mutate(
      corpus_id = if("corpus_id" %in% names(.)) as.character(corpus_id) else corpus_id_default,
      translation = if("translation" %in% names(.)) as.character(translation) else translation_default,
      corpus_family = if("corpus_family" %in% names(.)) as.character(corpus_family) else corpus_family_default,
      source_family = if("source_family" %in% names(.)) as.character(source_family) else source_family_default,
      doc_id = if("doc_id" %in% names(.)) as.character(doc_id) else paste0(source_family_default, "__", stringr::str_pad(row_number(), 3, pad = "0")),
      doc_id_original = if("doc_id_original" %in% names(.)) as.character(doc_id_original) else doc_id,
      book = if("book" %in% names(.)) as.character(book) else if("work" %in% names(.)) as.character(work) else book_default,
      chapter = if("chapter" %in% names(.)) suppressWarnings(as.integer(chapter)) else row_number(),
      section = if("section" %in% names(.)) as.character(section) else if("section_label" %in% names(.)) as.character(section_label) else paste0(book, "_", chapter),
      macro_genre = if("macro_genre" %in% names(.)) as.character(macro_genre) else macro_genre_default,
      mosiah_priority = if("mosiah_priority" %in% names(.)) as.character(mosiah_priority) else NA_character_,
      mosiah_priority_label = if("mosiah_priority_label" %in% names(.)) as.character(mosiah_priority_label) else NA_character_,
      chapter_date = if("chapter_date" %in% names(.)) as.character(chapter_date) else NA_character_,
      header = if("header" %in% names(.)) as.character(header) else if("chapter_title" %in% names(.)) as.character(chapter_title) else paste0(book, " ", chapter),
      n_verses = if("n_verses" %in% names(.)) suppressWarnings(as.integer(n_verses)) else NA_integer_,
      text = clean_text_for_analysis(text),
      doc_level = if("doc_level" %in% names(.)) as.character(doc_level) else doc_level_default
    ) %>%
    dplyr::select(
      corpus_id, translation, corpus_family, source_family,
      doc_id, doc_id_original, book, chapter, section, macro_genre,
      mosiah_priority, mosiah_priority_label, chapter_date, header,
      n_verses, text, doc_level
    )
}
```

## 5. Normalize corpora

```{r normalize-corpora}
kjv_docs <- normalize_general_docs(kjv_raw, "KJV", "KJV", "KJV", "KJV", "KJV", "Biblical")

bom_docs <- normalize_general_docs(bom_raw, "BOM", "1830", "BOM", "BOM_1830", "Book of Mormon", "Modern_Scripture") %>%
  dplyr::mutate(
    mosiah_priority = dplyr::if_else(
      is.na(mosiah_priority) | mosiah_priority == "",
      assign_mosiah_priority(book),
      mosiah_priority
    ),
    mosiah_priority_label = dplyr::if_else(
      is.na(mosiah_priority_label) | mosiah_priority_label == "",
      assign_mosiah_priority_label(mosiah_priority),
      mosiah_priority_label
    )
  )

boc_docs <- normalize_general_docs(boc_raw, "BOC", "1833", "BOC", "BOC_1833", "Book of Commandments", "Modern_Revelation")
dc1835_docs <- normalize_general_docs(dc1835_raw, "DC", "1835", "DC", "DC_1835", "Doctrine and Covenants", "Modern_Revelation")

jso_docs <- jso_raw %>%
  dplyr::mutate(
    author = as.character(author),
    author_code = as.character(author_code),
    corpus_id = "JSO",
    translation = "JSO",
    corpus_family = "JSO",
    source_family = "JSO",
    doc_id_original = as.character(doc_id),
    doc_id = paste0(author_code, "-", chapter),
    book = author,
    section = paste0(author_code, "-", chapter_original),
    macro_genre = "Early_LDS_Writing",
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    chapter_date = NA_character_,
    header = as.character(chapter_title),
    n_verses = NA_integer_,
    text = clean_text_for_analysis(text),
    doc_level = "chapter"
  ) %>%
  dplyr::select(
    corpus_id, translation, corpus_family, source_family,
    doc_id, doc_id_original, book, chapter, section, macro_genre,
    mosiah_priority, mosiah_priority_label, chapter_date, header,
    n_verses, text, doc_level,
    author, author_code, short_work, source_file, chapter_original, chapter_title
  )

corpus_1830_chapter_docs <- corpus_1830_chapter_raw %>%
  dplyr::mutate(
    corpus_id = "1830",
    translation = "1830",
    corpus_family = "1830",
    source_family = "1830_Secular",
    doc_id_original = as.character(doc_id),
    doc_id = paste0("1830_CH__", make_safe_id(author), "__", make_safe_id(short_work), "__", stringr::str_pad(chapter, 3, pad = "0")),
    book = paste(author, work, sep = ": "),
    section = paste0(short_work, "-", chapter_original),
    macro_genre = as.character(macro_genre),
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    chapter_date = NA_character_,
    header = as.character(chapter_title),
    n_verses = NA_integer_,
    text = clean_text_for_analysis(text),
    doc_level = "chapter"
  ) %>%
  dplyr::select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original, book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label, chapter_date, header, n_verses, text, doc_level)

corpus_1830_book_docs <- corpus_1830_book_raw %>%
  dplyr::mutate(
    corpus_id = "1830",
    translation = "1830",
    corpus_family = "1830",
    source_family = "1830_Secular_Book",
    doc_id_original = as.character(doc_id),
    doc_id = paste0("1830_BOOK__", make_safe_id(author), "__", make_safe_id(short_work)),
    book = paste(author, work, sep = ": "),
    chapter = as.integer(row_number()),
    section = "BOOK",
    macro_genre = as.character(macro_genre),
    mosiah_priority = NA_character_,
    mosiah_priority_label = NA_character_,
    chapter_date = NA_character_,
    header = paste(author, work, sep = ": "),
    n_verses = NA_integer_,
    text = clean_text_for_analysis(text),
    doc_level = "book"
  ) %>%
  dplyr::select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original, book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label, chapter_date, header, n_verses, text, doc_level)

analysis_docs <- dplyr::bind_rows(
  kjv_docs,
  bom_docs,
  boc_docs,
  dc1835_docs,
  jso_docs %>% dplyr::select(corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original, book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label, chapter_date, header, n_verses, text, doc_level),
  corpus_1830_chapter_docs,
  corpus_1830_book_docs
) %>%
  dplyr::mutate(
    text = clean_text_for_analysis(text),
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::filter(!is.na(text), text != "", n_words >= 50)

analysis_docs %>%
  dplyr::count(corpus_family, source_family, macro_genre, doc_level, name = "n_docs") %>%
  dplyr::arrange(corpus_family, source_family, macro_genre, doc_level) %>%
  print(n = 200)

bom_docs %>%
  dplyr::count(mosiah_priority, mosiah_priority_label, book, name = "n_chapters") %>%
  dplyr::arrange(mosiah_priority, book) %>%
  print(n = 100)

jso_docs %>%
  dplyr::count(author, author_code, name = "n_docs")
```

## 6. JSO author map

```{r jso-author-map}
jso_author_map <- jso_docs %>%
  dplyr::transmute(
    doc_id,
    author,
    author_code,
    short_name = paste0(author_code, "-", chapter),
    chapter,
    chapter_original,
    chapter_title,
    source_file
  )

jso_author_map %>%
  dplyr::arrange(author, chapter) %>%
  print(n = 100)
```

## 7. Framework A: Mclust MP composition

This section uses an existing Mclust classified object when available.

```{r mclust-classified-setup}
if(exists("classified")){
  classified_mp <- classified
} else if(exists("mc_all_boc_dc1835_jso_1830") && "classified" %in% names(mc_all_boc_dc1835_jso_1830)){
  classified_mp <- mc_all_boc_dc1835_jso_1830$classified
} else {
  classified_mp <- NULL
  warning("No classified Mclust object found. Mclust framework chunks will be skipped unless classified_mp is supplied.")
}

if(!is.null(classified_mp)){
  classified_mp <- classified_mp %>%
    dplyr::mutate(
      mosiah_priority = dplyr::if_else(
        corpus_family == "BOM" & (is.na(mosiah_priority) | mosiah_priority == ""),
        assign_mosiah_priority(book),
        as.character(mosiah_priority)
      ),
      mosiah_priority_label = dplyr::if_else(
        corpus_family == "BOM" & (is.na(mosiah_priority_label) | mosiah_priority_label == ""),
        assign_mosiah_priority_label(mosiah_priority),
        as.character(mosiah_priority_label)
      )
    )

  classified_mp %>%
    dplyr::count(corpus_family, source_family, name = "n_docs") %>%
    dplyr::arrange(corpus_family, source_family) %>%
    print(n = 100)
}
```

```{r mclust-mp-composition}
if(!is.null(classified_mp)){

  cluster_mp <- classified_mp %>%
    dplyr::filter(corpus_family %in% c("BOM", "JSO")) %>%
    dplyr::mutate(
      group_label = dplyr::case_when(
        corpus_family == "BOM" ~ mosiah_priority_label,
        corpus_family == "JSO" ~ book,
        TRUE ~ corpus_family
      )
    ) %>%
    dplyr::count(mclust_cluster, group_label, name = "n_docs") %>%
    dplyr::group_by(mclust_cluster) %>%
    dplyr::mutate(pct_cluster = n_docs / sum(n_docs)) %>%
    dplyr::ungroup() %>%
    dplyr::arrange(mclust_cluster, dplyr::desc(n_docs))

  cluster_mp %>% print(n = 200)
  readr::write_csv(cluster_mp, out_cluster_mp)
}
```

```{r sidney-cluster-overlap}
if(!is.null(classified_mp)){

  sidney_clusters <- classified_mp %>%
    dplyr::filter(corpus_family == "JSO") %>%
    dplyr::mutate(
      author_code = stringr::str_extract(doc_id, "^[A-Z]+"),
      author = dplyr::case_when(
        author_code == "JS" ~ "Joseph Smith",
        author_code == "SR" ~ "Sidney Rigdon",
        author_code == "OC" ~ "Oliver Cowdery",
        TRUE ~ "Unknown"
      )
    ) %>%
    dplyr::filter(author == "Sidney Rigdon") %>%
    dplyr::count(mclust_cluster, name = "sidney_docs")

  bom_mp_clusters <- classified_mp %>%
    dplyr::filter(corpus_family == "BOM", !is.na(mosiah_priority_label)) %>%
    dplyr::count(mclust_cluster, mosiah_priority, mosiah_priority_label, name = "bom_docs") %>%
    dplyr::group_by(mclust_cluster) %>%
    dplyr::mutate(
      bom_cluster_total = sum(bom_docs),
      pct_bom_cluster = bom_docs / bom_cluster_total
    ) %>%
    dplyr::ungroup()

  sidney_mp_cluster_overlap <- bom_mp_clusters %>%
    dplyr::left_join(sidney_clusters, by = "mclust_cluster") %>%
    dplyr::mutate(
      sidney_docs = tidyr::replace_na(sidney_docs, 0L),
      overlap_score = sidney_docs * pct_bom_cluster
    ) %>%
    dplyr::arrange(dplyr::desc(overlap_score), mclust_cluster)

  sidney_mp_cluster_overlap %>% print(n = 100)
}
```

```{r plot-mclust-mp-groups, fig.width=11, fig.height=6}
if(!is.null(classified_mp)){

  mclust_mp_plot_df <- classified_mp %>%
    dplyr::filter(
      corpus_family == "JSO" |
        (corpus_family == "BOM" & !is.na(mosiah_priority_label))
    ) %>%
    dplyr::mutate(
      plot_group = dplyr::case_when(
        corpus_family == "JSO" & stringr::str_detect(doc_id, "^SR") ~ "Sidney Rigdon",
        corpus_family == "JSO" & stringr::str_detect(doc_id, "^JS") ~ "Joseph Smith",
        corpus_family == "JSO" & stringr::str_detect(doc_id, "^OC") ~ "Oliver Cowdery",
        corpus_family == "BOM" ~ mosiah_priority_label,
        TRUE ~ corpus_family
      )
    ) %>%
    dplyr::count(mclust_cluster, plot_group, name = "n_docs")

  ggplot(
    mclust_mp_plot_df,
    aes(x = factor(mclust_cluster), y = n_docs, fill = plot_group)
  ) +
    geom_col() +
    theme_bw() +
    labs(
      title = "Mclust clusters: JSO authors and Book of Mormon Mosiah Priority groups",
      subtitle = "Testing whether Sidney's clusters overlap most with MP-1, then MP-2, then MP-3",
      x = "Mclust cluster",
      y = "Document count",
      fill = "Group"
    )
}
```

## 8. Framework B: ensemble nearest-neighbor votes

This is the primary analysis. The target set is restricted to BOM chapters so every vote is interpretable as MP-1, MP-2, or MP-3.

```{r ensemble-source-target}
source_docs_jso <- jso_docs %>%
  dplyr::select(doc_id, corpus_id, translation, corpus_family, source_family, book, chapter, section, macro_genre, header, text)

source_docs_sidney <- source_docs_jso %>%
  dplyr::semi_join(jso_author_map %>% dplyr::filter(author == "Sidney Rigdon"), by = "doc_id")

target_docs_bom <- bom_docs %>%
  dplyr::filter(!is.na(mosiah_priority)) %>%
  dplyr::select(doc_id, corpus_id, translation, corpus_family, source_family, book, chapter, section, macro_genre, mosiah_priority, mosiah_priority_label, header, text)

source_docs_jso %>% dplyr::count(book, name = "n_source_docs")
target_docs_bom %>% dplyr::count(mosiah_priority, mosiah_priority_label, name = "n_target_docs")
```

## 9. Feature classes

```{r feature-classes}
include_feature_classes <- c(
  "function_words", "punctuation", "sentence",
  "word_length", "lexical", "readability"
)


jso_config  <- default_config(include = include_feature_classes)


jso_config$weights <- c(
    function_words = 0.500,
    punctuation    = 0.075,
    sentence       = 0.175,
    word_length    = 0.125,
    lexical        = 0.075,
    readability    = 0.050
  )
```

## 10. Run ensemble: JSO to BOM

```{r run-jso-to-bom-ensemble}
result_jso_bom <- run_stylometry_ensemble(
  source_docs = source_docs_jso,
  target_docs = target_docs_bom,
  config = jso_config,
  text_col = "text",
  doc_id_col = "doc_id",
  k_neighbors = 20,
  use_cache = TRUE
)

names(result_jso_bom)
```

## 11. Run ensemble: Sidney to BOM

```{r run-sidney-to-bom-ensemble}
result_sidney_bom <- run_stylometry_ensemble(
  source_docs = source_docs_sidney,
  target_docs = target_docs_bom,
  config = jso_config,
  text_col = "text",
  doc_id_col = "doc_id",
  k_neighbors = 20,
  use_cache = TRUE
)

names(result_sidney_bom)
```

## 12. Built-in ensemble reports

```{r show-feature-votes-jso-bom}
show_feature_votes(result_jso_bom)
```

```{r show-consensus-jso-bom}
show_consensus(result_jso_bom)
```

```{r show-feature-votes-sidney-bom}
show_feature_votes(result_sidney_bom)
```

```{r show-consensus-sidney-bom}
show_consensus(result_sidney_bom)
```

## 13. Nearest-neighbor helpers

```{r nearest-helpers}
get_nearest_neighbors <- function(result){

  if(!"nearest_neighbors" %in% names(result)){
    stop("result does not contain nearest_neighbors. Available names: ", paste(names(result), collapse = ", "))
  }

  result$nearest_neighbors %>%
    dplyr::mutate(
      doc_id = as.character(doc_id),
      neighbor_doc_id = as.character(neighbor_doc_id),
      similarity = as.numeric(similarity),
      rank = as.integer(rank)
    )
}

label_nn_with_bom_mp <- function(nn, source_map, target_bom){
  nn %>%
    dplyr::left_join(
      source_map %>% dplyr::select(doc_id, author, author_code, short_name, chapter_title),
      by = "doc_id"
    ) %>%
    dplyr::left_join(
      target_bom %>%
        dplyr::select(
          neighbor_doc_id = doc_id,
          target_book = book,
          target_chapter = chapter,
          target_header = header,
          mosiah_priority,
          mosiah_priority_label
        ),
      by = "neighbor_doc_id"
    )
}
```

```{r label-nearest}
nn_jso_bom <- get_nearest_neighbors(result_jso_bom) %>%
  label_nn_with_bom_mp(jso_author_map, target_docs_bom)

nn_sidney_bom <- get_nearest_neighbors(result_sidney_bom) %>%
  label_nn_with_bom_mp(jso_author_map, target_docs_bom)

nn_sidney_bom %>%
  dplyr::select(short_name, chapter_title, rank, neighbor_doc_id, target_book, target_chapter, mosiah_priority, similarity) %>%
  dplyr::arrange(short_name, rank) 

head(nn_sidney_bom)

readr::write_csv(nn_sidney_bom, out_sidney_bom_neighbors)
```

## 14. Test 1: Sidney votes by Mosiah Priority

```{r sidney-mp-votes}
sidney_mp_votes <- nn_sidney_bom %>%
  dplyr::group_by(mosiah_priority, mosiah_priority_label) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    max_similarity = max(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes),
    mp_order = match(mosiah_priority, c("MP-1", "MP-2", "MP-3"))
  ) %>%
  dplyr::arrange(mp_order)

sidney_mp_votes

readr::write_csv(sidney_mp_votes, out_mp_votes)
```

```{r plot-sidney-mp-gradient, fig.width=8, fig.height=5}
ggplot(
  sidney_mp_votes,
  aes(
    x = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3")),
    y = weighted_pct_votes
  )
) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(
    title = "Sidney Rigdon Book of Mormon nearest-neighbor votes by Mosiah Priority",
    subtitle = "Hypothesis prediction: MP-1 > MP-2 > MP-3",
    x = "Mosiah Priority group",
    y = "Weighted vote share"
  )
```

## 15. Test 2: Sidney chapter-level MP heatmap

```{r sidney-chapter-mp-heatmap-data}
sidney_chapter_mp <- nn_sidney_bom %>%
  dplyr::group_by(short_name, chapter_title, mosiah_priority, mosiah_priority_label) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(short_name, chapter_title) %>%
  dplyr::mutate(weighted_pct_votes = weighted_votes / sum(weighted_votes)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    chapter_num = as.integer(stringr::str_extract(short_name, "\\d+$")),
    short_name = factor(short_name, levels = paste0("SR-", sort(unique(chapter_num)))),
    mosiah_priority = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3"))
  )

sidney_chapter_mp %>%
  dplyr::arrange(chapter_num, mosiah_priority) %>%
  print(n = 200)
```

```{r plot-sidney-chapter-mp-heatmap, fig.width=8, fig.height=10}
ggplot(
  sidney_chapter_mp,
  aes(x = mosiah_priority, y = fct_rev(short_name), fill = weighted_pct_votes)
) +
  geom_tile(color = "grey85") +
  scale_fill_gradient(low = "white", high = "darkgreen", labels = scales::percent) +
  theme_bw() +
  labs(
    title = "Sidney Rigdon chapters × Mosiah Priority groups",
    subtitle = "Row-normalized weighted nearest-neighbor vote share",
    x = "Book of Mormon group",
    y = "Sidney chapter",
    fill = "Weighted\nshare"
  )
```

## 16. Test 3: Sidney votes by Book of Mormon book

```{r sidney-bom-book-votes}
sidney_bom_book_votes <- nn_sidney_bom %>%
  dplyr::group_by(target_book, mosiah_priority, mosiah_priority_label) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(weighted_pct_votes = weighted_votes / sum(weighted_votes)) %>%
  dplyr::arrange(dplyr::desc(weighted_votes))

sidney_bom_book_votes %>% print(n = 100)
```

```{r plot-sidney-bom-book-votes, fig.width=10, fig.height=7}
ggplot(
  sidney_bom_book_votes,
  aes(x = reorder(target_book, weighted_votes), y = weighted_votes, fill = mosiah_priority)
) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  labs(
    title = "Sidney Rigdon nearest-neighbor votes by Book of Mormon book",
    subtitle = "Weighted votes, colored by Mosiah Priority",
    x = "Book of Mormon book",
    y = "Weighted votes",
    fill = "Mosiah Priority"
  )
```

## 17. Test 4: JSO author comparison

```{r author-mp-votes}
jso_author_mp_votes <- nn_jso_bom %>%
  dplyr::group_by(author, author_code, mosiah_priority, mosiah_priority_label) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(author, author_code) %>%
  dplyr::mutate(
    pct_votes = votes / sum(votes),
    weighted_pct_votes = weighted_votes / sum(weighted_votes),
    mp_order = match(mosiah_priority, c("MP-1", "MP-2", "MP-3"))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(author, mp_order)

jso_author_mp_votes

readr::write_csv(jso_author_mp_votes, out_mp_author_votes)
```

```{r plot-author-mp-comparison, fig.width=9, fig.height=5}
ggplot(
  jso_author_mp_votes,
  aes(
    x = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3")),
    y = weighted_pct_votes,
    fill = author
  )
) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(
    title = "JSO author comparison: BOM nearest-neighbor votes by Mosiah Priority",
    subtitle = "Does Sidney show the strongest MP-1 > MP-2 > MP-3 gradient?",
    x = "Mosiah Priority group",
    y = "Weighted vote share",
    fill = "Author"
  )
```

## 18. Statistical trend test

```{r trend-statistics}
sidney_trend <- sidney_mp_votes %>%
  dplyr::mutate(
    mp_numeric = dplyr::case_when(
      mosiah_priority == "MP-1" ~ 1,
      mosiah_priority == "MP-2" ~ 2,
      mosiah_priority == "MP-3" ~ 3,
      TRUE ~ NA_real_
    )
  )

sidney_trend

cor.test(
  sidney_trend$mp_numeric,
  sidney_trend$weighted_pct_votes,
  method = "spearman",
  exact = FALSE
)
```

## 19. Enrichment adjustment by target group size

```{r mp-enrichment}
mp_target_sizes <- target_docs_bom %>%
  dplyr::count(mosiah_priority, mosiah_priority_label, name = "target_chapters") %>%
  dplyr::mutate(expected_share = target_chapters / sum(target_chapters))

sidney_mp_enrichment <- sidney_mp_votes %>%
  dplyr::left_join(mp_target_sizes, by = c("mosiah_priority", "mosiah_priority_label")) %>%
  dplyr::mutate(enrichment = weighted_pct_votes / expected_share) %>%
  dplyr::arrange(mp_order)

sidney_mp_enrichment
```

```{r plot-mp-enrichment, fig.width=8, fig.height=5}
ggplot(
  sidney_mp_enrichment,
  aes(
    x = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3")),
    y = enrichment
  )
) +
  geom_col() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  labs(
    title = "Sidney Rigdon MP nearest-neighbor enrichment",
    subtitle = "Values above 1 mean more votes than expected from target group size",
    x = "Mosiah Priority group",
    y = "Enrichment ratio"
  )
```

## 20. Interpretation guide

The hypothesis is supported if:

```text
1. Sidney's weighted vote share is highest for MP-1.
2. Sidney's MP-1 enrichment is > 1.
3. Sidney's MP-1 share is stronger than Joseph and Oliver.
4. The pattern is visible across many Sidney chapters.
5. show_feature_votes(result_sidney_bom) and show_consensus(result_sidney_bom)
   repeatedly point to MP-1 targets.
```

The hypothesis is weakened if:

```text
1. MP-1, MP-2, and MP-3 shares are approximately equal.
2. MP-3 is as strong or stronger than MP-1.
3. Joseph or Oliver show the same gradient with equal or greater strength.
4. The MP-1 signal is driven by a small number of Sidney chapters.
```


```{r book-size norm}
## 17b. Normalize Sidney book votes by Book of Mormon book size

bom_book_sizes <- target_docs_bom %>%
  dplyr::count(
    book,
    mosiah_priority,
    mosiah_priority_label,
    name = "target_chapters"
  ) %>%
  dplyr::mutate(
    expected_share = target_chapters / sum(target_chapters)
  )

sidney_bom_book_enrichment <- sidney_bom_book_votes %>%
  dplyr::left_join(
    bom_book_sizes,
    by = c(
      "target_book" = "book",
      "mosiah_priority",
      "mosiah_priority_label"
    )
  ) %>%
  dplyr::mutate(
    observed_share = weighted_votes / sum(weighted_votes),
    enrichment = observed_share / expected_share
  ) %>%
  dplyr::arrange(dplyr::desc(enrichment))

sidney_bom_book_enrichment %>%
  dplyr::select(
    target_book,
    mosiah_priority,
    target_chapters,
    weighted_votes,
    observed_share,
    expected_share,
    enrichment
  ) %>%
  print(n = 100)

ggplot(
  sidney_bom_book_enrichment,
  aes(
    x = reorder(target_book, enrichment),
    y = enrichment,
    fill = mosiah_priority
  )
) +
  geom_col() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  theme_bw() +
  labs(
    title = "Sidney Rigdon Book of Mormon vote enrichment by book",
    subtitle = "Observed weighted vote share divided by expected share from book size",
    x = "Book of Mormon book",
    y = "Enrichment ratio",
    fill = "Mosiah Priority"
  )


### standardized residual Pearson residual
sidney_book_votes <- nn_sidney_bom %>%
  dplyr::group_by(target_book) %>%
  dplyr::summarise(
    observed = sum(similarity),
    .groups = "drop"
  )

bom_book_sizes <- target_docs_bom %>%
  dplyr::count(
    book,
    name = "chapters"
  )

total_votes <- sum(sidney_book_votes$observed)

expected_votes <- bom_book_sizes %>%
  dplyr::mutate(
    expected = total_votes * chapters / sum(chapters)
  )

sidney_stdres <- expected_votes %>%
  dplyr::left_join(
    sidney_book_votes,
    by = c("book" = "target_book")
  ) %>%
  dplyr::mutate(
    observed = tidyr::replace_na(observed, 0),
    stdres = (observed - expected) / sqrt(expected)
  ) %>%
  dplyr::arrange(desc(stdres))

sidney_stdres

ggplot(
  sidney_stdres,
  aes(
    x = reorder(book, stdres),
    y = stdres,
    fill = stdres > 0
  )
) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, color = "black") +
  geom_hline(yintercept = c(-2, 2),
             linetype = "dashed",
             color = "red") +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Sidney Rigdon standardized residuals by Book of Mormon book",
    subtitle = "Observed weighted nearest-neighbor votes versus expected from book size",
    x = "Book of Mormon book",
    y = "Standardized residual (O − E) / √E"
  )


sidney_stdres <- expected_votes %>%
  left_join(
    sidney_book_votes,
    by = c("book" = "target_book")
  ) %>%
  mutate(
    observed = replace_na(observed, 0),
    residual = observed - expected,
    stdres = residual / sqrt(expected * (1 - chapters / sum(chapters)))
  ) %>%
  arrange(desc(stdres))


sidney_counts <- nn_sidney_bom %>%
  dplyr::count(
    target_book,
    name = "observed"
  )

book_sizes <- target_docs_bom %>%
  dplyr::count(
    book,
    name = "chapters"
  ) %>%
  dplyr::mutate(
    expected = sum(sidney_counts$observed) *
      chapters / sum(chapters)
  )

sidney_residuals <- book_sizes %>%
  left_join(
    sidney_counts,
    by = c("book" = "target_book")
  ) %>%
  mutate(
    observed = tidyr::replace_na(observed, 0L),
    residual = observed - expected,
    pearson = residual / sqrt(expected),
    stdres = residual /
      sqrt(expected * (1 - chapters / sum(chapters)))
  ) %>%
  arrange(desc(stdres))

sidney_residuals

chisq_sidney <- chisq.test(
  x = sidney_residuals$observed,
  p = sidney_residuals$chapters /
      sum(sidney_residuals$chapters)
)

chisq_sidney


sidney_residuals$chisq_stdres <- chisq_sidney$stdres
sidney_residuals$chisq_residual <- chisq_sidney$residuals

sidney_residuals %>%
  arrange(desc(chisq_stdres))


export_root <- file.path(STYLO_ROOT, "export")

sidney_dir <- file.path(export_root, "sidney_mp_hypothesis")

sidney_resid_path <- file.path(sidney_dir, "sidney_bom_book_chisq_residuals.csv")

readr::write_csv(
  sidney_residuals,
  sidney_resid_path
)


ggplot(
  sidney_residuals,
  aes(
    x = reorder(book, chisq_stdres),
    y = chisq_stdres,
    fill = chisq_stdres > 0
  )
) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0) +
  geom_hline(
    yintercept = c(-2, 2),
    linetype = "dashed",
    color = "red"
  ) +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Sidney Rigdon standardized residuals by Book of Mormon book",
    subtitle = "Pearson standardized residuals from chi-square goodness-of-fit",
    x = "Book of Mormon book",
    y = "Standardized residual"
  )


```
```{r block to match Oliver and Joseph}
sidney_counts <- nn_sidney_bom %>%
  dplyr::count(
    target_book,
    name = "observed"
  )

book_sizes <- target_docs_bom %>%
  dplyr::count(
    book,
    mosiah_priority,
    mosiah_priority_label,
    name = "chapters"
  )

sidney_residuals <- book_sizes %>%
  dplyr::left_join(
    sidney_counts,
    by = c("book" = "target_book")
  ) %>%
  dplyr::mutate(
    observed = tidyr::replace_na(observed, 0L)
  ) %>%
  dplyr::arrange(book)

chisq_sidney <- chisq.test(
  x = sidney_residuals$observed,
  p = sidney_residuals$chapters / sum(sidney_residuals$chapters)
)

chisq_sidney

sidney_residuals <- sidney_residuals %>%
  dplyr::mutate(
    expected = as.numeric(chisq_sidney$expected),
    pearson_residual = as.numeric(chisq_sidney$residuals),
    chisq_stdres = as.numeric(chisq_sidney$stdres),
    residual = observed - expected,
    observed_share = observed / sum(observed),
    expected_share = chapters / sum(chapters),
    enrichment = observed_share / expected_share
  ) %>%
  dplyr::arrange(dplyr::desc(chisq_stdres))

sidney_residuals %>%
  dplyr::select(
    book,
    mosiah_priority,
    mosiah_priority_label,
    chapters,
    observed,
    expected,
    pearson_residual,
    chisq_stdres,
    residual,
    observed_share,
    expected_share,
    enrichment
  ) %>%
  print(n = 100)




sidney_dir <- file.path(export_root, "sidney_mp_hypothesis")

sidney_resid_path <- file.path(sidney_dir, "sidney_bom_book_chisq_residuals.csv")

readr::write_csv(
  sidney_residuals,
  sidney_resid_path
)



```


## 21. Save session objects

```{r save-rds}
saveRDS(
  list(
    result_jso_bom = result_jso_bom,
    result_sidney_bom = result_sidney_bom,
    nn_jso_bom = nn_jso_bom,
    nn_sidney_bom = nn_sidney_bom,
    sidney_mp_votes = sidney_mp_votes,
    sidney_mp_enrichment = sidney_mp_enrichment,
    jso_author_mp_votes = jso_author_mp_votes
  ),
  file.path(export_dir, "sidney_mp_hypothesis_objects.rds")
)
```
