---
title: "49 - Compare Joseph, Sidney, and Oliver Mosiah Priority Profiles"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 500
)

library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(tibble)
library(ggplot2)
library(scales)
library(forcats)
library(purrr)
```

## 1. Purpose

This notebook compares the Mosiah Priority affinity profiles for:

```text
Joseph Smith
Sidney Rigdon
Oliver Cowdery
```

The prior notebooks produced author-specific chi-square residual profiles:

```text
46_Sidney_MP_hypothesis.Rmd
47_Oliver_MP_hypothesis.Rmd
48_Joseph_MP_hypothesis.Rmd
```

This notebook compares those profiles directly using:

1. Standardized residual heatmap by Book of Mormon book.
2. Side-by-side residual bar chart.
3. Author-to-author residual correlations.
4. Difference profiles such as Sidney minus Joseph.
5. Mosiah Priority aggregate comparison.

## 2. Paths

```{r paths}
export_root <- file.path(STYLO_ROOT, "export")

sidney_dir <- file.path(export_root, "sidney_mp_hypothesis")
oliver_dir <- file.path(export_root, "oliver_mp_hypothesis")
joseph_dir <- file.path(export_root, "joseph_mp_hypothesis")

sidney_resid_path <- file.path(sidney_dir, "sidney_bom_book_chisq_residuals.csv")
oliver_resid_path <- file.path(oliver_dir, "oliver_bom_book_chisq_residuals.csv")
joseph_resid_path <- file.path(joseph_dir, "joseph_bom_book_chisq_residuals.csv")

sidney_mp_path <- file.path(sidney_dir, "sidney_mp_votes.csv")
oliver_mp_path <- file.path(oliver_dir, "oliver_mp_votes.csv")
joseph_mp_path <- file.path(joseph_dir, "joseph_mp_votes.csv")

compare_dir <- file.path(export_root, "compare_js_sr_oc_mp")
dir.create(compare_dir, recursive = TRUE, showWarnings = FALSE)
```

## 3. Read author residual profiles

```{r read-residuals}
required_files <- c(
  sidney_resid_path,
  oliver_resid_path,
  joseph_resid_path
)

missing_files <- required_files[!file.exists(required_files)]

if(length(missing_files) > 0){
  stop(
    "Missing residual CSV(s). Run notebooks 46, 47, and 48 first. Missing:\n",
    paste(missing_files, collapse = "\n")
  )
}

read_author_residuals <- function(path, author, author_code){

  readr::read_csv(path, show_col_types = FALSE) %>%
    dplyr::mutate(
      author = author,
      author_code = author_code,
      book = as.character(book),
      mosiah_priority = as.character(mosiah_priority),
      mosiah_priority_label = as.character(mosiah_priority_label),
      chisq_stdres = as.numeric(chisq_stdres),
      pearson_residual = as.numeric(pearson_residual),
      observed = as.numeric(observed),
      expected = as.numeric(expected),
      enrichment = as.numeric(enrichment)
    )

}

author_residuals <- dplyr::bind_rows(
  read_author_residuals(sidney_resid_path, "Sidney Rigdon", "SR"),
  read_author_residuals(oliver_resid_path, "Oliver Cowdery", "OC"),
  read_author_residuals(joseph_resid_path, "Joseph Smith", "JS")
)

author_residuals %>%
  dplyr::select(
    author,
    book,
    mosiah_priority,
    chapters,
    observed,
    expected,
    enrichment,
    chisq_stdres
  ) %>%
  dplyr::arrange(author, dplyr::desc(chisq_stdres)) %>%
  print(n = 100)

readr::write_csv(
  author_residuals,
  file.path(compare_dir, "author_bom_book_residuals_long.csv")
)
```

## 4. Book order

Use the canonical Book of Mormon order.

```{r book-order}
bom_book_order <- c(
  "1 Nephi",
  "2 Nephi",
  "Jacob",
  "Enos",
  "Jarom",
  "Omni",
  "Words of Mormon",
  "Mosiah",
  "Alma",
  "Helaman",
  "3 Nephi",
  "4 Nephi",
  "Mormon",
  "Ether",
  "Moroni"
)

author_order <- c(
  "Joseph Smith",
  "Sidney Rigdon",
  "Oliver Cowdery"
)

author_residuals <- author_residuals %>%
  dplyr::mutate(
    book = factor(book, levels = bom_book_order),
    author = factor(author, levels = author_order),
    mosiah_priority = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3"))
  )
```

## 5. Main heatmap: standardized residuals

```{r residual-heatmap, fig.width=8, fig.height=7}
ggplot(
  author_residuals,
  aes(
    x = author,
    y = fct_rev(book),
    fill = chisq_stdres
  )
) +
  geom_tile(color = "grey85") +
  geom_text(
    aes(label = round(chisq_stdres, 1)),
    size = 3
  ) +
  scale_fill_gradient2(
    low = "firebrick",
    mid = "white",
    high = "forestgreen",
    midpoint = 0
  ) +
  theme_bw() +
  labs(
    title = "Joseph, Sidney, and Oliver Book of Mormon affinity profiles",
    subtitle = "Pearson standardized residuals from author-specific chi-square goodness-of-fit tests",
    x = NULL,
    y = "Book of Mormon book",
    fill = "Std.\nresidual"
  )
```
```{r hclust}
mat <- author_residuals %>%
  select(author, book, chisq_stdres) %>%
  pivot_wider(names_from = author, values_from = chisq_stdres) %>%
  column_to_rownames("book") %>%
  as.matrix()

hc_books <- hclust(dist(mat))
hc_authors <- hclust(dist(t(mat)))

hc_books
hc_authors


# 3 authors × 15 books
author_mat <- residuals_wide %>%
  dplyr::select(
    book,
    `Joseph Smith`,
    `Sidney Rigdon`,
    `Oliver Cowdery`
  ) %>%
  tibble::column_to_rownames("book") %>%
  as.matrix()


hc_authors <- hclust(
  dist(t(author_mat)),
  method = "complete"
)

plot(
  hc_authors,
  main = "Hierarchical clustering of authors",
  xlab = "",
  sub = "",
  ylab = "Euclidean distance"
)




# Matrix: rows = books, columns = authors
book_mat <- residuals_wide %>%
  dplyr::select(
    book,
    `Joseph Smith`,
    `Sidney Rigdon`,
    `Oliver Cowdery`
  ) %>%
  tibble::column_to_rownames("book") %>%
  as.matrix()

# Correlation between books
book_cor <- cor(t(book_mat))

hc_books <- hclust(
  as.dist(1 - book_cor),
  method = "average"
)

plot(
  hc_books,
  hang = -1,
  cex = 0.9,
  main = "Hierarchical clustering of Book of Mormon books",
  xlab = "",
  sub = "",
  ylab = "1 - correlation"
)

rect.hclust(
  hc_books,
  k = 4,
  border = 2:5
)


hc_books <- hclust(
  dist(book_mat),
  method = "complete"
)

plot(
  hc_books,
  hang = -1,
  cex = 0.9,
  main = "Hierarchical clustering of Book of Mormon books",
  ylab = "Euclidean distance"
)

rect.hclust(
  hc_books,
  k = 4,
  border = 2:5
)


methods <- c(
  "complete",
  "average",
  "ward.D2",
  "single"
)

par(mfrow = c(2,2))

for(m in methods){

  hc <- hclust(
    dist(book_mat),
    method = m
  )

  plot(
    hc,
    main = m,
    hang = -1,
    cex = 0.8
  )
}

par(mfrow = c(1,1))

```


```{r more plots}

plot_book_hclust <- function(method = "complete", k = 4){

  hc <- hclust(
    dist(book_mat),
    method = method
  )

  plot(
    hc,
    hang = -1,
    cex = 0.9,
    main = paste("Book clustering -", method),
    xlab = "",
    sub = "",
    ylab = "Height"
  )

  rect.hclust(
    hc,
    k = k,
    border = 2:(k + 1)
  )

  invisible(hc)
}


hc_complete <- plot_book_hclust("complete", k = 4)
hc_average <- plot_book_hclust("average", k = 4)
hc_ward <- plot_book_hclust("ward.D2", k = 4)
hc_single <- plot_book_hclust("single", k = 4)


```

## 6. Side-by-side residual bar chart

```{r residual-bars, fig.width=12, fig.height=8}
ggplot(
  author_residuals,
  aes(
    x = fct_rev(book),
    y = chisq_stdres,
    fill = chisq_stdres > 0
  )
) +
  geom_col() +
  coord_flip() +
  facet_wrap(~author, nrow = 1) +
  geom_hline(yintercept = 0) +
  geom_hline(
    yintercept = c(-2, 2),
    linetype = "dashed",
    color = "red"
  ) +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Standardized residuals by author",
    subtitle = "Positive values indicate more nearest-neighbor votes than expected from book size",
    x = "Book of Mormon book",
    y = "Standardized residual"
  )
```

## 7. Wide comparison table

```{r wide-comparison}
residuals_wide <- author_residuals %>%
  dplyr::select(
    book,
    mosiah_priority,
    author,
    chisq_stdres
  ) %>%
  tidyr::pivot_wider(
    names_from = author,
    values_from = chisq_stdres
  ) %>%
  dplyr::mutate(
    SR_minus_JS = `Sidney Rigdon` - `Joseph Smith`,
    SR_minus_OC = `Sidney Rigdon` - `Oliver Cowdery`,
    OC_minus_JS = `Oliver Cowdery` - `Joseph Smith`
  ) %>%
  dplyr::arrange(book)

residuals_wide %>%
  print(n = 100)

readr::write_csv(
  residuals_wide,
  file.path(compare_dir, "author_bom_book_residuals_wide.csv")
)
```

## 8. Author-to-author correlations

```{r residual-correlations}
residual_matrix <- residuals_wide %>%
  dplyr::select(
    `Joseph Smith`,
    `Sidney Rigdon`,
    `Oliver Cowdery`
  )

cor_matrix <- cor(
  residual_matrix,
  use = "pairwise.complete.obs",
  method = "pearson"
)

cor_matrix

cor_long <- as.data.frame(as.table(cor_matrix)) %>%
  dplyr::rename(
    author_1 = Var1,
    author_2 = Var2,
    correlation = Freq
  )

cor_long



cor.test(
    residuals_wide$`Joseph Smith`,
    residuals_wide$`Oliver Cowdery`
)

cor.test(
    residuals_wide$`Joseph Smith`,
    residuals_wide$`Sidney Rigdon`
)

cor.test(
    residuals_wide$`Sidney Rigdon`,
    residuals_wide$`Oliver Cowdery`
)



dist(t(author_mat))


library(cocor)

cocor.dep.groups.overlap(
    r.jk = 0.86489,   # Joseph–Oliver
    r.jh = 0.67972,   # Joseph–Sidney
    r.kh = 0.74916,   # Oliver–Sidney
    n = 15
)

```


```{r pac pca}
pca <- prcomp(
    t(author_mat),
    scale. = TRUE
)

biplot(pca)
scores <- as.data.frame(pca$x)
scores$author <- rownames(scores)

ggplot(scores,
       aes(PC1, PC2, label = author)) +
    geom_point(size = 4) +
    geom_text(vjust = -0.5) +
    theme_bw()


library(ggplot2)

loadings <- as.data.frame(pca$rotation[,1:2])
loadings$book <- rownames(loadings)

ggplot(loadings,
       aes(PC1, PC2)) +
    geom_segment(
        aes(x=0,y=0,xend=PC1,yend=PC2),
        arrow=arrow(length=unit(.15,"cm"))
    ) +
    geom_text(aes(label=book),
              hjust=.5,
              vjust=-.4) +
    coord_equal() +
    theme_bw()

scores <- as.data.frame(pca$x)
scores$author <- rownames(scores)

ggplot(scores,
       aes(PC1,PC2,label=author)) +
    geom_point(size=5) +
    geom_text(vjust=-0.5) +
    coord_equal() +
    theme_bw()



pca$rotation[,2] %>%
  sort(decreasing = TRUE)

pca$rotation[,2] %>%
  sort()
```

```{r correlation-heatmap, fig.width=6, fig.height=5}
ggplot(
  cor_long,
  aes(
    x = author_1,
    y = author_2,
    fill = correlation
  )
) +
  geom_tile(color = "grey85") +
  geom_text(aes(label = round(correlation, 2)), size = 4) +
  scale_fill_gradient2(
    low = "firebrick",
    mid = "white",
    high = "forestgreen",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_bw() +
  labs(
    title = "Correlation of Book of Mormon residual profiles",
    subtitle = "Pearson correlation across Book of Mormon books",
    x = NULL,
    y = NULL,
    fill = "Correlation"
  )
```

## 9. Difference profiles

These plots show which books are more characteristic of one author than another.

### 9.1 Sidney minus Joseph

```{r plot-sr-minus-js, fig.width=9, fig.height=6}
ggplot(
  residuals_wide,
  aes(
    x = reorder(book, SR_minus_JS),
    y = SR_minus_JS,
    fill = SR_minus_JS > 0
  )
) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0) +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Sidney minus Joseph residual profile",
    subtitle = "Positive values mean stronger relative affinity for Sidney",
    x = "Book of Mormon book",
    y = "Difference in standardized residual"
  )
```

### 9.2 Sidney minus Oliver

```{r plot-sr-minus-oc, fig.width=9, fig.height=6}
ggplot(
  residuals_wide,
  aes(
    x = reorder(book, SR_minus_OC),
    y = SR_minus_OC,
    fill = SR_minus_OC > 0
  )
) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0) +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Sidney minus Oliver residual profile",
    subtitle = "Positive values mean stronger relative affinity for Sidney",
    x = "Book of Mormon book",
    y = "Difference in standardized residual"
  )
```

### 9.3 Oliver minus Joseph

```{r plot-oc-minus-js, fig.width=9, fig.height=6}
ggplot(
  residuals_wide,
  aes(
    x = reorder(book, OC_minus_JS),
    y = OC_minus_JS,
    fill = OC_minus_JS > 0
  )
) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0) +
  scale_fill_manual(
    values = c(
      "TRUE" = "forestgreen",
      "FALSE" = "firebrick"
    ),
    guide = "none"
  ) +
  theme_bw() +
  labs(
    title = "Oliver minus Joseph residual profile",
    subtitle = "Positive values mean stronger relative affinity for Oliver",
    x = "Book of Mormon book",
    y = "Difference in standardized residual"
  )
```

## 10. Mosiah Priority aggregate comparison

Read the MP aggregate vote files from notebooks 46, 47, and 48.

```{r read-mp-votes}
mp_files <- c(
  sidney_mp_path,
  oliver_mp_path,
  joseph_mp_path
)

missing_mp_files <- mp_files[!file.exists(mp_files)]

if(length(missing_mp_files) > 0){
  warning(
    "Missing MP vote CSV(s). MP aggregate comparison will be skipped. Missing:\n",
    paste(missing_mp_files, collapse = "\n")
  )
  author_mp_votes <- NULL
} else {

  read_author_mp <- function(path, author, author_code){
    readr::read_csv(path, show_col_types = FALSE) %>%
      dplyr::mutate(
        author = author,
        author_code = author_code,
        mosiah_priority = factor(mosiah_priority, levels = c("MP-1", "MP-2", "MP-3")),
        weighted_pct_votes = as.numeric(weighted_pct_votes),
        pct_votes = as.numeric(pct_votes)
      )
  }

  author_mp_votes <- dplyr::bind_rows(
    read_author_mp(sidney_mp_path, "Sidney Rigdon", "SR"),
    read_author_mp(oliver_mp_path, "Oliver Cowdery", "OC"),
    read_author_mp(joseph_mp_path, "Joseph Smith", "JS")
  ) %>%
    dplyr::mutate(
      author = factor(author, levels = author_order)
    )

  author_mp_votes %>%
    dplyr::arrange(author, mosiah_priority) %>%
    print(n = 100)

}
```

```{r plot-mp-aggregate-comparison, fig.width=9, fig.height=5}
if(!is.null(author_mp_votes)){

  ggplot(
    author_mp_votes,
    aes(
      x = mosiah_priority,
      y = weighted_pct_votes,
      fill = author
    )
  ) +
    geom_col(position = "dodge") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw() +
    labs(
      title = "Mosiah Priority nearest-neighbor vote shares by author",
      subtitle = "Weighted vote share across MP-1, MP-2, and MP-3",
      x = "Mosiah Priority group",
      y = "Weighted vote share",
      fill = "Author"
    )

}
```

## 11. Significant over- and under-representation

```{r significant-residuals}
significant_residuals <- author_residuals %>%
  dplyr::filter(abs(chisq_stdres) >= 2) %>%
  dplyr::mutate(
    direction = dplyr::if_else(
      chisq_stdres > 0,
      "Overrepresented",
      "Underrepresented"
    )
  ) %>%
  dplyr::arrange(author, direction, dplyr::desc(abs(chisq_stdres)))

significant_residuals %>%
  dplyr::select(
    author,
    book,
    mosiah_priority,
    observed,
    expected,
    chisq_stdres,
    direction
  ) %>%
  print(n = 100)

readr::write_csv(
  significant_residuals,
  file.path(compare_dir, "significant_author_book_residuals.csv")
)
```

## 12. Summary statistics by author

```{r author-summary-stats}
author_summary_stats <- author_residuals %>%
  dplyr::group_by(author) %>%
  dplyr::summarise(
    max_positive = max(chisq_stdres, na.rm = TRUE),
    max_negative = min(chisq_stdres, na.rm = TRUE),
    mean_abs_residual = mean(abs(chisq_stdres), na.rm = TRUE),
    n_over_2 = sum(chisq_stdres >= 2, na.rm = TRUE),
    n_under_minus_2 = sum(chisq_stdres <= -2, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(desc(mean_abs_residual))

author_summary_stats

author_summary_stats <- author_residuals %>%
  group_by(author) %>%
  summarise(
    max_positive = max(chisq_stdres),
    max_negative = min(chisq_stdres),
    mean_abs_residual = mean(abs(chisq_stdres)),
    sd_residual = sd(chisq_stdres),
    n_over_2 = sum(chisq_stdres >= 2),
    n_under_minus_2 = sum(chisq_stdres <= -2)
  )

author_summary_stats

```

## 13. Interpretation guide

Useful reading of the comparison:

```text
1. A large positive residual means the author's writings select that BOM book
   more often than expected by book size.

2. A large negative residual means the author's writings select that BOM book
   less often than expected by book size.

3. Shared positive residuals suggest a common affinity region.

4. Differences such as Sidney minus Joseph identify books where one author's
   affinity profile is especially distinctive.

5. If 3 Nephi is negative for all three authors, that suggests 3 Nephi may be
   intrinsically stylometrically distinct from the surviving JSO writings.
```

## 14. Save comparison object

```{r save-comparison}
saveRDS(
  list(
    author_residuals = author_residuals,
    residuals_wide = residuals_wide,
    cor_matrix = cor_matrix,
    cor_long = cor_long,
    author_mp_votes = author_mp_votes,
    significant_residuals = significant_residuals,
    author_summary_stats = author_summary_stats
  ),
  file.path(compare_dir, "compare_js_sr_oc_mp_objects.rds")
)
```
