---
title: "50 - Author Style Variance"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 500)

library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(tibble)
library(purrr)
library(ggplot2)
library(scales)
library(forcats)
```

## 1. Purpose

The MP notebooks showed that Sidney Rigdon has a more polarized Book of Mormon affinity profile than Joseph Smith or Oliver Cowdery.

This notebook tests a different question:

```text
Do Sidney Rigdon's own writings occupy a larger stylistic space?
```

That is a **within-author variance** question, not a Book of Mormon affinity question.

We compare Joseph, Sidney, and Oliver using:

1. Pairwise distances among each author's JSO chapters.
2. Distance from each chapter to its author's centroid.
3. PCA spread / convex hull area.
4. Feature-class-specific dispersion.
5. Optional permutation tests.

The core hypothesis is:

```text
Sidney has greater internal stylistic variance than Joseph or Oliver.
```

## 2. Paths

```{r paths}
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
export_dir <- file.path(STYLO_ROOT, "export", "author_style_variance")

dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)

jso_path <- file.path(prepared_dir, "jso_manual_chapter_docs.csv")

out_pairwise <- file.path(export_dir, "jso_author_pairwise_distances.csv")
out_centroid <- file.path(export_dir, "jso_author_centroid_distances.csv")
out_author_summary <- file.path(export_dir, "jso_author_style_variance_summary.csv")
out_feature_summary <- file.path(export_dir, "jso_author_feature_class_variance_summary.csv")
```

## 3. Read and normalize JSO corpus

```{r read-normalize-jso}
if(!file.exists(jso_path)){
  stop("Missing required file: ", jso_path)
}

clean_text_for_analysis <- function(x){
  x %>%
    as.character() %>%
    stringr::str_replace_all("\r", " ") %>%
    stringr::str_replace_all("\n", " ") %>%
    stringr::str_squish()
}

jso_raw <- readr::read_csv(jso_path, show_col_types = FALSE)

jso_docs <- jso_raw %>%
  dplyr::mutate(
    author = as.character(author),
    author_code = as.character(author_code),
    doc_id_original = as.character(doc_id),
    doc_id = paste0(author_code, "-", chapter),
    short_name = doc_id,
    book = author,
    header = as.character(chapter_title),
    text = clean_text_for_analysis(text),
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::filter(!is.na(text), text != "", n_words >= 50)

jso_docs %>%
  dplyr::count(author, author_code, name = "n_docs") %>%
  dplyr::arrange(author)

jso_docs %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )
```

## 4. Build transparent style feature matrix

This avoids content-heavy word n-grams and focuses on function words, punctuation, sentence length, word length, lexical richness, and readability-like metrics.

```{r feature-functions}
function_words <- c(
  "the","and","of","to","in","that","is","was","he","for","it","with",
  "as","his","on","be","at","by","i","this","had","not","are","but",
  "from","or","have","an","they","which","one","you","were","all","we",
  "there","can","their","has","would","what","will","if","so","no",
  "when","who","more","them","some","him","into","than","then","my",
  "out","up","upon","do","shall","may","me","our","us","your","ye"
)

count_sentences <- function(x){
  n <- stringr::str_count(x, "[.!?]+")
  pmax(n, 1)
}

tokenize_words <- function(x){
  x %>%
    stringr::str_to_lower() %>%
    stringr::str_extract_all("[a-z']+")
}

safe_div <- function(a, b){
  ifelse(b == 0 | is.na(b), NA_real_, a / b)
}

compute_style_features <- function(df){

  texts <- df$text
  words_list <- tokenize_words(texts)

  base <- tibble::tibble(
    doc_id = df$doc_id,
    author = df$author,
    author_code = df$author_code,
    chapter = df$chapter,
    short_name = df$short_name,
    chapter_title = df$chapter_title,
    n_words = purrr::map_int(words_list, length),
    n_sentences = purrr::map_int(texts, count_sentences),
    n_chars = nchar(texts)
  ) %>%
    dplyr::mutate(
      sentence_len_mean = n_words / n_sentences,
      char_per_word = n_chars / pmax(n_words, 1)
    )

  fw_mat <- purrr::map_dfr(seq_along(words_list), function(i){
    w <- words_list[[i]]
    counts <- tabulate(match(w, function_words), nbins = length(function_words))
    vals <- 1000 * counts / max(length(w), 1)

    tibble::as_tibble_row(
      stats::setNames(as.list(vals), paste0("fw_", function_words))
    ) %>%
      dplyr::mutate(doc_id = df$doc_id[[i]], .before = 1)
  })

  punct <- tibble::tibble(
    doc_id = df$doc_id,
    comma_per_1k = 1000 * stringr::str_count(texts, ",") / pmax(base$n_words, 1),
    semicolon_per_1k = 1000 * stringr::str_count(texts, ";") / pmax(base$n_words, 1),
    colon_per_1k = 1000 * stringr::str_count(texts, ":") / pmax(base$n_words, 1),
    question_per_1k = 1000 * stringr::str_count(texts, "\\?") / pmax(base$n_words, 1),
    exclamation_per_1k = 1000 * stringr::str_count(texts, "!") / pmax(base$n_words, 1),
    period_per_1k = 1000 * stringr::str_count(texts, "\\.") / pmax(base$n_words, 1)
  )

  sentence_bins <- purrr::map_dfr(seq_along(texts), function(i){
    sents <- stringr::str_split(texts[[i]], "[.!?]+")[[1]]
    sent_words <- stringr::str_count(sents, "\\S+")
    sent_words <- sent_words[sent_words > 0]

    if(length(sent_words) == 0){
      sent_words <- 0
    }

    tibble::tibble(
      doc_id = df$doc_id[[i]],
      sent_le_10 = mean(sent_words <= 10, na.rm = TRUE),
      sent_11_15 = mean(sent_words >= 11 & sent_words <= 15, na.rm = TRUE),
      sent_16_20 = mean(sent_words >= 16 & sent_words <= 20, na.rm = TRUE),
      sent_gt_20 = mean(sent_words > 20, na.rm = TRUE)
    )
  })

  lexical <- purrr::map_dfr(seq_along(words_list), function(i){
    w <- words_list[[i]]
    n <- length(w)
    types <- unique(w)
    word_lengths <- nchar(w)
    syllable_proxy <- stringr::str_count(w, "[aeiouy]+")
    syllable_proxy <- pmax(syllable_proxy, 1)

    tibble::tibble(
      doc_id = df$doc_id[[i]],
      type_token_ratio = safe_div(length(types), n),
      hapax_ratio = safe_div(sum(table(w) == 1), n),
      mean_word_length = mean(word_lengths, na.rm = TRUE),
      sd_word_length = sd(word_lengths, na.rm = TRUE),
      long_word_ratio = mean(word_lengths >= 7, na.rm = TRUE),
      syllables_per_word_proxy = mean(syllable_proxy, na.rm = TRUE),
      fk_like = 0.39 * safe_div(n, count_sentences(texts[[i]])) +
        11.8 * mean(syllable_proxy, na.rm = TRUE) - 15.59
    )
  })

  base %>%
    dplyr::left_join(fw_mat, by = "doc_id") %>%
    dplyr::left_join(punct, by = "doc_id") %>%
    dplyr::left_join(sentence_bins, by = "doc_id") %>%
    dplyr::left_join(lexical, by = "doc_id") %>%
    dplyr::mutate(
      dplyr::across(where(is.numeric), ~tidyr::replace_na(.x, 0))
    )
}

features <- compute_style_features(jso_docs)

features %>%
  dplyr::select(doc_id, author, n_words, sentence_len_mean, type_token_ratio, mean_word_length, fk_like) %>%
  dplyr::arrange(author, doc_id) %>%
  print(n = 100)
```

## 5. Scale feature matrix

```{r scale-feature-matrix}
feature_cols <- features %>%
  dplyr::select(where(is.numeric)) %>%
  names()

feature_cols <- setdiff(
  feature_cols,
  c("chapter", "n_words", "n_sentences", "n_chars")
)

X <- features %>%
  dplyr::select(dplyr::all_of(feature_cols)) %>%
  as.matrix()

rownames(X) <- features$doc_id

X_scaled <- scale(X)

X_scaled[is.nan(X_scaled)] <- 0

dim(X_scaled)
```

## 6. PCA overview

```{r pca}
pca <- prcomp(
  X_scaled,
  center = FALSE,
  scale. = FALSE
)

pca_scores <- as.data.frame(pca$x[, 1:5]) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    features %>%
      dplyr::select(doc_id, author, author_code, short_name, chapter, chapter_title),
    by = "doc_id"
  )

variance_explained <- tibble::tibble(
  PC = paste0("PC", seq_along(pca$sdev)),
  variance = pca$sdev^2 / sum(pca$sdev^2)
)

variance_explained %>%
  dplyr::slice_head(n = 10)
```

```{r plot-pca-author, fig.width=8, fig.height=6}
ggplot(
  pca_scores,
  aes(x = PC1, y = PC2, color = author)
) +
  geom_point(size = 3, alpha = 0.8) +
  theme_bw() +
  labs(
    title = "JSO writings in style-feature PCA space",
    subtitle = "Function words, punctuation, sentence, lexical, word-length, readability-like features",
    x = paste0("PC1 (", scales::percent(variance_explained$variance[[1]], accuracy = 0.1), ")"),
    y = paste0("PC2 (", scales::percent(variance_explained$variance[[2]], accuracy = 0.1), ")"),
    color = "Author"
  )
```

## 7. Pairwise within-author distances

```{r pairwise-distances}
D <- as.matrix(dist(X_scaled, method = "euclidean"))

pairwise_distances <- expand.grid(
  doc_id_1 = rownames(D),
  doc_id_2 = colnames(D),
  stringsAsFactors = FALSE
) %>%
  dplyr::filter(doc_id_1 < doc_id_2) %>%
  dplyr::mutate(distance = D[cbind(doc_id_1, doc_id_2)]) %>%
  dplyr::left_join(
    features %>% dplyr::select(doc_id_1 = doc_id, author_1 = author, short_name_1 = short_name),
    by = "doc_id_1"
  ) %>%
  dplyr::left_join(
    features %>% dplyr::select(doc_id_2 = doc_id, author_2 = author, short_name_2 = short_name),
    by = "doc_id_2"
  ) %>%
  dplyr::mutate(
    comparison_type = dplyr::if_else(author_1 == author_2, "within_author", "between_author"),
    author_pair = dplyr::if_else(
      author_1 == author_2,
      author_1,
      paste(pmin(author_1, author_2), pmax(author_1, author_2), sep = " / ")
    )
  )

within_author_distances <- pairwise_distances %>%
  dplyr::filter(comparison_type == "within_author")

within_author_summary <- within_author_distances %>%
  dplyr::group_by(author_pair) %>%
  dplyr::summarise(
    n_pairs = dplyr::n(),
    mean_distance = mean(distance),
    median_distance = median(distance),
    sd_distance = sd(distance),
    min_distance = min(distance),
    max_distance = max(distance),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_distance))

within_author_summary

readr::write_csv(pairwise_distances, out_pairwise)
```

```{r plot-within-pairwise-boxplot, fig.width=8, fig.height=5}
ggplot(
  within_author_distances,
  aes(
    x = fct_reorder(author_pair, distance, .fun = median),
    y = distance,
    fill = author_pair
  )
) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.35, show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "Within-author pairwise stylistic distances",
    subtitle = "Larger distances indicate greater internal stylistic spread",
    x = "Author",
    y = "Euclidean distance in scaled style-feature space"
  )
```

## 8. Distance from author centroid

```{r centroid-distances}
scaled_df <- as.data.frame(X_scaled) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    features %>% dplyr::select(doc_id, author, short_name, chapter),
    by = "doc_id"
  )

author_centroids <- scaled_df %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(feature_cols),
    names_to = "feature",
    values_to = "value"
  ) %>%
  dplyr::group_by(author, feature) %>%
  dplyr::summarise(centroid_value = mean(value), .groups = "drop")

centroid_distances <- scaled_df %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(feature_cols),
    names_to = "feature",
    values_to = "value"
  ) %>%
  dplyr::left_join(author_centroids, by = c("author", "feature")) %>%
  dplyr::mutate(squared_diff = (value - centroid_value)^2) %>%
  dplyr::group_by(doc_id, author, short_name, chapter) %>%
  dplyr::summarise(
    centroid_distance = sqrt(sum(squared_diff)),
    .groups = "drop"
  ) %>%
  dplyr::arrange(author, dplyr::desc(centroid_distance))

centroid_summary <- centroid_distances %>%
  dplyr::group_by(author) %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    mean_centroid_distance = mean(centroid_distance),
    median_centroid_distance = median(centroid_distance),
    sd_centroid_distance = sd(centroid_distance),
    max_centroid_distance = max(centroid_distance),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_centroid_distance))

centroid_summary

readr::write_csv(centroid_distances, out_centroid)
```

```{r plot-centroid-distances, fig.width=8, fig.height=5}
ggplot(
  centroid_distances,
  aes(
    x = fct_reorder(author, centroid_distance, .fun = median),
    y = centroid_distance,
    fill = author
  )
) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.5, show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "Distance from each author's stylistic centroid",
    subtitle = "Larger values indicate chapters farther from the author's average style",
    x = "Author",
    y = "Distance from author centroid"
  )
```
```{r distance}

centroid_info = centroid_distances %>%
  filter(author == "Sidney Rigdon") %>%
  left_join(
    jso_docs %>%
      select(doc_id, header),
    by = c("short_name" = "doc_id")
  ) %>%
  arrange(desc(centroid_distance)) %>%
  select(
    short_name,
    header,
    centroid_distance
  )

 centroid_info

```


## 9. PCA spread / convex hull area

```{r convex-hull-area}
polygon_area <- function(x, y){
  if(length(x) < 3) return(NA_real_)
  idx <- chull(x, y)
  x <- x[idx]
  y <- y[idx]
  abs(sum(x * c(y[-1], y[1]) - y * c(x[-1], x[1])) / 2)
}

pca_spread <- pca_scores %>%
  dplyr::group_by(author) %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    sd_PC1 = sd(PC1),
    sd_PC2 = sd(PC2),
    mean_radial_distance = mean(sqrt((PC1 - mean(PC1))^2 + (PC2 - mean(PC2))^2)),
    hull_area_PC1_PC2 = polygon_area(PC1, PC2),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(hull_area_PC1_PC2))

pca_spread
```

```{r plot-pca-hulls, fig.width=8, fig.height=6}
hull_points <- pca_scores %>%
  dplyr::group_by(author) %>%
  dplyr::slice(chull(PC1, PC2)) %>%
  dplyr::ungroup()

ggplot(
  pca_scores,
  aes(x = PC1, y = PC2, color = author, fill = author)
) +
  geom_polygon(
    data = hull_points,
    aes(group = author),
    alpha = 0.15,
    color = NA
  ) +
  geom_point(size = 3, alpha = 0.8) +
  theme_bw() +
  labs(
    title = "PCA style-space spread by author",
    subtitle = "Convex hulls show approximate occupied region in PC1-PC2 space",
    x = paste0("PC1 (", scales::percent(variance_explained$variance[[1]], accuracy = 0.1), ")"),
    y = paste0("PC2 (", scales::percent(variance_explained$variance[[2]], accuracy = 0.1), ")"),
    color = "Author",
    fill = "Author"
  )
```

## 10. Feature-class-specific dispersion

```{r feature-class-dispersion}
feature_class_map <- tibble::tibble(
  feature = feature_cols,
  feature_class = dplyr::case_when(
    stringr::str_starts(feature, "fw_") ~ "function_words",
    stringr::str_detect(feature, "comma|semicolon|colon|question|exclamation|period") ~ "punctuation",
    stringr::str_starts(feature, "sent_") | feature == "sentence_len_mean" ~ "sentence",
    stringr::str_detect(feature, "word_length|word|long_word|char_per_word|syllables") ~ "word_length",
    stringr::str_detect(feature, "type_token|hapax") ~ "lexical",
    stringr::str_detect(feature, "fk_like") ~ "readability",
    TRUE ~ "other"
  )
)

scaled_long <- scaled_df %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(feature_cols),
    names_to = "feature",
    values_to = "scaled_value"
  ) %>%
  dplyr::left_join(feature_class_map, by = "feature")

feature_class_dispersion <- scaled_long %>%
  dplyr::group_by(author, feature_class) %>%
  dplyr::summarise(
    n_features = dplyr::n_distinct(feature),
    mean_abs_deviation = mean(abs(scaled_value - mean(scaled_value)), na.rm = TRUE),
    sd_scaled_value = sd(scaled_value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(author, feature_class)

feature_class_dispersion

readr::write_csv(feature_class_dispersion, out_feature_summary)
```

```{r plot-feature-class-dispersion, fig.width=9, fig.height=5}
ggplot(
  feature_class_dispersion,
  aes(
    x = feature_class,
    y = mean_abs_deviation,
    fill = author
  )
) +
  geom_col(position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Within-author dispersion by feature class",
    subtitle = "Mean absolute deviation of scaled features within each author",
    x = "Feature class",
    y = "Mean absolute deviation",
    fill = "Author"
  )
```

## 11. Permutation test: is Sidney more dispersed?

This nonparametric test compares mean within-author pairwise distance. The null hypothesis is that author labels are exchangeable.

```{r permutation-test, cache=TRUE}
set.seed(1830)

observed_author_means <- within_author_distances %>%
  dplyr::group_by(author_pair) %>%
  dplyr::summarise(
    mean_distance = mean(distance),
    .groups = "drop"
  )

compute_label_mean_distance <- function(labels){

  tmp <- tibble::tibble(
    doc_id = rownames(D),
    perm_author = labels
  )

  expand.grid(
    doc_id_1 = rownames(D),
    doc_id_2 = colnames(D),
    stringsAsFactors = FALSE
  ) %>%
    dplyr::filter(doc_id_1 < doc_id_2) %>%
    dplyr::mutate(distance = D[cbind(doc_id_1, doc_id_2)]) %>%
    dplyr::left_join(
      tmp %>% dplyr::rename(doc_id_1 = doc_id, author_1 = perm_author),
      by = "doc_id_1"
    ) %>%
    dplyr::left_join(
      tmp %>% dplyr::rename(doc_id_2 = doc_id, author_2 = perm_author),
      by = "doc_id_2"
    ) %>%
    dplyr::filter(author_1 == author_2) %>%
    dplyr::group_by(author_1) %>%
    dplyr::summarise(
      mean_distance = mean(distance),
      .groups = "drop"
    )
}

n_perm <- 1000

perm_results <- purrr::map_dfr(seq_len(n_perm), function(i){
  compute_label_mean_distance(sample(features$author)) %>%
    dplyr::mutate(iter = i)
})

permutation_summary <- observed_author_means %>%
  dplyr::rename(author = author_pair, observed_mean_distance = mean_distance) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    p_greater = mean(
      perm_results$mean_distance[
        perm_results$author_1 == author
      ] >= observed_mean_distance
    )
  ) %>%
  dplyr::ungroup()

permutation_summary
```

## 12. Author style variance summary

```{r author-summary}
pairwise_summary <- within_author_distances %>%
  dplyr::group_by(author = author_pair) %>%
  dplyr::summarise(
    n_pairs = dplyr::n(),
    mean_pairwise_distance = mean(distance),
    median_pairwise_distance = median(distance),
    sd_pairwise_distance = sd(distance),
    max_pairwise_distance = max(distance),
    .groups = "drop"
  )

author_style_variance_summary <- pairwise_summary %>%
  dplyr::left_join(centroid_summary, by = "author") %>%
  dplyr::left_join(pca_spread, by = "author") %>%
  dplyr::left_join(permutation_summary, by = "author") %>%
  dplyr::arrange(dplyr::desc(mean_pairwise_distance))

author_style_variance_summary

readr::write_csv(author_style_variance_summary, out_author_summary)
```

## 13. Interpretation guide

Evidence supporting the "larger style capability / larger variance" interpretation would include:

```text
1. Sidney has the largest mean within-author pairwise distance.
2. Sidney has the largest mean distance from his author centroid.
3. Sidney occupies the largest PCA hull area.
4. Sidney is more dispersed across multiple feature classes, not only one.
5. Permutation tests show Sidney's dispersion is unusually high relative to random author labels.
```

Evidence against the interpretation would include:

```text
1. Sidney's Book of Mormon residual variance is high, but his own writings are internally compact.
2. Sidney's apparent variability is driven by one or two outlier chapters.
3. Greater dispersion appears only in document length or topic-related features rather than style-oriented features.
```

## 14. Save objects

```{r save-objects}
saveRDS(
  list(
    jso_docs = jso_docs,
    features = features,
    feature_cols = feature_cols,
    X_scaled = X_scaled,
    pca = pca,
    pca_scores = pca_scores,
    pairwise_distances = pairwise_distances,
    within_author_distances = within_author_distances,
    centroid_distances = centroid_distances,
    pca_spread = pca_spread,
    feature_class_dispersion = feature_class_dispersion,
    permutation_summary = permutation_summary,
    author_style_variance_summary = author_style_variance_summary
  ),
  file.path(export_dir, "author_style_variance_objects.rds")
)
```


yes, I don't understand this notebook. why is it not relying on normal features calculated by the ensemble


I think you've identified the fundamental problem. **Notebook 50 went off in a different direction than the rest of the project.**

Everything we've built so far has been based on the ensemble.

* Same feature sets.
* Same TF-IDF.
* Same cosine/Delta similarities.
* Same nearest-neighbor methodology.
* Same weighted voting.

Then I built Notebook 50 using a completely new handcrafted feature extractor (function words, punctuation, readability, etc.), which is **not** the same representation that `run_stylometry_ensemble()` uses. That means the results are not directly comparable to everything else you've done.

## I think Notebook 50 should instead be built entirely on the ensemble outputs.

Your ensemble already computes exactly what we need.

For every Sidney chapter, it already has

* character n-grams
* word n-grams
* function words
* POS n-grams
* punctuation
* sentence features
* lexical richness
* readability
* cosine similarities
* Burrows Delta
* weighted ensemble similarities
* nearest neighbors

Those are the author's "style coordinates."

---

## There are three possible notions of "style variance"

### Option A — Raw feature variance (what I mistakenly implemented)

```
chapter
   ↓
feature vector
   ↓
Euclidean distance
```

Pros:

* Simple.

Cons:

* Uses a different feature extraction pipeline.
* Doesn't use the ensemble.

I'd abandon this.

---

### Option B — Ensemble similarity-space variance ⭐

This is what I think you actually want.

Every chapter already has a similarity vector.

For example,

```
SR-12

Alma .......... .84
Mosiah ........ .81
Isaiah ........ .79
Jeremiah ...... .75
...
```

Every chapter is therefore represented by

```
(sim1, sim2, sim3, ..., simN)
```

Compute distances between **those vectors**.

Now we're measuring

> How differently do Sidney's chapters behave under the ensemble?

Everything stays inside the ensemble.

---

### Option C — Vote-profile variance ⭐⭐⭐ (my favorite)

Even better.

Every chapter has a vote distribution.

Example

```
SR-18

Alma      42%
Mosiah    26%
Jeremiah  10%
Isaiah     8%
Ether      5%
...
```

That vector itself becomes the style representation.

Then compute

```
distance(
   vote_vector(SR-1),
   vote_vector(SR-2)
)
```

If Sidney really shifts among stylistic modes, his vote vectors should vary substantially.

This is exactly the question you've been asking throughout the project.

---

## Even better: cluster the vote vectors

Instead of asking

> "How much variance?"

ask

> "How many distinct stylistic modes?"

For every chapter:

```
chapter
     ↓
nearest-neighbor votes
     ↓
vector over all books
     ↓
PCA
     ↓
hclust
     ↓
mclust
```

Then compare

| Author | Number of occupied modes |
| ------ | -----------------------: |
| Joseph |                        2 |
| Oliver |                        3 |
| Sidney |                        5 |

Now we're talking about **style capability**, not Euclidean variance.

---

## I think this is the strongest direction

I would scrap Notebook 50 and replace it with something like:

```
50_Author_Style_Modes.Rmd
```

using **only ensemble objects**.

1. Run `run_stylometry_ensemble()` once.
2. Extract every chapter's weighted vote vector (or full similarity vector).
3. Build a chapter × target-book matrix.
4. Compute distances between chapters in **ensemble space**.
5. PCA.
6. Hierarchical clustering.
7. mclust.
8. Compare Joseph, Sidney, and Oliver by:

   * average distance between chapters,
   * number of clusters occupied,
   * entropy of vote distributions,
   * within-author cluster diversity.

That notebook would fit seamlessly with the rest of the project because it never leaves the ensemble framework, and it directly tests the hypothesis you've been developing rather than introducing a new, incompatible feature representation. I think it will produce results that are much easier to interpret alongside your MP analyses.



