---
title: "51 - Author Style Modes Using Ensemble Outputs"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 500
)

library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(tibble)
library(purrr)
library(ggplot2)
library(scales)
library(forcats)
library(mclust)
```

## 1. Purpose

Notebook 50 used a handcrafted feature matrix. This notebook replaces that approach.

This notebook uses the **pre-defined ensemble engine** developed in the project:

```r
run_stylometry_ensemble()
```

The question is:

```text
Do Joseph Smith, Sidney Rigdon, and Oliver Cowdery occupy different
stylistic modes when represented by their ensemble nearest-neighbor behavior?
```

Instead of measuring raw feature variance, this notebook measures variation in:

1. **Book of Mormon vote profiles**
2. **Book-level weighted nearest-neighbor vectors**
3. **Mosiah Priority vote vectors**
4. **Chapter clustering in ensemble-space**
5. **Author cluster entropy / occupied style modes**

This keeps the analysis inside the same framework as notebooks 45-49.

## 2. Paths

```{r paths}
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
export_dir <- file.path(STYLO_ROOT, "export", "author_style_modes")

dir.create(export_dir, recursive = TRUE, showWarnings = FALSE)

jso_path <- file.path(prepared_dir, "jso_manual_chapter_docs.csv")
bom_path <- file.path(prepared_dir, "bom_chapter_docs.csv")

out_nn_jso_bom <- file.path(export_dir, "jso_to_bom_nearest_neighbors.csv")
out_book_profiles <- file.path(export_dir, "jso_chapter_bom_book_vote_profiles.csv")
out_mp_profiles <- file.path(export_dir, "jso_chapter_mp_vote_profiles.csv")
out_mode_assignments <- file.path(export_dir, "jso_author_style_mode_assignments.csv")
out_author_mode_summary <- file.path(export_dir, "jso_author_style_mode_summary.csv")
```

## 3. Read data

```{r read-data}
read_required <- function(path){
  if(!file.exists(path)){
    stop("Missing required file: ", path)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

jso_raw <- read_required(jso_path)
bom_raw <- read_required(bom_path)

jso_raw %>%
  dplyr::count(author, author_code, name = "n_docs") %>%
  dplyr::arrange(author)

dim(bom_raw)
```

## 4. Helpers

```{r helpers}
clean_text_for_analysis <- function(x){
  x %>%
    as.character() %>%
    stringr::str_replace_all("\r", " ") %>%
    stringr::str_replace_all("\n", " ") %>%
    stringr::str_squish()
}

assign_mosiah_priority <- function(book){
  dplyr::case_when(
    book %in% c("Mosiah", "Alma", "Helaman") ~ "MP-1",
    book %in% c("3 Nephi", "4 Nephi", "Mormon", "Ether", "Moroni") ~ "MP-2",
    book %in% c("1 Nephi", "2 Nephi", "Jacob", "Enos", "Jarom", "Omni", "Words of Mormon") ~ "MP-3",
    TRUE ~ NA_character_
  )
}

assign_mosiah_priority_label <- function(mp){
  dplyr::case_when(
    mp == "MP-1" ~ "MP-1_Mosiah_Alma_Helaman",
    mp == "MP-2" ~ "MP-2_3Nephi_to_Moroni",
    mp == "MP-3" ~ "MP-3_1Nephi_to_WordsOfMormon",
    TRUE ~ NA_character_
  )
}
```

## 5. Normalize JSO and BOM docs

```{r normalize-docs}
jso_docs <- jso_raw %>%
  dplyr::mutate(
    author = as.character(author),
    author_code = as.character(author_code),
    corpus_id = "JSO",
    translation = "JSO",
    corpus_family = "JSO",
    source_family = "JSO",
    doc_id_original = as.character(doc_id),
    doc_id = paste0(author_code, "-", chapter),
    short_name = doc_id,
    book = author,
    section = paste0(author_code, "-", chapter_original),
    macro_genre = "Early_LDS_Writing",
    header = as.character(chapter_title),
    text = clean_text_for_analysis(text),
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::filter(!is.na(text), text != "", n_words >= 50)

bom_docs <- bom_raw %>%
  dplyr::mutate(
    corpus_id = if("corpus_id" %in% names(.)) as.character(corpus_id) else "BOM",
    translation = if("translation" %in% names(.)) as.character(translation) else "1830",
    corpus_family = if("corpus_family" %in% names(.)) as.character(corpus_family) else "BOM",
    source_family = if("source_family" %in% names(.)) as.character(source_family) else "BOM_1830",
    doc_id = as.character(doc_id),
    book = as.character(book),
    chapter = as.integer(chapter),
    section = if("section" %in% names(.)) as.character(section) else paste0(book, "_", chapter),
    macro_genre = if("macro_genre" %in% names(.)) as.character(macro_genre) else "Modern_Scripture",
    header = if("header" %in% names(.)) as.character(header) else paste0(book, " ", chapter),
    text = clean_text_for_analysis(text),
    n_words = stringr::str_count(text, "\\S+"),
    mosiah_priority = if("mosiah_priority" %in% names(.)) as.character(mosiah_priority) else assign_mosiah_priority(book),
    mosiah_priority = dplyr::if_else(
      is.na(mosiah_priority) | mosiah_priority == "",
      assign_mosiah_priority(book),
      mosiah_priority
    ),
    mosiah_priority_label = if("mosiah_priority_label" %in% names(.)) as.character(mosiah_priority_label) else assign_mosiah_priority_label(mosiah_priority),
    mosiah_priority_label = dplyr::if_else(
      is.na(mosiah_priority_label) | mosiah_priority_label == "",
      assign_mosiah_priority_label(mosiah_priority),
      mosiah_priority_label
    )
  ) %>%
  dplyr::filter(!is.na(text), text != "", n_words >= 50)

jso_docs %>%
  dplyr::count(author, author_code, name = "n_docs") %>%
  dplyr::arrange(author)

bom_docs %>%
  dplyr::count(mosiah_priority, mosiah_priority_label, book, name = "n_chapters") %>%
  dplyr::arrange(mosiah_priority, book)
```

## 6. Author map

```{r author-map}
jso_author_map <- jso_docs %>%
  dplyr::transmute(
    doc_id,
    author,
    author_code,
    short_name,
    chapter,
    chapter_original,
    chapter_title,
    header
  )

jso_author_map %>%
  dplyr::arrange(author, chapter) %>%
  print(n = 100)
```

## 7. Run ensemble: JSO chapters to BOM chapters

Use the same style-oriented feature classes used in the MP hypothesis notebooks.

```{r feature-classes}
include_feature_classes <- c(
  "function_words", "punctuation", "sentence",
  "word_length", "lexical", "readability"
)


jso_config  <- default_config(include = include_feature_classes)


jso_config$weights <- c(
    function_words = 0.500,
    punctuation    = 0.075,
    sentence       = 0.175,
    word_length    = 0.125,
    lexical        = 0.075,
    readability    = 0.050
  )

```

```{r source-target-docs}
source_docs_jso <- jso_docs %>%
  dplyr::select(
    doc_id,
    corpus_id,
    translation,
    corpus_family,
    source_family,
    book,
    chapter,
    section,
    macro_genre,
    header,
    text
  )

target_docs_bom <- bom_docs %>%
  dplyr::select(
    doc_id,
    corpus_id,
    translation,
    corpus_family,
    source_family,
    book,
    chapter,
    section,
    macro_genre,
    mosiah_priority,
    mosiah_priority_label,
    header,
    text
  )

dim(source_docs_jso)
dim(target_docs_bom)
```

```{r run-jso-bom-ensemble}
result_jso_bom <- run_stylometry_ensemble(
  source_docs = source_docs_jso,
  target_docs = target_docs_bom,
  config = jso_config,
  text_col = "text",
  doc_id_col = "doc_id",
  k_neighbors = 20,
  use_cache = TRUE
)

names(result_jso_bom)
```

## 8. Built-in ensemble reports

```{r show-feature-votes}
show_feature_votes(result_jso_bom)
```

```{r show-consensus}
show_consensus(result_jso_bom)
```

## 9. Label nearest neighbors

```{r label-nearest}
nn_jso_bom <- result_jso_bom$nearest_neighbors %>%
  dplyr::mutate(
    doc_id = as.character(doc_id),
    neighbor_doc_id = as.character(neighbor_doc_id),
    similarity = as.numeric(similarity),
    rank = as.integer(rank)
  ) %>%
  dplyr::left_join(
    jso_author_map %>%
      dplyr::select(
        doc_id,
        author,
        author_code,
        short_name,
        source_chapter = chapter,
        source_title = chapter_title
      ),
    by = "doc_id"
  ) %>%
  dplyr::left_join(
    target_docs_bom %>%
      dplyr::select(
        neighbor_doc_id = doc_id,
        target_book = book,
        target_chapter = chapter,
        target_header = header,
        mosiah_priority,
        mosiah_priority_label
      ),
    by = "neighbor_doc_id"
  )

nn_jso_bom %>%
  dplyr::arrange(author, doc_id, rank) %>%
  dplyr::select(
    author,
    short_name,
    rank,
    neighbor_doc_id,
    target_book,
    target_chapter,
    mosiah_priority,
    similarity
  ) 

nn_jso_bom

readr::write_csv(nn_jso_bom, out_nn_jso_bom)
```

## 10. Chapter × BOM book vote profiles

Each JSO chapter is represented by its weighted nearest-neighbor vote profile across Book of Mormon books.

```{r book-vote-profiles}
chapter_book_profiles_long <- nn_jso_bom %>%
  dplyr::group_by(
    doc_id,
    author,
    author_code,
    short_name,
    source_title,
    target_book
  ) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    best_rank = min(rank, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(doc_id) %>%
  dplyr::mutate(
    vote_share = votes / sum(votes),
    weighted_vote_share = weighted_votes / sum(weighted_votes)
  ) %>%
  dplyr::ungroup()

chapter_book_profiles_long %>%
  dplyr::arrange(author, short_name, dplyr::desc(weighted_vote_share)) %>%
  print(n = 200)

readr::write_csv(chapter_book_profiles_long, out_book_profiles)
```

```{r book-profile-matrix}
book_profile_wide <- chapter_book_profiles_long %>%
  dplyr::select(
    doc_id,
    author,
    author_code,
    short_name,
    source_title,
    target_book,
    weighted_vote_share
  ) %>%
  tidyr::pivot_wider(
    names_from = target_book,
    values_from = weighted_vote_share,
    values_fill = 0
  )

book_cols <- setdiff(
  names(book_profile_wide),
  c("doc_id", "author", "author_code", "short_name", "source_title")
)

book_profile_matrix <- book_profile_wide %>%
  dplyr::select(dplyr::all_of(book_cols)) %>%
  as.matrix()

rownames(book_profile_matrix) <- book_profile_wide$doc_id

dim(book_profile_matrix)
```

## 11. Chapter × MP vote profiles

```{r mp-vote-profiles}
chapter_mp_profiles_long <- nn_jso_bom %>%
  dplyr::group_by(
    doc_id,
    author,
    author_code,
    short_name,
    source_title,
    mosiah_priority,
    mosiah_priority_label
  ) %>%
  dplyr::summarise(
    votes = dplyr::n(),
    weighted_votes = sum(similarity, na.rm = TRUE),
    mean_similarity = mean(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(doc_id) %>%
  dplyr::mutate(
    vote_share = votes / sum(votes),
    weighted_vote_share = weighted_votes / sum(weighted_votes)
  ) %>%
  dplyr::ungroup()

chapter_mp_profiles_long %>%
  dplyr::arrange(author, short_name, mosiah_priority) %>%
  print(n = 200)

readr::write_csv(chapter_mp_profiles_long, out_mp_profiles)
```

```{r mp-profile-matrix}
mp_profile_wide <- chapter_mp_profiles_long %>%
  dplyr::select(
    doc_id,
    author,
    author_code,
    short_name,
    source_title,
    mosiah_priority,
    weighted_vote_share
  ) %>%
  tidyr::pivot_wider(
    names_from = mosiah_priority,
    values_from = weighted_vote_share,
    values_fill = 0
  )

mp_cols <- intersect(c("MP-1", "MP-2", "MP-3"), names(mp_profile_wide))

mp_profile_matrix <- mp_profile_wide %>%
  dplyr::select(dplyr::all_of(mp_cols)) %>%
  as.matrix()

rownames(mp_profile_matrix) <- mp_profile_wide$doc_id

dim(mp_profile_matrix)
```

## 12. Entropy of each chapter's vote profile

Higher entropy means the chapter spreads its nearest-neighbor votes across many Book of Mormon books. Lower entropy means it concentrates on a few books.

```{r entropy}
entropy_bits <- function(p){
  p <- p[p > 0 & !is.na(p)]
  -sum(p * log2(p))
}

chapter_entropy <- chapter_book_profiles_long %>%
  dplyr::group_by(doc_id, author, author_code, short_name, source_title) %>%
  dplyr::summarise(
    book_entropy = entropy_bits(weighted_vote_share),
    effective_books = 2^book_entropy,
    n_books_with_votes = dplyr::n_distinct(target_book),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    chapter_mp_profiles_long %>%
      dplyr::group_by(doc_id) %>%
      dplyr::summarise(
        mp_entropy = entropy_bits(weighted_vote_share),
        effective_mp_groups = 2^mp_entropy,
        .groups = "drop"
      ),
    by = "doc_id"
  )

chapter_entropy %>%
  dplyr::arrange(author, dplyr::desc(book_entropy)) %>%
  print(n = 100)
```

```{r plot-author-entropy, fig.width=8, fig.height=5}
ggplot(
  chapter_entropy,
  aes(
    x = fct_reorder(author, book_entropy, .fun = median),
    y = book_entropy,
    fill = author
  )
) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.5, show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "Entropy of chapter-level Book of Mormon vote profiles",
    subtitle = "Higher entropy means nearest-neighbor votes are spread across more books",
    x = "Author",
    y = "Book-level vote entropy, bits"
  )
```

## 13. Within-author distances in ensemble vote-profile space

This is the direct replacement for Notebook 50's raw feature distance. Here each chapter is represented by its ensemble-derived Book of Mormon vote vector.

```{r profile-distances}
book_profile_dist <- as.matrix(
  dist(book_profile_matrix, method = "euclidean")
)

profile_pairwise_distances <- expand.grid(
  doc_id_1 = rownames(book_profile_dist),
  doc_id_2 = colnames(book_profile_dist),
  stringsAsFactors = FALSE
) %>%
  dplyr::filter(doc_id_1 < doc_id_2) %>%
  dplyr::mutate(
    distance = book_profile_dist[cbind(doc_id_1, doc_id_2)]
  ) %>%
  dplyr::left_join(
    book_profile_wide %>%
      dplyr::select(doc_id_1 = doc_id, author_1 = author, short_name_1 = short_name),
    by = "doc_id_1"
  ) %>%
  dplyr::left_join(
    book_profile_wide %>%
      dplyr::select(doc_id_2 = doc_id, author_2 = author, short_name_2 = short_name),
    by = "doc_id_2"
  ) %>%
  dplyr::mutate(
    comparison_type = dplyr::if_else(author_1 == author_2, "within_author", "between_author"),
    author_pair = dplyr::if_else(
      author_1 == author_2,
      author_1,
      paste(pmin(author_1, author_2), pmax(author_1, author_2), sep = " / ")
    )
  )

within_profile_distances <- profile_pairwise_distances %>%
  dplyr::filter(comparison_type == "within_author")

within_profile_summary <- within_profile_distances %>%
  dplyr::group_by(author = author_pair) %>%
  dplyr::summarise(
    n_pairs = dplyr::n(),
    mean_profile_distance = mean(distance),
    median_profile_distance = median(distance),
    sd_profile_distance = sd(distance),
    max_profile_distance = max(distance),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(mean_profile_distance))

within_profile_summary
```

```{r plot-within-profile-distances, fig.width=8, fig.height=5}
ggplot(
  within_profile_distances,
  aes(
    x = fct_reorder(author_pair, distance, .fun = median),
    y = distance,
    fill = author_pair
  )
) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.35, show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "Within-author distances in ensemble vote-profile space",
    subtitle = "Each chapter is represented by weighted BOM book nearest-neighbor shares",
    x = "Author",
    y = "Euclidean distance between chapter vote profiles"
  )
```

## 14. PCA of ensemble vote profiles

```{r vote-profile-pca}
vote_pca <- prcomp(
  book_profile_matrix,
  center = TRUE,
  scale. = TRUE
)

vote_scores <- as.data.frame(vote_pca$x[, 1:5]) %>%
  tibble::rownames_to_column("doc_id") %>%
  dplyr::left_join(
    book_profile_wide %>%
      dplyr::select(doc_id, author, author_code, short_name, source_title),
    by = "doc_id"
  )

vote_variance <- tibble::tibble(
  PC = paste0("PC", seq_along(vote_pca$sdev)),
  variance = vote_pca$sdev^2 / sum(vote_pca$sdev^2)
)

vote_variance %>% dplyr::slice_head(n = 10)
```

```{r plot-vote-pca, fig.width=8, fig.height=6}
ggplot(
  vote_scores,
  aes(
    x = PC1,
    y = PC2,
    color = author
  )
) +
  geom_point(size = 3, alpha = 0.85) +
  theme_bw() +
  labs(
    title = "JSO chapters in ensemble vote-profile PCA space",
    subtitle = "Coordinates are derived from BOM book weighted nearest-neighbor vote profiles",
    x = paste0("PC1 (", scales::percent(vote_variance$variance[[1]], accuracy = 0.1), ")"),
    y = paste0("PC2 (", scales::percent(vote_variance$variance[[2]], accuracy = 0.1), ")"),
    color = "Author"
  )
```

## 15. Cluster chapters into style modes

Here "style modes" are clusters of chapters with similar ensemble vote profiles.

```{r mclust-style-modes}
set.seed(1830)

apply(book_profile_matrix, 2, sd) %>%
  sort(decreasing = TRUE)

mc_modes <- mclust::Mclust(
  book_profile_matrix,
  G = 1:10
)

summary(mc_modes)

mode_assignments <- book_profile_wide %>%
  dplyr::select(doc_id, author, author_code, short_name, source_title) %>%
  dplyr::mutate(
    style_mode = as.integer(mc_modes$classification),
    uncertainty = as.numeric(mc_modes$uncertainty)
  ) %>%
  dplyr::left_join(chapter_entropy, by = c("doc_id", "author", "author_code", "short_name", "source_title"))

mode_assignments %>%
  dplyr::arrange(style_mode, author, short_name) %>%
  print(n = 200)

readr::write_csv(mode_assignments, out_mode_assignments)
```

```{r plot-style-modes-pca, fig.width=8, fig.height=6}
ggplot(
  mode_assignments %>%
    dplyr::left_join(
      vote_scores %>%
        dplyr::select(doc_id, PC1, PC2),
      by = "doc_id"
    ),
  aes(
    x = PC1,
    y = PC2,
    color = factor(style_mode),
    shape = author
  )
) +
  geom_point(size = 3, alpha = 0.85) +
  theme_bw() +
  labs(
    title = "Ensemble-derived style modes",
    subtitle = "Mclust clusters of chapter-level BOM book vote profiles",
    x = paste0("PC1 (", scales::percent(vote_variance$variance[[1]], accuracy = 0.1), ")"),
    y = paste0("PC2 (", scales::percent(vote_variance$variance[[2]], accuracy = 0.1), ")"),
    color = "Style mode",
    shape = "Author"
  )
```

## 16. Author occupancy of style modes

```{r author-mode-summary}
author_mode_counts <- mode_assignments %>%
  dplyr::count(author, style_mode, name = "n_docs") %>%
  dplyr::group_by(author) %>%
  dplyr::mutate(
    author_total = sum(n_docs),
    pct_author = n_docs / author_total
  ) %>%
  dplyr::ungroup()

author_mode_counts %>%
  dplyr::arrange(author, style_mode) %>%
  print(n = 100)

author_mode_entropy <- author_mode_counts %>%
  dplyr::group_by(author) %>%
  dplyr::summarise(
    n_style_modes_occupied = dplyr::n_distinct(style_mode),
    mode_entropy = entropy_bits(pct_author),
    effective_style_modes = 2^mode_entropy,
    .groups = "drop"
  )

author_style_mode_summary <- within_profile_summary %>%
  dplyr::left_join(
    chapter_entropy %>%
      dplyr::group_by(author) %>%
      dplyr::summarise(
        mean_book_entropy = mean(book_entropy),
        median_book_entropy = median(book_entropy),
        mean_effective_books = mean(effective_books),
        .groups = "drop"
      ),
    by = "author"
  ) %>%
  dplyr::left_join(author_mode_entropy, by = "author") %>%
  dplyr::arrange(dplyr::desc(effective_style_modes))

author_style_mode_summary

readr::write_csv(author_style_mode_summary, out_author_mode_summary)
```

```{r plot-author-mode-occupancy, fig.width=8, fig.height=5}
ggplot(
  author_mode_counts,
  aes(
    x = author,
    y = pct_author,
    fill = factor(style_mode)
  )
) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(
    title = "Author occupancy of ensemble-derived style modes",
    subtitle = "Mode assignments from Mclust over BOM book vote profiles",
    x = "Author",
    y = "Share of author's chapters",
    fill = "Style mode"
  )
```

## 17. Heatmap: chapter × BOM book profiles

```{r chapter-book-profile-heatmap, fig.width=12, fig.height=10}
top_order <- chapter_book_profiles_long %>%
  dplyr::group_by(target_book) %>%
  dplyr::summarise(total = sum(weighted_votes), .groups = "drop") %>%
  dplyr::arrange(dplyr::desc(total)) %>%
  dplyr::pull(target_book)

chapter_order <- mode_assignments %>%
  dplyr::arrange(author, style_mode, short_name) %>%
  dplyr::pull(short_name)

chapter_book_profiles_long %>%
  dplyr::mutate(
    target_book = factor(target_book, levels = top_order),
    short_name = factor(short_name, levels = rev(chapter_order))
  ) %>%
  ggplot(
    aes(
      x = target_book,
      y = short_name,
      fill = weighted_vote_share
    )
  ) +
  geom_tile(color = "grey85") +
  scale_fill_gradient(
    low = "white",
    high = "darkgreen",
    labels = scales::percent
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1)
  ) +
  labs(
    title = "JSO chapter × Book of Mormon book vote-profile heatmap",
    subtitle = "Row-normalized weighted nearest-neighbor vote shares",
    x = "Book of Mormon book",
    y = "JSO chapter",
    fill = "Weighted\nshare"
  )
```

## 18. Top style-mode signatures

Which Book of Mormon books define each ensemble style mode?

```{r mode-signatures}
mode_signatures <- chapter_book_profiles_long %>%
  dplyr::left_join(
    mode_assignments %>%
      dplyr::select(doc_id, style_mode),
    by = "doc_id"
  ) %>%
  dplyr::group_by(style_mode, target_book) %>%
  dplyr::summarise(
    mode_weighted_share = mean(weighted_vote_share, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(style_mode) %>%
  dplyr::slice_max(mode_weighted_share, n = 5, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(style_mode, dplyr::desc(mode_weighted_share))

mode_signatures %>%
  print(n = 100)
```

## 19. Optional: permutation test for profile distance

This tests whether each author's within-author profile dispersion is unusually high or low relative to random author labels.

```{r permutation-profile-distance, cache=TRUE}
set.seed(1830)

observed_profile_means <- within_profile_distances %>%
  dplyr::group_by(author_pair) %>%
  dplyr::summarise(
    observed_mean_profile_distance = mean(distance),
    .groups = "drop"
  )

compute_perm_profile_means <- function(labels){

  tmp <- tibble::tibble(
    doc_id = rownames(book_profile_dist),
    perm_author = labels
  )

  expand.grid(
    doc_id_1 = rownames(book_profile_dist),
    doc_id_2 = colnames(book_profile_dist),
    stringsAsFactors = FALSE
  ) %>%
    dplyr::filter(doc_id_1 < doc_id_2) %>%
    dplyr::mutate(distance = book_profile_dist[cbind(doc_id_1, doc_id_2)]) %>%
    dplyr::left_join(
      tmp %>% dplyr::rename(doc_id_1 = doc_id, author_1 = perm_author),
      by = "doc_id_1"
    ) %>%
    dplyr::left_join(
      tmp %>% dplyr::rename(doc_id_2 = doc_id, author_2 = perm_author),
      by = "doc_id_2"
    ) %>%
    dplyr::filter(author_1 == author_2) %>%
    dplyr::group_by(author_1) %>%
    dplyr::summarise(
      mean_distance = mean(distance),
      .groups = "drop"
    )
}

n_perm <- 1000

perm_profile_results <- purrr::map_dfr(seq_len(n_perm), function(i){
  compute_perm_profile_means(sample(book_profile_wide$author)) %>%
    dplyr::mutate(iter = i)
})

profile_permutation_summary <- observed_profile_means %>%
  dplyr::rename(author = author_pair) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    p_greater = mean(
      perm_profile_results$mean_distance[
        perm_profile_results$author_1 == author
      ] >= observed_mean_profile_distance
    ),
    p_less = mean(
      perm_profile_results$mean_distance[
        perm_profile_results$author_1 == author
      ] <= observed_mean_profile_distance
    )
  ) %>%
  dplyr::ungroup()

profile_permutation_summary
```

## 20. Interpretation guide

This notebook should be interpreted differently from Notebook 50.

Notebook 50 asked:

```text
Are an author's raw style features internally dispersed?
```

This notebook asks:

```text
Does an author's writing move among multiple ensemble-derived Book of Mormon
stylistic neighborhoods?
```

Evidence for broader style modes would include:

```text
1. Larger within-author distances in ensemble vote-profile space.
2. Higher entropy of chapter-level vote profiles.
3. More style modes occupied by an author.
4. Higher style-mode entropy / effective style modes.
5. Distinct clusters within an author's chapters, not one compact group.
```

Evidence against broader style modes would include:

```text
1. Low within-author vote-profile distance.
2. Low entropy.
3. Occupying few style modes.
4. A single dominant mode across most chapters.
```

## 21. Save objects

```{r save-objects}
saveRDS(
  list(
    result_jso_bom = result_jso_bom,
    nn_jso_bom = nn_jso_bom,
    chapter_book_profiles_long = chapter_book_profiles_long,
    book_profile_wide = book_profile_wide,
    book_profile_matrix = book_profile_matrix,
    chapter_mp_profiles_long = chapter_mp_profiles_long,
    mp_profile_wide = mp_profile_wide,
    chapter_entropy = chapter_entropy,
    profile_pairwise_distances = profile_pairwise_distances,
    within_profile_distances = within_profile_distances,
    within_profile_summary = within_profile_summary,
    vote_pca = vote_pca,
    vote_scores = vote_scores,
    mc_modes = mc_modes,
    mode_assignments = mode_assignments,
    author_mode_counts = author_mode_counts,
    author_style_mode_summary = author_style_mode_summary,
    mode_signatures = mode_signatures,
    profile_permutation_summary = profile_permutation_summary
  ),
  file.path(export_dir, "author_style_modes_objects.rds")
)
```



I think Notebook 51 should be rewritten around ensemble affinity space, not book summaries.

The book summaries are excellent for plots like the MP analyses, but they're too compressed for discovering latent style modes. You're throwing away roughly 94% of the information (239 chapters → 15 books) before clustering. Using the full chapter-level affinity space is much more likely to reveal meaningful structure if it exists.
