---
title: "61 - Prepare 1830 Comparator Corpus"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook prepares an external **1830-era comparator corpus** at the chapter/document level.

The target output is:

```text
data/prepared/corpus_1830_chapter_docs.csv
```

This corpus is intentionally separate from:

```text
KJV
BOM
BOC
SCR = Joseph Smith / Sidney Rigdon / Oliver Cowdery
```

The 1830 corpus serves as an external control group of roughly contemporary English, including fiction, historical romance, frontier narrative, biblical imitation, and religious prose.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
```

## 3. Directories

```{r directories}
raw_1830_dir <- file.path(
  STYLO_ROOT,
  "data",
  "raw",
  "1830"
)

prepared_dir <- file.path(
  STYLO_ROOT,
  "data",
  "prepared"
)

dir.create(raw_1830_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

out_path <- file.path(prepared_dir, "corpus_1830_chapter_docs.csv")
raw_index_path <- file.path(raw_1830_dir, "corpus_1830_raw_index.csv")
```

## 4. Text manifest

Project Gutenberg IDs are used where available. Spaulding, Campbell, Edwards, and View of the Hebrews may require local/manual files depending on source quality.

```{r manifest}
text_manifest <- tibble::tribble(
  ~author, ~work, ~short_work, ~year, ~source_type, ~gutenberg_id, ~local_file, ~doc_unit, ~expected_genre,

  "Jane Austen", "Pride and Prejudice", "Pride_and_Prejudice", 1813, "gutenberg", 1342, NA_character_, "chapter", "Fiction",
  "Jane Austen", "Emma", "Emma", 1815, "gutenberg", 158, NA_character_, "chapter", "Fiction",
  "Jane Austen", "Persuasion", "Persuasion", 1817, "gutenberg", 105, NA_character_, "chapter", "Fiction",

  "James Fenimore Cooper", "The Pioneers", "The_Pioneers", 1823, "gutenberg", 2275, NA_character_, "chapter", "Frontier_Fiction",
  "James Fenimore Cooper", "The Last of the Mohicans", "Last_of_the_Mohicans", 1826, "gutenberg", 940, NA_character_, "chapter", "Frontier_Fiction",
  "James Fenimore Cooper", "The Prairie", "The_Prairie", 1827, "gutenberg", 6450, NA_character_, "chapter", "Frontier_Fiction",

  "Walter Scott", "Ivanhoe", "Ivanhoe", 1819, "gutenberg", 82, NA_character_, "chapter", "Historical_Romance",
  "Walter Scott", "Rob Roy", "Rob_Roy", 1817, "gutenberg", 7025, NA_character_, "chapter", "Historical_Romance",
  "Walter Scott", "The Talisman", "The_Talisman", 1825, "gutenberg", 1377, NA_character_, "chapter", "Historical_Romance",

  "Ethan Smith", "View of the Hebrews", "View_of_the_Hebrews", 1823, "local", NA, "ethan_smith/view_of_the_hebrews.txt", "chapter", "Religious_Prose",

  "Solomon Spaulding", "Manuscript Story", "Manuscript_Story", 1812, "local", NA, "spaulding/manuscript_story.txt", "chapter", "Biblical_Imitation",
  "Solomon Spaulding", "Conneaut Creek Manuscript", "Conneaut_Creek", 1812, "local", NA, "spaulding/conneaut_creek.txt", "chapter", "Biblical_Imitation",

  "Jonathan Edwards", "Selected Sermons", "Selected_Sermons", 1741, "local", NA, "edwards/selected_sermons.txt", "sermon", "Sermon",

  "Alexander Campbell", "The Christian Baptist", "Christian_Baptist", 1823, "local_folder", NA, "campbell/christian_baptist", "article", "Restoration_Prose"
)

text_manifest

readr::write_csv(text_manifest, raw_index_path)
```

## 5. Download helpers

```{r download-helpers}
safe_slug <- function(x){
  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")
}

gutenberg_candidate_urls <- function(id){
  id <- as.integer(id)
  c(
    paste0("https://www.gutenberg.org/files/", id, "/", id, "-0.txt"),
    paste0("https://www.gutenberg.org/files/", id, "/", id, ".txt"),
    paste0("https://www.gutenberg.org/cache/epub/", id, "/pg", id, ".txt"),
    paste0("https://www.gutenberg.org/cache/epub/", id, "/pg", id, "-0.txt")
  )
}

download_gutenberg_text <- function(id, destfile){
  urls <- gutenberg_candidate_urls(id)

  for(u in urls){
    ok <- tryCatch({
      download.file(url = u, destfile = destfile, mode = "wb", quiet = TRUE)
      file.exists(destfile) && file.info(destfile)$size > 1000
    }, error = function(e){
      FALSE
    })

    if(isTRUE(ok)){
      return(destfile)
    }
  }

  warning("Could not download Gutenberg ID: ", id)
  NA_character_
}

local_raw_path <- function(local_file){
  file.path(raw_1830_dir, local_file)
}

get_raw_text_path <- function(row){
  if(row$source_type == "gutenberg"){
    author_dir <- file.path(raw_1830_dir, safe_slug(row$author))
    dir.create(author_dir, recursive = TRUE, showWarnings = FALSE)

    destfile <- file.path(author_dir, paste0(row$short_work, ".txt"))

    if(!file.exists(destfile)){
      download_gutenberg_text(row$gutenberg_id, destfile)
    }

    return(destfile)
  }

  if(row$source_type %in% c("local", "local_folder")){
    return(local_raw_path(row$local_file))
  }

  NA_character_
}
```

## 6. Cleanup helpers

```{r cleanup-helpers}
read_text_file <- function(path){
  if(!file.exists(path)){
    return(NA_character_)
  }

  paste(
    readr::read_lines(path, locale = readr::locale(encoding = "UTF-8"), progress = FALSE),
    collapse = "\n"
  )
}

clean_text_basic <- function(x){
  x %>%
    stringr::str_replace_all("\uFEFF", "") %>%
    stringr::str_replace_all("\r", "\n") %>%
    stringr::str_replace_all("\u00A0", " ") %>%
    stringr::str_replace_all("[ \t]+", " ") %>%
    stringr::str_replace_all("\n{3,}", "\n\n") %>%
    stringr::str_trim()
}

strip_gutenberg_boilerplate <- function(x){
  x <- clean_text_basic(x)

  start_loc <- stringr::str_locate(
    x,
    regex("\\*\\*\\* START OF (THE|THIS) PROJECT GUTENBERG EBOOK.*?\\*\\*\\*", ignore_case = TRUE, dotall = TRUE)
  )

  if(!any(is.na(start_loc))){
    x <- stringr::str_sub(x, start_loc[2] + 1)
  }

  end_loc <- stringr::str_locate(
    x,
    regex("\\*\\*\\* END OF (THE|THIS) PROJECT GUTENBERG EBOOK.*?\\*\\*\\*|End of the Project Gutenberg EBook.*$", ignore_case = TRUE, dotall = TRUE)
  )

  if(!any(is.na(end_loc))){
    x <- stringr::str_sub(x, 1, end_loc[1] - 1)
  }

  clean_text_basic(x)
}

normalize_document_text <- function(x){
  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()
}

make_1830_doc_id <- function(author, work, doc_number){
  paste0(
    "1830__",
    safe_slug(author),
    "__",
    safe_slug(work),
    "__",
    doc_number
  )
}
```

## 7. Chapter parsing helpers

```{r chapter-helpers}

ordinal_words <- c(
  FIRST = 1, SECOND = 2, THIRD = 3, FOURTH = 4, FIFTH = 5,
  SIXTH = 6, SEVENTH = 7, EIGHTH = 8, NINTH = 9, TENTH = 10,
  ELEVENTH = 11, TWELFTH = 12, THIRTEENTH = 13, FOURTEENTH = 14,
  FIFTEENTH = 15, SIXTEENTH = 16, SEVENTEENTH = 17,
  EIGHTEENTH = 18, NINETEENTH = 19, TWENTIETH = 20,
  `TWENTY-FIRST` = 21, `TWENTY-SECOND` = 22, `TWENTY-THIRD` = 23,
  `TWENTY-FOURTH` = 24, `TWENTY-FIFTH` = 25,
  `TWENTY-SIXTH` = 26, `TWENTY-SEVENTH` = 27,
  `TWENTY-EIGHTH` = 28, `TWENTY-NINTH` = 29,
  THIRTIETH = 30, `THIRTY-FIRST` = 31,
  `THIRTY-SECOND` = 32, `THIRTY-THIRD` = 33,
  `THIRTY-FOURTH` = 34, `THIRTY-FIFTH` = 35,
  `THIRTY-SIXTH` = 36, `THIRTY-SEVENTH` = 37,
  `THIRTY-EIGHTH` = 38, `THIRTY-NINTH` = 39,
  FORTIETH = 40, `FORTY-FIRST` = 41,
  `FORTY-SECOND` = 42, `FORTY-THIRD` = 43,
  `FORTY-FOURTH` = 44, `FORTY-FIFTH` = 45
)

word_ordinal_to_integer <- function(x){

  x <- toupper(
    str_squish(x)
  )

  ordinal_words[[x]]

}

roman_to_integer <- function(x){
  x <- toupper(stringr::str_squish(x))

  roman_values <- c(
    I = 1L, V = 5L, X = 10L, L = 50L,
    C = 100L, D = 500L, M = 1000L
  )

  chars <- strsplit(x, "")[[1]]

  if(any(!chars %in% names(roman_values))){
    return(NA_integer_)
  }

  vals <- roman_values[chars]
  total <- 0L

  for(i in seq_along(vals)){
    if(i < length(vals) && vals[i] < vals[i + 1]){
      total <- total - vals[i]
    } else {
      total <- total + vals[i]
    }
  }

  as.integer(total)
}

chapter_heading_pattern <- regex(
  paste0(
    "(?m)^\\s*",
    "(?:BOOK\\s+[A-ZIVXLCDM]+\\.?\\s*)?",
    "(?:VOLUME\\s+[A-ZIVXLCDM]+\\.?\\s*)?",
    "(?:CHAPTER|Chapter|CHAP\\.?|Chap\\.?)\\s+",
    "([0-9]+|[IVXLCDM]+|[A-Z]+(?:-[A-Z]+)?)",
    "\\.?",
    "(?:\\s*[-—.:]\\s*(.*))?",
    "\\s*$"
  ),
  ignore_case = FALSE
)

find_chapter_headings <- function(x){
  locs <- stringr::str_locate_all(x, chapter_heading_pattern)[[1]]
  matches <- stringr::str_match_all(x, chapter_heading_pattern)[[1]]

  if(nrow(locs) == 0){
    return(tibble::tibble())
  }

  tibble::tibble(
    match_start = locs[, "start"],
    match_end = locs[, "end"],
    heading = matches[, 1],
    chapter_raw = matches[, 2],
    chapter_title = matches[, 3]
  ) %>%
    dplyr::mutate(
      chapter = dplyr::case_when(
          stringr::str_detect(chapter_raw, "^[0-9]+$") ~
            as.integer(chapter_raw),
        
          stringr::str_detect(chapter_raw, "^[IVXLCDM]+$") ~
            roman_to_integer(chapter_raw),
        
          TRUE ~
            word_ordinal_to_integer(chapter_raw)
        ),
      chapter_title = stringr::str_squish(chapter_title)
    )
}

parse_chapters_from_text <- function(x){
  x <- clean_text_basic(x)
  headings <- find_chapter_headings(x)

  if(nrow(headings) == 0){
    return(
      tibble::tibble(
        chapter = 1L,
        chapter_title = NA_character_,
        text = normalize_document_text(x)
      )
    )
  }

  headings <- headings %>%
    dplyr::arrange(match_start) %>%
    dplyr::mutate(
      text_start = match_end + 1L,
      text_end = dplyr::lead(match_start, default = stringr::str_length(x) + 1L) - 1L
    )

  headings %>%
    dplyr::transmute(
      chapter = chapter,
      chapter_title = chapter_title,
      text = purrr::map2_chr(
        text_start,
        text_end,
        ~stringr::str_sub(x, .x, .y) %>% normalize_document_text()
      )
    ) %>%
    dplyr::filter(!is.na(text), text != "")
}
```

## 8. Sermon/article parsing helpers

For local sermon/article files, use one of these markers before each document:

```text
### Title of document
SERMON: Title
ARTICLE: Title
```

If no markers are found, the entire file becomes one document.

```{r document-parsers}
parse_marked_documents <- function(x, default_title = NA_character_){
  x <- clean_text_basic(x)
  lines <- stringr::str_split(x, "\n")[[1]]

  marker_idx <- which(
    stringr::str_detect(
      lines,
      regex("^\\s*(###\\s+|SERMON:\\s+|ARTICLE:\\s+)", ignore_case = TRUE)
    )
  )

  if(length(marker_idx) == 0){
    return(
      tibble::tibble(
        chapter = 1L,
        chapter_title = default_title,
        text = normalize_document_text(x)
      )
    )
  }

  marker_end <- c(marker_idx[-1] - 1L, length(lines))

  purrr::map2_dfr(
    marker_idx,
    marker_end,
    function(s, e){
      title <- lines[s] %>%
        stringr::str_replace(regex("^\\s*(###\\s+|SERMON:\\s+|ARTICLE:\\s+)", ignore_case = TRUE), "") %>%
        stringr::str_squish()

      body <- if(s < e){
        paste(lines[(s + 1L):e], collapse = "\n")
      } else {
        ""
      }

      tibble::tibble(
        chapter = NA_integer_,
        chapter_title = title,
        text = normalize_document_text(body)
      )
    }
  ) %>%
    dplyr::mutate(chapter = dplyr::row_number()) %>%
    dplyr::filter(!is.na(text), text != "")
}

parse_local_folder_articles <- function(folder_path){
  if(!dir.exists(folder_path)){
    return(tibble::tibble())
  }

  files <- list.files(folder_path, pattern = "\\.txt$", full.names = TRUE)

  if(length(files) == 0){
    return(tibble::tibble())
  }

  purrr::map_dfr(
    files,
    function(f){
      x <- read_text_file(f)

      tibble::tibble(
        chapter_title = tools::file_path_sans_ext(basename(f)),
        text = normalize_document_text(x)
      )
    }
  ) %>%
    dplyr::mutate(chapter = dplyr::row_number()) %>%
    dplyr::select(chapter, chapter_title, text)
}
```

## 9. Parse one manifest row

```{r parse-one}
parse_manifest_row <- function(row){
  raw_path <- get_raw_text_path(row)

  if(is.na(raw_path) || !file.exists(raw_path)){
    warning(
      "Missing raw file for ",
      row$author,
      " / ",
      row$work,
      ": ",
      raw_path
    )

    return(tibble::tibble())
  }

  if(row$source_type == "local_folder"){
    docs <- parse_local_folder_articles(raw_path)
  } else {
    raw_text <- read_text_file(raw_path)

    if(row$source_type == "gutenberg"){
      raw_text <- strip_gutenberg_boilerplate(raw_text)
    } else {
      raw_text <- clean_text_basic(raw_text)
    }

    docs <- if(row$doc_unit == "chapter"){
      parse_chapters_from_text(raw_text)
    } else {
      parse_marked_documents(raw_text, default_title = row$work)
    }
  }

  if(nrow(docs) == 0){
    return(tibble::tibble())
  }

  docs %>%
    dplyr::mutate(
      corpus_id = "1830",
      translation = as.character(row$year),
      corpus_family = "1830",
      source_family = paste0("1830_", safe_slug(row$author)),
      author = row$author,
      work = row$work,
      short_work = row$short_work,
      year = as.integer(row$year),
      book = paste0(row$author, " - ", row$work),
      section = row$expected_genre,
      macro_genre = row$expected_genre,
      doc_unit = row$doc_unit,
      chapter = as.integer(chapter),
      n_verses = NA_integer_,
      n_words = stringr::str_count(text, "\\S+"),
      doc_id = make_1830_doc_id(row$author, row$short_work, chapter),
      doc_id_original = doc_id
    ) %>%
    dplyr::select(
      corpus_id,
      translation,
      corpus_family,
      source_family,
      doc_id,
      doc_id_original,
      author,
      work,
      short_work,
      year,
      book,
      chapter,
      chapter_title,
      section,
      macro_genre,
      doc_unit,
      n_verses,
      n_words,
      text
    )
}
```

## 10. Download and parse corpus

```{r parse-corpus}
corpus_1830_docs <- purrr::pmap_dfr(
  text_manifest,
  function(author, work, short_work, year, source_type, gutenberg_id, local_file, doc_unit, expected_genre){
    row <- tibble::tibble(
      author = author,
      work = work,
      short_work = short_work,
      year = year,
      source_type = source_type,
      gutenberg_id = gutenberg_id,
      local_file = local_file,
      doc_unit = doc_unit,
      expected_genre = expected_genre
    )

    parse_manifest_row(row)
  }
)

corpus_1830_docs %>%
  dplyr::count(author, work, name = "n_docs") %>%
  dplyr::arrange(author, work)

corpus_1830_docs %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    n_authors = dplyr::n_distinct(author),
    n_works = dplyr::n_distinct(work),
    min_words = min(n_words, na.rm = TRUE),
    median_words = median(n_words, na.rm = TRUE),
    mean_words = mean(n_words, na.rm = TRUE),
    max_words = max(n_words, na.rm = TRUE)
  )
```

## 11. Filter tiny documents

```{r filter-tiny}
min_words_1830 <- 100

corpus_1830_docs_raw_parsed <- corpus_1830_docs

corpus_1830_chapter_docs <- corpus_1830_docs %>%
  dplyr::filter(n_words >= min_words_1830) %>%
  dplyr::arrange(author, work, chapter)

corpus_1830_docs_raw_parsed %>%
  dplyr::count(author, work, name = "n_docs_raw") %>%
  dplyr::left_join(
    corpus_1830_chapter_docs %>%
      dplyr::count(author, work, name = "n_docs_kept"),
    by = c("author", "work")
  ) %>%
  dplyr::mutate(
    n_docs_kept = dplyr::coalesce(n_docs_kept, 0L),
    n_dropped = n_docs_raw - n_docs_kept
  ) %>%
  dplyr::arrange(author, work)
```

## 12. Sanity checks

```{r sanity-checks}
corpus_1830_chapter_docs %>%
  dplyr::count(author, work, macro_genre, name = "n_docs") %>%
  dplyr::arrange(author, work)

corpus_1830_chapter_docs %>%
  dplyr::group_by(author, work) %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    min_chapter = min(chapter, na.rm = TRUE),
    max_chapter = max(chapter, na.rm = TRUE),
    min_words = min(n_words, na.rm = TRUE),
    median_words = median(n_words, na.rm = TRUE),
    mean_words = mean(n_words, na.rm = TRUE),
    max_words = max(n_words, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(author, work)

corpus_1830_chapter_docs %>%
  dplyr::arrange(n_words) %>%
  dplyr::select(author, work, chapter, chapter_title, n_words) %>%
  dplyr::slice_head(n = 30)

corpus_1830_chapter_docs %>%
  dplyr::arrange(dplyr::desc(n_words)) %>%
  dplyr::select(author, work, chapter, chapter_title, n_words) %>%
  dplyr::slice_head(n = 30)
```

## 13. Write outputs

```{r write-outputs}
raw_parsed_path <- file.path(raw_1830_dir, "corpus_1830_docs_raw_parsed.csv")

readr::write_csv(corpus_1830_docs_raw_parsed, raw_parsed_path)
readr::write_csv(corpus_1830_chapter_docs, out_path)

cat("Wrote raw parsed 1830 corpus to:\n")
cat(raw_parsed_path, "\n\n")

cat("Wrote prepared 1830 chapter corpus to:\n")
cat(out_path, "\n")
```

## 14. Manual-source instructions

If the local/manual sources are missing, create these files:

```text
data/raw/1830/ethan_smith/view_of_the_hebrews.txt
data/raw/1830/spaulding/manuscript_story.txt
data/raw/1830/spaulding/conneaut_creek.txt
data/raw/1830/edwards/selected_sermons.txt
```

For Campbell articles, create one `.txt` file per article in:

```text
data/raw/1830/campbell/christian_baptist/
```

For sermon/article collections, you may either use one file per document or use markers:

```text
### Document Title
SERMON: Document Title
ARTICLE: Document Title
```

The notebook will treat each marked section as a separate document.

## 15. Next step

After this notebook succeeds, add:

```text
data/prepared/corpus_1830_chapter_docs.csv
```

to the larger analysis notebook:

```text
KJV + BOM + BOC + SCR + 1830
```

Keep `SCR` separate from the `1830` external comparator corpus.
