---
title: "61 - Prepare 1830 Corpus: Washington Irving"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)

library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(tidyr)
```

## 1. Purpose

Prepare Washington Irving chapter/essay-level documents for the 1830 secular control corpus.

Yes: this notebook uses **Project Gutenberg** for the raw sources.

Recommended Irving corpus:

```text
The Sketch Book of Geoffrey Crayon, Gent.   1819-1820
Bracebridge Hall                            1822
Tales of a Traveller                        1824
```

These are all pre-1830 and give a strong American prose control set.

Outputs:

```text
data/prepared/corpus_1830_irving_chapter_docs.csv
data/prepared/corpus_1830_irving_manifest.csv
```

Metadata convention:

```text
author = "Washington Irving"
corpus_family = "1830"
source_family = "1830_Secular"
macro_genre = "American_Prose"
doc_unit = "chapter"
```

## 2. Directories

```{r dirs}
raw_1830_dir <- file.path(STYLO_ROOT, "data", "raw", "corpus_1830")
raw_irving_dir <- file.path(raw_1830_dir, "irving")
manual_irving_dir <- file.path(raw_irving_dir, "manual")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
diagnostic_dir <- file.path(raw_irving_dir, "diagnostics")

for(d in c(raw_1830_dir, raw_irving_dir, manual_irving_dir, prepared_dir, diagnostic_dir)){
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
}

irving_out_path <- file.path(prepared_dir, "corpus_1830_irving_chapter_docs.csv")
irving_manifest_out_path <- file.path(prepared_dir, "corpus_1830_irving_manifest.csv")
download_diag_path <- file.path(diagnostic_dir, "irving_download_results.csv")
parse_diag_path <- file.path(diagnostic_dir, "irving_parse_diagnostics.csv")
```

## 3. Manifest

Gutenberg IDs may vary by edition. These IDs are the common Project Gutenberg plain-text targets:

```text
2048   The Sketch Book
13515  Bracebridge Hall
13514  Tales of a Traveller
```

If a download fails, update the `gutenberg_id` value and rerun.

```{r manifest}
irving_manifest <- tibble::tribble(
  ~author,               ~work,                                      ~short_work,          ~year, ~gutenberg_id, ~expected_genre,
  "Washington Irving",   "The Sketch Book of Geoffrey Crayon, Gent.", "Sketch_Book",        1820,  2048,          "American_Prose",
  "Washington Irving",   "Bracebridge Hall",                         "Bracebridge_Hall",   1822,  13515,         "American_Prose",
  "Washington Irving",   "Tales of a Traveller",                     "Tales_Traveller",    1824,  13514,         "American_Prose"
)

irving_manifest
```

## 4. Download helpers

```{r download-helpers}
gutenberg_urls <- function(id){
  c(
    paste0("https://www.gutenberg.org/files/", id, "/", id, "-0.txt"),
    paste0("https://www.gutenberg.org/files/", id, "/", id, ".txt"),
    paste0("https://www.gutenberg.org/cache/epub/", id, "/pg", id, ".txt")
  )
}

raw_path_for_work <- function(short_work){
  file.path(raw_irving_dir, paste0(short_work, ".txt"))
}

manual_path_for_work <- function(short_work){
  file.path(manual_irving_dir, paste0(short_work, "_manual.txt"))
}

download_gutenberg_text <- function(gutenberg_id, short_work, overwrite = FALSE){

  raw_path <- raw_path_for_work(short_work)

  if(file.exists(raw_path) && !overwrite){
    return(
      tibble::tibble(
        short_work = short_work,
        gutenberg_id = gutenberg_id,
        raw_path = raw_path,
        url = NA_character_,
        ok = TRUE,
        status = "cached"
      )
    )
  }

  urls <- gutenberg_urls(gutenberg_id)

  for(u in urls){

    ok <- tryCatch({

      download.file(
        url = u,
        destfile = raw_path,
        mode = "wb",
        quiet = TRUE
      )

      file.exists(raw_path) && file.info(raw_path)$size > 1000

    }, error = function(e){
      FALSE
    })

    if(ok){
      return(
        tibble::tibble(
          short_work = short_work,
          gutenberg_id = gutenberg_id,
          raw_path = raw_path,
          url = u,
          ok = TRUE,
          status = "downloaded"
        )
      )
    }
  }

  tibble::tibble(
    short_work = short_work,
    gutenberg_id = gutenberg_id,
    raw_path = raw_path,
    url = paste(urls, collapse = " | "),
    ok = FALSE,
    status = "failed"
  )
}
```

## 5. Download raw sources

```{r download}
download_results <- purrr::pmap_dfr(

  irving_manifest,

  function(author,
           work,
           short_work,
           year,
           gutenberg_id,
           expected_genre){

    out <- download_gutenberg_text(
      gutenberg_id,
      short_work
    )

    bind_cols(

      tibble(
        author=author,
        work=work,
        short_work=short_work,
        year=year,
        gutenberg_id=gutenberg_id,
        expected_genre=expected_genre
      ),

      out

    )

  }

)


download_results <- download_results %>%
  rename(
    short_work = short_work...3,
    gutenberg_id = gutenberg_id...5
  ) %>%
  select(
    -short_work...7,
    -gutenberg_id...8
  )


download_results

readr::write_csv(download_results, download_diag_path)

if(any(!download_results$ok)){
  download_results %>% filter(!ok) %>% print(n = 20)
  stop("One or more Irving Gutenberg downloads failed. Update Gutenberg IDs or provide manual text files.")
}
```

## 6. Create manual copies

This mirrors your existing 1830 workflow: raw downloads are copied to `manual/` so you can edit chapter markers if needed.

```{r create-manual-copies}
manual_copy_results <- download_results %>%
  mutate(
    manual_path = manual_path_for_work(short_work),
    copied = purrr::map2_lgl(
      raw_path,
      manual_path,
      ~{
        if(!file.exists(.y)){
          file.copy(.x, .y, overwrite = FALSE)
        } else {
          TRUE
        }
      }
    )
  )

manual_copy_results %>%
  select(short_work, raw_path, manual_path, copied)

if(any(!manual_copy_results$copied)){
  stop("One or more manual copies failed.")
}

################ let's stop here and review manual...
```

## 7. Text cleaning helpers

```{r text-helpers}
normalize_doc_text <- function(x){

  x %>%
    stringr::str_replace_all("\r", "\n") %>%
    stringr::str_replace_all("\uFEFF", "") %>%
    stringr::str_replace_all("\u00A0", " ") %>%
    stringr::str_replace_all("[ \t]+", " ") %>%
    stringr::str_replace_all("\n{3,}", "\n\n") %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()

}

strip_gutenberg_boilerplate <- function(txt){

  start_patterns <- c(
    "\\*\\*\\* START OF (THE|THIS) PROJECT GUTENBERG EBOOK .*?\\*\\*\\*",
    "\\*\\*\\* START OF THE PROJECT GUTENBERG EBOOK .*?\\*\\*\\*",
    "START OF (THE|THIS) PROJECT GUTENBERG EBOOK"
  )

  end_patterns <- c(
    "\\*\\*\\* END OF (THE|THIS) PROJECT GUTENBERG EBOOK .*?\\*\\*\\*",
    "\\*\\*\\* END OF THE PROJECT GUTENBERG EBOOK .*?\\*\\*\\*",
    "END OF (THE|THIS) PROJECT GUTENBERG EBOOK"
  )

  for(p in start_patterns){
    loc <- stringr::str_locate(txt, regex(p, ignore_case = TRUE, dotall = TRUE))[1,]
    if(!is.na(loc[2])){
      txt <- substr(txt, loc[2] + 1, nchar(txt))
      break
    }
  }

  for(p in end_patterns){
    loc <- stringr::str_locate(txt, regex(p, ignore_case = TRUE, dotall = TRUE))[1,]
    if(!is.na(loc[1])){
      txt <- substr(txt, 1, loc[1] - 1)
      break
    }
  }

  txt
}

clean_heading <- function(x){

  x %>%
    stringr::str_replace_all("[\\*_]", "") %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_squish()

}
```

## 8. Chapter heading parser

This parser handles Irving essay/chapter headings in several forms:

```text
CHAPTER I.
CHAPTER FIRST.
THE AUTHOR'S ACCOUNT OF HIMSELF.
RIP VAN WINKLE.
THE LEGEND OF SLEEPY HOLLOW.
```

For Gutenberg Irving, not every unit is labeled `CHAPTER`. Therefore this parser also has a fallback list of known essay/story titles for each work.

```{r heading-parser}
roman_to_int <- function(x){

  if(is.na(x) || x == "") return(NA_integer_)

  vals <- c(I=1, V=5, X=10, L=50, C=100, D=500, M=1000)
  chars <- strsplit(toupper(x), "")[[1]]
  nums <- vals[chars]

  if(any(is.na(nums))) return(NA_integer_)

  total <- 0
  i <- 1

  while(i <= length(nums)){
    if(i < length(nums) && nums[i] < nums[i+1]){
      total <- total + nums[i+1] - nums[i]
      i <- i + 2
    } else {
      total <- total + nums[i]
      i <- i + 1
    }
  }

  as.integer(total)
}

ordinal_words <- c(
  FIRST=1, SECOND=2, THIRD=3, FOURTH=4, FIFTH=5,
  SIXTH=6, SEVENTH=7, EIGHTH=8, NINTH=9, TENTH=10,
  ELEVENTH=11, TWELFTH=12, THIRTEENTH=13, FOURTEENTH=14,
  FIFTEENTH=15, SIXTEENTH=16, SEVENTEENTH=17, EIGHTEENTH=18,
  NINETEENTH=19, TWENTIETH=20
)

chapter_heading_regex <- regex(
  paste0(
    "(?m)^\\s*",
    "(?:CHAPTER|Chapter)\\s+",
    "([0-9]+|[IVXLCDM]+|FIRST|SECOND|THIRD|FOURTH|FIFTH|SIXTH|SEVENTH|EIGHTH|NINTH|TENTH|ELEVENTH|TWELFTH|THIRTEENTH|FOURTEENTH|FIFTEENTH|SIXTEENTH|SEVENTEENTH|EIGHTEENTH|NINETEENTH|TWENTIETH)",
    "\\.?",
    "(?:\\s*[-—.:]\\s*(.*))?",
    "\\s*$"
  ),
  ignore_case = FALSE
)

parse_chapter_number_token <- function(x){

  x0 <- toupper(stringr::str_squish(x))

  if(stringr::str_detect(x0, "^[0-9]+$")){
    return(as.integer(x0))
  }

  if(x0 %in% names(ordinal_words)){
    return(as.integer(ordinal_words[[x0]]))
  }

  roman_to_int(x0)
}

known_irving_titles <- list(

  Sketch_Book = c(
    "THE AUTHOR'S ACCOUNT OF HIMSELF",
    "THE VOYAGE",
    "ROSCOE",
    "THE WIFE",
    "RIP VAN WINKLE",
    "ENGLISH WRITERS ON AMERICA",
    "RURAL LIFE IN ENGLAND",
    "THE BROKEN HEART",
    "THE ART OF BOOK-MAKING",
    "A ROYAL POET",
    "THE COUNTRY CHURCH",
    "THE WIDOW AND HER SON",
    "A SUNDAY IN LONDON",
    "THE BOAR'S HEAD TAVERN",
    "THE MUTABILITY OF LITERATURE",
    "RURAL FUNERALS",
    "THE INN KITCHEN",
    "THE SPECTRE BRIDEGROOM",
    "WESTMINSTER ABBEY",
    "CHRISTMAS",
    "THE STAGE COACH",
    "CHRISTMAS EVE",
    "CHRISTMAS DAY",
    "THE CHRISTMAS DINNER",
    "LONDON ANTIQUES",
    "LITTLE BRITAIN",
    "STRATFORD-ON-AVON",
    "TRAITS OF INDIAN CHARACTER",
    "PHILIP OF POKANOKET",
    "JOHN BULL",
    "THE PRIDE OF THE VILLAGE",
    "THE ANGLER",
    "THE LEGEND OF SLEEPY HOLLOW",
    "L'ENVOY"
  ),

  Bracebridge_Hall = c(
    "THE HALL",
    "THE BUSY MAN",
    "FAMILY SERVANTS",
    "THE WIDOW",
    "THE LOVERS",
    "FAMILY RELIQUES",
    "AN OLD SOLDIER",
    "THE WIDOW'S RETINUE",
    "READY-MONEY JACK",
    "BACHELORS",
    "WIVES",
    "STORY TELLING",
    "THE STOUT GENTLEMAN",
    "FOREST TREES",
    "A LITERARY ANTIQUARY",
    "THE FARM-HOUSE",
    "HORSEMANSHIP",
    "LOVE SYMPTOMS",
    "FALCONRY",
    "HAWKING",
    "ST. MARK'S EVE",
    "GENTILITY",
    "FORTUNE-TELLING",
    "LOVE-CHARMS",
    "THE LIBRARY",
    "THE STUDENT OF SALAMANCA",
    "ENGLISH COUNTRY GENTLEMEN",
    "A BACHELOR'S CONFESSION",
    "ENGLISH GRAVITY",
    "GIPSIES",
    "MAY-DAY",
    "THE MANUSCRIPT",
    "ANNETTE DELARBRE",
    "TRAVELLING",
    "POPULAR SUPERSTITIONS",
    "THE CULPRIT",
    "FAMILY MISFORTUNES",
    "LOVERS' TROUBLES",
    "THE HISTORIAN",
    "THE HAUNTED HOUSE",
    "DOLPH HEYLIGER",
    "THE STORM-SHIP",
    "THE WEDDING"
  ),

  Tales_Traveller = c(
    "STRANGE STORIES BY A NERVOUS GENTLEMAN",
    "THE GREAT UNKNOWN",
    "THE HUNTING DINNER",
    "THE ADVENTURE OF MY UNCLE",
    "THE ADVENTURE OF MY AUNT",
    "THE BOLD DRAGOON",
    "THE ADVENTURE OF THE GERMAN STUDENT",
    "THE ADVENTURE OF THE MYSTERIOUS PICTURE",
    "THE ADVENTURE OF THE MYSTERIOUS STRANGER",
    "THE STORY OF THE YOUNG ITALIAN",
    "BUCKTHORNE AND HIS FRIENDS",
    "A LITERARY LIFE",
    "A LITERARY DINNER",
    "THE CLUB OF QUEER FELLOWS",
    "THE POOR DEVIL AUTHOR",
    "BUCKTHORNE, OR THE YOUNG MAN OF GREAT EXPECTATIONS",
    "THE BOOBY SQUIRE",
    "THE STROLLING MANAGER",
    "THE CREOLE VILLAGE",
    "THE GREAT MISSISSIPPI BUBBLE",
    "DON JUAN",
    "BROEK",
    "SKETCHES IN PARIS",
    "THE DEVIL AND TOM WALKER",
    "KIDD THE PIRATE",
    "WOLFERT WEBBER",
    "THE ADVENTURE OF SAM, THE BLACK FISHERMAN",
    "THE KNIGHT OF MALTA",
    "THE STUDENT OF SALAMANCA",
    "THE LEGEND OF THE ENGULPHED CONVENT",
    "THE COUNT VAN HORN"
  )
)
```

## 9. Split by known headings

This is the preferred method for Irving because these works are essay/story collections.

```{r split-known-headings}
heading_pattern_for_titles <- function(titles){

  escaped <- stringr::str_replace_all(titles, "([\\^\\$\\.\\|\\?\\*\\+\\(\\)\\[\\]\\{\\}])", "\\\\\\1")

  regex(
    paste0(
      "(?m)^\\s*(",
      paste(escaped, collapse = "|"),
      ")\\s*$"
    ),
    ignore_case = FALSE
  )
}

split_text_by_headings <- function(txt, titles){

  pattern <- heading_pattern_for_titles(titles)

  locs <- stringr::str_locate_all(txt, pattern)[[1]]

  if(nrow(locs) == 0){
    return(tibble::tibble())
  }

  headings <- stringr::str_match_all(txt, pattern)[[1]][, 2]

  pieces <- purrr::map_dfr(seq_len(nrow(locs)), function(i){

    start <- locs[i, 2] + 1
    end <- if(i < nrow(locs)) locs[i + 1, 1] - 1 else nchar(txt)

    body <- substr(txt, start, end)

    tibble::tibble(
      chapter = i,
      chapter_title = clean_heading(headings[[i]]),
      text = normalize_doc_text(body)
    )

  })

  pieces %>%
    filter(text != "")

}

split_text_by_chapter_regex <- function(txt){

  locs <- stringr::str_locate_all(txt, chapter_heading_regex)[[1]]

  if(nrow(locs) == 0){
    return(tibble::tibble())
  }

  matches <- stringr::str_match_all(txt, chapter_heading_regex)[[1]]

  purrr::map_dfr(seq_len(nrow(locs)), function(i){

    token <- matches[i, 2]
    subtitle <- matches[i, 3]

    start <- locs[i, 2] + 1
    end <- if(i < nrow(locs)) locs[i + 1, 1] - 1 else nchar(txt)

    chapter <- parse_chapter_number_token(token)

    chapter_title <- ifelse(
      is.na(subtitle) || subtitle == "",
      paste0("Chapter ", token),
      clean_heading(subtitle)
    )

    tibble::tibble(
      chapter = chapter,
      chapter_title = chapter_title,
      text = normalize_doc_text(substr(txt, start, end))
    )

  }) %>%
    filter(!is.na(chapter), text != "")
}
```

## 10. Parse a manifest row

```{r parse-row}
parse_irving_row <- function(row){

  short_work <- row$short_work[[1]]
  manual_path <- manual_path_for_work(short_work)

  txt <- readr::read_file(manual_path)

  txt <- strip_gutenberg_boilerplate(txt)

  titles <- known_irving_titles[[short_work]]

  docs <- if(!is.null(titles)){
    split_text_by_headings(txt, titles)
  } else {
    tibble::tibble()
  }

  # Fallback to CHAPTER pattern.
  if(nrow(docs) == 0){
    docs <- split_text_by_chapter_regex(txt)
  }

  if(nrow(docs) == 0){
    warning("No Irving chapters parsed for: ", short_work)
    return(tibble::tibble())
  }

  docs %>%
    mutate(
      author = row$author[[1]],
      work = row$work[[1]],
      short_work = row$short_work[[1]],
      year = row$year[[1]],
      gutenberg_id = row$gutenberg_id[[1]],
      corpus_family = "1830",
      source_family = "1830_Secular",
      macro_genre = row$expected_genre[[1]],
      doc_unit = "chapter",
      doc_id = paste0(
        "1830__Washington_Irving__",
        short_work,
        "__",
        stringr::str_pad(chapter, width = 3, pad = "0")
      ),
      doc_id_original = doc_id,
      n_words = stringr::str_count(text, "\\S+")
    ) %>%
    select(
      corpus_family,
      source_family,
      author,
      work,
      short_work,
      year,
      gutenberg_id,
      doc_id,
      doc_id_original,
      macro_genre,
      doc_unit,
      chapter,
      chapter_title,
      n_words,
      text
    )

}
```

## 11. Parse all Irving works

```{r parse-all}
irving_docs <- purrr::pmap_dfr(
  irving_manifest,
  function(author, work, short_work, year, gutenberg_id, expected_genre){

    row <- tibble::tibble(
      author = author,
      work = work,
      short_work = short_work,
      year = year,
      gutenberg_id = gutenberg_id,
      expected_genre = expected_genre
    )

    parse_irving_row(row)
  }
)

irving_docs %>%
  count(short_work, name = "n_docs")

parse_diagnostics <- irving_docs %>%
  group_by(short_work, work) %>%
  summarise(
    n_docs = n(),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words),
    .groups = "drop"
  )

parse_diagnostics

readr::write_csv(parse_diagnostics, parse_diag_path)
```

## 12. Filter obvious non-text or tiny docs

For stylometry, extremely short units can be unstable. This keeps them visible but lets you decide.

```{r filter-check}
irving_docs %>%
  arrange(n_words) %>%
  select(short_work, chapter, chapter_title, n_words) %>%
  print(n = 50)

# Optional:
# irving_docs <- irving_docs %>% filter(n_words >= 100)
```

## 13. Write outputs

```{r write}
readr::write_csv(irving_docs, irving_out_path)
readr::write_csv(irving_manifest, irving_manifest_out_path)

cat("Wrote Irving docs to:\n")
cat(irving_out_path, "\n\n")

cat("Wrote Irving manifest to:\n")
cat(irving_manifest_out_path, "\n")
```

## 14. Integration note

Add to your master 1830 corpus after checking `doc_id` uniqueness:

```r
irving_docs <- readr::read_csv(
  file.path(STYLO_ROOT, "data", "prepared", "corpus_1830_irving_chapter_docs.csv"),
  show_col_types = FALSE
)

corpus_1830_docs <- bind_rows(
  corpus_1830_docs,
  irving_docs
)
```
