---
title: "61 - Prepare 1830 Comparator Corpus: Download, Manual, Parse"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.
source("./00_source_project.R")
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 300)
```

## 1. Purpose

This notebook separates 1830-corpus preparation into three explicit steps:

```text
STEP 1: Download raw source texts.
STEP 2: Copy each raw source into a manual-coding bin.
STEP 3: Parse only the manually coded files into chapter/document CSV.
```

The target output is:

```text
data/prepared/corpus_1830_chapter_docs.csv
```

The source texts remain untouched. Manual chapter boundaries are made only in the manual bin.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
```

## 3. Directories

```{r directories}
raw_1830_dir <- file.path(STYLO_ROOT, "data", "raw", "1830")
download_dir <- file.path(raw_1830_dir, "downloaded")
manual_dir <- file.path(raw_1830_dir, "manual")
diagnostic_dir <- file.path(raw_1830_dir, "diagnostics")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(download_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(manual_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

out_path <- file.path(prepared_dir, "corpus_1830_chapter_docs.csv")
manifest_path <- file.path(raw_1830_dir, "corpus_1830_manifest.csv")
```

## 4. Manifest

Project Gutenberg IDs are used where available. Local/manual works can be placed directly into the manual bin.

```{r manifest}
text_manifest <- tibble::tribble(
  ~author, ~work, ~short_work, ~year, ~source_type, ~gutenberg_id, ~doc_unit, ~expected_genre,

  "Jane Austen", "Pride and Prejudice", "Pride_and_Prejudice", 1813, "gutenberg", 1342, "chapter", "Fiction",
  "Jane Austen", "Emma", "Emma", 1815, "gutenberg", 158, "chapter", "Fiction",
  "Jane Austen", "Persuasion", "Persuasion", 1817, "gutenberg", 105, "chapter", "Fiction",

  "James Fenimore Cooper", "The Pioneers", "The_Pioneers", 1823, "gutenberg", 2275, "chapter", "Frontier_Fiction",
  "James Fenimore Cooper", "The Last of the Mohicans", "Last_of_the_Mohicans", 1826, "gutenberg", 940, "chapter", "Frontier_Fiction",
  "James Fenimore Cooper", "The Prairie", "The_Prairie", 1827, "gutenberg", 6450, "chapter", "Frontier_Fiction",

  "Walter Scott", "Ivanhoe", "Ivanhoe", 1819, "gutenberg", 82, "chapter", "Historical_Romance",
  "Walter Scott", "Rob Roy", "Rob_Roy", 1817, "gutenberg", 7025, "chapter", "Historical_Romance",
  "Walter Scott", "The Talisman", "The_Talisman", 1825, "gutenberg", 1377, "chapter", "Historical_Romance",

  "Ethan Smith", "View of the Hebrews", "View_of_the_Hebrews", 1823, "manual", NA, "chapter", "Religious_Prose",
  "Solomon Spaulding", "Manuscript Story", "Manuscript_Story", 1812, "manual", NA, "chapter", "Biblical_Imitation",
  "Solomon Spaulding", "Conneaut Creek Manuscript", "Conneaut_Creek", 1812, "manual", NA, "chapter", "Biblical_Imitation",
  "Jonathan Edwards", "Selected Sermons", "Selected_Sermons", 1741, "manual", NA, "sermon", "Sermon",
  "Alexander Campbell", "The Christian Baptist", "Christian_Baptist", 1823, "manual_folder", NA, "article", "Restoration_Prose"
) %>%
  dplyr::mutate(
    author_slug = stringr::str_replace_all(author, "[^A-Za-z0-9]+", "_") %>%
      stringr::str_replace_all("^_+|_+$", ""),
    raw_file = paste0(author_slug, "__", short_work, "__raw.txt"),
    manual_file = paste0(author_slug, "__", short_work, "__manual.txt")
  )

readr::write_csv(text_manifest, manifest_path)
text_manifest
```

## 5. General helpers

```{r helpers}
safe_slug <- function(x){
  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")
}

clean_text_basic <- function(x){
  x %>%
    stringr::str_replace_all("\uFEFF", "") %>%
    stringr::str_replace_all("\r", "\n") %>%
    stringr::str_replace_all("\u00A0", " ") %>%
    stringr::str_replace_all("[ \t]+", " ") %>%
    stringr::str_replace_all("\n{3,}", "\n\n") %>%
    stringr::str_trim()
}

normalize_document_text <- function(x){
  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()
}

read_text_file <- function(path){
  if(!file.exists(path)) return(NA_character_)
  paste(readr::read_lines(path, locale = readr::locale(encoding = "UTF-8"), progress = FALSE), collapse = "\n")
}

make_1830_doc_id <- function(author, work, doc_number){
  paste0("1830__", safe_slug(author), "__", safe_slug(work), "__", stringr::str_pad(doc_number, width = 3, pad = "0"))
}
```

## 6. STEP 1 — Download raw source texts

Downloaded texts go here:

```text
data/raw/1830/downloaded/
```

```{r step-1-download}
gutenberg_candidate_urls <- function(id){
  id <- as.integer(id)
  c(
    paste0("https://www.gutenberg.org/files/", id, "/", id, "-0.txt"),
    paste0("https://www.gutenberg.org/files/", id, "/", id, ".txt"),
    paste0("https://www.gutenberg.org/cache/epub/", id, "/pg", id, ".txt"),
    paste0("https://www.gutenberg.org/cache/epub/", id, "/pg", id, "-0.txt")
  )
}

download_gutenberg_text <- function(id, destfile){
  urls <- gutenberg_candidate_urls(id)
  for(u in urls){
    ok <- tryCatch({
      download.file(url = u, destfile = destfile, mode = "wb", quiet = TRUE)
      file.exists(destfile) && file.info(destfile)$size > 1000
    }, error = function(e) FALSE)
    if(isTRUE(ok)) return(tibble::tibble(path = destfile, url = u, ok = TRUE))
  }
  tibble::tibble(path = destfile, url = NA_character_, ok = FALSE)
}

download_results <- text_manifest %>%
  dplyr::filter(source_type == "gutenberg") %>%
  purrr::pmap_dfr(function(author, work, short_work, year, source_type, gutenberg_id,
                           doc_unit, expected_genre, author_slug, raw_file, manual_file){
    destfile <- file.path(download_dir, raw_file)
    if(file.exists(destfile) && file.info(destfile)$size > 1000){
      return(tibble::tibble(author = author, work = work, gutenberg_id = gutenberg_id,
                            path = destfile, url = NA_character_, ok = TRUE, status = "already_exists"))
    }
    download_gutenberg_text(gutenberg_id, destfile) %>%
      dplyr::mutate(author = author, work = work, gutenberg_id = gutenberg_id,
                    status = dplyr::if_else(ok, "downloaded", "failed")) %>%
      dplyr::select(author, work, gutenberg_id, path, url, ok, status)
  })

download_results
readr::write_csv(download_results, file.path(diagnostic_dir, "download_results.csv"))
```

## 7. STEP 2 — Create manual-coding bin

This copies raw downloaded files into:

```text
data/raw/1830/manual/
```

Manual coding should happen in the copied file, not the downloaded raw file.

Recommended manual marker format:

```text
### CHAPTER 1
Chapter text...

### CHAPTER 2
Chapter text...
```

```{r step-2-copy-to-manual}
manual_copy_results <- text_manifest %>%
  dplyr::mutate(
    downloaded_path = file.path(download_dir, raw_file),
    manual_path = file.path(manual_dir, manual_file)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    copied = {
      if(file.exists(manual_path)) FALSE
      else if(file.exists(downloaded_path)) file.copy(downloaded_path, manual_path, overwrite = FALSE)
      else FALSE
    },
    needs_manual_file = !file.exists(manual_path)
  ) %>%
  dplyr::ungroup()

manual_copy_results %>%
  dplyr::select(author, work, source_type, downloaded_path, manual_path, copied, needs_manual_file)

readr::write_csv(manual_copy_results, file.path(diagnostic_dir, "manual_copy_results.csv"))
```

## 8. Manual coding instructions

Edit files in:

```text
data/raw/1830/manual/
```

Insert chapter/document markers. The parser only trusts these markers.

Recommended marker format:

```text
### CHAPTER 1
### CHAPTER 2
### CHAPTER 3
```

Acceptable alternatives:

```text
### 1
### 2
### SERMON 1
### ARTICLE 1
### CHAPTER FIRST
### CHAPTER SECOND
```

If you want to exclude front matter, introductions, appendices, illustrations, editorial introductions, or Gutenberg boilerplate, either delete it from the manual file or leave it before the first `###` marker. The parser ignores text before the first marker.

## 9. STEP 3 — Parse manually coded files

```{r step-3-parser-helpers}
ordinal_words <- c(
  FIRST=1L, SECOND=2L, THIRD=3L, FOURTH=4L, FIFTH=5L,
  SIXTH=6L, SEVENTH=7L, EIGHTH=8L, NINTH=9L, TENTH=10L,
  ELEVENTH=11L, TWELFTH=12L, THIRTEENTH=13L, FOURTEENTH=14L,
  FIFTEENTH=15L, SIXTEENTH=16L, SEVENTEENTH=17L, EIGHTEENTH=18L, NINETEENTH=19L,
  TWENTIETH=20L, `TWENTY-FIRST`=21L, `TWENTY-SECOND`=22L, `TWENTY-THIRD`=23L,
  `TWENTY-FOURTH`=24L, `TWENTY-FIFTH`=25L, `TWENTY-SIXTH`=26L, `TWENTY-SEVENTH`=27L,
  `TWENTY-EIGHTH`=28L, `TWENTY-NINTH`=29L, THIRTIETH=30L, `THIRTY-FIRST`=31L,
  `THIRTY-SECOND`=32L, `THIRTY-THIRD`=33L, `THIRTY-FOURTH`=34L, `THIRTY-FIFTH`=35L,
  `THIRTY-SIXTH`=36L, `THIRTY-SEVENTH`=37L, `THIRTY-EIGHTH`=38L, `THIRTY-NINTH`=39L,
  FORTIETH=40L, `FORTY-FIRST`=41L, `FORTY-SECOND`=42L, `FORTY-THIRD`=43L,
  `FORTY-FOURTH`=44L, `FORTY-FIFTH`=45L, `FORTY-SIXTH`=46L, `FORTY-SEVENTH`=47L,
  `FORTY-EIGHTH`=48L, `FORTY-NINTH`=49L, FIFTIETH=50L, `FIFTY-FIRST`=51L,
  `FIFTY-SECOND`=52L, `FIFTY-THIRD`=53L, `FIFTY-FOURTH`=54L, `FIFTY-FIFTH`=55L,
  `FIFTY-SIXTH`=56L, `FIFTY-SEVENTH`=57L, `FIFTY-EIGHTH`=58L, `FIFTY-NINTH`=59L,
  SIXTIETH=60L, `SIXTY-FIRST`=61L
)

parse_marker_number <- function(marker_text, fallback_number){
  x <- marker_text %>%
    stringr::str_replace("^###\\s*", "") %>%
    stringr::str_replace(regex("^(CHAPTER|SERMON|ARTICLE)\\s+", ignore_case = TRUE), "") %>%
    stringr::str_replace("\\.$", "") %>%
    stringr::str_squish() %>%
    toupper()

  if(stringr::str_detect(x, "^[0-9]+$")) return(as.integer(x))
  if(x %in% names(ordinal_words)) return(as.integer(ordinal_words[[x]]))
  fallback_number
}

parse_manual_marked_file <- function(path){
  x <- read_text_file(path)
  if(is.na(x)) return(tibble::tibble())
  x <- clean_text_basic(x)
  lines <- stringr::str_split(x, "\n")[[1]]
  marker_idx <- which(stringr::str_detect(lines, "^\\s*###\\s+"))

  if(length(marker_idx) == 0){
    warning("No manual chapter markers found in: ", path)
    return(tibble::tibble())
  }

  marker_end <- c(marker_idx[-1] - 1L, length(lines))

  purrr::map2_dfr(marker_idx, marker_end, function(s, e){
    marker <- lines[s] %>% stringr::str_squish()
    body <- if(s < e) paste(lines[(s + 1L):e], collapse = "\n") else ""
    tibble::tibble(
      marker = marker,
      chapter_title = marker %>% stringr::str_replace("^###\\s*", "") %>% stringr::str_squish(),
      text = normalize_document_text(body)
    )
  }) %>%
    dplyr::mutate(
      doc_number = dplyr::row_number(),
      chapter = purrr::map2_int(marker, doc_number, parse_marker_number)
    ) %>%
    dplyr::filter(!is.na(text), text != "")
}
```

## 10. Parse corpus from manual bin

```{r step-3-parse-corpus}
parse_manifest_row_manual <- function(row){
  manual_path <- file.path(manual_dir, row$manual_file)

  if(!file.exists(manual_path)){
    warning("Missing manual file: ", manual_path)
    return(tibble::tibble())
  }

  docs <- parse_manual_marked_file(manual_path)
  if(nrow(docs) == 0) return(tibble::tibble())

  docs %>%
    dplyr::mutate(
      corpus_id = "1830",
      translation = as.character(row$year),
      corpus_family = "1830",
      source_family = paste0("1830_", safe_slug(row$author)),
      author = row$author,
      work = row$work,
      short_work = row$short_work,
      year = as.integer(row$year),
      book = paste0(row$author, " - ", row$work),
      section = row$expected_genre,
      macro_genre = row$expected_genre,
      doc_unit = row$doc_unit,
      n_verses = NA_integer_,
      n_words = stringr::str_count(text, "\\S+"),
      doc_id = make_1830_doc_id(row$author, row$short_work, chapter),
      doc_id_original = doc_id
    ) %>%
    dplyr::select(
      corpus_id, translation, corpus_family, source_family, doc_id, doc_id_original,
      author, work, short_work, year, book, chapter, doc_number, marker, chapter_title,
      section, macro_genre, doc_unit, n_verses, n_words, text
    )
}

corpus_1830_docs <- purrr::pmap_dfr(
  text_manifest,
  function(author, work, short_work, year, source_type, gutenberg_id,
           doc_unit, expected_genre, author_slug, raw_file, manual_file){
    row <- tibble::tibble(
      author = author, work = work, short_work = short_work, year = year,
      source_type = source_type, gutenberg_id = gutenberg_id,
      doc_unit = doc_unit, expected_genre = expected_genre,
      author_slug = author_slug, raw_file = raw_file, manual_file = manual_file
    )
    parse_manifest_row_manual(row)
  }
)

corpus_1830_docs %>%
  dplyr::count(author, work, macro_genre, name = "n_docs") %>%
  dplyr::arrange(author, work)
```

## 11. Filter tiny documents

```{r filter-tiny}
min_words_1830 <- 100

corpus_1830_docs_raw_parsed <- corpus_1830_docs

corpus_1830_chapter_docs <- corpus_1830_docs %>%
  dplyr::filter(n_words >= min_words_1830) %>%
  dplyr::arrange(author, work, chapter)

corpus_1830_docs_raw_parsed %>%
  dplyr::count(author, work, name = "n_docs_raw") %>%
  dplyr::left_join(
    corpus_1830_chapter_docs %>% dplyr::count(author, work, name = "n_docs_kept"),
    by = c("author", "work")
  ) %>%
  dplyr::mutate(
    n_docs_kept = dplyr::coalesce(n_docs_kept, 0L),
    n_dropped = n_docs_raw - n_docs_kept
  ) %>%
  dplyr::arrange(author, work)
```

## 12. Sanity checks

```{r sanity-checks}
corpus_1830_chapter_docs %>%
  dplyr::count(author, work, macro_genre, name = "n_docs") %>%
  dplyr::arrange(author, work)

corpus_1830_chapter_docs %>%
  dplyr::group_by(author, work) %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    min_chapter = min(chapter, na.rm = TRUE),
    max_chapter = max(chapter, na.rm = TRUE),
    min_words = min(n_words, na.rm = TRUE),
    median_words = median(n_words, na.rm = TRUE),
    mean_words = mean(n_words, na.rm = TRUE),
    max_words = max(n_words, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(author, work)

corpus_1830_chapter_docs %>%
  dplyr::arrange(n_words) %>%
  dplyr::select(author, work, chapter, doc_number, marker, n_words) %>%
  dplyr::slice_head(n = 30)

corpus_1830_chapter_docs %>%
  dplyr::arrange(dplyr::desc(n_words)) %>%
  dplyr::select(author, work, chapter, doc_number, marker, n_words) %>%
  dplyr::slice_head(n = 30)
```

## 13. Write outputs

```{r write-outputs}
raw_parsed_path <- file.path(raw_1830_dir, "corpus_1830_docs_raw_parsed.csv")

readr::write_csv(corpus_1830_docs_raw_parsed, raw_parsed_path)
readr::write_csv(corpus_1830_chapter_docs, out_path)

cat("Wrote raw parsed 1830 corpus to:\n")
cat(raw_parsed_path, "\n\n")

cat("Wrote prepared 1830 chapter corpus to:\n")
cat(out_path, "\n")
```

## 14. Notes

This notebook deliberately separates:

```text
downloaded source
manual-coded source
parsed corpus
```

That gives a clean audit trail and avoids fragile chapter-heading regexes.
