---
title: "62 - Parse 1830 Manual Corpus"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 300)

library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(tidyr)
library(digest)
```

## 1. Purpose

Parse every non-empty `*_manual.txt` / `*_incomplete.txt` file in:

```text
data/raw/1830/manual/
```

Ignore `_README.txt` and zero-byte files.

Manual chapter dividers:

```text
### CHAPTER 33: L’ENVOY.
### CHAPTER 2-24: THE AUTHOR'S FAREWELL.
```

Outputs:

```text
data/prepared/corpus_1830_manual_chapter_docs.csv
data/prepared/corpus_1830_manual_book_docs.csv
```

The parser preserves original chapter labels while creating clean sequential chapter numbers:

```text
chapter = 1, 2, 3, ...
chapter_original = "2-24"
chapter_title = "THE AUTHOR'S FAREWELL"
```

## 2. Directories

```{r dirs}
manual_dir <- file.path(STYLO_ROOT, "data", "raw", "1830", "manual")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
diagnostic_dir <- file.path(STYLO_ROOT, "data", "raw", "1830", "diagnostics")

dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)

chapter_out_path <- file.path(prepared_dir, "corpus_1830_manual_chapter_docs.csv")
book_out_path <- file.path(prepared_dir, "corpus_1830_manual_book_docs.csv")
manifest_out_path <- file.path(prepared_dir, "corpus_1830_manual_manifest.csv")
parse_diag_path <- file.path(diagnostic_dir, "corpus_1830_manual_parse_diagnostics.csv")
```

## 3. Enumerate manual files

```{r enumerate}
manual_files <- list.files(
  manual_dir,
  pattern = "\\.txt$",
  full.names = TRUE
)

manual_manifest <- tibble(
  path = manual_files,
  filename = basename(manual_files),
  file_size = file.info(manual_files)$size
) %>%
  filter(
    filename != "_README.txt",
    file_size > 0
  ) %>%
  arrange(filename)

manual_manifest

readr::write_csv(manual_manifest, manifest_out_path)
```

## 4. Filename metadata parser

Expected filename format:

```text
Author_Name_Work_Title_manual.txt
Author_Name_Work_Title_incomplete.txt
```

Known multi-token author names are handled explicitly.

```{r filename-parser}
known_author_prefixes <- c(
  "James_Fenimore_Cooper",
  "Washington_Irving",
  "Jane_Austen",
  "Walter_Scott",
  "Ethan_Smith",
  "Gilbert_Hunt",
  "Alexander_Campbell",
  "Jonathan_Edwards",
  "Solomon_Spaulding"
)

infer_macro_genre <- function(author, work){
  dplyr::case_when(
    author == "Jane Austen" ~ "Fiction",
    author == "James Fenimore Cooper" ~ "Frontier_Fiction",
    author == "Walter Scott" ~ "Historical_Romance",
    author == "Washington Irving" ~ "American_Prose",
    author %in% c("Ethan Smith", "Alexander Campbell", "Jonathan Edwards") ~ "Religious_Nonfiction",
    author %in% c("Gilbert Hunt", "Solomon Spaulding") ~ "American_Prose",
    TRUE ~ "1830_Prose"
  )
}

parse_filename_metadata <- function(path){
  fn <- basename(path)

  stem <- fn %>%
    str_remove("\\.txt$") %>%
    str_remove("_manual$") %>%
    str_remove("_incomplete$")

  matched_prefix <- known_author_prefixes[
    purrr::map_lgl(
      known_author_prefixes,
      ~str_starts(stem, paste0(.x, "_"))
    )
  ]

  if(length(matched_prefix) > 0){
    author_key <- matched_prefix[[which.max(nchar(matched_prefix))]]
    work_key <- stem %>% str_remove(paste0("^", fixed(author_key), "_"))
  } else {
    parts <- str_split(stem, "_")[[1]]

    if(length(parts) >= 3){
      author_key <- paste(parts[1:2], collapse = "_")
      work_key <- paste(parts[-c(1,2)], collapse = "_")
    } else {
      author_key <- parts[[1]]
      work_key <- paste(parts[-1], collapse = "_")
    }
  }

  author <- str_replace_all(author_key, "_", " ")
  work <- str_replace_all(work_key, "_", " ")
  short_work <- work_key

  tibble(
    author = author,
    work = work,
    short_work = short_work,
    source_file = fn,
    local_file = path,
    macro_genre = infer_macro_genre(author, work)
  )
}

file_metadata <- purrr::map_dfr(
  manual_manifest$path,
  parse_filename_metadata
)

file_metadata
```

## 5. Text helpers

```{r text-helpers}
normalize_doc_text <- function(x){
  x %>%
    str_replace_all("\r", "\n") %>%
    str_replace_all("\uFEFF", "") %>%
    str_replace_all("\u00A0", " ") %>%
    str_replace_all("[ \t]+", " ") %>%
    str_replace_all("\n{3,}", "\n\n") %>%
    str_replace_all("\\s+", " ") %>%
    str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    str_squish()
}

clean_heading <- function(x){
  x %>%
    str_replace_all("[\\*_]", "") %>%
    str_replace_all("\\s+", " ") %>%
    str_squish() %>%
    str_remove("\\s*\\.$")
}

make_safe_id <- function(x){
  x %>%
    str_replace_all("[^A-Za-z0-9]+", "_") %>%
    str_replace_all("^_+|_+$", "")
}
```

## 6. Chapter divider parser

Captures:

```text
### CHAPTER 33: L’ENVOY.
### CHAPTER 2-24: THE AUTHOR'S FAREWELL.
### CHAPTER 1
```

```{r chapter-parser}
chapter_divider_pattern <- regex(
  paste0(
    "(?m)^\\s*",
    "###\\s+",
    "CHAPTER\\s+",
    "([^:\\n]+?)",
    "(?:\\s*:\\s*(.*?))?",
    "\\s*$"
  ),
  ignore_case = TRUE
)

parse_manual_chapters <- function(path){
  txt <- readr::read_file(path)

  txt <- txt %>%
    str_replace_all("\r", "\n") %>%
    str_replace_all("\uFEFF", "")

  locs <- str_locate_all(txt, chapter_divider_pattern)[[1]]

  if(nrow(locs) == 0){
    return(
      tibble(
        chapter = 1L,
        chapter_original = "1",
        chapter_title = "Whole Work",
        chapter_heading_raw = NA_character_,
        text = normalize_doc_text(txt)
      ) %>%
        filter(text != "")
    )
  }

  matches <- str_match_all(txt, chapter_divider_pattern)[[1]]

  purrr::map_dfr(seq_len(nrow(locs)), function(i){
    heading_raw <- str_sub(txt, locs[i, 1], locs[i, 2])
    chapter_original <- clean_heading(matches[i, 2])
    title_raw <- matches[i, 3]

    chapter_title <- ifelse(
      is.na(title_raw) || str_squish(title_raw) == "",
      paste0("Chapter ", chapter_original),
      clean_heading(title_raw)
    )

    start <- locs[i, 2] + 1
    end <- if(i < nrow(locs)) locs[i + 1, 1] - 1 else nchar(txt)
    body <- str_sub(txt, start, end)

    tibble(
      chapter_original = chapter_original,
      chapter_title = chapter_title,
      chapter_heading_raw = heading_raw,
      text = normalize_doc_text(body)
    )
  }) %>%
    filter(text != "") %>%
    mutate(chapter = row_number(), .before = chapter_original)
}
```

## 7. Parse all manual files

```{r parse-all}
corpus_1830_manual_chapter_docs <- file_metadata %>%
  mutate(chapters = purrr::map(local_file, parse_manual_chapters)) %>%
  tidyr::unnest(chapters) %>%
  group_by(source_file) %>%
  mutate(
    chapter = row_number(),
    chapter_digest = digest::digest(
      paste(author, work, chapter_original, chapter_title, text, sep = "||"),
      algo = "xxhash64"
    ),
    duplicate_digest_in_work = duplicated(chapter_digest)
  ) %>%
  ungroup() %>%
  filter(!duplicate_digest_in_work) %>%
  group_by(source_file) %>%
  mutate(chapter = row_number()) %>%
  ungroup() %>%
  mutate(
    corpus_family = "1830",
    source_family = "1830_Secular",
    doc_unit = "chapter",
    doc_id = paste0(
      "1830__",
      make_safe_id(author),
      "__",
      make_safe_id(short_work),
      "__",
      str_pad(chapter, width = 3, pad = "0")
    ),
    doc_id_original = doc_id,
    n_words = str_count(text, "\\S+")
  ) %>%
  select(
    corpus_family,
    source_family,
    author,
    work,
    short_work,
    source_file,
    local_file,
    doc_id,
    doc_id_original,
    macro_genre,
    doc_unit,
    chapter,
    chapter_original,
    chapter_title,
    chapter_heading_raw,
    chapter_digest,
    n_words,
    text
  )

corpus_1830_manual_chapter_docs %>%
  count(author, work, name = "n_chapters") %>%
  arrange(author, work) %>%
  print(n = 100)
```

## 8. Diagnostics

```{r diagnostics}
parse_diagnostics <- corpus_1830_manual_chapter_docs %>%
  group_by(author, work, short_work, source_file, macro_genre) %>%
  summarise(
    n_chapters = n(),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words),
    .groups = "drop"
  ) %>%
  arrange(author, work)

parse_diagnostics

readr::write_csv(parse_diagnostics, parse_diag_path)

corpus_1830_manual_chapter_docs %>%
  arrange(n_words) %>%
  select(author, work, chapter, chapter_original, chapter_title, n_words) %>%
  print(n = 50)

corpus_1830_manual_chapter_docs %>%
  summarise(
    n_chapter_docs = n(),
    n_authors = n_distinct(author),
    n_works = n_distinct(paste(author, work)),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )
```

## 9. Build book-level CSV

```{r book-level}
corpus_1830_manual_book_docs <- corpus_1830_manual_chapter_docs %>%
  arrange(author, work, chapter) %>%
  group_by(
    corpus_family,
    source_family,
    author,
    work,
    short_work,
    source_file,
    local_file,
    macro_genre
  ) %>%
  summarise(
    n_chapters = n(),
    min_chapter = min(chapter),
    max_chapter = max(chapter),
    text = paste(text, collapse = " "),
    chapter_map = paste(
      paste0(chapter, "=", chapter_original, ": ", chapter_title),
      collapse = " | "
    ),
    .groups = "drop"
  ) %>%
  mutate(
    text = normalize_doc_text(text),
    doc_unit = "book",
    n_words = str_count(text, "\\S+"),
    doc_id = paste0(
      "1830__",
      make_safe_id(author),
      "__",
      make_safe_id(short_work),
      "__BOOK"
    ),
    doc_id_original = doc_id,
    book_digest = digest::digest(
      paste(author, work, text, sep = "||"),
      algo = "xxhash64"
    )
  ) %>%
  select(
    corpus_family,
    source_family,
    author,
    work,
    short_work,
    source_file,
    local_file,
    doc_id,
    doc_id_original,
    macro_genre,
    doc_unit,
    n_chapters,
    min_chapter,
    max_chapter,
    chapter_map,
    book_digest,
    n_words,
    text
  )

corpus_1830_manual_book_docs %>%
  select(author, work, n_chapters, n_words) %>%
  arrange(author, work) %>%
  print(n = 100)
```

## 10. Write outputs

```{r write}
readr::write_csv(corpus_1830_manual_chapter_docs, chapter_out_path)
readr::write_csv(corpus_1830_manual_book_docs, book_out_path)

cat("Wrote chapter CSV:\n")
cat(chapter_out_path, "\n\n")

cat("Wrote book CSV:\n")
cat(book_out_path, "\n")
```

## 11. Integration

Use the chapter CSV for stylometry unless you intentionally want book-level aggregation:

```r
corpus_1830_manual_chapter_docs <- readr::read_csv(
  file.path(STYLO_ROOT, "data", "prepared", "corpus_1830_manual_chapter_docs.csv"),
  show_col_types = FALSE
)
```
