---
title: "62 - Prepare DC and PoGP from LDS Scriptures CSV"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook prepares:

```text
Doctrine and Covenants = DC
Pearl of Great Price   = PoGP
```

from the public LDS scriptures CSV:

```text
https://raw.githubusercontent.com/beandog/lds-scriptures/master/csv/lds-scriptures.csv
```

Target outputs:

```text
data/prepared/dc_chapter_docs.csv
data/prepared/pogp_chapter_docs.csv
data/prepared/dc_pogp_chapter_docs.csv
```

The source appears to be the modern LDS text, so these corpora are labeled:

```text
edition = modern
```

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
library(purrr)
library(tibble)
```

## 3. Directories and paths

```{r directories}
raw_scripture_dir <- file.path(
  STYLO_ROOT,
  "data",
  "raw",
  "lds_scriptures"
)

prepared_dir <- file.path(
  STYLO_ROOT,
  "data",
  "prepared"
)

dir.create(raw_scripture_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

lds_csv_url <- "https://raw.githubusercontent.com/beandog/lds-scriptures/master/csv/lds-scriptures.csv"

lds_raw_path <- file.path(
  raw_scripture_dir,
  "lds-scriptures.csv"
)

dc_out_path <- file.path(
  prepared_dir,
  "dc_chapter_docs.csv"
)

pogp_out_path <- file.path(
  prepared_dir,
  "pogp_chapter_docs.csv"
)

combined_out_path <- file.path(
  prepared_dir,
  "dc_pogp_chapter_docs.csv"
)
```

## 4. Download source CSV

```{r download}
if(!file.exists(lds_raw_path)){

  download.file(
    url = lds_csv_url,
    destfile = lds_raw_path,
    mode = "wb",
    quiet = TRUE
  )

}

cat("Source CSV:\n")
cat(lds_raw_path, "\n")
```

## 5. Read and inspect source

```{r read-source}
lds_raw <- readr::read_csv(
  lds_raw_path,
  show_col_types = FALSE
)

names(lds_raw)

lds_raw %>%
  dplyr::count(volume_title, book_title, name = "n_verses") %>%
  dplyr::arrange(volume_title, book_title) %>%
  print(n = 200)
```

## 6. Normalize columns

The GitHub CSV has historically used fields like:

```text
volume_title
book_title
chapter_number
verse_number
scripture_text
```

```{r normalize-columns}
required_cols <- c(
  "volume_title",
  "book_title",
  "chapter_number",
  "verse_number",
  "scripture_text"
)

missing_cols <- setdiff(
  required_cols,
  names(lds_raw)
)

if(length(missing_cols) > 0){
  stop(
    "Missing expected columns: ",
    paste(missing_cols, collapse = ", ")
  )
}

lds <- lds_raw %>%
  dplyr::transmute(
    volume = volume_title,
    book = book_title,
    chapter = as.integer(chapter_number),
    verse = as.integer(verse_number),
    scripture_text = scripture_text
  ) %>%
  dplyr::mutate(
    volume = stringr::str_squish(volume),
    book = stringr::str_squish(book),
    scripture_text = stringr::str_squish(scripture_text)
  )

lds %>%
  dplyr::count(volume, book, name = "n_verses") %>%
  dplyr::arrange(volume, book) %>%
  print(n = 200)
```

## 7. Helpers

```{r helpers}
safe_slug <- function(x){

  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")

}

normalize_text <- function(x){

  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()

}

make_doc_id <- function(prefix, book, chapter){

  paste0(
    prefix,
    "__",
    safe_slug(book),
    "_",
    chapter
  )

}

make_chapter_docs <- function(df, corpus_family_value, source_family_value, edition_value){

  df %>%
    dplyr::arrange(book, chapter, verse) %>%
    dplyr::group_by(book, chapter) %>%
    dplyr::summarise(
      text = paste(scripture_text, collapse = " "),
      n_verses = dplyr::n(),
      min_verse = min(verse, na.rm = TRUE),
      max_verse = max(verse, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      text = normalize_text(text),
      n_words = stringr::str_count(text, "\\S+"),
      corpus_id = corpus_family_value,
      translation = edition_value,
      corpus_family = corpus_family_value,
      source_family = source_family_value,
      edition = edition_value,
      author = NA_character_,
      work = book,
      short_work = safe_slug(book),
      section = NA_character_,
      macro_genre = "Modern_Scripture",
      doc_unit = "chapter",
      chapter_title = paste0(book, " ", chapter),
      doc_id = make_doc_id(corpus_family_value, book, chapter),
      doc_id_original = doc_id
    ) %>%
    dplyr::select(
      corpus_id,
      translation,
      corpus_family,
      source_family,
      edition,
      doc_id,
      doc_id_original,
      author,
      work,
      short_work,
      book,
      chapter,
      chapter_title,
      section,
      macro_genre,
      doc_unit,
      n_verses,
      min_verse,
      max_verse,
      n_words,
      text
    )

}
```

## 8. Filter DC and PoGP

```{r filter-corpora}
lds %>%
  dplyr::distinct(volume) %>%
  dplyr::arrange(volume)

dc_verses <- lds %>%
  dplyr::filter(volume == "Doctrine and Covenants")

pogp_verses <- lds %>%
  dplyr::filter(volume == "Pearl of Great Price")

dc_verses %>%
  dplyr::count(book, name = "n_verses") %>%
  dplyr::arrange(book)

pogp_verses %>%
  dplyr::count(book, name = "n_verses") %>%
  dplyr::arrange(book)
```

## 9. Build chapter documents

```{r build-docs}
dc_chapter_docs <- make_chapter_docs(
  df = dc_verses,
  corpus_family_value = "DC",
  source_family_value = "DC",
  edition_value = "modern"
)

pogp_chapter_docs <- make_chapter_docs(
  df = pogp_verses,
  corpus_family_value = "PoGP",
  source_family_value = "PoGP",
  edition_value = "modern"
)

dc_chapter_docs %>%
  dplyr::count(book, name = "n_chapters") %>%
  dplyr::arrange(book)

pogp_chapter_docs %>%
  dplyr::count(book, name = "n_chapters") %>%
  dplyr::arrange(book)
```

## 10. Add PoGP subfamilies

```{r pogp-subfamilies}
pogp_chapter_docs <- pogp_chapter_docs %>%
  dplyr::mutate(
    source_family = dplyr::case_when(
      book %in% c("Moses", "Abraham") ~ "PoGP_Translation",
      book %in% c("Joseph Smith--Matthew", "Joseph Smith--History") ~ "PoGP_History",
      book %in% c("Articles of Faith") ~ "PoGP_Doctrine",
      TRUE ~ "PoGP"
    )
  )

dc_pogp_chapter_docs <- dplyr::bind_rows(
  dc_chapter_docs,
  pogp_chapter_docs
)

pogp_chapter_docs %>%
  dplyr::count(book, source_family, name = "n_chapters") %>%
  dplyr::arrange(source_family, book)

dc_pogp_chapter_docs %>%
  dplyr::count(corpus_family, source_family, edition, name = "n_docs")
```

## 11. Sanity checks

```{r sanity-checks}
dc_pogp_chapter_docs %>%
  dplyr::summarise(
    n_docs = dplyr::n(),
    n_corpora = dplyr::n_distinct(corpus_family),
    n_books = dplyr::n_distinct(book),
    min_words = min(n_words, na.rm = TRUE),
    median_words = median(n_words, na.rm = TRUE),
    mean_words = mean(n_words, na.rm = TRUE),
    max_words = max(n_words, na.rm = TRUE)
  )

dc_pogp_chapter_docs %>%
  dplyr::arrange(n_words) %>%
  dplyr::select(corpus_family, book, chapter, n_verses, n_words, doc_id) %>%
  dplyr::slice_head(n = 30)

dc_pogp_chapter_docs %>%
  dplyr::arrange(dplyr::desc(n_words)) %>%
  dplyr::select(corpus_family, book, chapter, n_verses, n_words, doc_id) %>%
  dplyr::slice_head(n = 30)

dc_pogp_chapter_docs %>%
  dplyr::count(corpus_family, source_family, book, name = "n_chapters") %>%
  dplyr::arrange(corpus_family, source_family, book)
```

## 12. Optional: verse gaps

```{r verse-gaps}
verse_gaps <- lds %>%
  dplyr::filter(volume %in% c("Doctrine and Covenants", "Pearl of Great Price")) %>%
  dplyr::group_by(volume, book, chapter) %>%
  dplyr::summarise(
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_verses = dplyr::n_distinct(verse),
    expected_n = max_verse - min_verse + 1L,
    has_gap = n_verses != expected_n,
    missing_verses = paste(
      setdiff(seq(min_verse, max_verse), sort(unique(verse))),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  dplyr::filter(has_gap)

verse_gaps
```

## 13. Write outputs

```{r write-outputs}
readr::write_csv(
  dc_chapter_docs,
  dc_out_path
)

readr::write_csv(
  pogp_chapter_docs,
  pogp_out_path
)

readr::write_csv(
  dc_pogp_chapter_docs,
  combined_out_path
)

cat("Wrote DC chapter docs to:\n")
cat(dc_out_path, "\n\n")

cat("Wrote PoGP chapter docs to:\n")
cat(pogp_out_path, "\n\n")

cat("Wrote combined DC + PoGP chapter docs to:\n")
cat(combined_out_path, "\n")
```

## 14. Next step

Add these prepared files to the larger analysis framework:

```text
KJV + BOM + BOC + DC + PoGP + SCR + 1830
```

Recommended top-level corpus families:

```text
KJV
BOM
BOC
DC
PoGP
SCR
1830
```
