---
title: "63 - Prepare 1835 Doctrine and Covenants"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook downloads, caches, parses, and prepares the **1835 Doctrine and Covenants** from:

```text
https://doctrineandcovenants.com/sections/1835dcs/section-1/
```

The target output is a common chapter/verse-style CSV:

```text
data/prepared/dc_1835_chapter_docs.csv
data/prepared/dc_1835_verse_docs.csv
```

The resulting corpus uses:

```text
corpus_family = "DC"
source_family = "DC_1835"
edition = "1835"
```

For this project, each **section** is treated as the common imported `chapter` unit.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(rvest)
library(xml2)
library(tidyr)
```

## 3. Directories

```{r directories}
raw_dc1835_dir <- file.path(STYLO_ROOT, "data", "raw", "dc_1835")
html_cache_dir <- file.path(raw_dc1835_dir, "html")
diagnostic_dir <- file.path(raw_dc1835_dir, "diagnostics")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(raw_dc1835_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(html_cache_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

dc1835_verse_out_path <- file.path(prepared_dir, "dc_1835_verse_docs.csv")
dc1835_chapter_out_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")
dc1835_parse_diag_path <- file.path(diagnostic_dir, "dc_1835_parse_diagnostics.csv")
dc1835_link_diag_path <- file.path(diagnostic_dir, "dc_1835_links_discovered.csv")
```

## 4. Download / cache helpers

The site uses section pages like:

```text
https://doctrineandcovenants.com/sections/1835dcs/section-1/
```

This notebook first crawls from Section 1 using same-domain links that contain:

```text
/sections/1835dcs/section-
```

If the crawl misses anything, a fixed-range supplement tries `section-1` through `section-120`.

```{r download-helpers}
base_section_url <- "https://doctrineandcovenants.com/sections/1835dcs/section-1/"
section_url_pattern <- "/sections/1835dcs/section-[^/]+/?$"

safe_slug <- function(x){
  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")
}

normalize_url <- function(url){
  url <- stringr::str_replace(url, "#.*$", "")

  if(stringr::str_starts(url, "/")){
    url <- paste0("https://doctrineandcovenants.com", url)
  }

  if(!stringr::str_detect(url, "/$")){
    url <- paste0(url, "/")
  }

  url
}

section_id_from_url <- function(url){
  stringr::str_match(
    url,
    regex("/section-([^/]+)/?$", ignore_case = TRUE)
  )[, 2]
}

section_number_from_url <- function(url){
  id <- section_id_from_url(url)

  suppressWarnings(
    as.integer(
      stringr::str_extract(id, "[0-9]+")
    )
  )
}

cache_path_for_url <- function(url){
  section_id <- section_id_from_url(url)

  if(is.na(section_id) || section_id == ""){
    section_id <- safe_slug(url)
  }

  file.path(html_cache_dir, paste0("section_", section_id, ".html"))
}

download_section_html <- function(url, overwrite = FALSE){
  url <- normalize_url(url)
  cache_path <- cache_path_for_url(url)

  if(file.exists(cache_path) && !overwrite){
    return(
      tibble::tibble(
        url = url,
        cache_path = cache_path,
        downloaded = FALSE,
        ok = TRUE,
        status = "cached"
      )
    )
  }

  ok <- tryCatch({
    download.file(
      url = url,
      destfile = cache_path,
      mode = "wb",
      quiet = TRUE
    )

    file.exists(cache_path) && file.info(cache_path)$size > 1000
  }, error = function(e){
    FALSE
  })

  tibble::tibble(
    url = url,
    cache_path = cache_path,
    downloaded = ok,
    ok = ok,
    status = dplyr::if_else(ok, "downloaded", "failed")
  )
}

read_cached_html <- function(cache_path){
  xml2::read_html(cache_path, encoding = "UTF-8")
}
```

## 5. Crawl available 1835 section links

```{r crawl-links}
discover_section_links <- function(start_url, max_pages = 200){
  start_url <- normalize_url(start_url)

  queue <- start_url
  visited <- character()
  discovered <- tibble::tibble()

  while(length(queue) > 0 && length(visited) < max_pages){
    current_url <- queue[[1]]
    queue <- queue[-1]

    if(current_url %in% visited){
      next
    }

    dl <- download_section_html(current_url)
    visited <- c(visited, current_url)

    if(!isTRUE(dl$ok[[1]])){
      next
    }

    page <- read_cached_html(dl$cache_path[[1]])

    links <- page %>%
      rvest::html_elements("a") %>%
      rvest::html_attr("href") %>%
      purrr::discard(is.na) %>%
      purrr::map_chr(normalize_url) %>%
      unique()

    section_links <- links[
      stringr::str_detect(
        links,
        fixed("https://doctrineandcovenants.com/sections/1835dcs/section-")
      )
    ]

    section_links <- section_links[
      stringr::str_detect(
        section_links,
        regex(section_url_pattern, ignore_case = TRUE)
      )
    ]

    discovered <- dplyr::bind_rows(
      discovered,
      tibble::tibble(
        source_url = current_url,
        discovered_url = section_links,
        discovered_section_id = section_id_from_url(section_links),
        discovered_section_number = section_number_from_url(section_links)
      )
    )

    new_links <- setdiff(section_links, visited)
    queue <- unique(c(queue, new_links))
  }

  discovered %>%
    dplyr::distinct(discovered_url, .keep_all = TRUE) %>%
    dplyr::arrange(discovered_section_number, discovered_section_id)
}

discovered_links <- discover_section_links(
  base_section_url,
  max_pages = 200
)

discovered_links

readr::write_csv(
  discovered_links,
  dc1835_link_diag_path
)
```

## 6. Optional fixed-range supplement

The website may use special page names or omit some next links. This supplement is safe because failed URLs are recorded and ignored.

```{r fixed-range-supplement}
fixed_section_urls <- tibble::tibble(
  discovered_url = paste0(
    "https://doctrineandcovenants.com/sections/1835dcs/section-",
    1:120,
    "/"
  ),
  source_url = "fixed_range",
  discovered_section_id = as.character(1:120),
  discovered_section_number = 1:120
)

section_url_index <- dplyr::bind_rows(
  discovered_links,
  fixed_section_urls
) %>%
  dplyr::distinct(discovered_url, .keep_all = TRUE) %>%
  dplyr::arrange(discovered_section_number, discovered_section_id)

section_url_index %>%
  print(n = 150)
```

## 7. Download all indexed section pages

```{r download-all}
download_results <- section_url_index %>%
  dplyr::mutate(
    download_info = purrr::map(discovered_url, download_section_html)
  ) %>%
  tidyr::unnest(download_info)

download_results %>%
  dplyr::count(status, ok)

download_results %>%
  dplyr::filter(ok) %>%
  dplyr::arrange(discovered_section_number) %>%
  dplyr::select(discovered_section_number, discovered_section_id, discovered_url, cache_path) %>%
  print(n = 150)

readr::write_csv(
  download_results,
  file.path(diagnostic_dir, "dc_1835_download_results.csv")
)
```

## 8. Parser helpers

The parser is defensive. It tries to:

1. remove navigation, scripts, style, footer, and comment text,
2. extract the main content if possible,
3. detect verse lines with forms like `1 ...`, `1. ...`, or `1: ...`,
4. fall back to paragraph numbering if explicit verse numbers are absent.

```{r parser-helpers}
clean_text <- function(x){
  x %>%
    stringr::str_replace_all("\uFEFF", "") %>%
    stringr::str_replace_all("\r", "\n") %>%
    stringr::str_replace_all("\u00A0", " ") %>%
    stringr::str_replace_all("[ \t]+", " ") %>%
    stringr::str_replace_all("\n{3,}", "\n\n") %>%
    stringr::str_trim()
}

normalize_doc_text <- function(x){
  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()
}

drop_bad_lines <- function(lines){
  bad <- stringr::str_detect(
    lines,
    regex(
      paste(
        c(
          "^Previous$",
          "^Next$",
          "^Home$",
          "^Contents$",
          "^Doctrine and Covenants$",
          "^1835 Doctrine and Covenants$",
          "^Section \\d+$",
          "^\\s*$",
          "Privacy Policy",
          "Terms",
          "Copyright",
          "Leave a Reply",
          "Comments",
          "Search"
        ),
        collapse = "|"
      ),
      ignore_case = TRUE
    )
  )

  lines[!bad]
}

extract_title <- function(page, fallback_section){
  title_candidates <- c(
    page %>% rvest::html_elements("h1") %>% rvest::html_text2(),
    page %>% rvest::html_elements("h2") %>% rvest::html_text2(),
    page %>% rvest::html_elements("title") %>% rvest::html_text2()
  )

  title_candidates <- title_candidates %>%
    stringr::str_squish()

  title_candidates <- title_candidates[
    !is.na(title_candidates) &
      title_candidates != "" &
      !stringr::str_detect(title_candidates, regex("Doctrine and Covenants.com", ignore_case = TRUE))
  ]

  if(length(title_candidates) > 0){
    return(title_candidates[[1]])
  }

  paste0("Section ", fallback_section)
}

extract_main_node <- function(page){
  selectors <- c(
    "article",
    "main",
    ".entry-content",
    ".post-content",
    ".content",
    "#content"
  )

  for(sel in selectors){
    node <- page %>% rvest::html_element(sel)

    if(length(node) > 0 && !inherits(node, "xml_missing")){
      return(node)
    }
  }

  page %>% rvest::html_element("body")
}

extract_page_lines <- function(page){
  xml2::xml_find_all(page, ".//script|.//style|.//nav|.//footer|.//header|.//aside") %>%
    xml2::xml_remove()

  main_node <- extract_main_node(page)

  text <- main_node %>%
    rvest::html_text2() %>%
    clean_text()

  lines <- stringr::str_split(text, "\n")[[1]] %>%
    stringr::str_squish()

  lines <- lines[lines != ""]
  lines <- drop_bad_lines(lines)

  lines
}

is_verse_start <- function(line){
  stringr::str_detect(
    line,
    regex("^\\s*([0-9]{1,3})\\s*([\\.:])\\s+")
  ) |
    stringr::str_detect(
      line,
      regex("^\\s*([0-9]{1,3})\\s+[A-Z“\"']")
    )
}

parse_verse_number <- function(line){
  m1 <- stringr::str_match(
    line,
    regex("^\\s*([0-9]{1,3})\\s*([\\.:])\\s+(.*)$")
  )

  if(!is.na(m1[1, 2])){
    return(as.integer(m1[1, 2]))
  }

  m2 <- stringr::str_match(
    line,
    regex("^\\s*([0-9]{1,3})\\s+(.*)$")
  )

  if(!is.na(m2[1, 2])){
    return(as.integer(m2[1, 2]))
  }

  NA_integer_
}

strip_verse_number <- function(line){
  line %>%
    stringr::str_replace(regex("^\\s*[0-9]{1,3}\\s*([\\.:])\\s+"), "") %>%
    stringr::str_replace(regex("^\\s*[0-9]{1,3}\\s+"), "") %>%
    stringr::str_squish()
}

parse_verses_from_lines <- function(lines){
  if(length(lines) == 0){
    return(tibble::tibble())
  }

  verse_start <- purrr::map_lgl(lines, is_verse_start)

  if(sum(verse_start) >= 2){
    verse_group <- cumsum(verse_start)

    tibble::tibble(
      line = lines,
      verse_start = verse_start,
      verse_group = verse_group
    ) %>%
      dplyr::filter(verse_group > 0) %>%
      dplyr::group_by(verse_group) %>%
      dplyr::summarise(
        first_line = dplyr::first(line),
        text = paste(line, collapse = " "),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        verse = purrr::map_int(first_line, parse_verse_number),
        text = purrr::map_chr(text, strip_verse_number),
        text = normalize_doc_text(text)
      ) %>%
      dplyr::filter(!is.na(verse), text != "") %>%
      dplyr::select(verse, text)
  } else {
    tibble::tibble(
      verse = seq_along(lines),
      text = normalize_doc_text(lines)
    ) %>%
      dplyr::filter(text != "")
  }
}

parse_dc1835_page <- function(cache_path, url){
  page <- read_cached_html(cache_path)

  section_number <- section_number_from_url(url)
  section_id <- section_id_from_url(url)

  title <- extract_title(page, section_number)

  lines <- extract_page_lines(page)

  lines_for_verses <- lines[
    !stringr::str_detect(
      lines,
      regex("^Section\\s+[0-9A-Za-z-]+", ignore_case = TRUE)
    )
  ]

  verses <- parse_verses_from_lines(lines_for_verses)

  if(nrow(verses) == 0){
    return(
      tibble::tibble(
        section = section_number,
        section_id = section_id,
        title = title,
        verse = integer(),
        text = character(),
        source_url = url,
        cache_path = cache_path
      )
    )
  }

  verses %>%
    dplyr::mutate(
      section = section_number,
      section_id = section_id,
      title = title,
      source_url = url,
      cache_path = cache_path
    ) %>%
    dplyr::select(
      section,
      section_id,
      title,
      verse,
      text,
      source_url,
      cache_path
    )
}
```

## 9. Parse cached pages

```{r parse-pages}
parseable_pages <- download_results %>%
  dplyr::filter(ok) %>%
  dplyr::distinct(discovered_url, cache_path, .keep_all = TRUE) %>%
  dplyr::arrange(discovered_section_number)

dc1835_verses_raw <- purrr::map2_dfr(
  parseable_pages$cache_path,
  parseable_pages$discovered_url,
  parse_dc1835_page
)

dc1835_verses_raw %>%
  dplyr::count(section, section_id, title, name = "n_verses") %>%
  dplyr::arrange(section, section_id) %>%
  print(n = 150)

parse_diagnostics <- dc1835_verses_raw %>%
  dplyr::group_by(section, section_id, title, source_url, cache_path) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_words = sum(stringr::str_count(text, "\\S+"), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(section, section_id)

parse_diagnostics

readr::write_csv(
  parse_diagnostics,
  dc1835_parse_diag_path
)
```

## 10. Build common verse-level structure

```{r verse-structure}
dc1835_verse_docs <- dc1835_verses_raw %>%
  dplyr::filter(!is.na(section), !is.na(verse), text != "") %>%
  dplyr::mutate(
    corpus_id = "DC",
    translation = "1835",
    corpus_family = "DC",
    source_family = "DC_1835",
    edition = "1835",
    author = NA_character_,
    work = "Doctrine and Covenants",
    short_work = "DC_1835",
    book = "Doctrine and Covenants",
    chapter = as.integer(section),
    chapter_title = title,
    verse = as.integer(verse),
    verse_id = paste0("DC_1835__", chapter, "_", verse),
    doc_id = verse_id,
    doc_id_original = doc_id,
    section_label = paste0("Section ", section_id),
    macro_genre = "Modern_Scripture",
    doc_unit = "verse",
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    chapter_title,
    verse,
    verse_id,
    section_id,
    section_label,
    macro_genre,
    doc_unit,
    n_words,
    text,
    source_url,
    cache_path
  )

dc1835_verse_docs %>%
  dplyr::arrange(chapter, verse) %>%
  dplyr::slice_head(n = 20)
```

## 11. Build common chapter-level structure

Each 1835 D&C section is treated as one imported `chapter`.

```{r chapter-structure}
dc1835_chapter_docs <- dc1835_verse_docs %>%
  dplyr::arrange(chapter, verse) %>%
  dplyr::group_by(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    author,
    work,
    short_work,
    book,
    chapter,
    chapter_title,
    section_id,
    section_label
  ) %>%
  dplyr::summarise(
    text = paste(text, collapse = " "),
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    source_url = dplyr::first(source_url),
    cache_path = dplyr::first(cache_path),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    text = normalize_doc_text(text),
    n_words = stringr::str_count(text, "\\S+"),
    macro_genre = "Modern_Scripture",
    doc_unit = "chapter",
    doc_id = paste0(
      "DC_1835__",
      stringr::str_pad(chapter, width = 3, pad = "0")
    ),
    doc_id_original = doc_id
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    chapter_title,
    section_id,
    section_label,
    macro_genre,
    doc_unit,
    n_verses,
    min_verse,
    max_verse,
    n_words,
    text,
    source_url,
    cache_path
  )

dc1835_chapter_docs %>%
  dplyr::arrange(chapter) %>%
  dplyr::select(chapter, section_id, chapter_title, n_verses, n_words, source_url) %>%
  print(n = 150)
```

## 12. Verse gap diagnostics

```{r verse-gaps}
verse_gaps <- dc1835_verse_docs %>%
  dplyr::group_by(chapter, section_id, chapter_title) %>%
  dplyr::summarise(
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_verses = dplyr::n_distinct(verse),
    expected_n = max_verse - min_verse + 1L,
    has_gap = n_verses != expected_n,
    missing_verses = paste(
      setdiff(seq(min_verse, max_verse), sort(unique(verse))),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  dplyr::filter(has_gap)

verse_gaps

readr::write_csv(
  verse_gaps,
  file.path(diagnostic_dir, "dc_1835_verse_gaps.csv")
)
```

## 13. Duplicate / odd section diagnostics

```{r oddities}
dc1835_chapter_docs %>%
  dplyr::count(chapter, name = "n_pages") %>%
  dplyr::filter(n_pages > 1)

dc1835_chapter_docs %>%
  dplyr::filter(is.na(chapter) | n_words < 50) %>%
  dplyr::select(chapter, section_id, chapter_title, n_verses, n_words, source_url) %>%
  print(n = 100)

missing_sections <- tibble::tibble(
  section = seq(
    min(dc1835_chapter_docs$chapter, na.rm = TRUE),
    max(dc1835_chapter_docs$chapter, na.rm = TRUE)
  )
) %>%
  dplyr::anti_join(
    dc1835_chapter_docs %>% dplyr::distinct(section = chapter),
    by = "section"
  )

missing_sections
```

## 14. Write outputs

```{r write-outputs}
readr::write_csv(
  dc1835_verse_docs,
  dc1835_verse_out_path
)

readr::write_csv(
  dc1835_chapter_docs,
  dc1835_chapter_out_path
)

cat("Wrote 1835 DC verse docs to:\n")
cat(dc1835_verse_out_path, "\n\n")

cat("Wrote 1835 DC chapter docs to:\n")
cat(dc1835_chapter_out_path, "\n")
```

## 15. Next step

Use:

```text
data/prepared/dc_1835_chapter_docs.csv
```

in the larger corpus:

```text
KJV + BOM + BOC + DC_1835 + DC_modern + PoGP + JSO + 1830
```

Recommended metadata distinction:

```text
corpus_family = DC
source_family = DC_1835 or DC_modern
edition = 1835 or modern
```
