---
title: "63 - Prepare 1835 Doctrine and Covenants v6"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run from notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)

library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(rvest)
library(xml2)
library(tidyr)
```

## 1. Purpose

This version deliberately separates the workflow into debugging stages:

```text
1. Crawl URL list from section 1 to section 134.
2. Sanity-check the URL list count and stopping point.
3. Download only the confirmed URLs.
4. Sanity-check downloaded page count.
5. Parse verses from cached HTML.
6. Sanity-check parsed pages and verse counts.
7. Save verse-level and chapter-level CSV files.
```

This avoids false fixed-range URLs and avoids redirects into other editions.

The crawler follows only the site's actual `rel="next"` link.

It allows both paths used by the site:

```text
/sections/1835dcs/
/sections/1835dc/
```

and stops when the parsed website section id is `134`.

Outputs:

```text
data/prepared/dc_1835_verse_docs.csv
data/prepared/dc_1835_chapter_docs.csv
```

## 2. Directories

```{r directories}
raw_dc1835_dir <- file.path(STYLO_ROOT, "data", "raw", "dc_1835")
html_cache_dir <- file.path(raw_dc1835_dir, "html")
diagnostic_dir <- file.path(raw_dc1835_dir, "diagnostics")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(raw_dc1835_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(html_cache_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

url_index_path <- file.path(diagnostic_dir, "dc_1835_v6_url_index.csv")
download_diag_path <- file.path(diagnostic_dir, "dc_1835_v6_download_results.csv")
parse_diag_path <- file.path(diagnostic_dir, "dc_1835_v6_parse_diagnostics.csv")
raw_verses_path <- file.path(diagnostic_dir, "dc_1835_v6_raw_verses.csv")

verse_out_path <- file.path(prepared_dir, "dc_1835_verse_docs.csv")
chapter_out_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")
```

## 3. URL helpers

```{r url-helpers}
start_url <- "https://doctrineandcovenants.com/sections/1835dcs/section-1/"
stop_section_id <- "134"

normalize_url <- function(url){

  url <- stringr::str_replace(url, "#.*$", "")

  url <- stringr::str_replace(
    url,
    "^http://doctrineandcovenants\\.com",
    "https://doctrineandcovenants.com"
  )

  if(stringr::str_starts(url, "/")){
    url <- paste0("https://doctrineandcovenants.com", url)
  }

  if(!stringr::str_detect(url, "/$")){
    url <- paste0(url, "/")
  }

  url
}

website_section_id_from_url <- function(url){

  stringr::str_match(
    url,
    regex("/sections?-([^/]+)/?$", ignore_case = TRUE)
  )[, 2]

}

section_numbers_from_id <- function(section_id){

  nums <- stringr::str_extract_all(section_id, "[0-9]+")[[1]]

  as.integer(nums)

}

section_numbers_from_url <- function(url){

  sid <- website_section_id_from_url(url)

  if(is.na(sid)){
    return(integer())
  }

  section_numbers_from_id(sid)

}

chapter_number_from_url <- function(url){

  nums <- section_numbers_from_url(url)

  if(length(nums) == 0){
    return(NA_integer_)
  }

  nums[[1]]

}

is_1835_url <- function(url){

  stringr::str_detect(
    url,
    regex("doctrineandcovenants\\.com/sections/1835dcs?/", ignore_case = TRUE)
  )

}

cache_path_for_url <- function(url){

  sid <- website_section_id_from_url(url)

  if(is.na(sid) || sid == ""){
    sid <- stringr::str_replace_all(url, "[^A-Za-z0-9]+", "_")
  }

  file.path(
    html_cache_dir,
    paste0("sections_", sid, ".html")
  )

}
```

## 4. Crawl URL list only

This stage does **not** parse verses. It only builds the authoritative URL list.

```{r crawl-url-list}
download_for_crawl <- function(url){

  url <- normalize_url(url)
  cache_path <- cache_path_for_url(url)

  if(!file.exists(cache_path)){

    download.file(
      url = url,
      destfile = cache_path,
      mode = "wb",
      quiet = TRUE
    )

  }

  cache_path

}

extract_next_url_from_cache <- function(cache_path){

  page <- xml2::read_html(cache_path, encoding = "UTF-8")

  href <- page %>%
    rvest::html_element("a.tessera-post-nav[rel='next']") %>%
    rvest::html_attr("href")

  if(length(href) == 0 || is.na(href)){
    return(NA_character_)
  }

  href <- normalize_url(href)

  if(!is_1835_url(href)){
    return(NA_character_)
  }

  href

}

crawl_urls_1_to_134 <- function(start_url, stop_section_id = "134", max_pages = 200){

  current_url <- normalize_url(start_url)
  visited <- character()
  rows <- list()

  for(page_order in seq_len(max_pages)){

    if(current_url %in% visited){
      stop("Crawler loop detected at: ", current_url)
    }

    cache_path <- download_for_crawl(current_url)

    website_section_id <- website_section_id_from_url(current_url)
    section_numbers <- section_numbers_from_url(current_url)

    rows[[page_order]] <- tibble::tibble(
      page_order = page_order,
      source_url = current_url,
      website_section_id = website_section_id,
      section_numbers = paste(section_numbers, collapse = ","),
      chapter = chapter_number_from_url(current_url),
      cache_path = cache_path
    )

    visited <- c(visited, current_url)

    if(website_section_id == stop_section_id){
      message("Reached stop section id: ", stop_section_id)
      break
    }

    next_url <- extract_next_url_from_cache(cache_path)

    if(is.na(next_url)){
      stop("No next URL found before reaching section ", stop_section_id, ". Last URL: ", current_url)
    }

    current_url <- next_url
  }

  dplyr::bind_rows(rows)

}

url_index <- crawl_urls_1_to_134(
  start_url = start_url,
  stop_section_id = stop_section_id,
  max_pages = 200
)

readr::write_csv(url_index, url_index_path)

url_index %>%
  dplyr::select(page_order, chapter, website_section_id, section_numbers, source_url) %>%
  print(n = 200)
```

## 5. Sanity check URL list

This chunk should pass before moving on.

```{r sanity-url-list}
url_summary <- url_index %>%
  dplyr::summarise(
    n_pages = dplyr::n(),
    first_url = dplyr::first(source_url),
    last_url = dplyr::last(source_url),
    first_section_id = dplyr::first(website_section_id),
    last_section_id = dplyr::last(website_section_id),
    min_chapter = min(chapter, na.rm = TRUE),
    max_chapter = max(chapter, na.rm = TRUE),
    duplicate_urls = sum(duplicated(source_url))
  )

url_summary

if(url_summary$first_section_id != "1"){
  stop("URL crawl did not start at section id 1.")
}

if(url_summary$last_section_id != stop_section_id){
  stop("URL crawl did not end at section id ", stop_section_id, ".")
}

if(url_summary$duplicate_urls > 0){
  stop("URL crawl contains duplicate URLs.")
}

# Useful, but not fatal: the 1835 site uses grouped pages.
parsed_section_numbers <- url_index %>%
  tidyr::separate_rows(section_numbers, sep = ",") %>%
  dplyr::mutate(section_number = as.integer(section_numbers)) %>%
  dplyr::distinct(section_number) %>%
  dplyr::arrange(section_number)

parsed_section_numbers %>%
  print(n = 150)

missing_1_to_134 <- setdiff(1:134, parsed_section_numbers$section_number)

missing_1_to_134
```

## 6. Download confirmed pages

At this point, `url_index` is the authoritative download list.

```{r download-confirmed-pages}
download_confirmed_page <- function(source_url, cache_path){

  source_url <- normalize_url(source_url)

  ok <- tryCatch({

    if(!file.exists(cache_path)){
      download.file(
        url = source_url,
        destfile = cache_path,
        mode = "wb",
        quiet = TRUE
      )
    }

    file.exists(cache_path) && file.info(cache_path)$size > 1000

  }, error = function(e){
    FALSE
  })

  tibble(
    ok = ok,
    file_size = ifelse(
      file.exists(cache_path),
      file.info(cache_path)$size,
      NA_real_
    )
  )

}

download_results <- url_index %>%
  dplyr::mutate(
    download = purrr::map2(source_url, cache_path, download_confirmed_page)
  ) %>%
  tidyr::unnest(download)

readr::write_csv(download_results, download_diag_path)

download_results %>%
  dplyr::count(ok)

if(any(!download_results$ok)){
  download_results %>%
    dplyr::filter(!ok) %>%
    print(n = 200)

  stop("One or more confirmed URLs failed to download.")
}

if(nrow(download_results) != nrow(url_index)){
  stop("Downloaded page count does not equal URL index count.")
}
```

## 7. Parse helpers

The parser is intentionally simple:

1. Read cached HTML as a string.
2. Extract `<h1>`.
3. Extract all verse blocks that begin with `<h4 ...>NUMBER</h4>`.
4. Strip HTML from the body.
5. Return verse rows.

No fixed-range URLs. No broad DOM walking.

```{r parse-helpers}
normalize_doc_text <- function(x){

  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()

}

html_to_text <- function(x){

  if(is.na(x) || x == ""){
    return("")
  }

  xml2::read_html(paste0("<div>", x, "</div>")) %>%
    rvest::html_text2() %>%
    normalize_doc_text()

}

extract_h1_title <- function(html, fallback_title){

  title_html <- stringr::str_match(
    html,
    regex("<h1[^>]*>(.*?)</h1>", dotall = TRUE, ignore_case = TRUE)
  )[, 2]

  title <- html_to_text(title_html)

  if(is.na(title) || title == ""){
    return(fallback_title)
  }

  title

}

extract_verse_rows <- function(html){

  page <- xml2::read_html(html)

  nodes <- rvest::html_elements(
    page,
    xpath = "//h4[normalize-space(text()) and number(normalize-space(text())) = number(normalize-space(text()))]"
  )

  purrr::map_dfr(nodes, function(h4){

    verse <- h4 %>%
      rvest::html_text2() %>%
      stringr::str_squish() %>%
      as.integer()

    p <- xml2::xml_find_first(
      h4,
      "following-sibling::p[1]"
    )

    text <- p %>%
      rvest::html_text2() %>%
      normalize_doc_text()

    tibble::tibble(
      verse = verse,
      text = text
    )

  }) %>%
    dplyr::filter(!is.na(verse), text != "")
}



parse_dc1835_page_simple <- function(cache_path, source_url, page_order){

  html <- readr::read_file(cache_path)

  fallback_title <- paste0(
    "Section ",
    website_section_id_from_url(source_url)
  )

  chapter_title <- extract_h1_title(
    html,
    fallback_title
  )

  extract_verse_rows(html) %>%
    dplyr::mutate(
      page_order = page_order,
      chapter = chapter_number_from_url(source_url),
      website_section_id = website_section_id_from_url(source_url),
      section_numbers = paste(section_numbers_from_url(source_url), collapse = ","),
      chapter_title = chapter_title,
      source_url = source_url,
      cache_path = cache_path
    ) %>%
    dplyr::select(
      page_order,
      chapter,
      website_section_id,
      section_numbers,
      chapter_title,
      verse,
      text,
      source_url,
      cache_path
    )
}
```

## 8. Parse confirmed pages

```{r parse-confirmed-pages}
dc1835_verses_raw <- purrr::pmap_dfr(
  url_index %>%
    dplyr::select(cache_path, source_url, page_order),
  function(cache_path, source_url, page_order){

    parse_dc1835_page_simple(
      cache_path = cache_path,
      source_url = source_url,
      page_order = page_order
    )

  }
)

readr::write_csv(dc1835_verses_raw, raw_verses_path)

dc1835_verses_raw %>%
  dplyr::count(page_order, chapter, website_section_id, chapter_title, name = "n_verses") %>%
  dplyr::arrange(page_order) %>%
  print(n = 200)
```

## 9. Sanity check parsed pages

```{r sanity-parsed-pages}
parse_diagnostics <- dc1835_verses_raw %>%
  dplyr::group_by(page_order, chapter, website_section_id, section_numbers, chapter_title, source_url, cache_path) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_words = sum(stringr::str_count(text, "\\S+"), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(page_order)

readr::write_csv(parse_diagnostics, parse_diag_path)

parse_diagnostics %>%
  print(n = 200)

parsed_page_count <- parse_diagnostics %>%
  dplyr::summarise(n_pages = dplyr::n()) %>%
  dplyr::pull(n_pages)

if(parsed_page_count != nrow(url_index)){
  stop("Parsed page count does not equal URL index count.")
}

zero_verse_pages <- parse_diagnostics %>%
  dplyr::filter(n_verses == 0 | is.na(n_verses))

if(nrow(zero_verse_pages) > 0){
  zero_verse_pages %>% print(n = 200)
  stop("At least one page parsed zero verses.")
}

tiny_pages <- parse_diagnostics %>%
  dplyr::filter(n_words < 20)

tiny_pages

section_1_check <- parse_diagnostics %>%
  dplyr::filter(page_order == 1)

section_1_check

# Section 1 should have 7 verses on this site.
if(nrow(section_1_check) == 1 && section_1_check$n_verses != 7){
  warning("Section 1 did not parse as 7 verses. Inspect page_order 1.")
}
```

## 10. Build common verse-level CSV

```{r build-verse-docs}
dc1835_verse_docs <- dc1835_verses_raw %>%
  dplyr::mutate(
    corpus_id = "DC",
    translation = "1835",
    corpus_family = "DC",
    source_family = "DC_1835",
    edition = "1835",
    author = NA_character_,
    work = "Doctrine and Covenants",
    short_work = "DC_1835",
    book = "Doctrine and Covenants",
    chapter = as.integer(chapter),
    verse = as.integer(verse),
    verse_id = paste0(
      "DC_1835__",
      stringr::str_pad(page_order, width = 3, pad = "0"),
      "_",
      stringr::str_pad(verse, width = 3, pad = "0")
    ),
    doc_id = verse_id,
    doc_id_original = doc_id,
    section_label = dplyr::if_else(
      stringr::str_detect(website_section_id, "-"),
      paste0("Sections ", website_section_id),
      paste0("Section ", website_section_id)
    ),
    macro_genre = "Modern_Scripture",
    doc_unit = "verse",
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    verse,
    verse_id,
    website_section_id,
    section_numbers,
    section_label,
    macro_genre,
    doc_unit,
    n_words,
    text,
    source_url,
    cache_path
  )

readr::write_csv(dc1835_verse_docs, verse_out_path)

dc1835_verse_docs %>%
  dplyr::slice_head(n = 20)
```

## 11. Build common chapter-level CSV

Each downloaded 1835 page is one stylometric unit.

```{r build-chapter-docs}
dc1835_chapter_docs <- dc1835_verse_docs %>%
  dplyr::arrange(page_order, verse) %>%
  dplyr::group_by(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    website_section_id,
    section_numbers,
    section_label,
    source_url,
    cache_path
  ) %>%
  dplyr::summarise(
    text = paste(text, collapse = " "),
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    text = normalize_doc_text(text),
    n_words = stringr::str_count(text, "\\S+"),
    macro_genre = "Modern_Scripture",
    doc_unit = "chapter",
    doc_id = paste0(
      "DC_1835__",
      stringr::str_pad(page_order, width = 3, pad = "0"),
      "__",
      dplyr::if_else(
        stringr::str_detect(website_section_id, "-"),
        paste0("sections_", website_section_id),
        paste0("section_", website_section_id)
      )
    ),
    doc_id_original = doc_id
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    website_section_id,
    section_numbers,
    section_label,
    macro_genre,
    doc_unit,
    n_verses,
    min_verse,
    max_verse,
    n_words,
    text,
    source_url,
    cache_path
  )

readr::write_csv(dc1835_chapter_docs, chapter_out_path)

dc1835_chapter_docs %>%
  dplyr::select(page_order, chapter, website_section_id, chapter_title, n_verses, n_words, source_url) %>%
  dplyr::arrange(page_order) %>%
  print(n = 200)
```

## 12. Final checks

```{r final-checks}
cat("URL pages:", nrow(url_index), "\n")
cat("Parsed chapter docs:", nrow(dc1835_chapter_docs), "\n")
cat("Verse docs:", nrow(dc1835_verse_docs), "\n")

if(nrow(dc1835_chapter_docs) != nrow(url_index)){
  stop("Final chapter docs do not match URL index count.")
}

dc1835_chapter_docs %>%
  dplyr::summarise(
    n_pages = dplyr::n(),
    total_verses = sum(n_verses),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )


dc1835_chapter_docs %>%
  arrange(n_words) %>%
  select(
    page_order,
    chapter,
    website_section_id,
    n_verses,
    n_words,
    chapter_title
  ) %>%
  print(n = 20)

dc1835_chapter_docs %>%
  arrange(desc(n_words)) %>%
  select(
    page_order,
    chapter,
    website_section_id,
    n_verses,
    n_words,
    chapter_title
  ) %>%
  print(n = 20)

```
