---
title: "63 - Prepare 1835 Doctrine and Covenants v4"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
# Run this notebook from inside the notebooks/ folder.

source("./00_source_project.R")

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

options(
  tibble.width = Inf,
  tibble.print_max = 300
)
```

## 1. Purpose

This notebook downloads, caches, parses, and prepares the **1835 Doctrine and Covenants** from:

```text
https://doctrineandcovenants.com/sections/1835dcs/section-1/
```

It follows the actual site navigation by repeatedly reading the `rel="next"` link:

```html
<a class="tessera-post-nav" href="..." rel="next">...</a>
```

This avoids false fixed-range URLs that redirect to other editions.

The crawl stops after parsing:

```text
https://doctrineandcovenants.com/sections/1835dcs/section-134/
```

Target outputs:

```text
data/prepared/dc_1835_chapter_docs.csv
data/prepared/dc_1835_verse_docs.csv
```

Metadata:

```text
corpus_family = "DC"
source_family = "DC_1835"
edition = "1835"
```

Each downloaded 1835 page is treated as one imported `chapter` / stylometric unit.

The key v4 change is that verses are parsed from the DOM using:

```html
<h4 class="wp-block-heading">7</h4>
```

as the authoritative verse marker, instead of line-based text parsing. The parser stops after the last verse when it reaches post-navigation, dates, reading-time metadata, or next-post content.

## 2. Packages

```{r packages}
library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(rvest)
library(xml2)
library(tidyr)
```

## 3. Directories

```{r directories}
raw_dc1835_dir <- file.path(STYLO_ROOT, "data", "raw", "dc_1835")
html_cache_dir <- file.path(raw_dc1835_dir, "html")
diagnostic_dir <- file.path(raw_dc1835_dir, "diagnostics")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")

dir.create(raw_dc1835_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(html_cache_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)

dc1835_verse_out_path <- file.path(prepared_dir, "dc_1835_verse_docs.csv")
dc1835_chapter_out_path <- file.path(prepared_dir, "dc_1835_chapter_docs.csv")

crawl_diag_path <- file.path(diagnostic_dir, "dc_1835_v4_crawl_results.csv")
parse_diag_path <- file.path(diagnostic_dir, "dc_1835_v4_parse_diagnostics.csv")
gap_diag_path <- file.path(diagnostic_dir, "dc_1835_v4_verse_gaps.csv")
raw_verse_diag_path <- file.path(diagnostic_dir, "dc_1835_v4_raw_verses.csv")
```

## 4. URL and cache helpers

```{r url-cache-helpers}
start_url <- "https://doctrineandcovenants.com/sections/1835dcs/section-1/"
stop_url <- "https://doctrineandcovenants.com/sections/1835dcs/section-134/"

safe_slug <- function(x){
  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("^_+|_+$", "")
}

normalize_url <- function(url){
  url <- stringr::str_replace(url, "#.*$", "")

  url <- stringr::str_replace(
    url,
    "^http://doctrineandcovenants\\.com",
    "https://doctrineandcovenants.com"
  )

  if(stringr::str_starts(url, "/")){
    url <- paste0("https://doctrineandcovenants.com", url)
  }

  if(!stringr::str_detect(url, "/$")){
    url <- paste0(url, "/")
  }

  url
}

start_url <- normalize_url(start_url)
stop_url <- normalize_url(stop_url)

website_section_id_from_url <- function(url){
  stringr::str_match(
    url,
    regex("/sections?-([^/]+)/?$", ignore_case = TRUE)
  )[, 2]
}

section_numbers_from_id <- function(section_id){
  nums <- stringr::str_extract_all(section_id, "[0-9]+")[[1]]
  as.integer(nums)
}

section_numbers_from_url <- function(url){
  section_id <- website_section_id_from_url(url)

  if(is.na(section_id)){
    return(integer())
  }

  section_numbers_from_id(section_id)
}

chapter_number_from_url <- function(url){
  nums <- section_numbers_from_url(url)

  if(length(nums) == 0){
    return(NA_integer_)
  }

  nums[[1]]
}

cache_path_for_url <- function(url){
  url <- normalize_url(url)
  website_section_id <- website_section_id_from_url(url)

  if(is.na(website_section_id) || website_section_id == ""){
    website_section_id <- safe_slug(url)
  }

  file.path(
    html_cache_dir,
    paste0("sections_", website_section_id, ".html")
  )
}

download_section_html <- function(url, overwrite = FALSE){
  url <- normalize_url(url)
  cache_path <- cache_path_for_url(url)

  if(file.exists(cache_path) && !overwrite){
    return(
      tibble::tibble(
        url = url,
        cache_path = cache_path,
        downloaded = FALSE,
        ok = TRUE,
        status = "cached"
      )
    )
  }

  ok <- tryCatch({
    download.file(
      url = url,
      destfile = cache_path,
      mode = "wb",
      quiet = TRUE
    )

    file.exists(cache_path) && file.info(cache_path)$size > 1000
  }, error = function(e){
    FALSE
  })

  tibble::tibble(
    url = url,
    cache_path = cache_path,
    downloaded = ok,
    ok = ok,
    status = dplyr::if_else(ok, "downloaded", "failed")
  )
}

read_cached_html <- function(cache_path){
  xml2::read_html(cache_path, encoding = "UTF-8")
}
```

## 5. Next-link crawler

Rules:

1. Start at `section-1/`.
2. Download/cache the current page.
3. Record the current page.
4. If current page is `section-134/`, stop immediately.
5. Otherwise extract the `rel="next"` link and continue.
6. Only accept next links inside `/sections/1835dcs/`.

```{r crawl-by-next}
extract_next_1835_url <- function(page){

  next_href <- page %>%
    rvest::html_element("a.tessera-post-nav[rel='next']") %>%
    rvest::html_attr("href")

  if(length(next_href) == 0 || is.na(next_href)){
    return(NA_character_)
  }

  next_url <- normalize_url(next_href)

  if(!stringr::str_detect(
    next_url,
    fixed("doctrineandcovenants.com/sections/1835dcs/")
  )){
    return(NA_character_)
  }

  next_url
}

crawl_1835_by_next <- function(start_url, stop_url, max_pages = 150, overwrite = FALSE){

  current_url <- normalize_url(start_url)
  stop_url <- normalize_url(stop_url)

  out <- tibble::tibble()
  visited <- character()

  for(i in seq_len(max_pages)){

    if(current_url %in% visited){
      warning("Loop detected at: ", current_url)
      break
    }

    dl <- download_section_html(current_url, overwrite = overwrite)

    if(!isTRUE(dl$ok[[1]])){
      warning("Failed to download: ", current_url)
      break
    }

    page <- read_cached_html(dl$cache_path[[1]])

    website_section_id <- website_section_id_from_url(current_url)
    section_numbers <- section_numbers_from_url(current_url)

    out <- dplyr::bind_rows(
      out,
      tibble::tibble(
        page_order = i,
        discovered_url = current_url,
        website_section_id = website_section_id,
        section_numbers = paste(section_numbers, collapse = ","),
        chapter = chapter_number_from_url(current_url),
        cache_path = dl$cache_path[[1]],
        downloaded = dl$downloaded[[1]],
        status = dl$status[[1]],
        ok = dl$ok[[1]]
      )
    )

    visited <- c(visited, current_url)

    if(current_url == stop_url){
      message("Reached stop URL: ", stop_url)
      break
    }

    next_url <- extract_next_1835_url(page)

    if(is.na(next_url)){
      warning("No next 1835 link found after: ", current_url)
      break
    }

    current_url <- next_url
  }

  out
}

section_url_index <- crawl_1835_by_next(
  start_url = start_url,
  stop_url = stop_url,
  max_pages = 150,
  overwrite = FALSE
)

section_url_index %>%
  print(n = 150)

readr::write_csv(section_url_index, crawl_diag_path)
```

## 6. Text helpers

```{r text-helpers}
clean_text <- function(x){
  x %>%
    stringr::str_replace_all("\uFEFF", "") %>%
    stringr::str_replace_all("\r", "\n") %>%
    stringr::str_replace_all("\u00A0", " ") %>%
    stringr::str_replace_all("[ \t]+", " ") %>%
    stringr::str_replace_all("\n{3,}", "\n\n") %>%
    stringr::str_trim()
}

normalize_doc_text <- function(x){
  x %>%
    stringr::str_replace_all("\\s+", " ") %>%
    stringr::str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    stringr::str_squish()
}

extract_title <- function(page, fallback_label){
  title_candidates <- c(
    page %>% rvest::html_elements("h1") %>% rvest::html_text2(),
    page %>% rvest::html_elements("h2") %>% rvest::html_text2(),
    page %>% rvest::html_elements("title") %>% rvest::html_text2()
  )

  title_candidates <- title_candidates %>%
    stringr::str_squish()

  title_candidates <- title_candidates[
    !is.na(title_candidates) &
      title_candidates != "" &
      !stringr::str_detect(title_candidates, regex("Doctrine and Covenants.com", ignore_case = TRUE))
  ]

  if(length(title_candidates) > 0){
    return(title_candidates[[1]])
  }

  fallback_label
}

extract_main_node <- function(page){
  selectors <- c(
    "article",
    "main",
    ".entry-content",
    ".post-content",
    ".content",
    "#content"
  )

  for(sel in selectors){
    node <- page %>% rvest::html_element(sel)

    if(length(node) > 0 && !inherits(node, "xml_missing")){
      return(node)
    }
  }

  page %>% rvest::html_element("body")
}
```

## 7. DOM verse parser

The 1835 pages encode verse numbers as heading nodes, for example:

```html
<h4 class="wp-block-heading">7</h4>
```

This parser walks the direct child nodes of the content area in order.

It starts a verse when it sees a numeric `h4`.

It appends later paragraph/list/text blocks to that verse.

It stops when it reaches navigation or later post metadata/content.

```{r dom-verse-parser}
is_verse_heading <- function(node){
  tag <- xml2::xml_name(node)

  txt <- rvest::html_text2(node) %>%
    stringr::str_squish()

  tag == "h4" &&
    stringr::str_detect(txt, "^[0-9]{1,3}$")
}

is_stop_node <- function(node){

  cls <- xml2::xml_attr(node, "class")
  id <- xml2::xml_attr(node, "id")
  tag <- xml2::xml_name(node)

  txt <- rvest::html_text2(node) %>%
    stringr::str_squish()

  combined <- paste(
    tag,
    cls,
    id,
    txt,
    sep = " "
  )

  stringr::str_detect(
    combined,
    regex(
      paste(
        c(
          "tessera-post-nav",
          "post-navigation",
          "nav-links",
          "wp-block-post-date",
          "wp-block-post-title",
          "Reading Time",
          "^h\\d\\s+Introduction to the 1835 Doctrine and Covenants$",
          "Introduction to the 1835 Doctrine and Covenants",
          "^ON PRIESTHOOD AND CALLING",
          "^Revelation to Joseph Smith",
          "^December [0-9]{1,2}, [0-9]{4}$",
          "^January [0-9]{1,2}, [0-9]{4}$",
          "^February [0-9]{1,2}, [0-9]{4}$",
          "^March [0-9]{1,2}, [0-9]{4}$",
          "^April [0-9]{1,2}, [0-9]{4}$",
          "^May [0-9]{1,2}, [0-9]{4}$",
          "^June [0-9]{1,2}, [0-9]{4}$",
          "^July [0-9]{1,2}, [0-9]{4}$",
          "^August [0-9]{1,2}, [0-9]{4}$",
          "^September [0-9]{1,2}, [0-9]{4}$",
          "^October [0-9]{1,2}, [0-9]{4}$",
          "^November [0-9]{1,2}, [0-9]{4}$"
        ),
        collapse = "|"
      ),
      ignore_case = TRUE
    )
  )
}

is_non_content_node <- function(node){

  tag <- xml2::xml_name(node)

  cls <- xml2::xml_attr(node, "class")
  txt <- rvest::html_text2(node) %>%
    stringr::str_squish()

  combined <- paste(tag, cls, txt)

  stringr::str_detect(
    combined,
    regex(
      paste(
        c(
          "^script",
          "^style",
          "sharedaddy",
          "jp-relatedposts",
          "comments",
          "comment-respond",
          "wp-block-separator",
          "tessera-post-nav",
          "post-navigation",
          "Previous",
          "Next",
          "Share this",
          "Like this"
        ),
        collapse = "|"
      ),
      ignore_case = TRUE
    )
  )
}


parse_verses_dom <- function(main_node){

  nodes <- rvest::html_elements(
    main_node,
    xpath = "./*"
  )

  verses <- list()

  current_verse <- NULL
  current_text <- character()

  started <- FALSE

  for(node in nodes){

    # Stop once we've started verses and encounter
    # navigation / metadata / next article
    if(started && is_stop_node(node)){
      break
    }

    # Ignore obvious non-content nodes
    if(is_non_content_node(node)){
      next
    }

    txt <- rvest::html_text2(node) %>%
      stringr::str_squish()

    if(txt == ""){
      next
    }

    # Verse marker:
    # <h4 class="wp-block-heading">7</h4>

    if(is_verse_heading(node)){

      # Save previous verse
      if(!is.null(current_verse)){

        verses[[length(verses) + 1L]] <-
          tibble::tibble(
            verse = current_verse,
            text = paste(
              current_text,
              collapse = " "
            ) %>%
              normalize_doc_text()
          )

      }

      current_verse <- as.integer(txt)

      current_text <- character()

      started <- TRUE

      next
    }

    # Append text to current verse

    if(!is.null(current_verse)){

      current_text <- c(
        current_text,
        txt
      )

    }

  }

  # Save last verse

  if(!is.null(current_verse)){

    verses[[length(verses) + 1L]] <-
      tibble::tibble(
        verse = current_verse,
        text = paste(
          current_text,
          collapse = " "
        ) %>%
          normalize_doc_text()
      )

  }

  # IMPORTANT:
  # bind_rows(list()) gives a 0-column tibble,
  # so explicitly return verse/text columns.

  if(length(verses) == 0){

    return(
      tibble::tibble(
        verse = integer(),
        text = character()
      )
    )

  }

  result <- dplyr::bind_rows(verses)

  result %>%
    dplyr::filter(
      !is.na(verse),
      text != ""
    ) %>%
    dplyr::arrange(verse)

}


parse_dc1835_page <- function(cache_path, url, page_order){

  page <- read_cached_html(cache_path)

  # Remove obvious non-content containers before parsing.
  xml2::xml_find_all(
    page,
    ".//script|.//style|.//nav|.//footer|.//header|.//aside"
  ) %>%
    xml2::xml_remove()

  website_section_id <- website_section_id_from_url(url)
  section_numbers <- section_numbers_from_url(url)
  chapter_number <- chapter_number_from_url(url)

  fallback_label <- if(length(section_numbers) > 1){
    paste0("Sections ", paste(section_numbers, collapse = ", "))
  } else {
    paste0("Section ", website_section_id)
  }

  title <- extract_title(page, fallback_label)

  main_node <- extract_main_node(page)

  verses <- parse_verses_dom(main_node)

  if(nrow(verses) == 0){

    warning("No DOM verses parsed for: ", url)

    return(
      tibble::tibble(
        page_order = page_order,
        chapter = chapter_number,
        website_section_id = website_section_id,
        section_numbers = paste(section_numbers, collapse = ","),
        chapter_title = title,
        verse = integer(),
        text = character(),
        source_url = url,
        cache_path = cache_path
      )
    )
  }

  verses %>%
    dplyr::mutate(
      page_order = page_order,
      chapter = chapter_number,
      website_section_id = website_section_id,
      section_numbers = paste(section_numbers, collapse = ","),
      chapter_title = title,
      source_url = url,
      cache_path = cache_path
    ) %>%
    dplyr::select(
      page_order,
      chapter,
      website_section_id,
      section_numbers,
      chapter_title,
      verse,
      text,
      source_url,
      cache_path
    )
}
```

## 8. Parse cached pages

```{r parse-pages}
parseable_pages <- section_url_index %>%
  dplyr::filter(ok) %>%
  dplyr::arrange(page_order)

dc1835_verses_raw <- purrr::pmap_dfr(
  parseable_pages %>%
    dplyr::select(cache_path, discovered_url, page_order),
  function(cache_path, discovered_url, page_order){
    parse_dc1835_page(
      cache_path = cache_path,
      url = discovered_url,
      page_order = page_order
    )
  }
)

readr::write_csv(dc1835_verses_raw, raw_verse_diag_path)

dc1835_verses_raw %>%
  dplyr::count(page_order, chapter, website_section_id, section_numbers, chapter_title, name = "n_verses") %>%
  dplyr::arrange(page_order) %>%
  print(n = 150)

parse_diagnostics <- dc1835_verses_raw %>%
  dplyr::group_by(page_order, chapter, website_section_id, section_numbers, chapter_title, source_url, cache_path) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_words = sum(stringr::str_count(text, "\\S+"), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(page_order)

parse_diagnostics

readr::write_csv(parse_diagnostics, parse_diag_path)
```

## 9. Section 1 check

Section 1 should parse as 7 verses.

```{r section-1-check}
dc1835_verses_raw %>%
  dplyr::filter(page_order == 1) %>%
  dplyr::select(verse, text) %>%
  print(n = 20)

dc1835_verses_raw %>%
  dplyr::filter(page_order == 1) %>%
  dplyr::summarise(
    n_verses = dplyr::n(),
    min_verse = min(verse),
    max_verse = max(verse)
  )
```

## 10. Build verse-level structure

```{r verse-structure}
dc1835_verse_docs <- dc1835_verses_raw %>%
  dplyr::filter(!is.na(chapter), !is.na(verse), text != "") %>%
  dplyr::mutate(
    corpus_id = "DC",
    translation = "1835",
    corpus_family = "DC",
    source_family = "DC_1835",
    edition = "1835",
    author = NA_character_,
    work = "Doctrine and Covenants",
    short_work = "DC_1835",
    book = "Doctrine and Covenants",
    chapter = as.integer(chapter),
    verse = as.integer(verse),
    verse_id = paste0(
      "DC_1835__",
      stringr::str_pad(chapter, width = 3, pad = "0"),
      "_",
      stringr::str_pad(verse, width = 3, pad = "0")
    ),
    doc_id = verse_id,
    doc_id_original = doc_id,
    section_label = dplyr::if_else(
      stringr::str_detect(website_section_id, "-"),
      paste0("Sections ", website_section_id),
      paste0("Section ", website_section_id)
    ),
    macro_genre = "Modern_Scripture",
    doc_unit = "verse",
    n_words = stringr::str_count(text, "\\S+")
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    verse,
    verse_id,
    website_section_id,
    section_numbers,
    section_label,
    macro_genre,
    doc_unit,
    n_words,
    text,
    source_url,
    cache_path
  )

dc1835_verse_docs %>%
  dplyr::arrange(page_order, verse) %>%
  dplyr::slice_head(n = 20)
```

## 11. Build chapter-level structure

Each downloaded 1835 page is treated as one chapter/stylometric unit.

```{r chapter-structure}
dc1835_chapter_docs <- dc1835_verse_docs %>%
  dplyr::arrange(page_order, verse) %>%
  dplyr::group_by(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    website_section_id,
    section_numbers,
    section_label
  ) %>%
  dplyr::summarise(
    text = paste(text, collapse = " "),
    n_verses = dplyr::n(),
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    source_url = dplyr::first(source_url),
    cache_path = dplyr::first(cache_path),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    text = normalize_doc_text(text),
    n_words = stringr::str_count(text, "\\S+"),
    macro_genre = "Modern_Scripture",
    doc_unit = "chapter",
    doc_id = paste0(
      "DC_1835__",
      stringr::str_pad(page_order, width = 3, pad = "0"),
      "__",
      dplyr::if_else(
        stringr::str_detect(website_section_id, "-"),
        paste0("sections_", website_section_id),
        paste0("section_", website_section_id)
      )
    ),
    doc_id_original = doc_id
  ) %>%
  dplyr::select(
    corpus_id,
    translation,
    corpus_family,
    source_family,
    edition,
    doc_id,
    doc_id_original,
    author,
    work,
    short_work,
    book,
    chapter,
    page_order,
    chapter_title,
    website_section_id,
    section_numbers,
    section_label,
    macro_genre,
    doc_unit,
    n_verses,
    min_verse,
    max_verse,
    n_words,
    text,
    source_url,
    cache_path
  )

dc1835_chapter_docs %>%
  dplyr::arrange(page_order) %>%
  dplyr::select(page_order, chapter, website_section_id, section_numbers, chapter_title, n_verses, n_words, source_url) %>%
  print(n = 150)
```

## 12. Diagnostics

```{r diagnostics}
verse_gaps <- dc1835_verse_docs %>%
  dplyr::group_by(page_order, chapter, website_section_id, chapter_title) %>%
  dplyr::summarise(
    min_verse = min(verse, na.rm = TRUE),
    max_verse = max(verse, na.rm = TRUE),
    n_verses = dplyr::n_distinct(verse),
    expected_n = max_verse - min_verse + 1L,
    has_gap = n_verses != expected_n,
    missing_verses = paste(
      setdiff(seq(min_verse, max_verse), sort(unique(verse))),
      collapse = ", "
    ),
    .groups = "drop"
  ) %>%
  dplyr::filter(has_gap)

verse_gaps

readr::write_csv(verse_gaps, gap_diag_path)

dc1835_chapter_docs %>%
  dplyr::filter(is.na(chapter) | n_words < 50) %>%
  dplyr::select(page_order, chapter, website_section_id, chapter_title, n_verses, n_words, source_url) %>%
  print(n = 100)

parsed_section_numbers <- dc1835_chapter_docs %>%
  tidyr::separate_rows(section_numbers, sep = ",") %>%
  dplyr::mutate(section_number = as.integer(section_numbers)) %>%
  dplyr::distinct(section_number) %>%
  dplyr::arrange(section_number)

parsed_section_numbers %>%
  print(n = 150)

setdiff(1:134, parsed_section_numbers$section_number)
```

## 13. Write outputs

```{r write-outputs}
readr::write_csv(dc1835_verse_docs, dc1835_verse_out_path)
readr::write_csv(dc1835_chapter_docs, dc1835_chapter_out_path)

cat("Wrote 1835 DC verse docs to:\n")
cat(dc1835_verse_out_path, "\n\n")

cat("Wrote 1835 DC chapter docs to:\n")
cat(dc1835_chapter_out_path, "\n")
```

## 14. Next step

Use:

```text
data/prepared/dc_1835_chapter_docs.csv
```

in the larger corpus:

```text
KJV + BOM + BOC + DC_1835 + DC_modern + PoGP + JSO + 1830
```

Recommended metadata:

```text
corpus_family = DC
source_family = DC_1835 or DC_modern
edition = 1835 or modern
```
