---
title: "64 - Prepare JSO Manual Corpus"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
source("./00_source_project.R")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(tibble.width = Inf, tibble.print_max = 300)

library(dplyr)
library(stringr)
library(readr)
library(purrr)
library(tibble)
library(tidyr)
library(digest)
```

## 1. Purpose

Parse the JSO manual corpus from:

```text
data/raw/JSO/manual/
```

Only files ending with:

```text
__manual.txt
```

are considered. Each file is one author bundle. There is no separate book/work designation in the filename.

Chapter/document dividers follow the same convention used for the 1830 manual corpus:

```text
### CHAPTER 1: Title
### CHAPTER 2-24: Title
```

Outputs:

```text
data/prepared/jso_manual_chapter_docs.csv
data/prepared/jso_manual_book_docs.csv
data/prepared/jso_manual_manifest.csv
```

## 2. Directories

```{r dirs}
manual_dir <- file.path(STYLO_ROOT, "data", "raw", "JSO", "manual")
prepared_dir <- file.path(STYLO_ROOT, "data", "prepared")
diagnostic_dir <- file.path(STYLO_ROOT, "data", "raw", "JSO", "diagnostics")

dir.create(prepared_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)

chapter_out_path <- file.path(prepared_dir, "jso_manual_chapter_docs.csv")
book_out_path <- file.path(prepared_dir, "jso_manual_book_docs.csv")
manifest_out_path <- file.path(prepared_dir, "jso_manual_manifest.csv")

marker_diag_path <- file.path(diagnostic_dir, "jso_manual_marker_counts.csv")
parse_diag_path <- file.path(diagnostic_dir, "jso_manual_parse_diagnostics.csv")
diff_diag_path <- file.path(diagnostic_dir, "jso_manual_marker_vs_parsed_diff.csv")
```

## 3. Enumerate manual files

```{r enumerate}
manual_files <- list.files(
  manual_dir,
  pattern = "__manual\\.txt$",
  full.names = TRUE
)

manual_manifest <- tibble(
  path = manual_files,
  filename = basename(manual_files),
  file_size = file.info(manual_files)$size
) %>%
  filter(file_size > 0) %>%
  arrange(filename)

manual_manifest

readr::write_csv(manual_manifest, manifest_out_path)
```

### Sanity check: file count

```{r sanity-file-count}
manual_manifest %>%
  summarise(
    n_files = n(),
    min_size = min(file_size),
    median_size = median(file_size),
    max_size = max(file_size)
  )

if(nrow(manual_manifest) == 0){
  stop("No non-empty JSO __manual.txt files found in: ", manual_dir)
}
```

## 4. Filename metadata parser

Expected form:

```text
Sidney_Rigdon__manual.txt
Joseph_Smith__manual.txt
Oliver_Cowdery__manual.txt
```

```{r filename-parser}
make_safe_id <- function(x){
  x %>%
    str_replace_all("[^A-Za-z0-9]+", "_") %>%
    str_replace_all("^_+|_+$", "")
}

infer_jso_author_code <- function(author){
  case_when(
    str_detect(author, regex("^Joseph\\s+Smith", ignore_case = TRUE)) ~ "JS",
    str_detect(author, regex("^Sidney\\s+Rigdon", ignore_case = TRUE)) ~ "SR",
    str_detect(author, regex("^Oliver\\s+Cowdery", ignore_case = TRUE)) ~ "OC",
    TRUE ~ make_safe_id(author)
  )
}

parse_jso_filename_metadata <- function(path){
  fn <- basename(path)

  author_key <- fn %>%
    str_remove("__manual\\.txt$")

  author <- author_key %>%
    str_replace_all("_", " ") %>%
    str_squish()

  author_code <- infer_jso_author_code(author)

  tibble(
    author = author,
    author_code = author_code,
    work = paste0(author, " Manual Writings"),
    short_work = paste0(author_code, "_Manual"),
    source_file = fn,
    local_file = path,
    macro_genre = "Early_LDS_Writing"
  )
}

file_metadata <- purrr::map_dfr(
  manual_manifest$path,
  parse_jso_filename_metadata
)

file_metadata
```

### Sanity check: author mapping

```{r sanity-author-mapping}
file_metadata %>%
  count(author, author_code, source_file, name = "n_files") %>%
  arrange(author)

if(any(is.na(file_metadata$author) | file_metadata$author == "")){
  stop("At least one JSO file has a missing author.")
}

if(any(duplicated(file_metadata$source_file))){
  stop("Duplicate source_file entries detected.")
}
```

## 5. Text helpers

```{r text-helpers}
normalize_doc_text <- function(x){
  x %>%
    str_replace_all("\r", "\n") %>%
    str_replace_all("\uFEFF", "") %>%
    str_replace_all("\u00A0", " ") %>%
    str_replace_all("[ \t]+", " ") %>%
    str_replace_all("\n{3,}", "\n\n") %>%
    str_replace_all("\\s+", " ") %>%
    str_replace_all("\\s+([,.;:!?])", "\\1") %>%
    str_squish()
}

clean_heading <- function(x){
  x %>%
    str_replace_all("[\\*_]", "") %>%
    str_replace_all("\\s+", " ") %>%
    str_squish() %>%
    str_remove("\\s*\\.$")
}
```

## 6. Count chapter markers before parsing

```{r marker-counts}
chapter_divider_pattern <- regex(
  "(?m)^\\s*###\\s+CHAPTER\\b.*$",
  ignore_case = TRUE
)

marker_counts <- manual_manifest %>%
  mutate(
    n_markers = purrr::map_int(
      path,
      ~{
        txt <- readr::read_file(.x)
        stringr::str_count(txt, chapter_divider_pattern)
      }
    )
  ) %>%
  select(filename, path, n_markers) %>%
  arrange(desc(n_markers), filename)

marker_counts

readr::write_csv(marker_counts, marker_diag_path)
```

### Sanity check: marker totals

```{r sanity-marker-counts}
marker_counts %>%
  summarise(
    n_files = n(),
    total_markers = sum(n_markers),
    min_markers = min(n_markers),
    median_markers = median(n_markers),
    max_markers = max(n_markers)
  )

zero_marker_files <- marker_counts %>%
  filter(n_markers == 0)

zero_marker_files

if(nrow(zero_marker_files) > 0){
  warning("At least one JSO manual file has zero chapter markers.")
}
```

## 7. Chapter parser

```{r chapter-parser}
parse_manual_chapters <- function(path){

  txt <- readr::read_file(path) %>%
    str_replace_all("\r", "\n") %>%
    str_replace_all("\uFEFF", "")

  locs <- str_locate_all(txt, chapter_divider_pattern)[[1]]
  headings <- str_extract_all(txt, chapter_divider_pattern)[[1]]
  
  
  
  
  
  
  
  
  
  
  tibble(

  heading = headings,

  start = locs[,1],

  stop = locs[,2],

  matched_text =

    purrr::map2_chr(

      locs[,1],

      locs[,2],

      ~str_sub(
        txt,
        .x,
        .y
      )

    )

)

  if(length(headings) == 0){
    return(
      tibble(
        chapter = integer(),
        chapter_original = character(),
        chapter_title = character(),
        chapter_heading_raw = character(),
        text = character()
      )
    )
  }

  purrr::map_dfr(seq_along(headings), function(i){

    heading_raw <- headings[[i]]

    after_chapter <- heading_raw %>%
      str_remove(regex("^\\s*###\\s+CHAPTER\\s*", ignore_case = TRUE)) %>%
      str_squish()

    chapter_original <- after_chapter %>%
      str_replace("\\s*:.*$", "") %>%
      str_remove("\\s*\\.$") %>%
      str_squish()

    chapter_title <- after_chapter %>%
      str_replace("^[^:]+:?\\s*", "") %>%
      str_remove("\\s*\\.$") %>%
      str_squish()

    if(is.na(chapter_title) || chapter_title == ""){
      chapter_title <- paste0("Chapter ", chapter_original)
    }

    start <- locs[i, 2] + 1
    end <- if(i < nrow(locs)) locs[i + 1, 1] - 1 else nchar(txt)

    tibble(
      chapter = i,
      chapter_original = chapter_original,
      chapter_title = chapter_title,
      chapter_heading_raw = heading_raw,
      text = normalize_doc_text(str_sub(txt, start, end))
    )
  }) %>%
    filter(text != "")
}
```

## 8. Parse all JSO manual files

```{r parse-all}
jso_manual_chapter_docs <- file_metadata %>%
  mutate(
    chapters = purrr::map(local_file, parse_manual_chapters)
  ) %>%
  tidyr::unnest(chapters) %>%

  mutate(

    text = text %>%
      stringr::str_replace_all("\r", " ") %>%
      stringr::str_replace_all("\n", " ") %>%
      stringr::str_squish(),

    chapter_title = chapter_title %>%
      stringr::str_squish(),

    chapter_heading_raw = chapter_heading_raw %>%
      stringr::str_squish(),

    chapter_original = chapter_original %>%
      stringr::str_squish()

  ) %>%

  group_by(source_file) %>%
  mutate(
    chapter = row_number()
  ) %>%
  ungroup() %>%

  mutate(
    chapter_digest = purrr::pmap_chr(
      list(author, work, chapter_original, chapter_title, text),
      ~digest::digest(
        paste(..., sep = "||"),
        algo = "xxhash64"
      )
    )
  ) %>%

  group_by(source_file) %>%
  mutate(
    duplicate_digest_in_work = duplicated(chapter_digest)
  ) %>%
  ungroup() %>%

  filter(
    !duplicate_digest_in_work
  ) %>%

  group_by(source_file) %>%
  mutate(
    chapter = row_number()
  ) %>%
  ungroup() %>%

  mutate(
    corpus_family = "JSO",

    source_family = "JSO",

    doc_unit = "chapter",

    doc_id = paste0(
      "JSO__",
      make_safe_id(author_code),
      "__",
      str_pad(chapter, width = 3, pad = "0")
    ),

    doc_id_original = doc_id,

    n_words = stringr::str_count(
      text,
      "\\S+"
    )

  ) %>%

  select(
    corpus_family,
    source_family,

    author,
    author_code,

    work,
    short_work,

    source_file,
    local_file,

    doc_id,
    doc_id_original,

    macro_genre,

    doc_unit,

    chapter,
    chapter_original,

    chapter_title,

    chapter_heading_raw,

    chapter_digest,

    n_words,

    text
  )

jso_manual_chapter_docs %>%
  count(author, source_file, name = "n_chapters") %>%
  arrange(author, source_file) %>%
  print(n = 100)
```

## 9. Sanity check: marker count versus parsed count

```{r sanity-marker-vs-parsed}
parsed_counts <- jso_manual_chapter_docs %>%
  count(source_file, name = "n_parsed")

marker_vs_parsed <- marker_counts %>%
  left_join(
    parsed_counts,
    by = c("filename" = "source_file")
  ) %>%
  mutate(
    n_parsed = ifelse(is.na(n_parsed), 0L, n_parsed),
    diff = n_markers - n_parsed
  ) %>%
  arrange(desc(abs(diff)), filename)

marker_vs_parsed

readr::write_csv(marker_vs_parsed, diff_diag_path)

marker_vs_parsed %>%
  summarise(
    total_markers = sum(n_markers),
    total_parsed = sum(n_parsed),
    total_diff = sum(diff),
    files_with_diff = sum(diff != 0)
  )
```

## 10. Parse diagnostics

```{r parse-diagnostics}
jso_parse_diagnostics <- jso_manual_chapter_docs %>%
  group_by(author, author_code, source_file, macro_genre) %>%
  summarise(
    n_chapters = n(),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words),
    .groups = "drop"
  ) %>%
  arrange(author, source_file)

jso_parse_diagnostics

readr::write_csv(jso_parse_diagnostics, parse_diag_path)
```

### Sanity check: tiny and large documents

```{r sanity-doc-lengths}
jso_manual_chapter_docs %>%
  arrange(n_words) %>%
  select(author, source_file, chapter, chapter_original, chapter_title, n_words) %>%
  print(n = 50)

jso_manual_chapter_docs %>%
  arrange(desc(n_words)) %>%
  select(author, source_file, chapter, chapter_original, chapter_title, n_words) %>%
  print(n = 30)

jso_manual_chapter_docs %>%
  summarise(
    n_chapter_docs = n(),
    n_authors = n_distinct(author),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )
```

## 11. Build book-level CSV

Each source file becomes one book-level / author-bundle document.

```{r book-level}
jso_manual_book_docs <- jso_manual_chapter_docs %>%
  arrange(author, source_file, chapter) %>%
  group_by(
    corpus_family,
    source_family,
    author,
    author_code,
    work,
    short_work,
    source_file,
    local_file,
    macro_genre
  ) %>%
  summarise(
    n_chapters = n(),
    min_chapter = min(chapter),
    max_chapter = max(chapter),
    text = paste(text, collapse = " "),
    chapter_map = paste(
      paste0(chapter, "=", chapter_original, ": ", chapter_title),
      collapse = " | "
    ),
    .groups = "drop"
  ) %>%
  mutate(
    text = normalize_doc_text(text),
    doc_unit = "book",
    n_words = str_count(text, "\\S+"),
    doc_id = paste0(
      "JSO__",
      make_safe_id(author_code),
      "__BOOK"
    ),
    doc_id_original = doc_id,
    book_digest = digest::digest(
      paste(author, work, text, sep = "||"),
      algo = "xxhash64"
    )
  ) %>%
  select(
    corpus_family,
    source_family,
    author,
    author_code,
    work,
    short_work,
    source_file,
    local_file,
    doc_id,
    doc_id_original,
    macro_genre,
    doc_unit,
    n_chapters,
    min_chapter,
    max_chapter,
    chapter_map,
    book_digest,
    n_words,
    text
  )

jso_manual_book_docs %>%
  select(author, source_file, n_chapters, n_words) %>%
  arrange(author, source_file) %>%
  print(n = 100)
```

## 12. Final sanity checks

```{r final-sanity}
cat("Manual files:", nrow(manual_manifest), "\n")
cat("Chapter docs:", nrow(jso_manual_chapter_docs), "\n")
cat("Book docs:", nrow(jso_manual_book_docs), "\n")

if(nrow(jso_manual_chapter_docs) == 0){
  stop("No JSO chapter docs were parsed.")
}

if(nrow(jso_manual_book_docs) != nrow(file_metadata)){
  warning("Book-doc count does not equal file count. This can happen if a file has zero parsed chapters.")
}

jso_manual_book_docs %>%
  summarise(
    n_books = n(),
    total_chapters = sum(n_chapters),
    total_words = sum(n_words),
    min_words = min(n_words),
    median_words = median(n_words),
    mean_words = mean(n_words),
    max_words = max(n_words)
  )
```

## 13. Write outputs

```{r write}
readr::write_csv(
  jso_manual_chapter_docs,
  chapter_out_path
)

readr::write_csv(
  jso_manual_book_docs,
  book_out_path
)

cat("Wrote JSO chapter CSV:\n")
cat(chapter_out_path, "\n\n")

cat("Wrote JSO book CSV:\n")
cat(book_out_path, "\n")
```

## 14. Integration

Use the chapter CSV for stylometry:

```r
jso_manual_chapter_docs <- readr::read_csv(
  file.path(STYLO_ROOT, "data", "prepared", "jso_manual_chapter_docs.csv"),
  show_col_types = FALSE
)
```
