---
title: "KJV English Stylometry: Isaiah and Control Chapters"
author: "Monte J. Shaffer"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup, message=FALSE, warning=FALSE}
# ============================================================
# KJV English Stylometry Notebook
# Goal:
#   Compare Isaiah sections using KJV English text, with controls
#   such as Samuel, Kings, Jeremiah, Ezekiel, Psalms, etc.
#
# Main methods:
#   - Chapter-level feature matrices
#   - Most frequent words
#   - Function-word style
#   - Character n-grams
#   - PCA visualization
#   - Hierarchical clustering
#   - Cosine similarity nearest neighbors
#   - Optional POS tagging with UDPipe
# ============================================================

required_packages <- c(
  "tidyverse", "tidytext", "stringr", "readr", "janitor",
  "widyr", "proxy", "FactoMineR", "factoextra", "stylo",
  "udpipe", "mclust"
)

install_if_missing <- function(pkgs) {
  missing <- pkgs[!pkgs %in% installed.packages()[, "Package"]]
  if (length(missing) > 0) install.packages(missing)
}

install_if_missing(required_packages)

library(tidyverse)
library(tidytext)
library(stringr)
library(readr)
library(janitor)
library(widyr)
library(proxy)
library(FactoMineR)
library(factoextra)
library(stylo)
library(udpipe)
library(mclust)
```


# 0. Data prep
```text
translation, book, chapter, verse, text
```

The translation would be KJV, WLC (Hebrew), SEPT (Septuagint).

Use this R Markdown block. It downloads `kjv.json` from `bibleapi-bibles-json`, filters Isaiah plus your controls, and writes `data/kjv.csv`. The repo includes KJV JSON and is intended for JSON Bible imports. ([GitHub][1])


```{r prepare-kjv-csv}
# Prepare KJV CSV for Isaiah stylometry + controls

library(dplyr)
library(readr)
library(stringr)
library(scriptuRs)

dir.create("data", showWarnings = FALSE)

target_books <- c(
  "Isaiah",
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings",
  "Psalms", "Job", "Proverbs"
)

kjv_selected <- kjv_bible() %>%
  filter(book_title %in% target_books) %>%
  transmute(
    translation = "KJV",
    book = book_title,
    chapter = as.integer(chapter_number),
    verse = as.integer(verse_number),
    text = str_squish(text)
  ) %>%
  arrange(match(book, target_books), chapter, verse)

write_csv(kjv_selected, "data/kjv.csv")

cat("Wrote: data/kjv.csv\n")
cat("Rows:", nrow(kjv_selected), "\n")
table(kjv_selected$book)
```

## Chapters
```{r}
library(dplyr)
library(stringr)
library(readr)

kjv_chapters <- kjv_selected %>%
  group_by(book, chapter) %>%
  summarize(
    text = str_c(text, collapse = " "),
    n_verses = n(),
    .groups = "drop"
  ) %>%
  mutate(
    doc_id = paste(book, chapter, sep="_")
  ) %>%
  select(doc_id, book, chapter, n_verses, text)

write_csv(kjv_chapters, "data/kjv_chapters.csv")

dim(kjv_chapters)

(kjv_chapters)
```

## Deutero?
```{r}
library(dplyr)
library(readr)

kjv_chapters <- kjv_chapters %>%
  mutate(
    section = case_when(
      book == "Isaiah" & chapter <= 35 ~ "Isaiah_1_35",
      book == "Isaiah" & chapter <= 39 ~ "Isaiah_36_39",
      book == "Isaiah" & chapter <= 55 ~ "Isaiah_40_55",
      book == "Isaiah" & chapter <= 66 ~ "Isaiah_56_66",
      TRUE ~ book
    )
  )

table(kjv_chapters$section)

write_csv(kjv_chapters, "data/kjv_chapters.csv")
```

## POS tagging
```{r}
# install.packages("quanteda")
library(quanteda)
library(dplyr)

corp <- corpus(
  kjv_chapters,
  text_field = "text"
)

docvars(corp, "section") <- kjv_chapters$section

toks <- tokens(
  corp,
  remove_punct = TRUE,
  remove_numbers = TRUE
)

chars <- tokens_ngrams(
  toks,
  n = 3,
  concatenator = ""
)

dfm3 <- dfm(chars)

dim(dfm3)


dfm3_trim <- dfm_trim(
  dfm3,
  min_docfreq = 5
)

dim(dfm3_trim)

topfeatures(dfm3_trim, 50)

dfm_word3 <- dfm3_trim



corp <- corpus(
  kjv_chapters,
  text_field="text"
)

chars <- tokens(
  corp,
  what = "character",
  remove_punct=TRUE,
  remove_numbers=TRUE
)

char3 <- tokens_ngrams(
  chars,
  n=3,
  concatenator=""
)

dfm_char3 <- dfm(char3)

dfm_char3 <- dfm_trim(
  dfm_char3,
  min_docfreq=5
)

topfeatures(dfm_char3,50)




chars4 <- tokens_ngrams(
    chars,
    n=4,
    concatenator=""
)

dfm_char4 <- dfm(chars4)

dfm_char4 <- dfm_trim(
    dfm_char4,
    min_docfreq=5
)

topfeatures(dfm_char4,20)





library(stopwords)

fw <- stopwords("en")

toks_fw <- tokens_select(
    toks,
    pattern=fw
)

dfm_fw <- dfm(toks_fw)

topfeatures(dfm_fw,50)




phrases <- c(
"thus saith the lord",
"the word of the lord",
"lord of hosts",
"and it came to pass",
"the house of",
"the land of",
"the children of"
)

library(stringr)

formula_counts <- kjv_chapters %>%
  mutate(

    thus_saith =
      str_count(
        str_to_lower(text),
        "thus saith the lord"
      ),

    word_of_lord =
      str_count(
        str_to_lower(text),
        "the word of the lord"
      ),

    lord_of_hosts =
      str_count(
        str_to_lower(text),
        "lord of hosts"
      )

  )









library(proxy)

m <- as.matrix(dfm_weight(dfm3_trim, "prop"))

cos_sim <- 1 - proxy::dist(
  m,
  method = "cosine"
)

cos_sim[1:10,1:10]





isaiah40 <- which(kjv_chapters$section=="Isaiah_40_55")

sort(
  cos_sim[isaiah40[1],],
  decreasing=TRUE
)[1:20]





library(FactoMineR)
library(factoextra)

X <- as.matrix(dfm_weight(dfm3_trim, "prop"))

pca <- PCA(
  X,
  graph=FALSE
)

fviz_pca_ind(
  pca,
  habillage = kjv_chapters$section,
  repel=TRUE
)











X <- as.matrix(
    dfm_weight(
        dfm_char3,
        scheme="prop"
    )
)

library(proxy)

D <- as.matrix(
    proxy::dist(
        X,
        method="cosine"
    )
)

rownames(D) <- kjv_chapters$doc_id
colnames(D) <- kjv_chapters$doc_id

sort(
    D["Isaiah_40",]
)[1:20]





nearest_neighbors <- function(doc){

    sort(
      D[doc,]
    )[1:10]

}

nearest_neighbors("Isaiah_1")
nearest_neighbors("Isaiah_35")
nearest_neighbors("Isaiah_40")
nearest_neighbors("Isaiah_55")
nearest_neighbors("Isaiah_56")
nearest_neighbors("Isaiah_66")


nearest_neighbors("Isaiah_35")
nearest_neighbors("Isaiah_36")
nearest_neighbors("Isaiah_39")
nearest_neighbors("Isaiah_40")
nearest_neighbors("Isaiah_55")
nearest_neighbors("Isaiah_56")


mean_distance <- function(group1, group2, D, metadata){

  idx1 <- which(metadata$section == group1)
  idx2 <- which(metadata$section == group2)

  mean(
    D[idx1, idx2]
  )

}



mean_distance(
  "Isaiah_1_35",
  "Isaiah_40_55",
  D,
  kjv_chapters
)

mean_distance(
  "Isaiah_40_55",
  "Isaiah_56_66",
  D,
  kjv_chapters
)

mean_distance(
  "Isaiah_40_55",
  "Jeremiah",
  D,
  kjv_chapters
)

mean_distance(
  "Isaiah_40_55",
  "Ezekiel",
  D,
  kjv_chapters
)

mean_distance(
  "Isaiah_1_35",
  "Jeremiah",
  D,
  kjv_chapters
)






```





### full working code?
---
title: "KJV Isaiah Stylometry"
output: html_notebook
---

```{r setup}
library(dplyr)
library(stringr)
library(readr)
library(quanteda)
library(proxy)
library(scriptuRs)

dir.create("data", showWarnings = FALSE)
```

```{r prepare-kjv}
target_books <- c(
  "Isaiah",
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings",
  "Psalms", "Job", "Proverbs"
)

kjv_selected <- kjv_bible() %>%
  filter(book_title %in% target_books) %>%
  transmute(
    translation = "KJV",
    book = book_title,
    chapter = as.integer(chapter_number),
    verse = as.integer(verse_number),
    text = str_squish(text)
  ) %>%
  arrange(match(book, target_books), chapter, verse)

write_csv(kjv_selected, "data/kjv.csv")

table(kjv_selected$book)
```

```{r make-chapters}
kjv_chapters <- kjv_selected %>%
  group_by(book, chapter) %>%
  summarize(
    text = str_c(text, collapse = " "),
    n_verses = n(),
    .groups = "drop"
  ) %>%
  mutate(
    doc_id = paste(book, chapter, sep = "_"),
    section = case_when(
      book == "Isaiah" & chapter <= 35 ~ "Isaiah_1_35",
      book == "Isaiah" & chapter <= 39 ~ "Isaiah_36_39",
      book == "Isaiah" & chapter <= 55 ~ "Isaiah_40_55",
      book == "Isaiah" & chapter <= 66 ~ "Isaiah_56_66",
      TRUE ~ book
    )
  ) %>%
  select(doc_id, book, chapter, section, n_verses, text)

write_csv(kjv_chapters, "data/kjv_chapters.csv")

table(kjv_chapters$section)
```

```{r corpus}
corp <- corpus(kjv_chapters, text_field = "text")

docvars(corp, "doc_id") <- kjv_chapters$doc_id
docvars(corp, "book") <- kjv_chapters$book
docvars(corp, "chapter") <- kjv_chapters$chapter
docvars(corp, "section") <- kjv_chapters$section
```

```{r character-trigrams}
chars <- tokens(
  corp,
  what = "character",
  remove_punct = TRUE,
  remove_numbers = TRUE
)

char3 <- tokens_ngrams(
  chars,
  n = 3,
  concatenator = ""
)

dfm_char3 <- dfm(char3) %>%
  dfm_trim(min_docfreq = 5)

dim(dfm_char3)
topfeatures(dfm_char3, 50)
```

```{r character-fourgrams}
char4 <- tokens_ngrams(
  chars,
  n = 4,
  concatenator = ""
)

dfm_char4 <- dfm(char4) %>%
  dfm_trim(min_docfreq = 5)

dim(dfm_char4)
topfeatures(dfm_char4, 50)
```

```{r word-trigrams}
toks_words <- tokens(
  corp,
  remove_punct = TRUE,
  remove_numbers = TRUE
)

word3 <- tokens_ngrams(
  toks_words,
  n = 3,
  concatenator = "_"
)

dfm_word3 <- dfm(word3) %>%
  dfm_trim(min_docfreq = 5)

dim(dfm_word3)
topfeatures(dfm_word3, 50)
```

```{r distance-functions}
make_distance <- function(dfm_object) {
  
  X <- as.matrix(
    dfm_weight(
      dfm_object,
      scheme = "prop"
    )
  )
  
  D <- as.matrix(
    proxy::dist(
      X,
      method = "cosine"
    )
  )
  
  rownames(D) <- kjv_chapters$doc_id
  colnames(D) <- kjv_chapters$doc_id
  
  return(D)
}

nearest_neighbors <- function(doc, D, n = 20) {
  sort(D[doc, ])[1:n]
}

mean_distance <- function(group1, group2, D, metadata) {
  
  idx1 <- which(metadata$section == group1)
  idx2 <- which(metadata$section == group2)
  
  mean(D[idx1, idx2])
}

section_distance_matrix <- function(D, metadata) {
  
  groups <- unique(metadata$section)
  
  M <- matrix(
    NA,
    nrow = length(groups),
    ncol = length(groups)
  )
  
  rownames(M) <- groups
  colnames(M) <- groups
  
  for (i in seq_along(groups)) {
    for (j in seq_along(groups)) {
      M[i, j] <- mean_distance(
        groups[i],
        groups[j],
        D,
        metadata
      )
    }
  }
  
  round(M, 4)
}
```

```{r distances-char3}
D_char3 <- make_distance(dfm_char3)

nearest_neighbors("Isaiah_35", D_char3, 10)
nearest_neighbors("Isaiah_36", D_char3, 10)
nearest_neighbors("Isaiah_39", D_char3, 10)
nearest_neighbors("Isaiah_40", D_char3, 20)
nearest_neighbors("Isaiah_55", D_char3, 10)
nearest_neighbors("Isaiah_56", D_char3, 10)

section_distance_matrix(D_char3, kjv_chapters)
```

```{r key-comparisons-char3}
char3_comparisons <- tibble(
  comparison = c(
    "Isaiah_1_35 vs Isaiah_40_55",
    "Isaiah_40_55 vs Isaiah_56_66",
    "Isaiah_40_55 vs Jeremiah",
    "Isaiah_40_55 vs Ezekiel",
    "Isaiah_1_35 vs Jeremiah"
  ),
  distance = c(
    mean_distance("Isaiah_1_35", "Isaiah_40_55", D_char3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Isaiah_56_66", D_char3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Jeremiah", D_char3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Ezekiel", D_char3, kjv_chapters),
    mean_distance("Isaiah_1_35", "Jeremiah", D_char3, kjv_chapters)
  )
)

char3_comparisons %>%
  arrange(distance)
```

```{r compare-feature-sets}
D_char4 <- make_distance(dfm_char4)
D_word3 <- make_distance(dfm_word3)

feature_comparison <- bind_rows(
  tibble(
    feature_set = "Character 3-grams",
    comparison = char3_comparisons$comparison,
    distance = char3_comparisons$distance
  ),
  tibble(
    feature_set = "Character 4-grams",
    comparison = char3_comparisons$comparison,
    distance = c(
      mean_distance("Isaiah_1_35", "Isaiah_40_55", D_char4, kjv_chapters),
      mean_distance("Isaiah_40_55", "Isaiah_56_66", D_char4, kjv_chapters),
      mean_distance("Isaiah_40_55", "Jeremiah", D_char4, kjv_chapters),
      mean_distance("Isaiah_40_55", "Ezekiel", D_char4, kjv_chapters),
      mean_distance("Isaiah_1_35", "Jeremiah", D_char4, kjv_chapters)
    )
  ),
  tibble(
    feature_set = "Word 3-grams",
    comparison = char3_comparisons$comparison,
    distance = c(
      mean_distance("Isaiah_1_35", "Isaiah_40_55", D_word3, kjv_chapters),
      mean_distance("Isaiah_40_55", "Isaiah_56_66", D_word3, kjv_chapters),
      mean_distance("Isaiah_40_55", "Jeremiah", D_word3, kjv_chapters),
      mean_distance("Isaiah_40_55", "Ezekiel", D_word3, kjv_chapters),
      mean_distance("Isaiah_1_35", "Jeremiah", D_word3, kjv_chapters)
    )
  )
)

feature_comparison %>%
  arrange(feature_set, distance)
```

```{r pca-char3}
X_char3 <- as.matrix(
  dfm_weight(dfm_char3, scheme = "prop")
)

pca_char3 <- prcomp(
  X_char3,
  center = TRUE,
  scale. = FALSE
)

pca_df <- as_tibble(pca_char3$x[, 1:2]) %>%
  mutate(
    doc_id = kjv_chapters$doc_id,
    book = kjv_chapters$book,
    chapter = kjv_chapters$chapter,
    section = kjv_chapters$section
  )

head(pca_df)
```

```{r plot-pca-char3}
library(ggplot2)

#install.packages("ggrepel")
library(ggrepel)

var_explained <- pca_char3$sdev^2 /
                 sum(pca_char3$sdev^2)

pc1_pct <- round(
  100 * var_explained[1],
  1
)

pc2_pct <- round(
  100 * var_explained[2],
  1
)



ggplot(
  pca_df,
  aes(
    x = PC1,
    y = PC2,
    color = section
  )
) +
  geom_point(
    size = 2,
    alpha = 0.7
  ) +

  geom_text_repel(
    data =
      subset(
        pca_df,
        grepl("^Isaiah", doc_id)
      ),

    aes(
      label = doc_id
    ),

    size = 3,
    max.overlaps = 100
  ) +

  theme_minimal() +

  labs(
    title = "KJV Stylometry: Character 3-gram PCA",

    subtitle =
      "Isaiah chapters labeled",

    x =
      paste0(
        "PC1 (",
        pc1_pct,
        "%)"
      ),

    y =
      paste0(
        "PC2 (",
        pc2_pct,
        "%)"
      )
  )
```


```{r save-results}
write_csv(char3_comparisons, "data/char3_comparisons.csv")
write_csv(feature_comparison, "data/feature_comparison.csv")
write_csv(pca_df, "data/pca_char3.csv")
```


```{r isaiah only}
isaiah_pca <- pca_df %>%
  filter(grepl("^Isaiah", doc_id))

ggplot(
  isaiah_pca,
  aes(
    PC1,
    PC2,
    color=section
  )
)+

geom_point(size=3)+

geom_path(
  aes(group=1),
  color="gray70",
  alpha=.5
)+

geom_text(
  aes(label=chapter),
  size=3,
  check_overlap=TRUE,
  vjust=-.5
)+

theme_minimal()+

labs(
  title="Isaiah Character 3-gram PCA",
  subtitle="Chapters connected sequentially",
  x=paste0(
    "PC1 (",
    pc1_pct,
    "%)"
  ),
  y=paste0(
    "PC2 (",
    pc2_pct,
    "%)"
  )
)


















dir.create("corpus", showWarnings = FALSE)

for(i in 1:nrow(kjv_chapters)){

  writeLines(

    kjv_chapters$text[i],

    paste0(
      "corpus/",
      kjv_chapters$doc_id[i],
      ".txt"
    )

  )

}



library(stylo)
stylo(

  corpus.dir="corpus",

  analysis.type="CA",

  mfw.min=100,

  mfw.max=1000,

  distance.measure="delta",

  gui=FALSE

)


```






































```{r pos-tagging}
library(udpipe)
library(dplyr)
library(stringr)
library(quanteda)


dir.create(
  "cache",
  showWarnings=FALSE
)



# Download once
model_file <- udpipe_download_model(language = "english-ewt")
ud_model <- udpipe_load_model(model_file$file_model)

# Tag chapter texts  ## this takes awhile...
if(file.exists("cache/pos_df.rds")){

  pos_df <-

    readRDS(
      "cache/pos_df.rds"
    )

} else {

  pos_annot <-

    udpipe_annotate(

      ud_model,

      x=kjv_chapters$text,

      doc_id=
        kjv_chapters$doc_id

    )

  pos_df <-

    as.data.frame(
      pos_annot
    )

  saveRDS(

    pos_df,

    "cache/pos_df.rds"

  )

}
```



```{r pos-trigrams}
pos_docs <- pos_df %>%
  filter(!is.na(upos)) %>%
  group_by(doc_id) %>%
  summarize(
    pos_text = str_c(upos, collapse = " "),
    .groups = "drop"
  )

pos_corp <- corpus(pos_docs, text_field = "pos_text")

toks_pos <- tokens(
  pos_corp,
  what = "word"
)

pos3 <- tokens_ngrams(
  toks_pos,
  n = 3,
  concatenator = "_"
)

dfm_pos3 <- dfm(pos3) %>%
  dfm_trim(min_docfreq = 5)

dim(dfm_pos3)
topfeatures(dfm_pos3, 50)







D_pos3 <- make_distance(dfm_pos3)

nearest_neighbors("Isaiah_40", D_pos3, 20)

pos3_comparisons <- tibble(
  feature_set = "POS 3-grams",
  comparison = char3_comparisons$comparison,
  distance = c(
    mean_distance("Isaiah_1_35", "Isaiah_40_55", D_pos3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Isaiah_56_66", D_pos3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Jeremiah", D_pos3, kjv_chapters),
    mean_distance("Isaiah_40_55", "Ezekiel", D_pos3, kjv_chapters),
    mean_distance("Isaiah_1_35", "Jeremiah", D_pos3, kjv_chapters)
  )
)

pos3_comparisons %>%
  arrange(distance)















```

































# 1. Data input

This notebook assumes you have a KJV chapter-level file with at least these columns:

```text
book, chapter, verse, text
```

A verse-level CSV is ideal. Each row should be one verse.

Example:

```text
book,chapter,verse,text
Isaiah,1,1,"The vision of Isaiah the son of Amoz..."
Isaiah,1,2,"Hear, O heavens, and give ear, O earth..."
```

```{r load-data}
# Change this path to your local KJV CSV file.
# Example: "data/kjv.csv"
kjv_path <- "data/kjv.csv"

kjv <- read_csv(kjv_path, show_col_types = FALSE) %>%
  clean_names() %>%
  mutate(
    book = str_squish(book),
    chapter = as.integer(chapter),
    verse = as.integer(verse),
    text = str_squish(text)
  )

glimpse(kjv)
```

# 2. Select Isaiah and control books

Recommended control groups:

- **Prophetic controls**: Jeremiah, Ezekiel, Hosea, Amos
- **Narrative controls**: 1 Samuel, 2 Samuel, 1 Kings, 2 Kings
- **Poetic controls**: Psalms, Job, Proverbs

```{r select-books}
control_books <- c(
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings",
  "Psalms", "Job", "Proverbs"
)

books_to_keep <- c("Isaiah", control_books)

corpus_verses <- kjv %>%
  filter(book %in% books_to_keep)

corpus_chapters <- corpus_verses %>%
  group_by(book, chapter) %>%
  summarise(text = paste(text, collapse = " "), .groups = "drop") %>%
  mutate(
    doc_id = paste(book, chapter, sep = "_"),
    isaiah_section = case_when(
      book == "Isaiah" & chapter >= 1  & chapter <= 35 ~ "Isaiah_1_35",
      book == "Isaiah" & chapter >= 36 & chapter <= 39 ~ "Isaiah_36_39_narrative",
      book == "Isaiah" & chapter >= 40 & chapter <= 55 ~ "Isaiah_40_55",
      book == "Isaiah" & chapter >= 56 & chapter <= 66 ~ "Isaiah_56_66",
      TRUE ~ "Control"
    ),
    genre_group = case_when(
      book %in% c("1 Samuel", "2 Samuel", "1 Kings", "2 Kings") ~ "Narrative",
      book %in% c("Psalms", "Job", "Proverbs") ~ "Poetic_Wisdom",
      book %in% c("Jeremiah", "Ezekiel", "Hosea", "Amos", "Isaiah") ~ "Prophetic",
      TRUE ~ "Other"
    )
  )

corpus_chapters %>% count(book, genre_group)
```

# 3. Basic tokenization

```{r tokenize}
tokens <- corpus_chapters %>%
  select(doc_id, book, chapter, isaiah_section, genre_group, text) %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z]"))

tokens %>% count(word, sort = TRUE) %>% head(30)
```

# 4. Chapter-length diagnostics

Short chapters can behave erratically in stylometry.

```{r chapter-lengths}
chapter_lengths <- tokens %>%
  count(doc_id, book, chapter, isaiah_section, genre_group, name = "word_count")

chapter_lengths %>%
  ggplot(aes(x = word_count)) +
  geom_histogram(bins = 30) +
  labs(title = "Chapter word counts", x = "Words per chapter", y = "Number of chapters")

chapter_lengths %>% arrange(word_count) %>% head(20)
```

Optional filter:

```{r filter-short-chapters}
min_words <- 150
valid_docs <- chapter_lengths %>%
  filter(word_count >= min_words) %>%
  pull(doc_id)

tokens_valid <- tokens %>% filter(doc_id %in% valid_docs)
corpus_valid <- corpus_chapters %>% filter(doc_id %in% valid_docs)
```

# 5. Most frequent word matrix

This is the classic stylometric starting point.

```{r mfw-matrix}
num_mfw <- 300

mfw <- tokens_valid %>%
  count(word, sort = TRUE) %>%
  slice_head(n = num_mfw) %>%
  pull(word)

word_matrix <- tokens_valid %>%
  filter(word %in% mfw) %>%
  count(doc_id, word) %>%
  group_by(doc_id) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(doc_id, word, freq) %>%
  pivot_wider(names_from = word, values_from = freq, values_fill = 0)

word_metadata <- corpus_valid %>% select(doc_id, book, chapter, isaiah_section, genre_group)

word_features <- word_matrix %>%
  left_join(word_metadata, by = "doc_id") %>%
  relocate(doc_id, book, chapter, isaiah_section, genre_group)

word_features %>% glimpse()
```

# 6. PCA on most frequent words

```{r pca-mfw}
feature_cols <- setdiff(names(word_features), c("doc_id", "book", "chapter", "isaiah_section", "genre_group"))

pca_input <- word_features %>%
  select(all_of(feature_cols)) %>%
  as.data.frame()

rownames(pca_input) <- word_features$doc_id

pca_result <- prcomp(pca_input, center = TRUE, scale. = TRUE)

pca_scores <- as_tibble(pca_result$x, rownames = "doc_id") %>%
  left_join(word_metadata, by = "doc_id")

pca_scores %>%
  ggplot(aes(x = PC1, y = PC2, label = doc_id, color = isaiah_section, shape = genre_group)) +
  geom_point(size = 2.5) +
  labs(title = "PCA: KJV chapter style using most frequent words")
```

# 7. PCA: Isaiah only

```{r pca-isaiah-only}
pca_scores %>%
  filter(book == "Isaiah") %>%
  ggplot(aes(x = PC1, y = PC2, label = chapter, color = isaiah_section)) +
  geom_point(size = 3) +
  geom_text(nudge_y = 0.15, size = 3) +
  labs(title = "PCA: Isaiah chapters only")
```

# 8. Hierarchical clustering

```{r hierarchical-clustering}
dist_matrix <- dist(pca_input, method = "euclidean")
hc <- hclust(dist_matrix, method = "ward.D2")

plot(hc, labels = rownames(pca_input), main = "Hierarchical clustering: KJV chapters")
```

# 9. Cosine similarity nearest neighbors

This is useful for your proposed “top-k nearest neighbors” approach.

```{r cosine-similarity}
cosine_dist <- proxy::dist(pca_input, method = "cosine")
cosine_sim <- 1 - as.matrix(cosine_dist)

top_k_neighbors <- function(doc, sim_matrix, k = 5) {
  sims <- sort(sim_matrix[doc, ], decreasing = TRUE)
  sims <- sims[names(sims) != doc]
  tibble(
    doc_id = doc,
    neighbor = names(sims)[1:k],
    similarity = as.numeric(sims[1:k])
  )
}

isaiah_docs <- word_metadata %>%
  filter(book == "Isaiah") %>%
  pull(doc_id)

neighbors <- map_dfr(isaiah_docs, top_k_neighbors, sim_matrix = cosine_sim, k = 5) %>%
  left_join(word_metadata, by = c("neighbor" = "doc_id")) %>%
  rename(
    neighbor_book = book,
    neighbor_chapter = chapter,
    neighbor_section = isaiah_section,
    neighbor_genre = genre_group
  )

neighbors %>%
  filter(str_detect(doc_id, "Isaiah")) %>%
  arrange(doc_id, desc(similarity))
```

# 10. Function-word analysis

For English KJV, function words can reduce content/topic effects.

```{r function-words}
function_words <- c(
  "the", "and", "of", "to", "that", "in", "he", "shall", "for", "unto",
  "i", "his", "a", "they", "be", "is", "not", "with", "him", "them",
  "it", "my", "me", "thy", "thou", "ye", "you", "your", "we", "our",
  "as", "but", "or", "if", "from", "by", "upon", "therefore", "then", "when",
  "which", "who", "whom", "whose", "this", "these", "those", "all", "no", "neither"
)

function_matrix <- tokens_valid %>%
  filter(word %in% function_words) %>%
  count(doc_id, word) %>%
  group_by(doc_id) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(doc_id, word, freq) %>%
  pivot_wider(names_from = word, values_from = freq, values_fill = 0)

function_features <- function_matrix %>%
  left_join(word_metadata, by = "doc_id") %>%
  relocate(doc_id, book, chapter, isaiah_section, genre_group)

function_input <- function_features %>%
  select(-doc_id, -book, -chapter, -isaiah_section, -genre_group) %>%
  as.data.frame()
rownames(function_input) <- function_features$doc_id

function_pca <- prcomp(function_input, center = TRUE, scale. = TRUE)

function_scores <- as_tibble(function_pca$x, rownames = "doc_id") %>%
  left_join(word_metadata, by = "doc_id")

function_scores %>%
  ggplot(aes(x = PC1, y = PC2, color = isaiah_section, shape = genre_group)) +
  geom_point(size = 2.5) +
  labs(title = "PCA: KJV function-word style")
```

# 11. Character n-grams

Character n-grams often perform well in authorship work.

```{r char-ngrams}
make_char_ngrams <- function(text, n = 4) {
  text <- str_to_lower(text)
  text <- str_replace_all(text, "[^a-z ]", " ")
  text <- str_squish(text)
  chars <- strsplit(text, "")[[1]]
  if (length(chars) < n) return(character(0))
  map_chr(seq_len(length(chars) - n + 1), ~ paste0(chars[.x:(.x+n-1)], collapse = ""))
}

char_n <- 4
num_char_features <- 500

char_tokens <- corpus_valid %>%
  select(doc_id, book, chapter, isaiah_section, genre_group, text) %>%
  mutate(char_ngram = map(text, make_char_ngrams, n = char_n)) %>%
  select(-text) %>%
  unnest(char_ngram) %>%
  filter(str_detect(char_ngram, "[a-z]"))

top_char <- char_tokens %>%
  count(char_ngram, sort = TRUE) %>%
  slice_head(n = num_char_features) %>%
  pull(char_ngram)

char_matrix <- char_tokens %>%
  filter(char_ngram %in% top_char) %>%
  count(doc_id, char_ngram) %>%
  group_by(doc_id) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(doc_id, char_ngram, freq) %>%
  pivot_wider(names_from = char_ngram, values_from = freq, values_fill = 0)

char_features <- char_matrix %>%
  left_join(word_metadata, by = "doc_id") %>%
  relocate(doc_id, book, chapter, isaiah_section, genre_group)

char_input <- char_features %>%
  select(-doc_id, -book, -chapter, -isaiah_section, -genre_group) %>%
  as.data.frame()
rownames(char_input) <- char_features$doc_id

char_pca <- prcomp(char_input, center = TRUE, scale. = TRUE)

char_scores <- as_tibble(char_pca$x, rownames = "doc_id") %>%
  left_join(word_metadata, by = "doc_id")

char_scores %>%
  ggplot(aes(x = PC1, y = PC2, color = isaiah_section, shape = genre_group)) +
  geom_point(size = 2.5) +
  labs(title = "PCA: KJV character 4-gram style")
```

# 12. mclust mixture modeling

This estimates the number of clusters from the feature space. Treat this as exploratory.

```{r mclust-model}
# Use PCA scores rather than hundreds of raw features.
# Usually PC1-PC10 is enough for exploratory clustering.

mclust_input <- pca_scores %>%
  select(PC1:PC10) %>%
  as.data.frame()
rownames(mclust_input) <- pca_scores$doc_id

mc <- Mclust(mclust_input)
summary(mc)
plot(mc, what = "classification")

mclust_classes <- tibble(
  doc_id = rownames(mclust_input),
  mclust_class = mc$classification
) %>%
  left_join(word_metadata, by = "doc_id")

mclust_classes %>%
  count(mclust_class, book, isaiah_section, genre_group) %>%
  arrange(mclust_class, book)
```

# 13. Optional POS tagging with UDPipe

For KJV English, this is feasible. It may take time on a full corpus.

```{r udpipe-pos, eval=FALSE}
# Download once. Then reuse the downloaded model path.
english_model_info <- udpipe_download_model(language = "english-ewt")
english_model <- udpipe_load_model(english_model_info$file_model)

pos_annotations <- udpipe_annotate(
  english_model,
  x = corpus_valid$text,
  doc_id = corpus_valid$doc_id
) %>%
  as.data.frame() %>%
  as_tibble()

pos_annotations %>% select(doc_id, token, lemma, upos, xpos, feats) %>% head(20)
```

# 14. POS n-gram stylometry

```{r pos-ngrams, eval=FALSE}
pos_ngrams <- pos_annotations %>%
  filter(!is.na(upos)) %>%
  group_by(doc_id) %>%
  summarise(pos_sequence = paste(upos, collapse = " "), .groups = "drop") %>%
  unnest_tokens(pos_bigram, pos_sequence, token = "ngrams", n = 2)

num_pos_features <- 150

top_pos_bigrams <- pos_ngrams %>%
  count(pos_bigram, sort = TRUE) %>%
  slice_head(n = num_pos_features) %>%
  pull(pos_bigram)

pos_matrix <- pos_ngrams %>%
  filter(pos_bigram %in% top_pos_bigrams) %>%
  count(doc_id, pos_bigram) %>%
  group_by(doc_id) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup() %>%
  select(doc_id, pos_bigram, freq) %>%
  pivot_wider(names_from = pos_bigram, values_from = freq, values_fill = 0)

pos_features <- pos_matrix %>%
  left_join(word_metadata, by = "doc_id") %>%
  relocate(doc_id, book, chapter, isaiah_section, genre_group)

pos_input <- pos_features %>%
  select(-doc_id, -book, -chapter, -isaiah_section, -genre_group) %>%
  as.data.frame()
rownames(pos_input) <- pos_features$doc_id

pos_pca <- prcomp(pos_input, center = TRUE, scale. = TRUE)

pos_scores <- as_tibble(pos_pca$x, rownames = "doc_id") %>%
  left_join(word_metadata, by = "doc_id")

pos_scores %>%
  ggplot(aes(x = PC1, y = PC2, color = isaiah_section, shape = genre_group)) +
  geom_point(size = 2.5) +
  labs(title = "PCA: KJV POS bigram style")
```

# 15. Ensemble nearest-neighbor voting

This combines multiple feature sets by asking: for each Isaiah chapter, what documents repeatedly appear among its nearest neighbors?

```{r ensemble-neighbors}
nearest_neighbors_from_matrix <- function(feature_matrix, metadata, feature_name, k = 5) {
  cosine_dist <- proxy::dist(feature_matrix, method = "cosine")
  cosine_sim <- 1 - as.matrix(cosine_dist)
  docs <- rownames(feature_matrix)

  map_dfr(docs, function(doc) {
    sims <- sort(cosine_sim[doc, ], decreasing = TRUE)
    sims <- sims[names(sims) != doc]
    tibble(
      doc_id = doc,
      neighbor = names(sims)[1:k],
      similarity = as.numeric(sims[1:k]),
      feature_set = feature_name
    )
  }) %>%
    left_join(metadata, by = c("neighbor" = "doc_id")) %>%
    rename(
      neighbor_book = book,
      neighbor_chapter = chapter,
      neighbor_section = isaiah_section,
      neighbor_genre = genre_group
    )
}

neighbors_mfw <- nearest_neighbors_from_matrix(pca_input, word_metadata, "MFW", k = 5)
neighbors_function <- nearest_neighbors_from_matrix(function_input, word_metadata, "Function_Words", k = 5)
neighbors_char <- nearest_neighbors_from_matrix(char_input, word_metadata, "Character_4grams", k = 5)

ensemble_neighbors <- bind_rows(neighbors_mfw, neighbors_function, neighbors_char)

isaiah_ensemble_summary <- ensemble_neighbors %>%
  filter(doc_id %in% isaiah_docs) %>%
  count(doc_id, neighbor_book, neighbor_section, neighbor_genre, sort = TRUE)

isaiah_ensemble_summary %>%
  arrange(doc_id, desc(n))
```

# 16. Suggested interpretation logic

Use this notebook to look for patterns such as:

1. Do Isaiah 40–55 chapters cluster together across multiple feature sets?
2. Do Isaiah 36–39 cluster with Kings/Samuel because they are narrative?
3. Do Isaiah 40–55 cluster with other prophetic controls such as Jeremiah/Ezekiel?
4. Do content-free features tell the same story as content-heavy features?
5. Does the result persist when chapter size thresholds change?
6. Does the result persist when using windows instead of chapters?

# 17. Optional next step: rolling windows instead of chapters

Chapter divisions are not always ideal. A stronger method is fixed-length windows.

```{r rolling-window-template, eval=FALSE}
# Template only.
# Build documents of approximately 500-word windows inside each book.

window_size <- 500

rolling_windows <- tokens %>%
  group_by(book) %>%
  mutate(token_index = row_number()) %>%
  mutate(window_id = ceiling(token_index / window_size)) %>%
  group_by(book, window_id) %>%
  summarise(text = paste(word, collapse = " "), .groups = "drop") %>%
  mutate(doc_id = paste(book, "window", window_id, sep = "_"))
```

