---
title: "Generic Stylometry Ensemble Feature Pipeline"
author: "Monte J. Shaffer"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(tibble)

# Optional packages used by specific feature classes.
# Install if needed:
# install.packages(c("SnowballC", "udpipe", "quanteda", "quanteda.textstats"))

has_snowball <- requireNamespace("SnowballC", quietly = TRUE)
has_udpipe <- requireNamespace("udpipe", quietly = TRUE)
has_quanteda <- requireNamespace("quanteda", quietly = TRUE)
has_quanteda_textstats <- requireNamespace("quanteda.textstats", quietly = TRUE)
```

# 1. Purpose

This notebook builds a generic stylometry ensemble pipeline. It assumes the source data can be prepared with the following fields:

```text
corpus_id, translation, doc_id, book, chapter, section, macro_genre, n_verses, text
```

The main comparison unit is `section`. For Isaiah, sections may be `Isaiah_1_35`, `Isaiah_36_39`, `Isaiah_40_55`, and `Isaiah_56_66`. For other books, the section can simply be the book name or any custom grouping.

# 2. Data Preparation

## 2.1 Target Books

```{r target-books}
target_books <- c(
  "Isaiah",
  "Jeremiah", "Ezekiel", "Hosea", "Amos",
  "1 Samuel", "2 Samuel", "1 Kings", "2 Kings",
  "Psalms", "Job", "Proverbs"
)
```

## 2.2 Section Assignment

This function is Isaiah-aware but not Isaiah-bound. All non-Isaiah books default to their book name as the section.

```{r assign-sections}
assign_sections_kjv <- function(data){

  data %>%
    mutate(
      section = case_when(
        book == "Isaiah" & chapter <= 35 ~ "Isaiah_1_35",
        book == "Isaiah" & chapter <= 39 ~ "Isaiah_36_39",
        book == "Isaiah" & chapter <= 55 ~ "Isaiah_40_55",
        book == "Isaiah" & chapter <= 66 ~ "Isaiah_56_66",
        TRUE ~ book
      ),

      macro_genre = case_when(
        book %in% c("1 Samuel", "2 Samuel", "1 Kings", "2 Kings") ~ "Narrative",
        book %in% c("Psalms", "Job", "Proverbs") ~ "Poetic/Wisdom",
        TRUE ~ "Prophetic"
      )
    )
}
```

## 2.3 Build Section Documents

```{r build-section-docs}
build_section_docs <- function(data, target_books = NULL){

  if(!is.null(target_books)){
    data <- data %>% filter(book %in% target_books)
  }

  grouping_vars <- c("corpus_id", "translation", "section", "macro_genre")
  missing_vars <- setdiff(grouping_vars, names(data))

  if(length(missing_vars) > 0){
    stop("Missing required columns: ", paste(missing_vars, collapse = ", "))
  }

  data %>%
    group_by(corpus_id, translation, section, macro_genre) %>%
    summarize(
      text = str_c(text, collapse = " "),
      n_verses = if("n_verses" %in% names(data)) sum(n_verses, na.rm = TRUE) else n(),
      n_chapters = n_distinct(chapter),
      books = str_c(unique(book), collapse = "; "),
      .groups = "drop"
    ) %>%
    mutate(
      doc_id = paste(corpus_id, translation, section, sep = "__")
    ) %>%
    select(corpus_id, translation, doc_id, section, macro_genre, books, n_chapters, n_verses, text)
}
```

## 2.4 Example Setup

If your verse-level data frame is named `kjv_selected`, run:

```{r example-data-prep, eval=FALSE}
section_docs <- kjv_selected %>%
  assign_sections_kjv() %>%
  build_section_docs(target_books = target_books)

docs <- section_docs$text
names(docs) <- section_docs$section
```

For testing this notebook without external data, use the small demo below.

```{r demo-docs}
demo_docs <- c(
  isa1  = "The vision of Isaiah the son of Amoz.",
  isa2  = "And it shall come to pass in the last days.",
  isa40 = "Comfort ye, comfort ye my people."
)

# Use demo docs unless you have already created docs.
if(!exists("docs")){
  docs <- demo_docs
}
```

# 3. Shared Utility Functions

## 3.1 Tokenizer

```{r tokenizer}
tokenize_text <- function(text, stem = FALSE, lowercase = TRUE, remove_punct = TRUE){

  if(lowercase){
    text <- tolower(text)
  }

  if(remove_punct){
    text <- gsub("[^a-z ]", " ", text)
  }

  text <- gsub("\\s+", " ", text)
  text <- trimws(text)

  if(nchar(text) == 0){
    return(character(0))
  }

  words <- unlist(strsplit(text, " "))
  words <- words[words != ""]

  if(stem){
    if(!has_snowball){
      stop("Package 'SnowballC' is required for stemming. Install it with install.packages('SnowballC').")
    }
    words <- SnowballC::wordStem(words, language = "en")
  }

  words
}
```

## 3.2 TF-IDF Matrix

```{r tfidf-matrix}
tfidf_matrix <- function(features_by_doc){

  vocab <- sort(unique(unlist(features_by_doc)))

  if(length(vocab) == 0){
    stop("No features found. Check input text or feature extractor.")
  }

  mat <- matrix(
    0,
    nrow = length(features_by_doc),
    ncol = length(vocab),
    dimnames = list(names(features_by_doc), vocab)
  )

  for(i in seq_along(features_by_doc)){
    tab <- table(features_by_doc[[i]])
    if(length(tab) > 0){
      mat[i, names(tab)] <- as.numeric(tab)
    }
  }

  rs <- rowSums(mat)
  rs[rs == 0] <- 1

  tf <- mat / rs

  df <- colSums(tf > 0)
  N <- nrow(tf)
  idf <- log((N + 1) / (df + 1)) + 1

  sweep(tf, 2, idf, "*")
}
```

## 3.3 Cosine Similarity

```{r cosine-sim}
cosine_sim <- function(mat){

  norm <- sqrt(rowSums(mat^2))
  norm[norm == 0] <- 1

  mat2 <- mat / norm
  mat2 %*% t(mat2)
}
```

## 3.4 Top Features

```{r top-features}
top_features <- function(mat, doc, k = 20){

  if(!doc %in% rownames(mat)){
    stop(
      paste0(
        "Document '", doc, "' not found. Available docs: ",
        paste(rownames(mat), collapse = ", ")
      )
    )
  }

  sort(mat[doc, ], decreasing = TRUE)[1:min(k, ncol(mat))]
}
```

## 3.5 Matrix Averaging

```{r matrix-average}
average_matrices <- function(mats){
  Reduce("+", mats) / length(mats)
}
```

# 4. Feature Class: Character N-Grams

```{r character-ngrams}
char_ngrams <- function(text, n){

  text <- tolower(text)
  text <- gsub("[^a-z ]", " ", text)
  text <- gsub("\\s+", " ", text)
  text <- trimws(text)

  chars <- strsplit(text, "")[[1]]

  if(length(chars) < n){
    return(character(0))
  }

  sapply(
    seq_len(length(chars) - n + 1),
    function(i) paste(chars[i:(i + n - 1)], collapse = "")
  )
}

char_tfidf <- function(docs, n){
  features_by_doc <- lapply(docs, char_ngrams, n = n)
  tfidf_matrix(features_by_doc)
}
```

# 5. Feature Class: Word and Stem N-Grams

```{r word-stem-ngrams}
word_ngrams <- function(text, n = 1, stem = FALSE){

  words <- tokenize_text(text, stem = stem)

  if(length(words) < n){
    return(character(0))
  }

  sapply(
    seq_len(length(words) - n + 1),
    function(i) paste(words[i:(i + n - 1)], collapse = "_")
  )
}

ngram_tfidf <- function(docs, n = 1, stem = FALSE){
  features_by_doc <- lapply(docs, word_ngrams, n = n, stem = stem)
  tfidf_matrix(features_by_doc)
}
```

# 6. Feature Class: Function Words

## 6.1 Function Word Sets

```{r function-word-sets}
function_word_sets <- list(

  basic = c(
    "the","and","of","to","in",
    "that","which","for","with","by",
    "on","at","from","as","if",
    "or","but","not","is","are",
    "was","were","be","been","being"
  ),

  kjv = c(
    "the","and","of","to","in",
    "that","which","shall","unto",
    "ye","thou","thee","thy","thine",
    "hath","hast","had",
    "doth","didst",
    "saith",
    "yea","nay","verily",
    "for","with","by","from"
  ),

  pronouns = c(
    "i","me","my","mine",
    "thou","thee","thy","thine",
    "he","him","his",
    "she","her","hers",
    "we","us","our","ours",
    "ye","you","your","yours",
    "they","them","their","theirs"
  ),

  conjunctions = c(
    "and","or","but","nor",
    "for","yet","so",
    "if","because","therefore",
    "though","although",
    "wherefore","whereas"
  ),

  prepositions = c(
    "of","to","in","unto",
    "by","with","from","for",
    "on","upon","at","into",
    "through","over","under",
    "before","after","against",
    "among","between"
  ),

  auxiliaries = c(
    "shall","will","should","would",
    "may","might","can","could","must",
    "is","are","was","were",
    "be","been","being",
    "hath","has","had",
    "doth","does","did"
  )
)
```

## 6.2 Function Word Frequency and TF-IDF

```{r function-word-features}
function_word_freq <- function(docs, vocab){

  vocab <- sort(unique(vocab))

  mat <- matrix(
    0,
    nrow = length(docs),
    ncol = length(vocab),
    dimnames = list(names(docs), vocab)
  )

  for(i in seq_along(docs)){
    words <- tokenize_text(docs[i], stem = FALSE)
    n_words <- length(words)

    if(n_words == 0){
      next
    }

    tab <- table(words)
    common <- intersect(names(tab), vocab)

    mat[i, common] <- as.numeric(tab[common])
    mat[i, ] <- mat[i, ] / n_words
  }

  mat
}

function_word_tfidf <- function(docs, vocab){

  vocab <- sort(unique(vocab))

  features_by_doc <- lapply(docs, function(x){
    words <- tokenize_text(x, stem = FALSE)
    words[words %in% vocab]
  })

  tfidf_matrix(features_by_doc)
}
```

# 7. Feature Class: POS N-Grams

```{r pos-functions}
load_udpipe_model <- function(model_file = "english-ewt-ud-2.5-191206.udpipe"){

  if(!has_udpipe){
    stop("Package 'udpipe' is required for POS features. Install it with install.packages('udpipe').")
  }

  if(!file.exists(model_file)){
    model_info <- udpipe::udpipe_download_model(language = "english")
    model_file <- model_info$file_model
  }

  udpipe::udpipe_load_model(model_file)
}

pos_ngrams_from_pos <- function(pos, n){

  pos <- pos[!is.na(pos)]

  if(length(pos) < n){
    return(character(0))
  }

  sapply(
    seq_len(length(pos) - n + 1),
    function(i) paste(pos[i:(i + n - 1)], collapse = "_")
  )
}

pos_tfidf_set <- function(docs, n = 3, ud_model){

  anno <- udpipe::udpipe_annotate(
    ud_model,
    x = docs,
    doc_id = names(docs)
  )

  anno <- as.data.frame(anno)

  pos_by_doc <- split(anno$upos, anno$doc_id)

  # Ensure every document appears, even if udpipe drops an empty text.
  pos_by_doc <- pos_by_doc[names(docs)]

  features_by_doc <- lapply(pos_by_doc, pos_ngrams_from_pos, n = n)
  names(features_by_doc) <- names(docs)

  tfidf_matrix(features_by_doc)
}
```

# 8. Feature Class: Punctuation Frequency

```{r punctuation-features}
count_regex <- function(pattern, text){
  m <- gregexpr(pattern, text, perl = TRUE)[[1]]
  if(m[1] == -1) 0 else length(m)
}

punctuation_features <- function(docs){

  cols <- c(
    "comma_per_word",
    "semicolon_per_word",
    "colon_per_word",
    "quote_per_word",
    "period_per_word",
    "punct_total_per_word"
  )

  mat <- matrix(
    0,
    nrow = length(docs),
    ncol = length(cols),
    dimnames = list(names(docs), cols)
  )

  for(i in seq_along(docs)){

    text <- docs[i]
    words <- tokenize_text(text)
    n_words <- length(words)
    if(n_words == 0) n_words <- 1

    comma_n <- count_regex(",", text)
    semicolon_n <- count_regex(";", text)
    colon_n <- count_regex(":", text)
    quote_n <- count_regex("[\"'“”‘’]", text)
    period_n <- count_regex("\\.", text)

    total_n <- comma_n + semicolon_n + colon_n + quote_n + period_n

    mat[i, ] <- c(
      comma_n / n_words,
      semicolon_n / n_words,
      colon_n / n_words,
      quote_n / n_words,
      period_n / n_words,
      total_n / n_words
    )
  }

  mat
}
```

# 9. Feature Class: Sentence Length

```{r sentence-length-features}
sentence_length_bins <- function(docs){

  bins <- c(
    "sent_lt10",
    "sent_10_19",
    "sent_20_29",
    "sent_30_39",
    "sent_ge40"
  )

  mat <- matrix(
    0,
    nrow = length(docs),
    ncol = length(bins),
    dimnames = list(names(docs), bins)
  )

  for(i in seq_along(docs)){

    sents <- unlist(strsplit(docs[i], "[.!?]+"))
    sents <- trimws(sents)
    sents <- sents[sents != ""]

    if(length(sents) == 0){
      next
    }

    lens <- sapply(sents, function(s) length(tokenize_text(s)))

    mat[i, "sent_lt10"]  <- mean(lens < 10)
    mat[i, "sent_10_19"] <- mean(lens >= 10 & lens < 20)
    mat[i, "sent_20_29"] <- mean(lens >= 20 & lens < 30)
    mat[i, "sent_30_39"] <- mean(lens >= 30 & lens < 40)
    mat[i, "sent_ge40"]  <- mean(lens >= 40)
  }

  mat
}

sentence_length_tfidf <- function(docs){

  tf <- sentence_length_bins(docs)

  df <- colSums(tf > 0)
  N <- nrow(tf)
  idf <- log((N + 1) / (df + 1)) + 1

  sweep(tf, 2, idf, "*")
}
```

# 10. Feature Class: Lexical Richness

```{r lexical-features}
lexical_features <- function(docs){

  cols <- c(
    "type_token_ratio",
    "hapax_ratio",
    "avg_word_length"
  )

  mat <- matrix(
    0,
    nrow = length(docs),
    ncol = length(cols),
    dimnames = list(names(docs), cols)
  )

  for(i in seq_along(docs)){

    words <- tokenize_text(docs[i])
    N <- length(words)

    if(N == 0){
      next
    }

    tab <- table(words)

    mat[i, "type_token_ratio"] <- length(unique(words)) / N
    mat[i, "hapax_ratio"] <- sum(tab == 1) / N
    mat[i, "avg_word_length"] <- mean(nchar(words))
  }

  mat
}
```

# 11. Feature Class: Readability

```{r readability-features}
readability_features <- function(docs){

  if(!has_quanteda || !has_quanteda_textstats){
    warning("Packages 'quanteda' and 'quanteda.textstats' are required for readability features. Returning NULL.")
    return(NULL)
  }

  corp <- quanteda::corpus(docs)

  r <- quanteda.textstats::textstat_readability(
    corp,
    measure = c(
      "Flesch",
      "Flesch.Kincaid",
      "ARI",
      "Coleman.Liau",
      "SMOG"
    )
  )

  mat <- as.matrix(r[, -1, drop = FALSE])
  rownames(mat) <- names(docs)

  mat
}
```

# 12. Build All Feature Sets

```{r build-all-features}
build_feature_sets <- function(
    docs,
    include_pos = FALSE,
    include_readability = FALSE,
    ud_model = NULL
){

  feature_sets <- list(

    character = list(
      char3 = char_tfidf(docs, 3),
      char4 = char_tfidf(docs, 4),
      char5 = char_tfidf(docs, 5)
    ),

    word = list(
      word1 = ngram_tfidf(docs, 1, stem = FALSE),
      word2 = ngram_tfidf(docs, 2, stem = FALSE),
      word3 = ngram_tfidf(docs, 3, stem = FALSE)
    ),

    stem = list(
      stem1 = ngram_tfidf(docs, 1, stem = TRUE),
      stem2 = ngram_tfidf(docs, 2, stem = TRUE),
      stem3 = ngram_tfidf(docs, 3, stem = TRUE)
    ),

    function_words_freq = lapply(
      function_word_sets,
      function(vocab) function_word_freq(docs, vocab)
    ),

    function_words_tfidf = lapply(
      function_word_sets,
      function(vocab) function_word_tfidf(docs, vocab)
    ),

    punctuation = list(
      punctuation_freq = punctuation_features(docs)
    ),

    sentence_length = list(
      sentence_freq = sentence_length_bins(docs),
      sentence_tfidf = sentence_length_tfidf(docs)
    ),

    lexical = list(
      lexical_richness = lexical_features(docs)
    )
  )

  names(feature_sets$function_words_freq) <- paste0("fw_freq_", names(function_word_sets))
  names(feature_sets$function_words_tfidf) <- paste0("fw_tfidf_", names(function_word_sets))

  if(include_pos){
    if(is.null(ud_model)){
      ud_model <- load_udpipe_model()
    }

    feature_sets$pos <- list(
      pos3 = pos_tfidf_set(docs, 3, ud_model),
      pos4 = pos_tfidf_set(docs, 4, ud_model),
      pos5 = pos_tfidf_set(docs, 5, ud_model)
    )
  }

  if(include_readability){
    read_mat <- readability_features(docs)
    if(!is.null(read_mat)){
      feature_sets$readability <- list(
        readability = read_mat
      )
    }
  }

  feature_sets
}
```

Build features. For a quick run, POS is off by default because it requires model download/loading.

```{r run-feature-build}
feature_sets <- build_feature_sets(
  docs,
  include_pos = FALSE,
  include_readability = FALSE
)
```

To include POS:

```{r run-feature-build-pos, eval=FALSE}
ud_model <- load_udpipe_model()

feature_sets <- build_feature_sets(
  docs,
  include_pos = TRUE,
  include_readability = FALSE,
  ud_model = ud_model
)
```

# 13. Count Feature Variants and Raw Columns

```{r count-features}
feature_variant_count <- sum(lengths(feature_sets))
feature_column_count <- sum(
  unlist(
    lapply(feature_sets, function(class_group){
      lapply(class_group, ncol)
    })
  )
)

feature_summary <- tibble(
  class = names(feature_sets),
  variants = lengths(feature_sets),
  raw_columns = sapply(feature_sets, function(class_group){
    sum(sapply(class_group, ncol))
  })
)

feature_variant_count
feature_column_count
feature_summary
```

# 14. Cosine Similarity for Every Variant

```{r all-cosines}
feature_cosines <- lapply(
  feature_sets,
  function(class_group){
    lapply(class_group, cosine_sim)
  }
)
```

# 15. Class-Level Similarity

Each feature class may contain several variants. We average those variants within the class first.

```{r class-similarity}
class_similarity <- lapply(
  feature_cosines,
  average_matrices
)
```

# 16. Weighted Ensemble Similarity

## 16.1 Default Class Weights

These are starting weights. They can be changed after validation.

```{r class-weights}
class_weights <- c(
  character = 0.25,
  function_words_freq = 0.12,
  function_words_tfidf = 0.08,
  pos = 0.15,
  word = 0.10,
  stem = 0.10,
  sentence_length = 0.07,
  punctuation = 0.05,
  lexical = 0.04,
  readability = 0.04
)
```

## 16.2 Ensemble Function

```{r ensemble-function}
build_overall_similarity <- function(class_similarity, class_weights){

  active_weights <- class_weights[names(class_similarity)]
  active_weights[is.na(active_weights)] <- 0

  if(sum(active_weights) == 0){
    active_weights <- rep(1 / length(class_similarity), length(class_similarity))
    names(active_weights) <- names(class_similarity)
  } else {
    active_weights <- active_weights / sum(active_weights)
  }

  overall <- Reduce(
    "+",
    Map(
      function(mat, w) mat * w,
      class_similarity,
      active_weights[names(class_similarity)]
    )
  )

  list(
    overall_similarity = overall,
    active_weights = active_weights
  )
}
```

```{r run-ensemble}
ensemble <- build_overall_similarity(
  class_similarity,
  class_weights
)

overall_similarity <- ensemble$overall_similarity
active_weights <- ensemble$active_weights

active_weights
round(overall_similarity, 3)
```

# 17. Nearest-Neighbor Tables

```{r nearest-neighbors}
similarity_to_long <- function(sim_mat){

  as.data.frame(as.table(sim_mat)) %>%
    as_tibble() %>%
    rename(
      doc_a = Var1,
      doc_b = Var2,
      similarity = Freq
    ) %>%
    filter(doc_a != doc_b) %>%
    arrange(doc_a, desc(similarity))
}

nearest_neighbors <- function(sim_mat, k = 5){

  similarity_to_long(sim_mat) %>%
    group_by(doc_a) %>%
    slice_max(similarity, n = k, with_ties = FALSE) %>%
    ungroup()
}

nn_overall <- nearest_neighbors(overall_similarity, k = 5)

nn_overall
```

# 18. Class-Level Nearest Neighbors

```{r class-nearest-neighbors}
class_nn <- imap_dfr(
  class_similarity,
  function(mat, class_name){
    nearest_neighbors(mat, k = 5) %>%
      mutate(class = class_name, .before = 1)
  }
)

class_nn
```

# 19. Export Results

```{r export-results}
out_dir <- "stylometry_outputs"
if(!dir.exists(out_dir)) dir.create(out_dir)

write.csv(
  round(overall_similarity, 6),
  file.path(out_dir, "overall_similarity.csv"),
  row.names = TRUE
)

write.csv(
  nn_overall,
  file.path(out_dir, "overall_nearest_neighbors.csv"),
  row.names = FALSE
)

write.csv(
  class_nn,
  file.path(out_dir, "class_nearest_neighbors.csv"),
  row.names = FALSE
)

write.csv(
  feature_summary,
  file.path(out_dir, "feature_summary.csv"),
  row.names = FALSE
)
```

# 20. Notes on Interpretation

The ensemble computes similarity in three levels:

1. **Variant level**: each feature variant produces a document-by-document cosine similarity matrix.
2. **Class level**: variants inside the same class are averaged.
3. **Overall level**: class-level similarities are combined using class weights.

The default weights should be treated as starting assumptions. For a research pipeline, weights should eventually be tuned by validation tasks, known-author controls, genre controls, or leave-one-out experiments.
